### 2015-04-09 draft.tex
add copy of submitted draft and two pdf images necessary for compiling

\documentclass[11pt,letterpaper, notitlepage]{article}
\usepackage{epsf,graphicx,psfrag}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage[reqno]{amsmath}
\usepackage{endnotes}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{amssymb,enumerate}
\usepackage{dcolumn}
\usepackage{verbatim}
%\usepackage{hypertex}
\usepackage[pdftex, bookmarks, colorlinks=true, plainpages = false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[hmargin=3.0cm,vmargin=3.0cm]{geometry}
%\usepackage{geometry}
\usepackage{nicefrac}
\usepackage{pdfsync}
\usepackage{url}
\usepackage{pst-all}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{graphicx,verbatim,scalefnt}

\newcommand{\ta}{\theta}
\newcommand{\Ta}{\Theta}
\newcommand{\ga}{\gamma}
\newcommand{\un}{\underline}
\newcommand{\ov}{\overline}
\newcommand{\de}{\delta}
\linespread{1.05} % Palatino needs more leading (space between lines)
\title{Unrecognized States: A Theory of Self-Determination and Foreign Influence}

\author{Kristy Buzard\\
Syracuse University\\
\\
Benjamin A.T. Graham \\
University of Southern California\\
\\
Ben Horne\\ 
University of California, San Diego}

\date{\today}
\begin{document}
\begin{titlepage}
\maketitle
\thispagestyle{empty}
\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}

\begin{abstract}
\noindent Unrecognized states are characterized by stagnant or crumbling economies and political instability, often serve as havens for illicit trade, and challenge the territorial sovereignty of recognized states. Their persistence is both intellectually puzzling and normatively problematic, but unrecognized statehood can be a remarkably stable outcome, persisting for decades. Our four-player model reveals that unrecognized statehood emerges as an equilibrium outcome when a patron state is willing and able to persistently invest resources to sustain it. %Counterintuitively, patrons will make these investments even when unrecognized statehood is not their most preferred outcome. 
We assess options available to actors in the international community who seek to impose their preferred outcomes in these disputes and find that, although sanctions are the most frequently employed, they can often lead to renewed conflict instead of the intended resolution.
\end{abstract}

\end{titlepage}
\setcounter{page}{1}

\doublespacing
\section{Introduction}
South Ossetia is an archetypical unrecognized state -- characteristic of those regions of the world in which non-state actors control territory and govern populations. From 1990 to 1992, Ossetian rebels fought a successful secessionist civil war against the Georgian government that ended with a ceasefire and left the rebels in \textit{de facto} control of much of the region of South Ossetia, which sits along Georgia's northern border with Russia. In the 22 years since the ceasefire was signed, South Ossetia has functioned as an unrecognized state, governing its own affairs but unrecognized by foreign nations.\footnote{South Ossetia, along with Abkhazia, was recognized by Russia and Nicaragua in 2008, Venezuela and Nauru in 2009, and Tuvalu in 2011.} The Georgian government maintains its claim to the territory of South Ossetia, while the South Ossetians continue to seek international recognition of their independence. In 2004, the Georgian government began intermittent efforts to close trade with the separatist region, and in 2008, following escalating provocations from the Ossetian side, Georgia attempted to reclaim the territory by military force. Russian troops acting in support of the Ossetians quickly crushed the would-be reconquest, and the status quo was restored.%\footnote{Good sources on the history of the conflict include Lynch 2004, Asmus 2010, and a series of reports by the International Crisis Group.}

The international system is characterized by a norm under which seceding entities are rarely recognized by foreign states unless they are first recognized by the home state.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states. In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).}%The higher the costs of secession, the greater the security of existing states. 
Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.

Persistent unrecognized statehood is both intellectually puzzling and normatively undesirable. It is an extremely costly outcome for the secessionists, the home states from which they are attempting to secede, and the international community more broadly. Nonetheless, unrecognized statehood has been a recurring phenomenon since WWII, with half a dozen current unrecognized states having existed for more than 20 years.\footnote{Abkhazia, Ngorno-Karabakh, Transnistria, Somaliland, South Ossetia, and the Turkish Republic of Northern Cyprus.} Much of the existing literature treats unrecognized statehood as either a transient phenomenon or simply as a failure to reach recognized statehood. In contrast, we argue that unrecognized statehood is an important outcome in its own right, and one that is potentially extremely stable over time. A major contribution of this paper is the demonstration of unrecognized statehood as a long-run equilibrium outcome of the strategic interaction between not just the secessionists and the home government, but also the international community and states or non-state actors who act as patrons of the secessionists. We show that unrecognized statehood can be a stable outcome despite its persistently high costs to all actors involved.

The costs to the secessionists are arguably the highest. Non-recognition locks unrecognized states out of a states-only club whose members enjoy benefits in terms of both security and economic integration. Unrecognized states do not benefit from the norm of territorial integrity that reduces the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Non-recognition also prevents entry into multilateral and bilateral trade agreements, dramatically reduces access to foreign aid, and excludes their territory from international legal frameworks, making it harder to secure foreign investment (Milhalkanian 2004; Caspersen 2013: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King 2001). However, as Moldova deepens ties with the EU, Transnistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu 2011).} %In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).

The home state bears costs of non-settlement as well, although they are not as high as those facing the unrecognized state. Instability in the border region and the diversion of military resources to monitor the \textit{de facto} border are costly, as is maintenance of economic sanctions against the unrecognized state. In Azerbaijan, where an oil boom sent GDP soaring in the mid-2000s, the unresolved secessionist conflict in Nagorno-Karabakh led a massive military buildup to take precedence over other government spending.\footnote{Military spending increased 51\% in 2004-2005, and went up another 82\% in 2006 (International Crisis Group 2007)} Foreign investors are more wary of investing in states that do not effectively control their own territory, and there are diplomatic costs as well. For Georgia, failure to resolve its outstanding secessionist conflicts has hindered progress toward NATO accession.

%Unrecognized statehood is a thorn in the side of the state system. Many unrecognized states are havens for smuggling and trafficking illicit goods, they are prone to violent conflict, and their very existence challenges the norm of territorial sovereignty (King 2001; Stanislawski 2008). Like most unrecognized states, South Ossetia pays high costs for non-recognition. Its political survival is tenuous, its licit economy is in shambles after years of isolation and instability, and yet, if the history of similar entities are any guide, its chances of eventual recognition are slim. Stalemate imposes similar, though less severe, costs on home states (e.g. Georgia) as well.

%\subsection{Non-Recognition and Its Costs}
%Since WWII, the path to recognized statehood has been narrow, and securing territorial control is not sufficient to gain recognition. The international community has been exceedingly clear that a national right of self-determination applies only to cases of decolonization: it does not extend to cases of secession.  In 1970, UN secretary general U Thant said, ``The United Nations has never accepted and does not accept and I do not believe will ever accept the principle of secession of a part of its member states."\footnote{Pegg 1998: 8} In cases where the home state's sovereignty is disputed or where the home state government commits mass atrocity crimes, some states may choose to grant recognition, but there remains a strong norm of home-state veto.That is, secessionist entities are rarely recognized by foreign states unless they are first recognized by the home state from which they are attempting to secede.  This has lead to the existence of an odd group of state-like entities that control territory and govern population but are not recognized as states and cannot participate in the international system. We refer to these entities as unrecognized states, though, variously specified, they are also referred to states-within-states\footnote{Kingston and Spears 2004} \textit{de facto} states,\footnote{Pegg 1998, Lynch 2004} or almost-states.\footnote{Stanislawski 2008}

%Since WWII, the path to recognized statehood has been narrow, and securing territorial control is not sufficient to gain recognition. Seceding entities are rarely recognized by foreign states unless they are first recognized by the home state from which they are attempting to secede.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states. In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm of home state veto emerges because almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby making secession less likely.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize."} The higher the costs of secession, the greater the security of existing states.

%Unrecognized statehood is a particularly compelling outcome from secession in part because while the secessionists have succeeded in securing control of (most of) the territory they aspire to rule they are not recognized by most other states, trapping them in a costly limbo.

%Unrecognized states are subject to high costs of non-recognition generated by an international system extremely hostile to non-state territorial units. The modern state system is based on institutions designed to facilitate peaceful and economically beneficial relations between like actors, and on the empowerment and legitimation of states through mutual recognition (Spruyt 1996). The norm challenged most directly by unrecognized states is that which, in the post-WWII era, has become most fundamental to it -- the norm of territorial sovereignty. Almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, and thereby makes secession less likely.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize."} The higher the costs of secession, the greater the security of existing states.

%The most direct manifestation of this systemic hostility to secession is the norm of states refusing to recognize a seceding entity unless the home state recognizes it first.\footnote{For a thorough international legal discussion of the issue, see Crawford 2006.} This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states. In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty. However, the norm is strong and gives recognition by the home state its significance: without recognition by the home state, the post-WWII norm is for other states is to withhold recognition as well.\footnote{Caspersen (2012) traces the history of collective non-recognition, which begins after WWI and solidifies after WWII. In terms of the model presented below, the payoff from territorial control before WWI was equivalent to the payoff from recognized statehood. Players did not need to participate in the game described in the model. We limit our discussion to the post-WWII era, when the norm of home state veto is fully developed.}

%Abkhazia, Ngorno-Karabakh, Transnistria, Somaliland, South Ossetia, and the Turkish Republic of Northern Cyprus have all controlled territory for over a decade without gaining recognition by more than four other states.

%Non-recogition locks unrecognized states out of a states-only club whose members enjoy benefits in terms of both security and economic integration. For recognized states, the norm of territorial integrity lowers the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Recognition also allows entry into multilateral and bilateral trade agreements, dramatically bolsters access to foreign aid, and incorporates territory into international legal frameworks that make it easier to secure foreign investment (Milhalkanian (2004); Caspersen 2012: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King 2001). However, as Moldova deepens ties with the EU, Transinistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu 2011).} In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).

%Most unrecognized states are low income countries, and the costs of non-recognition grow over time as their economies further decay. Because the resources of unrecognized state governments are primarily focused on security, unrecognized states are generally characterized by a lack of public investment in infrastructure and education. They also face severe brain drain and a lack of private investment caused by the absence of security and restrictions on trade.

%Non-settlement has costs for the home state as well, but they are simply not as high as those facing the unrecognized state. Instability in the border region and the diversion of military resources to monitor the \textit{de facto} border are costly, as is maintenance of economic sanctions against the unrecognized state. In Azerbaijan, where an oil boom sent GDP soaring in the mid-2000s, the unresolved secessionist conflict in Nagorno-Karabakh lead a massive military buildup to take precedence over other government spending.\footnote{Military spending increased 51\% in 2004-2005, and went up another 82\% in 2006 (International Crisis Group 2007)} Foreign investors may be more wary of investing in states that do not effectively control their own territory, and there are diplomatic costs as well. For Georgia, failure to resolve its outstanding secessionist conflicts has hindered progress toward NATO accession.

The purpose of this paper is to identify the mechanisms that sustain unrecognized statehood as a stable equilibrium and to evaluate potential strategies through which actors may preserve the stability of this equilibrium, or disrupt it by inducing war or negotiated settlement. While there are many game theoretic treatments of civil conflict, we believe ours is the first formal model in which unrecognized statehood is addressed as an outcome.

We develop a novel, four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome. These outside actors are a patron, which prefers independence to reunification, and a fourth player we denote as player $c$ who represents one or more members of the international community. Player $c$ is assumed to prefer reunification to independence.

The game is infinitely repeated and proceeds as follows. At the beginning of each period, the payoffs for the secessionists from unrecognized statehood---to which we refer as the status quo---deteriorate by a fixed amount, representing the ongoing costs of non-recognition. Next, the patron and international community can, in turn, make investments to augment the payoffs of the home state government and the secessionists. Finally, the home government and secessionists play a simultaneous stage game in which each decides whether to initiate conflict, maintain the status quo of unrecognized statehood, or cede the issue of status.

We provide a set of conditions under which perpetual unrecognized statehood is the outcome of a subgame-perfect equilibrium of this game. %This requires that the patron is both willing and able to make investments sufficient to (a) offset the deterioration in the secessionists' status quo payoffs and (b) deter the international community from making investments that would induce a negotiated settlement. That is, it must be that the prospect of reunification is more important for the patron to avoid than for the international community to achieve. Additionally, in order for the status quo to persist, preferences must be such that recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve; otherwise, the international community will not be able to deter the patron from making investments to instigate recognition.
Somewhat counterintuitively, in this equilibrium a patron will expend resources every period if necessary to sustain the status quo of unrecognized statehood even when the patron would prefer that the secessionists achieve recognized independence. This occurs when the international community's preferences against recognition are sufficiently strong---as is almost always likely to the case---to make it prohibitively expensive for the patron to facilitate recognition in the face of counteractivity by the international community.

In the end, the status quo outcome is observed when the patron---who faces a much tighter budget constraint---outspends an international community with access to a much larger pool of resources. This occurs because each outside actor's willingness to pay to achieve its most preferred outcome is outstripped by the other's desire to avoid its least desired outcome. We are left with an unresolved conflict with no clear winner, and even if there is no official loser, all parties pay significant costs.

We go on to analyze a tool that has been one of the most frequently employed by the international community in these types of territorial disputes: economic and military sanctions. We show that, although the goal of such sanctions is presumably to encourage secessionists to cede the issue of status, their most pronounced effect is to increase the range of conditions under which war occurs. %Instead, the patron may compensate for the additional cost of sanctions through increased investments or the secessionists may turn to fighting instead of ceding if the gap between the payoffs from the status quo and reunification are too large. Alternatively, if the secessionists' military capabilities are sufficiently weakened, the home state government may be emboldened to initiate fighting itself.
%Thus while sanctions can be effective in only a limited range of cases, they can counterproductively lead to increased conflict in others. 
In contrast, we argue that members of the international community (player $c$) are capable of intervening through other means to overcome the influence of the patron and induce settlement if they are sufficiently motivated to do so. Policy options include the enforcement of bargains and the granting of positive inducements for settlement, rather than direct pressure on the unrecognized state via sanctions.

\subsection{The Empirical Landscape}

The contributions of this paper are primarily theoretical, but it is useful to introduce the empirical landscape that informs our model. %We reference individual cases throughout to illustrate our theory and to justify its assumptions. 
Table 1 catalogs the post-WWII cases of unrecognized statehood, separated according to their current status. We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state.\footnote{This excludes, for example, the territories governed by FARC in Colombia because FARC does not seek recognized statehood and includes cases, like South Ossetia, where recognition is obtained by only a handful of other states. We also exclude entities, like the Slovak Republic, whose secession was not opposed by the state from which it seceded. We acknowledge that other scholars might differ with us with regard to the inclusion or exclusion of specific cases. We assert that our model is a useful tool for analyzing each of the cases in this table; it may be useful in explaining some of the ambiguous cases we exclude as well.} All six current unrecognized states have now been unrecognized for over twenty years.

%\pagebreak
\begin{center}
\includegraphics[width=6.25in]{Table_1_082214.pdf}
\end{center}

Of those cases in the table, all but Moheli controlled territory and govern(ed) populations for at least two years. These cases represent the most successful cases of attempted secession in the Post-WWII era, and yet eventual military defeat by the home state is still the modal form of resolution. Recognition by the home state is similarly rare, occurring in only three cases and never except as a direct result of concessions won on the battlefield. In cases where recognition by the home state or the right to a referendum on independence is not secured as part of the initial peace agreement, it has not historically been forthcoming. Only four cases of negotiated reunification are observed: secessionists who are strong enough to secure and retain territorial control are rarely willing to surrender their independence at the bargaining table, even though the chances of eventual recognition are vanishingly slim. Thus, the number of long-running, costly stalemates has been substantial, most of them eventually ending in military reconquest by the home state.

\subsection{Unrecognized States in the Literature}

%Political scientists care a great deal about both civil conflict and about state formation, but unrecognized states do not sit neatly in either literature. The literature on civil war is focused primarily on war onset (e.g. Fearon and Laitin 2003; Hegre and Sambanis 2006), war intensity and duration (e.g. Collier and Hoeffler 2004; Cunningham 2006), and the durability of post-conflict peace (e.g. Hartzell, Hoddie and Rothchild 2001). Unrecognized statehood does not fit neatly into these areas of study because, while unrecognized states begin, and often end, through violent conflict, periods of unrecognized statehood generally contain little, if any, fighting. %Unrecognized statehood often represents a relatively stable end to civil war, but the cessation of fighting is not coterminous with resolution of the conflict in any meaningful way. We argue that persistent unrecognized statehood is not a successful resolution to secessionist civil war; it is a costly and normatively bad outcome in its own right, and it needs to be analyzed as such. 
%Unrecognized states also fall outside most treatments of state formation, because most unrecognized states never achieve recognition (or have not achieved it yet).\footnote{For example, Roeder (2007) treats unrecognized states simply as failures to gain recognized statehood, not as outcomes to be analyzed in their own right.}

The political science literature on civil war is focused primarily on war onset, %\footnote{e.g. Fearon and Laitin 2003, Hegre and Sambanis 2006}
war intensity and duration, %\footnote{e.g. Collier and Hoeffler 2004; Cunningham 2006} 
and the durability of post-conflict peace. %\footnote{e.g. Hartzell, Hoddie and Rothchild 2001} 
Unrecognized statehood does not fit neatly into these areas of study because, while unrecognized states begin, and often end, through violent conflict, periods of unrecognized statehood generally contain little, if any, fighting. Unrecognized states also fall outside most treatments of state formation, because most unrecognized states never achieve recognition (or have not achieved it yet).\footnote{For example, Roeder (2007) treats unrecognized states simply as failures to gain recognized statehood, not as outcomes to be analyzed in their own right.}

The first literature to address unrecognized states directly was grounded in comparative politics, and a robust area-studies literature exists around each of the current cases of unrecognized statehood.\footnote{Lynch (2004), King (2001), Stanislawski (2008), and Bakke et al. (2013) provide notable treatments of the Former Soviet cases, and a pair of edited volumes by Bahcheli, Bartmann, and Srebrnik (2004) and Kingston and Spears (2004) each compile broader sets of case studies.} More recent literature has addressed wider ranges of cases and made important conceptual progress identifying patterns and commonalities across cases (e.g. Pegg 1998; Kolst\o \ 2006; Geldenhuys 2009; Caspersen and Stansfield 2011; Caspersen 2013). However, the literature continues to lack a clear general theory specifying the conditions under which this outcome emerges and persists.

One of the major contributions of this article is to provide a unified analytic framework for understanding the mechanisms sustaining unrecognized statehood as a stable equilibrium. By modeling unrecognized statehood formally, we move away from a case-by-case treatment toward development of a rigorous general theory. The model allows us to assess the conditions under which unrecognized statehood persists, and those under which war and negotiated settlement occur. It also allows us to evaluate various strategies available to actors, particularly states and coalitions of states who want to facilitate a peaceful and permanent resolution.

Although we are not aware of any formal models of unrecognized statehood there is a growing economics literature that employs game theoretic models of conflict. An early model of territorial disputes in which the divisibility of the contested territory is an important parameter is Grossman (2004). In Hirschleifer, Boldrin and Levin (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions. In our context, the contested territory is indivisible; this implies that peace would require very large concessions so that this pathway of using small compromises to avoid conflict is not available.

In contrast to Bueno de Mesquita (2013), which exemplifies a class of literature that examines how rebel groups make decisions about resistance strategies, we begin our analysis at the point where initial rebellion has already been successful to the point of gaining \textit{de facto} control of the disputed territory.

Another strand of the literature explicitly models the bargaining process between potential adversaries, either with asymmetric (Yared 2010) or complete information (Jackson and Morelli 2007, Schwartz and Sonin 2008, Bevi\'{a} and Corch\'{o}n 2010, McBride and Skaperdas 2014). Acemoglu, Golosov, Tsyvinski and Yared (2012) develop a dynamic model of resource wars featuring limited commitment. We abstract from questions of resource allocation and bargaining in order to focus on the dynamics created by the involvement of outside players: in the case of unrecognized states, the international community and patron states.

We next turn to presenting the model. In Section 3, we present a set of conditions under which unrecognized statehood is a stable, long-run equilibrium outcome. In Section 4, we analyze the potential effects of sanctions on the resolution of territorial disputes while Section 5 explores extensions of the model. Section 6 evaluates policy implications of the model, assessing various strategies available to outside actors that seek to induce their preferred outcomes. Section 7 concludes.

\section{A Model of Unrecognized Statehood} 
We model a dispute over a piece of territory that is controlled by a secessionist group and also claimed by a home state. The central issue of contention, independence vs. reunification, is both difficult to divide and highly valued by both sides. The secessionists seek recognized statehood, the home state seeks reunification, and these demands do not vary over time. The side payments that can be offered in exchange for the opponent's surrender of the independence/reunification issue are sharply limited by the absence of large concessions that can credibly be made (Walter 1997, 2002; Schultz 2010).

A major innovation of our model is the introduction of international actors. While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis 2000; Regan 2002; Balch-Lindsay, Enterline, and Joyce 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter 1997, 2002).

The model presented here allows us to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}

\subsection{The Players}

We construct a model with four players: the secessionist elite ($s$); the central government of the home state ($g$) from which $s$ is attempting to secede; and two other players, the patron ($p$) and the international community ($c$). The latter two are outside actors---states, groups of states and/or individuals acting in concert---that have interests in the outcome of the attempted secession.

Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders. As discussed earlier, most states in the international system prefer any given secessionist conflict to be resolved by reunification. In practice we believe that in most cases there are many states that may act as player $c$ in our model, and often we observe groups of states like the OECD or the UN acting in this capacity.

Player $p$ most prefers recognized independence, aligning its interests with the secessionists. We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium. Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan, Gleditsch and Cunningham 2012), e.g. as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g. Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g. Armenia's support of Nagorno-Karabakh).\footnote{While annexation is appealing to many patrons (and some unrecognized states), annexation is not an outcome we model. International norms against irredentism are strong, and the costs of annexing an unrecognized state appear to be prohibitively large in most cases (e.g. Zacher 2001). Prior to Russia's annexation of Crimea, all post-WWII cases of annexation involved the annexation of colonial territory, rather than parts of the metropole. Israel's occupation of the Golan Heights and the West Bank are possible exceptions here.}

We acknowledge that there may exist patrons whose most-preferred outcome is the status quo. This naturally makes a status quo equilibrium easier to achieve, as there is an actor who strictly prefers the status quo and can expend resources to make it more likely. We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely. Even in this circumstance, the status quo remains an equilibrium outcome.

\subsection{Timing and Structure of Interaction}

The game begins at a status quo in which the secessionist elite controls at least some of the disputed territory, but cannot gain international recognition unless the central government cedes its claim to the territory. This condition is archetypical of cases in which a militarily successful war of secession ends in a ceasefire.

The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur. There are an infinite number of periods $n=1,2,\ldots$, with play proceeding until an absorbing state is reached.

Future payoffs are discounted with parameters $\delta_i$, where 1/$\delta_i$ is player $i$'s discount rate for $i \in (g, s, c, p)$, $0\leq\delta_i\leq1$. Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \delta_i^{n-1}$.

In each period, play proceeds as follows:

\begin{figure}
\includegraphics{Timeline2.pdf}
\caption{Timeline}
\end{figure}

\begin{enumerate} 
\item The status quo payoffs for $s$ are reduced by $\mu$.

\item $p$ chooses a level of resource expenditure $R_{pn}\in \mathbb{R}^+$ in period $n$ and invests it to alter the payoffs of players $s$ and/or $g$.

\item $c$ chooses a level of resource expenditure $R_{cn}\in \mathbb{R}^+$ in period $n$ and invests it to alter the payoffs of players $s$ and/or $g$.

\item $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
\end{enumerate}

In order to fully describe the structure of the game, we first introduce the stage game payoffs.

\subsubsection{Stage Game Payoffs}

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $\Omega_n$ &$\Omega_n$&$W_{gn}, L_{sn}$ \\ \hline
	Status Quo& $\Omega_n$& $Q_{gn}, Q_{sn}$ &$W_{gn}, L_{sn}$ \\ \hline
	Cede&$L_{gn}, W_{sn}$& $L_{gn}, W_{sn}$ &$Q_{gn}, Q_{sn}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e. the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2). We interpret one player ceding as that player ceding the issue of status (independence vs. reunification) in exchange for some set of (relatively small) payments from the opposing player. For example, if the secessionists cede, the secessionist region is reunified with the home state and the payoffs are $L_s$ for the secessionists and $W_g$ for the home state. Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.

If both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period. In this case neither player has demonstrated a willingness to give up more than the other. Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}

There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously. We denote the payoffs of war as a lottery $\Omega$.\footnote{Because the unrecognized state already controls territory and the \emph{de facto} borders are armed, there is likely only a small advantage to be gained by attacking first for either side. Therefore, we argue it is not essential to differentiate between these war scenarios analytically.} This lottery determines one of three potential outcomes: outright victory by either of the players, which are both absorbing states, and an indecisive war where costs of war are borne but then the game returns to the status quo in which the game continues to the next period. Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government. Therefore, outright victory requires more than simply securing control of the disputed territory (which the secessionists have already done when the game begins) and requires the ability to impose the terms of settlement. In practice, this likely involves the capture of the home state capital and/or the overthrow of the home state government.

For probabilities $p_1$ of outright victory, $p_2$ of loss and $1-p_1-p_2$ of non-decisive war, player $i \in \{s, g\}$ in period $n$ with a fixed cost of war $\zeta_i$
faces war lottery $ \omega_{in} \equiv (p_1 (W_{in}-\zeta_i), p_2( L_{in}-\zeta_i), (1-p_1-p_2)(Q_{in}-\zeta_i))$. $\Omega_n \equiv (\omega_{gn}, \omega_{sn}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.} Players are assumed to approach war lotteries as expected values.

If both $s$ and $g$ play Status Quo, then the status quo persists. The status quo payoffs for the secessionists are modeled as steadily declining due to the costs of non-recognition discussed in the introduction. As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists. Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_{s,n+1} = Q_{s,n} - \mu$. Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments. Other payoffs remain unchanged from period to period; they can only be altered by the investments of players $p$ or $c$, e.g. $L_{s,n} = l_{s,n} + R_{p,n} + R_{c,n}$ and $l_{s,n+1} = L_{s,n}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$. %who can expend resources ($R$) to increase a payoff at the rate $\gamma$.  For example, if $p$ expends resources to aid the secessionists, the payoff $Q_s$ becomes $Q_s+\gamma R_p$ in the round that the resources are expended.

\subsubsection{Payoffs for the Patron and International Community}

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_{pn}$ to increase the secessionists status quo payoffs, the beginning-of-period payoff $q_{sn}$ becomes end-of-period payoff $Q_{sn} = q_{sn} + R_{pn}$. We can write analogous expressions for $q_{gn}, \ l_{sn},\ l_{gn},\ w_{sn}$, and $w_{sn}$, and the international community can also make investments to increase whichever of these payoffs they like. $B_{pn}$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

We assume player $c$ prefers peace to war. For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war. This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.

We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$. $X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state. Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.

Player $p$'s payoffs in period $n$ are $U_{pn}= \alpha X+\lambda Y -R_{pn}$, while the payoffs for player $c$ are $U_{cn}= \beta X+\nu Y -R_{cn}$, with both payoffs denoted in currency units.

The patron opposes reunification while player $c$ prefers it, so $\alpha$ is negative and $\beta$ is positive. The patron prefers recognized independence so $\lambda$ is positive. Player $c$ is averse to the creation of new states so $\nu$ is negative.

The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources. As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.

%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.  If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.  Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.

\subsubsection{The Low Payoffs from Ceding}
Aside from the central assumption that the government of the home state most prefers reunification and the secessionists most prefer independence accompanied by international recognition, we will for the most part leave the model general enough to incorporate a range of preferences and capabilities for players $g$ and $s$. %Our results will specify parameter ranges for these preferences over which outcomes of interest---such as perpetual unrecognized statehood---can be supported in equilibrium.
We do, however, assert that the payoffs for the party that cedes the issue of status (independence vs. reunification) are consistently low. This reflects a combination of two factors. First, the issue of status is indivisible\footnote{Either the unrecognized state has sovereignty over its territory and is co-equal with the home state, or the secessionist region (and its government) are subordinate to the central government.} and highly valued by each side, making its surrender undesirable. Various forms of ethno-nationalism often motivate secession, and the values attached to independence by secessionists (and to reunification by citizens of the home state) are generally large relative to the values placed on economic prosperity and other goals. Second, many of the payments that could be offered are not credible (e.g. Schultz 2010).

The difficulty of making credible payments in exchange for status is one clearly demonstrated in the civil war literature (e.g. Licklider 1995; Walter 1997, 2002; Fearon and Laitin 2007; Doyle and Sambanis 2006). Unrecognized states generally constitute ``sons of the soil" conflicts in which the central government cannot credibly commit to preserving the local demographic and political dominance of the secessionist elite once the disputed territory reverts to central government control (Weimer 1978; Fearon 2004). While the central government might initially grant the secessionist elite a high level of autonomy in exchange for agreeing to reunification, the level of autonomy is likely to decrease over time, perhaps quite quickly. The payments that can be offered by the unrecognized state to the central government are similarly small or unenforceable: once recognition is achieved, the secessionists have little reason to uphold any prior commitments.

%At the time of secession, ethnic Akbhaz made up a minority of the population of Abkhazia, but by the late 1990s they controlled almost all political posts in the \textit{de facto} government of the region (Cornell 2001; Wooleh 2006). In 2004, the basket of payments offered by the Georgians in exchange for reunification included a provision guaranteeing that ethnic Abkhaz would retain a majority in the regional parliament, even if the return of internally displaced persons (IDPs) once again placed ethnic Abkhaz in a minority demographic position in the region. The promise, however, was not very meaningful. First, even if the promise were upheld, it would still mean a step back from the total dominance the ethnic Abkhaz currently enjoy in the region. Second, if Georgian IDPs returned, they may demand and receive a more equitable system of representation. These concerns are not abstract; this type of reneging has already occurred in cases that did reach settlement.

The case of Gagauzia illustrates the commitment problem nicely. Gagauzia achieved \emph{de facto} independence at the time of the Soviet Union's collapse, but agreed to rejoin Moldova in 1994 as an autonomous region. While Gagauzia was granted substantial autonomy under the Moldovan Law on the Special Legal Status of Gagauzia, when the governor of Gagauzia, Dmitrii Croiter, moved to assert these powers in 1999, the Moldovan government balked. By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government. The Moldovan government jailed a number of other Gagauz politicians, and while \emph{de jure} Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the \emph{de facto} level of autonomy has been limited by continued central government meddling in less-than-free regional elections. The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.
%WE CAN CUT THE GAGAUZ EXAMPLE IF NECESSARY -- BG July 2014

%The payments that can be offered by the unrecognized state to the central government are similarly small or unenforceable. To return to the Abkhaz example: Under a scenario in which Georgia recognizes Abkhazia as an independent state, the ethnically Georgian region of Gali, currently under Abkhaz control, would likely rejoin Georgia, Russian troops would be (at least temporarily) expelled, and compensation might be paid to displaced Georgians, but other side payments are difficult to picture. Abkhazia might promise Georgia privileged access to Abkhaz ports, or promise to keep Russian troops out of its territory permanently, but once recognition is granted, any such promises could easily be reneged upon.

%Furthermore, while relinquishing territorial claims relieves the home state of a persistent source of instability, conceivably supplementing $g$'s payoffs from ceding, it also would increase the probability of future attempts at secession by other regions (Walter 2006). The lack of enforceable side payments and the ambiguous impact on stability make ceding the issue of status a low-payoff option for $g$, just as it is for $s$.

%Among the means through which $p$ or $c$ can spend resources to alter the payoffs to $g$ and $s$ is to enforce bargains and guarantee future concessions by either party, raising the payoffs from ceding. For example, if both the home state and the secessionists prefer reunification with autonomy rights to both war and continued stalemate, $c$ can spend resources to enforce an agreement in which the secessionists are promised specific autonomy rights and $c$ agrees to ensure that these rights are later retained. Should the home state later attempt to revoke the promised autonomy, $c$ can levee sanctions or employ other coercive measures to deter this action. The promises and pitfalls of this approach are discussed in the section on policy implications of the model.

\section{Explaining Outcomes of Secessionist Conflicts} 
\label{sec:main}
%The game has three outcomes that we will consider, each of which can be characterized by a class of equilibria. The outcome of greatest interest is both players choosing the status quo in perpetuity; we will also examine the two absorbing states, reunification and recognition by the home state (i.e. recognized statehood).\footnote{Other classes of equilibria are possible, for example it is theoretically possible for the players to agree to a lottery between the absorbing states which would yield a higher expected payoff than the status quo. However, in practice there is a credibility issue: the losing party does not have an incentive to cede if they lose the lottery. Such strategies are, in any case, not analyzed here.}

Despite the preferences of the international community for peace, the most common outcome of secessionist conflicts in the post-WWII period has been reunification with the home state via outright military reconquest. By contrast, reunification through negotiated settlement has been very rare. We will explore potential explanations for this seeming inability of the international community to achieve its aims in Sections~\ref{sec:alt} and \ref{sec:ext}. First, we turn to the outcome that we find the most puzzling: perpetual unrecognized statehood.

\subsection{Analysis of the ``Status Quo'' Equilibrium} 
\label{sec:mainproof}
Unrecognized states are frequently viewed as temporary phenomena or as non-equilibrium outcomes attributable to players' misperceptions of the strategic situation, or their fundamental irrationality. Our central result shows that unrecognized statehood can be an equilibrium outcome capable of being sustained in perpetuity by fully rational, perfectly informed actors. This is true even when there is no actor that prefers unrecognized statehood as a first-best outcome.

\begin{quotation}
\noindent \bf \emph{Proposition 1:} \rm  \emph{There exists an equilibrium in which the outcome is perpetual unrecognized statehood. The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\delta_c}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}

\emph{The following are sufficient conditions for such an equilibrium: }

\begin{enumerate}
\item \textit{For both players $g$ and $s$, $Q_{in} \geq L_{in} \ \forall n$: in each period, remaining in the status quo is better than ceding.}

\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta_i} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta_i} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}

\item \textit{$\frac{-\alpha}{1-\de_p} \geq \frac {\beta}{1-\delta_c}$: reunification is more important for the patron to avoid than for the international community to achieve.}

\item  \textit{$\frac {-\nu\cdot p_{1s}}{1-\delta_c} \geq \frac{\lambda \cdot p_{1s} + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}

\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}

\item \textit{$B_{pn} \geq \mu, \ \forall n>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6). If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $n> 1$ compared to period $1$, (6) could be binding.}

\end{enumerate}

\end{quotation}

These conditions for the status quo are sufficient but not necessary. For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run. The inequality \emph{will} bind for some set of periods $n \geq 1$ because the secessionist payoffs from the status quo decrease over time. In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.

There are many other potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}). As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity. We will use the solution concept of subgame perfect equilibrium to show that, given the sufficient conditions in Proposition 1, at least one status quo equilibrium will exist.\\

\noindent {\bf Proof of Proposition 1} \\
Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3). Although period $n$ may be reached because in period $n-1$ fighting resulted in a stalemate, both players ceded, or the status quo had been maintained, the strategic landscape in period $n$ is the same. Thus all periods in which the players are able to move are in an equivalence class in which the only possible strategic difference is in the value of the state variables.

Analysis of the posited equilibrium proceeds most naturally by backward induction within each period. Thus we begin with the stage game between the government and secessionists.

\begin{quotation}
\noindent \textit{{\bf Lemma 1:} Conditions 1 and 2 are sufficient for status quo to be the outcome of Stage 3 in any period.} 
\end{quotation}

The proofs of the lemmas are in the Appendix.

Lemma 1 establishes ranges for the payoffs for players $g$ and $s$---gross of investments by the outside players---in which the status quo outcome can occur. The rest of this section tackles the more difficult task of determining the incentives and actions of the outside actors---that is, the patron state and the international community---to impact those payoffs.

To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{sn},q_{gn},w_{sn},w_{gn},l_{sn},$ and $l_{gn}$.\footnote{In the case of $w_{sn}$ and $w_{gn}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e. \emph{Recognition, Status Quo}, or \emph{Reunification}---they will target given that they will choose the most efficient way to alter the payoffs of players $s$ and $g$.

%Given their preference orderings, some of these can be ruled out. Player $p$ will not invest to increase $s$'s payoffs from ceding ($l_{sn}$) or $g$'s payoffs from war when it could do so for $s$. It will also not invest in $q_{gn}$ when it can instead increase $l_{gn}$, leaving the list of rational investment choices for $p$ as $q_{sn},l_{gn}$, and $w_{sn}$. Symmetrically, $c$ will not invest in $q_{sn},L_{gn}$, or $w_{sn}$. Given the assumption that the international community follows a norm of encouraging peaceful resolution, it also will not invest in $w_{gn}$, leaving only $l_{sn}$ and $q_{gn}$ as rational investment options for player $c$.

Lemma 2 addresses potential efforts by the international community to influence the outcome toward \emph{Reunification}:

\begin{quotation}
\noindent \textit{{\bf Lemma 2:} When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.}
\end{quotation}

\noindent Condition 3 provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome. To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

Consider period 1. When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\delta_c}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest. That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\delta_c} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment. Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.

In periods $n > 1$, there are two cases to consider. Either (a) the buffer created in period $n-1$ was precisely the necessary $\frac{\beta}{1-\delta_c}$, or (b) the buffer is larger than $\frac{\beta}{1-\delta_c}$. In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\delta_c}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\delta_c}$ provides lower welfare to player $p$. Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $n$. However, because in each period $q_{sn}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a). Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $n > 1$(Condition 6).

With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending for or against recognition of the secessionist state:

\begin{quotation}
\noindent \textit{{\bf Lemma 3:} When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.}
\end{quotation}

In the equilibrium under consideration here, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome. This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved. The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Conditions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}

One last possibility is ruled out by Lemma 4: that the patron would invest to encourage the secessionists to fight.

\begin{quotation}
\noindent \textit{{\bf Lemma 4:} Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.}
\end{quotation}

Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community. All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.

Thus, given Conditions (1)-(6) hold, equilibrium strategies for each period in this status quo equilibrium are:
\begin{itemize}
	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $n$ whenever this is affordable. That is, $p$ invests $R_{pn} = \max\left\{0,\frac{\beta}{1-\delta_c} -(q_{sn} - l_{sn})\right\}\leq B_{pn}$. If this inequality is violated, $p$ invests nothing.
	\item Player $c$ invests $R_{cn} = \frac{\beta}{1-\delta_c} -(Q_{sn} - l_{sn})$ if $\frac{\beta}{1-\delta_c} > (Q_{sn} - l_{sn})$. Otherwise, it invests nothing. 
	\item Players $s$ and $g$ play Status Quo as long as it yields the highest continuation value, and play Cede or Fight if continuation values from either exceeds the status quo continuation value. 
\end{itemize}

Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de_c}$ (with a possible lump sum investment at $n=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period. \hfill $\blacksquare$

%Our status quo equilibrium requires that $p$ invests in every period enough to maintain $F^*$ by offsetting the $\mu$ decline in the secessionists' status quo payoffs. The total per-period equilibrium investment $R_{pn}$, a flow payment, in the long run is thus $\mu$ per period in this steady state. In a status quo equilibrium $c$ need not invest at all since stage game payoffs of $g$ do not deteriorate.\footnote{ In the absence of the assumption in the model setup above that player $c$ does not want war, the patron would need to retain a second buffer against $c$ funding war.  The expected value of war for the home state would need to be maintained at a level lower than the status quo payoff by a buffer of $F*=\frac{\beta}{1-\delta_c}$. If this buffer did not exist in the first round of the game, the patron would fund the alteration of payoffs to create the buffer, thus assuring payments by $c$ would be ineffective at trying to change the home state's payoffs to make war more attractive than the status quo.}

By contributing $\mu$ in each period, the patron supplies sufficient resources to the unrecognized state to ensure that the secessionist elite prefers the status quo to surrendering independence.

\subsection{Discussion}

The existence and durability of this not-infrequently observed status quo equilibrium is counterintuitive on two levels. First, the large, relatively rich international community is outspent by a relatively small, less-resourced patron; second, unrecognized statehood is a stable equilibrium in spite of being undesirable to all players. The key condition leading to this outcome is that each outside actor's willingness to pay to achieve its most preferred outcome is outweighed by the other's desire to avoid it's least desired outcome. An ongoing unresolved conflict results.

Despite its high costs, this equilibrium is quite robust. Because player $c$ and the patron can adjust contributions to reflect changing conditions on the ground, exogenous shocks that might otherwise have the potential to alter the equilibrium have their strategic impact nullified. For example, while a drought in the unrecognized state might decrease the secessionist elite's payoffs from the status quo and increase their need for international trade and assistance, additional humanitarian and economic assistance from the patron can offset the effects of the shock and preserve the status quo. Likewise, if the home state gains military strength, altering the probabilities in the war lottery, the patron can offset these changes by providing arms or otherwise investing in the defenses of the unrecognized state. See further discussion along these lines in Section~\ref{sec:sanctions}.

\subsection{Alternative Outcomes}
\label{sec:alt}

The conditions of Proposition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome. In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.

If the payoffs from Fight are strictly greater than the payoffs from Cede for both players, then (Fight,Fight) will be the only additional equilibrium outcome. If this inequality holds for just one player, then we have only the additional outcome in which that player chooses Fight and its opponent cedes. If both players strictly prefer Cede to Fight, then both the (Fight, Cede) and (Cede,Fight) outcomes are added to the status quo outcome.\footnote{In the case of any indifference, we get the relevant combination of (Fight, Fight), (Fight,Cede) and (Cede, Fight).}

There are at least two takeaways from the multiplicity of equilibrium outcomes. First, it indicates that there may be an important role for external actors to play in coordinating expectations about which equilibrium will be played, and in the absence of such coordination, equilibrium switching from the status quo equilibrium to one of the other outcomes is possible. Second, most of the types of outcomes that we observe in the post-WWII era are consistent with the set of parameters outlined in Proposition 1 that support the status quo outcome. In the next section, we turn to the use of sanctions by player $c$ and the home state to attempt to force reunification.

\section{The Impact of Economic Sanctions}
\label{sec:sanctions}
%It is useful to look more specifically at how spending by the patron and player $c$ can alter the game's payoffs, and potentially its outcomes.  By analyzing comparative statics in the normal form game, we can evaluate the different strategies through which these players pursue their desired outcomes and the conditions under which they might be successful.

%Let us first consider what occurs if the patron supplies the secessionists with weapons or other military support, changing the expected payoffs from war. This would make war more appealing to the secessionist elite and less appealing to the central government, deterring attempts at reconquest. At a certain point, if the central government (and $c$) do not invest in the home state military to counteract this support, due either to their preferences or budget constraint, the expected payoff of war for the secessionists can surpass the status quo payoff, and the secessionists would have the incentives to fight:\footnote{Arrows indicate the direction of change in payoffs due to the outside action. A blank cell or "-" indicates no change.}

%\begin{center}
%\begin{tabular}{|  c| c | c | c |}
% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
%	Cede& $$& $$ &$$ \\ \hline
%	Status Quo& $(\downarrow),(\uparrow)$& $$&$$ \\ \hline
%	Fight & $(\downarrow),(\uparrow)$ &$(\downarrow),(\uparrow)$&$$\\ \hline
% \multicolumn{4}{c} {\textit{Figure 2: Additional Military Support to the Secessionists}}
%\end{tabular}
%\end{center}
%\vspace{5pt}

%Alternatively, the patron state may supply humanitarian support (such as providing passports to citizens of the unrecognized state or funding schools), making the status quo more appealing and stable:

%\begin{center}
%\begin{tabular}{|  c| c | c | c |}
% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	%Cede& $$& $$ &$(-),(\uparrow)$ \\ \hline
%	Status Quo& $$& $(-),(\uparrow)$&$$ \\ \hline
	%Fight & $$ &$$&$$\\ \hline
% \multicolumn{4}{c} { \emph{Figure 3: The Effects of Humanitarian Assistance to the Secessionists}}
%\end{tabular}
%\end{center}
%\vspace{5pt}

%Similarly, $c$ can contribute resources to make ceding more likely by either party. However, because $c$ generally prefers reunification to independence, these resources are most likely to be committed to encouraging ceding by the secessionist elite (reunification), instead of ceding by the home state (recognition).

%Changes in payoffs can be made either by carrots or by sticks.  In one option $c$ provides the secessionist elite with positive inducements, like aid, in exchange for rejoining the home state. $c$ may also expend resources to make payments by the home state, such as various autonomy rights, more credible.  The effect is the same: for a price, $c$ can increase the secessionists' payoffs from ceding.

%\begin{center}
%\begin{tabular}{|  c| c | c | c |}
% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
%	Cede& $$& $$ &$$ \\ \hline
%	Status Quo& $$& $$&$(-),(\uparrow)$ \\ \hline
%	Fight & $$ &$$&$(-),(\uparrow)$\\ \hline
% \multicolumn{4}{c} {\emph{Figure 4: Positive Inducements From Player $c$}}
%\end{tabular}
%\end{center}
%\vspace{5pt}

In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. The international community, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_{cn} < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

Let us begin with the simplest case, in which the sanctions affect only the secessionists' status quo payoffs, as when the imposition of sanctions has a negative impact on the economy of the unrecognized state. In the event that the secessionists cede or are defeated militarily, we presume the sanctions would be immediately lifted -- the home state would not want to sanction itself. In the event the secessionists achieve military victory, we presume they are able to force the home state to lift the sanctions.\footnote{Recall that our definition of military victory includes the ability to dictate the terms of settlement.} Sanctions are not expected to affect the payoffs to either ceding or military victory and neither should the cost of fighting itself be negatively impacted.

Thus, sanctions serve only two purposes: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede. If the patron wishes to maintain the status quo, its per-period investment must increase to compensate for the additional degradation of the status quo payoffs caused by the sanctions. All of this implies that the effect of sanctions on the unrecognized state's choice is not unambiguous. Proposition 2 lists necessary conditions for the existence of an equilibrium in which ``ceding'' is the equilibrium outcome once sanctions are introduced.

%Unfortunately, while the intended effect of sanctions is to reduce the unrecognized state's payoffs from the status quo, sanctions may also reduce the secessionists' payoffs from ceding.  When the home state collaborates with $c$ to enforce sanctions and impose economic suffering on the residents of the secessionist region, this can have the unintended consequence of increasing the hostility of the secessionists toward reunification.

\begin{quotation}
\noindent \bf \emph{Proposition 2:} \rm  \emph{Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo. In order for sanctions to lead to ceding by the secessionists, the following are required:}

\begin{enumerate}
\item \textit{The patron must either be unable or find that it is not worthwhile to invest the additional amount now required to maintain the status quo.}

\item \textit{The patron must either be unable or find that it is not worthwhile to invest to instigate fighting by the secessionists.}

%\item \textit{The secessionists' continuation values from playing \emph{Cede} must be higher than their continuation values from playing \emph{Fight}.}

%\item \textit{The sanctions must overcome the gap between the \emph{Status Quo} and \emph{Cede} continuation values faster than the gap between the \emph{Status Quo} and \emph{Fight} continuation values closes.}

\item \textit{The continuation value from playing Status Quo must fall below that from playing Cede before it falls below that from playing Fight. }
\end{enumerate}

\end{quotation}

The proof of Proposition 2 is in the Appendix.

If condition (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1. If conditions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of condition 3. Note here from condition (2) that sanctions can induce investment behavior by the patron that was ruled out under the conditions of Proposition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.

Thus, even if we do not consider the sanctions to impose any direct costs on the home state (though in practice they likely do), it remains ambiguous whether sanctions will benefit the home state. Recall that in the status quo equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo. If sanctions induce the secessionists to play Fight, this is a worse outcome for the home state than if the status quo had been allowed to persist.

%The situations where condition 2 or 3 fails and the secessionists initiate war follow the same general logic as political science models in the power transitions literature that predict pre-emptive war as the response to the increasing relative capabilities of a rival (Levy 1997; Powell 1999b). The expected payoffs from war are not increasing, but the alternatives are getting worse. The parallels with preemptive war become more direct if we increase the realism of our assumptions and allow sanctions to have a negative effect not only on the economy (the Status Quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).

Moving beyond this simple case, we can add realism by allowing sanctions to have a negative effect not only on the economy (the status quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war). This is an important extension because one motivation for sanctions is often precisely that -- to weaken the military capability of the secessionists.

In the model, this is represented as reducing (increasing) the secessionists' probability of victory (loss) in the war lottery. This should serve to increase the range of parameters over which the conditions of Proposition 2 hold. However, at the same time, the home government experiences changes of the same magnitude and opposite sign in its war lottery, increasing its payoffs from playing Fight. The effect of sanctions on the home state's strategic considerations is clear cut:

\begin{quotation}
\noindent \bf \emph{Proposition 3:} \rm  \emph{Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities. The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.}

\end{quotation}

The proof of Proposition 3 is immediate. Although under the conditions of Proposition 1 the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting. The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.

Thus, propositions 2 and 3 imply that sanctions are both wealth destroying and violence increasing. The sanctions destroy wealth directly by damaging the economy of the secessionist region and lowering the secessionists' payoffs from the status quo. If the degradation of status quo payoffs are not offset by the patron and if the secessionists' continuation value from fighting exceeds that from the status quo before the continuation value from ceding does, the secessionists will initiate war. Conversely, if the sanctions degrade the secessionists military capabilities sufficiently, it induces the home state to fight.

This logic is well illustrated by the case of Tamil Eelam, a territory in Northern Sri Lanka that existed as an unrecognized state from 1987-2009. Throughout the conflict, the Sri Lankan government used sanctions and blockades to reduce the secessionists' status quo payoffs and degrade their military capabilities. However, the effects of these sanctions were offset by two patrons: India, during the early stages of the conflict, and the Tamil diaspora throughout. The disruption of the status quo equilibrium began in 2006, when the United States, European Union, Canada, and India formally designated the leading secessionist organization, the Liberation Tigers of Tamil Eelam (LTTE), as terrorists.\footnote{Other states then followed suit, including Australia in 2008.} This designation greatly strengthened the sanctions against the secessionists and led to a sharp decline in both the quality of life in Tamil Eelam and the military capabilities of the LTTE. In January 2008, the Sri Lankan government abrogated the existing ceasefire agreement and in 2009 it launched a full-scale military offensive that ended with a decisive victory over the LTTE and reunification of Tamil Eelam and Sri Lanka. While there were many factors at play, the strengthening of sanctions increased the home state's probability of victory and thereby played a role in its decision to initiate a return to war.

%The Tamil war of independence began in 1983, with India initially serving as a patron of the secessionists. India withdrew its support following a failed peacekeeping intervention in 1989, after which the secessionists relied on economic support from the Tamil diaspora -- i.e. ethnic Tamils living abroad.

%An empirical investigation of post-WWII cases provides illustration of these propositions. We observe in practice that, when the home state imposes sanctions on the secessionists, it is most often the home state that initiates the return to violence. Table X provides 13 cases in which a period of unrecognized statehood (\emph{de facto} territorial control by the secessionists) was ended through decisive military victory by one side or the other.  In five of these cases, sanctions by the home state were present against the secessionists.  In four of the cases the home state initiated the fighting; in one case the secessionists initiated; in six cases either the initial conflict through which the secessionists achieved \emph{de facto} territorial control never actually ceased; and in two cases

%weakening the unrecognized state militarily is equivalent to strengthening the home government militarily

%\vspace{3pt}
%\begin{center}
%\begin{tabular}{|  c| c | c | c |}
% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
%	Cede& $$& $$ &$(-),(\downarrow)$ \\ \hline
%	Status Quo& $(\uparrow), (\downarrow)$& $(-),(\downarrow)$&$(-), (\downarrow)$ \\ \hline
%	Fight & $(\uparrow), (\downarrow)$ &$(\uparrow), (\downarrow)$&$(-),(\downarrow)$\\ \hline
% \multicolumn{4}{c} {\emph{Figure 5: Economic Sanctions by the Home State and Player $c$}}
%\end{tabular}
%\end{center}
%\vspace{5pt}

Despite this potential for perverse effects, sanctions are a common tool of outside actors ($c$), more common than aid and other positive inducements. A flow payment of carrots, even backed by the promises of a ``neutral" third party, may not be credible in the eyes of the secessionists, which could explain the frequent resort to sticks.

%As we consider tools such as sanctions aimed at disrupting the status quo equilibrium, it is important to note that not every force exerted on the situation will lead to a change in strategy. Only if a knife-edge condition exists or if large enough investments are made or sanctions applied to overcome a buffer will the perturbations of payoffs lead to a change in the actions of the decision makers. Some situations very much favor the status quo because of relatively high status quo payoffs compared to the alternatives. These situations are hard to perturb even if significant pressures are exerted from the outside players. The stability of the game can be amplified by strong preferences of both $p$ and $c$ which lead them to contribute rather than allow their least-preferred outcome.

\section{Extensions}
\label{sec:ext}
\subsection{Introducing Uncertainty}
The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$. Payoffs as described in the stage game are those perceived by $g$ and $s$. If these payoff values are not precisely known by $c$ and $p$, enough uncertainty may be present in order for both those players to contribute resources in equilibrium.

Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs. After viewing these (uncertain) payoffs, $c$ and $p$ each invest some level of resources, altering the payoff structure before it is observed (accurately) by $s$ and $g$. Based on this altered payoff structure, $c$ and $g$ choose their strategies. Uncertainty can lead to outcomes where either $c$ or $p$ invests too much, wastes resources, or makes a more severe misstep, such as $p$ investing too little and allowing reunification to occur.

In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups. This makes over-contribution by $c$ more likely than under-contribution by the patron. %Of particular note here are the resources the European Union and its members have devoted toward promoting a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.

In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources. By adding some uncertainty about payoffs, we can account for the observed fact that $c$ sometimes expends resources unsuccessfully. This type of spending can also be explained as non-strategic spending -- i.e. spending aimed at goals other than promoting settlement, like pure humanitarianism.

%\subsection{Partial Recognition}
%The norm of home state veto gives recognition by the home state its significance: recognition by the home state is the core demand of the secessionists in our model. In cases where the secessionists can gain recognition from a large number of states without first gaining recognition by the home state, the status quo is a less costly outcome for the secessionists. This increases the patience of the secessionists. In the model, if the payoffs to the secessionist leaders for the status quo rise, the equilibrium decision by the secessionist elite to play "status quo" becomes more stable. Less support from a patron is needed and deterioration of conditions for the unrecognized state does not immediately disrupt the status quo equilibrium. Because the home state is aware of this payoff change, it may realize that holding out is less likely to be fruitful -- reabsorption of the secessionists by the home state is less likely to occur through the secessionists ceding. Additionally, if the international community's preferences shift toward protection of the secessionists following mass atrocity crimes, expected payoffs from a military conflict would decrease for the home state. It is possible that the status quo payoffs are lowered for the home state as well, due to lack of support for its cause by the international community. If these dynamics lead to a sufficiently low home state payoff for both status quo and war, the the home state will cede.

%With the official adoption of the Responsibility to Protect doctrine by the United Nations and the precedent of Kosovo, home states are put on notice that the commission of mass atrocity crimes against the residents of seceding entities may lead to recognition of that entity by other countries.\footnote{For a discussion of the sanction theory of recognition, see Berlin (2009).} However, the norm of home state veto remains strong, and so long as further atrocities are not committed, the Responsibility to Protect doctrine is unlikely to affect recognition of the unrecognized states already in existence.

\subsection{Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
The game that we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible. We do not model any direct exchange of resources or imposition of harm between $c$ and the patron, but we take the implications of these possible outside interactions seriously. The willingness and ability of either party to contribute resources within the game we model may be affected by their interaction with one another in other contexts. As we will discuss in greater detail below, in a number of cases a patron has withdrawn support for an unrecognized state in response to pressure exerted by actors in the international community ($c$) in other venues.

\subsection{If the Patron Withdraws Support}
Support by a foreign patron is, in almost all cases, necessary for the persistence of unrecognized statehood (Kolst\o \ 2006; Caspersen 2013). When there is no patron, or when the patron withdraws its support, military reconquest by the home state is likely. The only unrecognized state currently in existence which has been able to survive without a patron is Somaliland. Somaliland has been able to survive as long as it has because of the extreme weakness of the home state (Somalia).

%In the model, $g$ or $s$ has an incentive to fight rather than stay in the status quo if the expected discounted stream of payoffs is greater by fighting. Specifically, if condition (2) for the status quo equilibrium is not met, then one player will prefer war to both ceding and the status quo. In the case of most prolonged stalemates, strategic spending by the patron averts these outcomes. The patron provides military assistance to the secessionists at such a level that $g$ does not prefer to initiate conflict, and provides sufficient economic and humanitarian assistance to prevent $s$ from preferring war to a continuation of the status quo. However, if there is no patron or if the patron is not sufficiently interested, war may become a more attractive outcome than the status quo for at least one of the parties.

%The cases where there is no patron, such as in Chechnya, this is relatively easy to explain. 
The case of Chechnya is quite typical of unrecognized states without patrons. Chechnya fought and won its initial war of secession at a time when the home state (Russia) was in political and economic disarray following the collapse of the Soviet Union. As Russia gradually recovered and strengthened, Chechnya had no patron support to offset the relative decline in its military capabilities. Over time the war lottery became progressively more skewed in favor of Russian victory, the payoffs to ceding for the Chechen secessionists remained extremely low, and the Russian government invaded and reconquered Chechnya in 1999.

It is worth exploring, however, the reasons why a patron might support a secessionist group during its initial rebellion and then withdraw support at a later date. Patrons' strategic interests in the unrecognized state vary from patron to patron, and both budget constraints and salience of interest vary over time. Returning to the case of Tamil Eelam, domestic political concerns (primarily Tamil ethnic solidarity) induced a modest level of Indian support for the secessionists  from 1983-1987. These domestic political concerns were eventually outweighed by broader strategic security concerns and a desire for regional stability. In 1987 the Indian government signed a peace accord with Sri Lanka (the home state) and largely withdrew their support from the secessionists.

%, even sending in peacekeepers that later clashed with the secessionists militarily (Singer 1992). The Tamil diaspora funded the secessionist military until 2009, when this support proved insufficient and the territory was reconquered by the Sri Lankan government.\footnote{On a smaller scale, the Isaaq diaspora has also fulfilled some of the roles of the patron in Somaliland (Galipo 2011).}

As noted in the section on outside interactions between the patron and player $c$, the patron's decision to withdraw support for the secessionists is sometimes motivated by interactions between the patron and $c$ that we do not model directly. Empirically, we observe a number of cases in which $c$ applies pressure directly to encourage the patron to withdraw support from the unrecognized state. In an extreme example involving both sanctions and direct military confrontation, NATO coerced Serbia into, among other things, withdrawing its support from Republika Srpska and Republika Srpska Krajina, both of which were then reconquered militarily.\footnote{For an excellent discussion of the case of Republika Srpska, see Zahar 2004.} %In the final section of the paper, we consider direct coercion of the patron among the strategies available to player $c$.

\section{Policy Implications: Options for Player \texorpdfstring{$c$}{c}}

Sanctions imposed by the home state, often co-enforced by player $c$, aim to disrupt the status quo equilibrium and force the secessionists to cede. However, as Section \ref{sec:sanctions} demonstrates, sanctions also increase the risk of war, something player $c$ prefers to avoid.

Fortunately, there is a better way. If instead of enforcing sanctions, player $c$ tries to promote settlement by supplementing the payoffs from unification, it is able to induce negotiated settlement without simultaneously increasing the risk of war. This can be done either through promises of benefits to the unrecognized state provided directly by $c$, like aid, or by a commitment from $c$ to serve as a third-party guarantor of side payments promised by the ceding side: both directly increase $L_i$ with the goal of raising it above $Q_i$ for the player who will cede in the settlement. In the case of contingent promises of aid, the calculation is relatively straightforward: 1) the promise of aid must be credibly contingent on negotiated settlement, and 2) the aid offered must be valued more highly than the concessions required to reach an agreement. It is the second condition that is most problematic. Because both sides place such a high value on status (independence vs. reunification), even large amounts of aid are likely to be valued less than the concessions necessary to reach an agreement.

Serving as a third-party guarantor of autonomy rights is a way for player $c$ to potentially overcome problems of indivisibility and commitment and help the parties reach a credible compromise on status (Walter 2002). However, this strategy is only tenable when the only impediment to settlement is the unenforceability of a bargain, and when $c$ is credible as an enforcer of that bargain.

%Credible enforcement of future autonomy rights can be viewed either as increasing the value of available side payments or as making the central issue of contention divisible. In either view, a range of previously untenable agreements are made possible.

In Southern Sudan, third-party actors, including the UN, invested substantial resources to help negotiate a settlement to the initial war of secession and to ensure that the Sudanese government both allowed the promised referendum and respected its results. While the UN and others invested resources in Southern Sudan to enforce independence, not autonomy, they have demonstrated that outside actors are capable of enforcing difficult concessions by the home state government. This bodes well for the future credibility of organizations like the UN as third-party enforcers. However, the role of outside actors in enforcing other past agreements might give secessionists pause. For example, a referendum on independence in Western Sahara, which the UN ruled to be necessary more than thirty years ago, has never come to pass.\footnote{For a thorough analysis of the Western Sahara case, see Zunes and Mundy (2010).} Nonetheless, it is possible for an outside state or coalition to invest resources to enforce agreements, allowing for negotiated settlements that would otherwise be impossible to reach.

To show that it is possible for an actor like player $c$ to enforce the terms of negotiated agreements at a reasonable cost is not sufficient to imply that such an outcome is likely. The political will necessary to achieve success in Southern Sudan was motivated largely by the magnitude of the atrocities that accompanied the war of secession, and enforcement was made credible, in part, due to the weakness of Sudan relative to the outside states involved. Enforcing the terms of an agreement between Georgia, South Ossetia, and Russia, for example, would be more difficult.

It is also possible for actors like $c$ to affect the payoffs of the patron through interactions in other games outside of our model. Such actions would manifest themselves within the model as reductions in the patron's willingness to pay to sustain the status quo. If the patron is unwilling to pay to sustain the status quo, the status quo payoffs of the secessionists will decline over time, eventually leading to either war or negotiated settlement. Under these conditions, the within-game costs to $c$ of inducing negotiated reunification also fall.

In this section we have argued that outside states are capable of imposing their preferred outcome, including peaceful reunification. The key, however, is motivation: actors like player $c$ are capable of inducing peaceful reunification when they are willing to invest the resources necessary. However, strong preferences of secessionists against reunification and the opposing intervention of the patron make the costs of such interventions prohibitively high in most cases. Unrecognized statehood is a stable equilibrium because, while there are many actors in the international community that share the preferences we ascribe to $c$, they are usually unwilling to invest sufficient resources to outspend the patron and induce their preferred outcome.

\section{Conclusions}
In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives. Current events, particularly, Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon. While the Ukrainian situation remains fluid, the status quo equilibrium we model is a plausible future. The equilibrium could emerge as follows: with Russian support, the secessionists solidify \emph{de facto} control of parts of Eastern Ukraine, but risk of direct confrontation with the West is sufficiently high to prevent Russian annexation of Eastern Ukraine.\footnote{Annexation of Crimea was achieved at a reasonable cost to Russia. The costs of annexation Eastern Ukraine would likely be substantially higher.} Further Russian economic aid keeps the status quo payoffs of the secessionists sufficiently high that they decline negotiated reunification. If necessary, contributions to the home state by U.S. and the EU (player $c$) strengthen the home state enough to prevent the secessionists from overthrowing the Ukrainian government and forcing recognition. In twenty years, Eastern Ukraine may look a good deal like South Ossetia does today.

While the importance of outside actors in civil conflict has been widely acknowledged in the empirical literature, it is rarely modeled formally. We introduce a unique four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome. This allows us both to examine the means through which patron states sustain unrecognized statehood as a stable equilibrium, and to rigorously analyze the strategies available to other outside actors to pursue reunification. It is not always in the interests of outside states to bear the costs of inducing their preferred outcome, but we identify the mechanisms through which this is possible and the thresholds that must be overcome.

In the model we present, we show that unrecognized statehood can emerge as an equilibrium outcome even when it is a terrible outcome for all players involved. The patron, even when it prefers outright independence, is willing to bear costs in every period to uphold an outcome that is far short of its ideal, and which imposes high costs on others as well. The international community, though wealthier than the patron, does does not outspend the patron to force reunification because the patron's desire to avoid reunification is stronger than the international community's desire to achieve it.

By providing economic and other aid to the unrecognized state, the patron keeps the secessionist elite's payoffs from the status quo high enough to prevent a negotiated reunification. The stability of this equilibrium is abetted by the indivisible nature of independence and the difficulty of enforcing autonomy as a condition of reunification. In cases where there is no patron or where the patron eventually becomes unwilling to continue its support, the result has almost always been violent reconquest by the home state.

Our model also suggests, however, that the historical pattern of costly stalemate followed by violent resolution is not the only possible path. We show that the stabilizing effect of the patron can be overcome by a sufficiently motivated international community. While the most frequently employed means through which outside actors attempt to induce settlement---sanctions---also increases the risk of war, we show that it is possible for outside actors to induce their preferred outcome without running this risk. In particular, we suggest that they can provide positive inducements for settlement and serve as a third-party guarantor of negotiated settlements in which unrecognized states rejoin the home state as autonomous regions. It is often not the lack of available means that prevents outside actors from inducing their preferred outcome, but rather the lack of will.

\pagebreak

\section{References}\label{A:references}
\hangindent=1cm

\begin{enumerate}[1.]

\item \hangindent=1cm Acemoglu, Daron, Golosov, Mikhail, Tsyvinski, Aleh, and Pierre Yared. 2012. "A Dynamic Theory of Resource Wars." 
\emph{The Quarterly Journal of Economics} 127: 283-331.

\item \hangindent=1cm Bahcheli, Tozun, Barry Martmann, and Henry Srebrnik, eds. 2004 "De Facto States: The Quest for Sovereignty." New York, NY: Routledge.

\item \hangindent=1cm Bakke, Kristin M., John O'Loughlin, Gerard Toal, and Michael D. Ward.  2013.  "Convincing State-Builders? Disaggregating Internal Legitimacy in Abkhazia."  \emph{International Studies Quarterly.}

%\item \hangindent=1cm Balch-Lindsay, Dylan, and Andrew J. Enterline. 2000. "Killing Time: The World Politics of Civil War Duration, 1820-1992." \emph {International Studies Quarterly} 44 (4):615-642.

\item \hangindent=1cm Balch-Lindsay, Dylan, Andrew J. Enterline, and Kyle A. Joyce.  2008.  "Third-Party Intervention and the Civil War Process."  \emph{Journal of Peace Research} 45: 345-63.

\item \hangindent=1cm Baliga, Sandeep, and Tomas Sjostrom. 2012. "The Strategy of Manipulating Conflict." \emph{The American Economic Review} 102 (6): 2897-2922.

\item \hangindent=1cm Bevi\'{a}, Carmen and Luis C. Corch\'{o}n. 2010. ``Peace Agreements without Commitment.'' \emph{Games and Economic Behavior} 68 (2): 469-487.

\item \hangindent=1cm Bueno de Mesquita, Ethan. 2013. ``Rebel tactics.'' \emph{Journal of Political Economy} 121 (2): 323-357.

\item \hangindent=1cm Caspersen, Nina. 2013. "Unrecognized States: The Struggle for Sovereignty in the Modern International System." Malden, MA: Polity Press.

\item \hangindent=1cm Caspersen, Nina, and Gareth Stansfield. 2011. "Unrecognized States in the International System." New York, NY: Routledge.

%\item \hangindent=1cm Christin, Thomas, and Simon Hug. 2012. "Federalism, the Geographic Location of Groups, and Conflict." \emph {Conflict Management and Peace Science} 29 (1): 93-122.

\item \hangindent=1cm Coggins, Bridget. 2011. "Friends in High Places: International Politics and the Emergence of States from Secessionism." \emph {International Organization} 65 (03): 433-467.

%\item \hangindent=1cm Collier, Paul, and Anke Hoeffler. 2004. "Greed and Grievance in Civil War." \emph {Oxford Economic Papers} 56 (4): 563-595.

%\item \hangindent=1cm Collier, Paul and Anke Hoeffler. 2005. "Resource Rents, Governance, and Conflict." \emph {Journal of Conflict Resolution} 49 (4): 625-633.

\item \hangindent=1cm Crawford, James. 2006. "The Creation of States in International Law." Oxford: Oxford University Press.

%\item \hangindent=1cm Cunningham, David E. 2006. "Veto Players and Civil War Duration." \emph {American Journal of Political Science} 50 (4): 875-892.

\item \hangindent=1cm Doyle, Michael W., and Nicholas Sambanis. 2006. "Making War and Building Peace: United Nations Peace Operations." Princeton: Princeton University Press.

\item \hangindent=1cm Elbadawi, Ibrahim, and Nicholas Sambanis. 2000. "External Interventions and the Duration of Civil Wars." 2433.

%\item \hangindent=1cm Fazal, Tanisha. 2007. "State Death: The Politics and Geography of Conquest, Occupation, and Annexation." Princeton: Princeton University Press.

%\item \hangindent=1cm Fearon, James D. 1995. "Rationalist explanations for war." \emph {International Organization} 49 (03): 379-414.

\item \hangindent=1cm Fearon, James D., and David D. Laitin. 2003. "Ethnicity, Insurgency, and Civil War." \emph {American Political Science Review} 97 (01): 75-90.

\item \hangindent=1cm Fearon, James D. 2004. "Why Do Some Civil Wars Last So Much Longer than Others?" \emph {Journal of Peace Research} 41 (3): 275-301.

\item \hangindent=1cm Geldenhuys, Deon. 2009. "Contested States in World Politics." Basingstoke: Palgrave McMillan.

%\item \hangindent=1cm Ghai, Yash, and Anthony Regan. 2006. "Unitary State, Devolution, Autonomy, Aecession: State Building and Nation Building in Bouganville, Papua New Guinea." 95 (386).

%\item \hangindent=1cm Hartzell, Caroline, Matthew Hoddie, and Donald Rothchild. 2001. "Stabilizing the peace after civil war: An investigation of some key variables." \emph {International Organization} 55 (01): 183-208.

%\item \hangindent=1cm Hegre, Havard, and Nicholas Sambanis. 2006. "Sensitivity Analysis of Empirical Results on Civil War Onset." \emph {Journal of Peace Research} 50 (4): 508-535.

\item \hangindent=1cm Grossman, Herschel I. 2004. ''Peace and war in territorial disputes.'' No. w10601. National Bureau of Economic Research.

\item \hangindent=1cm Hirshleifer, Jack, Boldrin, Michele, and David K Levine. 2009. ''The slippery slope of concession.'' \emph{Economic Inquiry} 47 (2): 197-205.

\item \hangindent=1cm Jackson, Matthew O., and Massimo Morelli. 2007. "Political bias and war." \emph{The American Economic Review} 97 (4): 1353-1373.

\item \hangindent=1cm King, Charles. 2001. "The Benefits of Ethnic War: Understanding Eurasia's Unrecognized States." \emph{World Politics} 53 (4): 524-552.

\item \hangindent=1cm Kingston, Paul, and Ian S. Spears, eds. 2004. "States-Within-States." New York, NY: Palgrave McMillan.

\item \hangindent=1cm Kolst\o, P{\aa}l. 2006. "The Sustainability and Future of Unrecognized Quasi-States." \emph {Journal of Peace Research} 43 (6): 723-740.

\item \hangindent=1cm Licklider, Roy. 1995. "The Consequences of Negotiated Settlements in Civil Wars, 1945-1993." \emph {American Political Science Review} (89): 681-90.

%\item \hangindent=1cm Levy, Jack S. 1987. "Declining Power and the Preventive Motivation for War." \emph{World Politics} 40: 82-107.

\item \hangindent=1cm Lynch, Dov. 2004. "Engaging Eurasia's Separatist States." Washington, DC: U.S. Institute of Peace.

\item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, settlement, and the shadow of the future." \emph{Journal of Economic Behavior $\&$ Organization} 105: 75-89.

\item \hangindent=1cm Pegg, Scott. 1998. "International society and the de facto state." New York: Ashgate.

%\item \hangindent=1cm Powell, Robert. 1999a. "The Modeling Enterprise and Security Studies." \emph {International Security} 24 (2): 97-106.

%\item \hangindent=1cm Powell, Robert. 1999b. \emph{In the Shadow of Power: States and Strategies in International Politics.} Princeton: Princeton University Press.

\item \hangindent=1cm Regan, P.M. 2002. "Civil wars and foreign powers: Outside intervention in intrastate conflict." Ann Arbor: University of Michigan Press.

\item \hangindent=1cm Roeder, Philip G. 2007. "Where nation-states come from: institutional change in the age of nationalism." Princeton: Princeton University Press.

\item \hangindent=1cm Salehyan, Idean, Kristian Skrede Gleditsch, and David E. Cunningham. 2011. "Explaining External Support for Insurgent Groups." \emph {International Organization} 65 (04): 709-744.

\item \hangindent=1cm Schwarz, Michael, and Konstantin Sonin. 2008. "A theory of brinkmanship, conflicts, and commitments." \emph{Journal of Law, Economics, and Organization} 24 (1): 163-183.

\item \hangindent=1cm Schultz, Kenneth A. 2010. "The Enforcement Problem in Coercive Bargaining: Interstate Conflict over Rebel Support in Civil Wars." \emph {International Organization} 64 (02): 281-312.

%\item \hangindent=1cm Singer, Marshall R. 1992. "Sri Lanka's Tamil Sinhalese Ethnic Conflict: Alternative Solutions." \emph {Asian Survey} 32 (8): 712-722.

%\item \hangindent=1cm Spears, Ian S. 2004. "States-Within-States: An Introduction to Their Empirical Attributes." In States-Within-States, eds. Paul Kingston and Ian S. Spears. New York: Palgrave McMillan.

%\item \hangindent=1cm Spruyt, Hendrik. 1996. "The Sovereign State and its Competitors." Princeton, NJ: Princeton University Press.

\item \hangindent=1cm Stanislawski, Bartosz H. 2008. "Para-States, Quasi-States, and Black Spots: Perhaps Not States, But Not `Ungoverned Territories,' Either." \emph {International Studies Review} 10 (2): 366-396.

%\item \hangindent=1cm Tir, Jaroslav. 2006. "Domestic-Level Territorial Disputes: Conflict Management via Secession." \emph {Conflict Management and Peace Science} 23 (4): 309-328.

\item \hangindent=1cm van Houten, Pieter. 1998. "The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives EuropŽennes de Sociologie} 39 (01): 110-146.

\item \hangindent=1cm Walter, Barbara F. 1997. "The Critical Barrier to Civil War Settlement." \emph {International Organization} 51 (03): 335-364.

\item \hangindent=1cm Walter, Barbara F.. 2002. "Committing to Peace: The successful settlement of civil wars." Princeton: Princeton University Press.

%\item \hangindent=1cm Walter, Barbara F. 2006. "Building Reputation: Why Governments Fight Some Separatists but Not Others." \emph {American Journal of Political Science} 50 (2): 313-330.

\item \hangindent=1cm Yared, Pierre. 2010. ``A Dynamic Theory of War and Peace.'' \emph{Journal of Economic Theory} 145: 1921-1950.

\item \hangindent=1cm Zacher, Mark W. 2001. "The Territorial Integrity Norm: International Boundaries and the Use of Force." \emph {International Organization} 55 (02): 215-250.

\end{enumerate}

\pagebreak
\appendix
\section{Appendix}

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{(1-\delta_s)}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{(1-\delta_i)}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$. Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta_i}$.

In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2). The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
}{1 -\de_p}-R_p^{RU}$. For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de_c} -R_c^{RU}$. Neither player will be willing to invest in its least preferred state. That is, $R_p^{RU} = R_c^{REC} =0$. So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de_c}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ} 
\right)}{1 -\de_p}$ each period toward maintaining it.

Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g. counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de_c} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de_p} > \frac{\beta+ R_c^{SQ}}{1 -\de_c}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\frac{-\alpha}{1 -\de_p}  > \frac{\beta}{1 -\de_c}$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de_p} -R_p^{REC}$. For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de_c}-R_c^{REC}$. As noted above, $R_c^{REC} =0$. So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de_p}$ toward recognition. We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de_c}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$.

Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $-\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.

Thus $\frac{-\nu}{1 -\de_c} > \frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary. A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{\nu p_1 + \beta p_2}{1 -\de_c} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_1 + \beta p_2}{1 -\de_c}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).

$p$ is willing to invest up to $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}$ as this is most restrictive. In this only case of interest, we have $\frac{\alpha p_2}{1 -\de_p} +\frac{\lambda p_1}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0$. Note that the first term is negative while the rest are positive.

Multiplying Condition 3 by $-p_{2s}$, we have $\frac{-\beta \cdot p_{2s}}{1-\de_c} \geq \frac{\alpha \cdot p_{2s}}{1-\de_p}$. Using this with Condition 4, we have
\[
  \textstyle -\left(\frac{\nu p_{1s} + \beta p_{2s}}{1 -\de_c}\right) = - \frac{\nu p_{1s}}{1 -\de_c} - \frac{\beta p_{2s}}{1 -\de_c} \geq \frac{\alpha p_{2s}}{1 -\de_p} +\frac{\lambda p_{1s}}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0 \ \forall p_{1s},p_{2s}
\]

Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude. Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $n+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $n$. As the only alteration is a reduction in $Q_{sn}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

Therefore if sanctions are to disrupt the Status Quo equilibrium and potentially lead to ceding by the secessionists, it must be that the patron is either not able or not willing to make the investment required to maintain the status quo under player $c$'s sanctions (Condition 1).

Assume then that Condition 1 holds so that the patron makes no investments in the Status Quo payoffs of player $s$. We also need the patron to withhold investments toward encouraging the secessionists to fight. Lemma 4 of Proposition 1 is not sufficient to guarantee this behavior since no investments are being made in the Status Quo. In addition, the reference outcome changes: players $p$ and $c$ must now compare their payoffs in the war scenario to the reunification outcome. To be precise, the least stringent condition is $m$ periods of Status Quo followed by the Reunification outcome, where $m$ is the maximal number of periods that it takes for the sanctions to degrade the status quo payoffs below the level of the payoffs from ceding for player $s$.

Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de_p^m}{1-de_p} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de_p} + \frac{\de_p}{1-\de_p}\alpha (1-p_{1s}-p_{2s})$. For player $c$, it is $0 + \frac{\de_p^m}{1-de_p} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de_p} + \frac{\de_p}{1-\de_p}\beta (1-p_{1s}-p_{2s})$. Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
\[
  \textstyle \frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de_p} - \frac{\de_p \alpha}{1-\de_p} (p_{1s}+p_{2s}) + \alpha \sum_{i=1}^{m-1} \de_p^i > \frac{\nu p_{1s} + \beta p_{2s}}{1-\de_c} - \frac{\de_c \beta}{1-\de_c} (p_{1s}+p_{2s}) + \beta \sum_{i=1}^{m-1} \de_c^i
\]

If this condition holds, then there will be no investments following sanctions. But it is still not guaranteed that the secessionists will choose to cede. Status Quo is no longer a long-run equilibrium strategy, but even without outside investments, it is possible that Fight could be a more attractive action than Cede. Degrading the status quo payoffs is not sufficient to make the continuation value from Cede the highest available.

To see why, note that the sanctions only have a partial effect in reducing the Fight continuation value through the war lottery, while their entire impact is felt on the Status Quo continuation value. Thus the gap between the Status Quo and Fight continuation values narrows at the same time as the corresponding gap between the Status Quo and Cede continuation values. Although the former gap narrows more slowly, it is likely to be smaller at the outset so sanctions do not necessarily make Cede the most preferred action for player $s$.

Therefore, in order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight. \hfill $\blacksquare$

%Several possibilities exist. First, the sanctions could degrade the payoffs to ceding more than they degrade the status quo payoffs. If $S_L > \mu + S_Q$ where $S_L$ and $S_Q$ are the per-period reduction in the payoffs to ceding and maintaining the Status Quo, respectively, both payoffs will degrade until eventually player $s$ will choose \emph{Fight}. This occurs because the continuation value from \emph{Cede} will remain below that from playing \emph{Status Quo}, but the sanctions only have a partial effect in reducing the \emph{Fight} continuation value through the war lottery, while their entire impact is felt on the \emph{Status Quo} continuation value. The gap between the Status Quo and Fight continuation values that was positive without sanctions will eventually become negative, leading to conflict.

%Thus we must have $\mu + S_Q > S_L$ if there is any chance that the secessionists will choose to cede.

\end{document}

### 2015-04-09 to_do_list.tex
add to_do_list with kristy's items for the revision

\documentclass[12pt]{article}

\addtolength{\textwidth}{1.4in}
\addtolength{\oddsidemargin}{-.7in} %left margin
\addtolength{\evensidemargin}{-.7in}
\setlength{\textheight}{8.5in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}
\setlength{\headheight}{0.0in}
\setlength{\footskip}{.5in}
\renewcommand{\baselinestretch}{1.0}
\setlength{\parindent}{0pt}
\linespread{1.1}

\usepackage{amssymb, amsmath, amsthm, bm}
\usepackage{graphicx,csquotes,verbatim}
\usepackage[backend=biber,block=space,style=authoryear]{biblatex}
\setlength{\bibitemsep}{\baselineskip}
\usepackage[american]{babel}
%dell laptop
\addbibresource{C:/Users/Kristy/Dropbox/Research/xBibs/tradeagreements.bib}
%\addbibresource{C:/Users/Kristy/Documents/Dropbox/Research/xBibs/tradeagreements.bib}
\renewcommand{\newunitpunct}{,}
\renewbibmacro{in:}{}

\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{xcolor}
\hbadness=10000

\newcommand{\ve}{\varepsilon}
\newcommand{\ov}{\overline}
\newcommand{\un}{\underline}
\newcommand{\ta}{\theta}
\newcommand{\al}{\alpha}
\newcommand{\Ta}{\Theta}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\de}{\delta}

\begin{document}
\begin{center}
Unrecognized States To-do List (JLEO Revision)
\end{center}

\vskip.3in
Kristy
\begin{itemize}
	\item Think through reviewer 2's suggestion about uncertainty
		\begin{itemize}
			\item How would model actually work with asymmetric information?
		\end{itemize}
	\item Take indecisive war to decisive war
	\item Change sign on payoff functions for $c$ and $p$
	\item Homogenize discount factors
	\item Decide on Reviewer 1's suggestion to recast Proposition 1 as a ``class of games'' instead of ``sufficient conditions''
		\begin{itemize}
			\item ``I understand why the set of sufficient conditions in this proposition is the key result. However, these mechanisms are essentially just assumptions on payoffs; and therefore I would be inclined to interpret these conditions as characterizations of “a class” of games, where this class would satisfy all the assumptions on payoffs (instead of saying sufficient conditions). \\
				I would recommend clearing up Section 3.1 in terms of packaging the results: 
					\begin{enumerate}
						\item First, I would identify a class of games that satisfy the assumptions on payoffs (sufficient conditions (1) through (6) in Proposition 1). Denote this class as G for example. 
						\item Then, formally define the “status quo” equilibrium – an equilibrium in which the equilibrium outcome is perpetual unrecognized statehood with all the actors’ strategies specified. 
						\item Finally, I would state a much clearer and simpler proposition, e.g., For any game G, there exists a status quo equilibrium. ... Then subsection 3.2 would basically be a discussion of this class of games. 
					\end{enumerate}
		\end{itemize}
	\item Clearly specify punishments / equilibrium path to satisfy editor
		\begin{itemize}
			\item ``Clear and detailed description of the noncooperative game''
			\item ``I would also like to see the key logic laid out better in the text, and the full
equilibrium strategy profiles noted so that it is clear how the players are rewarding and punishing each other on and off the equilibrium path''
		\end{itemize}
	\item State process governing status-quo payoffs clearly
		\begin{itemize}
			\item ``If I understand correctly, expenditure R by the patron is essentially a long-term investment that raises the status-quo payoff of the secessionist elite by R in every future period. I suggest stating it this way: The per-period status quo payoff is a state variable; its transition from one period to the next involves an automatic reduction by mu (for the reasons you discuss) and an increase by the amount of the patron's investment.
		\end{itemize}
	\item Make sure equilibrium concept is clearly stated and defined
		\begin{itemize}
			\item ``Because the model is not a repeated game, due to the state variable just described, it may be a bit confusing to some readers, and so it is important to provide formal definitions and to clearly describe the equilibrium concept being utilized. It seems to me that, as Referee 2 states, you are using the Markov perfect equilibrium concept.''
		\end{itemize}
\end{itemize}

\end{document}

### 2015-06-19 JLEO_eqm.tex
create JLEO_eqm document to hold work on streamlined model and eqm

--- 2015-04-09 to_do_list.tex
@@
 \begin{center}
-Unrecognized States To-do List (JLEO Revision)
+Unrecognized States -- JLEO Streamlined Model and Equilibrium
 \end{center}
@@
 \begin{itemize}
-	\item Think through reviewer 2's suggestion about uncertainty
-		\begin{itemize}
-			\item How would model actually work with asymmetric information?
-		\end{itemize}
 	\item Take indecisive war to decisive war
@@
 		\end{itemize}
-	\item State process governing status-quo payoffs clearly
-		\begin{itemize}
-			\item ``If I understand correctly, expenditure R by the patron is essentially a long-term investment that raises the status-quo payoff of the secessionist elite by R in every future period.
-I suggest stating it this way: The per-period status quo payoff is a state variable; its transition from one period to the next involves an automatic reduction by mu (for the reasons you discuss) and an increase by the amount of the patron's investment.
-		\end{itemize}
 	\item Make sure equilibrium concept is clearly stated and defined
 		\begin{itemize}
-			\item ``Because the model is not a repeated game, due to the state variable just described, it may be a bit confusing to some readers, and so it is important to provide formal definitions and to clearly describe the equilibrium concept being utilized.
+			\item ``it is important to provide formal definitions and to clearly describe the equilibrium concept being utilized.
 It seems to me that, as Referee 2 states, you are using the Markov perfect equilibrium concept.''

### 2015-06-19 JLEO_eqm.tex
add main result and associated proofs from draft to JLEO_eqm

@@
 
+\newpage
+\begin{quotation}
+\noindent \bf \emph{Proposition 1:} \rm  \emph{There exists an equilibrium in which the outcome is perpetual unrecognized statehood.
+The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\delta_c}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}
+
+\emph{The following are sufficient conditions for such an equilibrium: }
+
+\begin{enumerate}
+\item \textit{For both players $g$ and $s$, $Q_{in} \geq L_{in} \ \forall n$: in each period, remaining in the status quo is better than ceding.}
+
+\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta_i} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta_i} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
+
+\item \textit{$\frac{-\alpha}{1-\de_p} \geq \frac {\beta}{1-\delta_c}$: reunification is more important for the patron to avoid than for the international community to achieve.}
+
+\item  \textit{$\frac {-\nu\cdot p_{1s}}{1-\delta_c} \geq \frac{\lambda \cdot p_{1s} + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
+
+\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+
+\item \textit{$B_{pn} \geq \mu, \ \forall n>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
+If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $n> 1$ compared to period $1$, (6) could be binding.}
+
+\end{enumerate}
+
+\end{quotation}
+
+These conditions for the status quo are sufficient but not necessary.
+For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run.
+The inequality \emph{will} bind for some set of periods $n \geq 1$ because the secessionist payoffs from the status quo decrease over time.
+In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
+
+There are many other potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
+As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity.
+We will use the solution concept of subgame perfect equilibrium to show that, given the sufficient conditions in Proposition 1, at least one status quo equilibrium will exist.\\
+
+\noindent {\bf Proof of Proposition 1} \\
+Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3).
+Although period $n$ may be reached because in period $n-1$ fighting resulted in a stalemate, both players ceded, or the status quo had been maintained, the strategic landscape in period $n$ is the same.
+Thus all periods in which the players are able to move are in an equivalence class in which the only possible strategic difference is in the value of the state variables.
+
+Analysis of the posited equilibrium proceeds most naturally by backward induction within each period. Thus we begin with the stage game between the government and secessionists.
+
+\begin{quotation}
+\noindent \textit{{\bf Lemma 1:} Conditions 1 and 2 are sufficient for status quo to be the outcome of Stage 3 in any period.} 
+\end{quotation}
+
+The proofs of the lemmas are in the Appendix.
+
+Lemma 1 establishes ranges for the payoffs for players $g$ and $s$---gross of investments by the outside players---in which the status quo outcome can occur.
+The rest of this section tackles the more difficult task of determining the incentives and actions of the outside actors---that is, the patron state and the international community---to impact those payoffs.
+
+To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{sn},q_{gn},w_{sn},w_{gn},l_{sn},$ and $l_{gn}$.\footnote{In the case of $w_{sn}$ and $w_{gn}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e.
+\emph{Recognition, Status Quo}, or \emph{Reunification}---they will target given that they will choose the most efficient way to alter the payoffs of players $s$ and $g$.
+
+%Given their preference orderings, some of these can be ruled out.
+Player $p$ will not invest to increase $s$'s payoffs from ceding ($l_{sn}$) or $g$'s payoffs from war when it could do so for $s$.
+It will also not invest in $q_{gn}$ when it can instead increase $l_{gn}$, leaving the list of rational investment choices for $p$ as $q_{sn},l_{gn}$, and $w_{sn}$.
+Symmetrically, $c$ will not invest in $q_{sn},L_{gn}$, or $w_{sn}$.
+Given the assumption that the international community follows a norm of encouraging peaceful resolution, it also will not invest in $w_{gn}$, leaving only $l_{sn}$ and $q_{gn}$ as rational investment options for player $c$.
+
+Lemma 2 addresses potential efforts by the international community to influence the outcome toward \emph{Reunification}:
+
+\begin{quotation}
+\noindent \textit{{\bf Lemma 2:} When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.}
+\end{quotation}
+
+\noindent Condition 3 provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
+To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}
+
+Consider period 1.
+When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\delta_c}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
+That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\delta_c} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
+Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
+
+In periods $n > 1$, there are two cases to consider.
+Either (a) the buffer created in period $n-1$ was precisely the necessary $\frac{\beta}{1-\delta_c}$, or (b) the buffer is larger than $\frac{\beta}{1-\delta_c}$.
+In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\delta_c}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\delta_c}$ provides lower welfare to player $p$.
+Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $n$.
+However, because in each period $q_{sn}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
+Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $n > 1$(Condition 6).
+
+With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending for or against recognition of the secessionist state:
+
+\begin{quotation}
+\noindent \textit{{\bf Lemma 3:} When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.}
+\end{quotation}
+
+In the equilibrium under consideration here, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome.
+This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved.
+The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Conditions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
+
+One last possibility is ruled out by Lemma 4: that the patron would invest to encourage the secessionists to fight.
+
+\begin{quotation}
+\noindent \textit{{\bf Lemma 4:} Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.}
+\end{quotation}
+
+Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
+All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.
+
+Thus, given Conditions (1)-(6) hold, equilibrium strategies for each period in this status quo equilibrium are:
+\begin{itemize}
+	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $n$ whenever this is affordable.
+That is, $p$ invests $R_{pn} = \max\left\{0,\frac{\beta}{1-\delta_c} -(q_{sn} - l_{sn})\right\}\leq B_{pn}$.
+If this inequality is violated, $p$ invests nothing.
+	\item Player $c$ invests $R_{cn} = \frac{\beta}{1-\delta_c} -(Q_{sn} - l_{sn})$ if $\frac{\beta}{1-\delta_c} > (Q_{sn} - l_{sn})$. Otherwise, it invests nothing. 
+	\item Players $s$ and $g$ play Status Quo as long as it yields the highest continuation value, and play Cede or Fight if continuation values from either exceeds the status quo continuation value. 
+\end{itemize}
+
+Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de_c}$ (with a possible lump sum investment at $n=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
+\hfill $\blacksquare$
+
+%Our status quo equilibrium requires that $p$ invests in every period enough to maintain $F^*$ by offsetting the $\mu$ decline in the secessionists' status quo payoffs.
+The total per-period equilibrium investment $R_{pn}$, a flow payment, in the long run is thus $\mu$ per period in this steady state.
+In a status quo equilibrium $c$ need not invest at all since stage game payoffs of $g$ do not deteriorate.\footnote{ In the absence of the assumption in the model setup above that player $c$ does not want war, the patron would need to retain a second buffer against $c$ funding war.
+The expected value of war for the home state would need to be maintained at a level lower than the status quo payoff by a buffer of $F*=\frac{\beta}{1-\delta_c}$.
+If this buffer did not exist in the first round of the game, the patron would fund the alteration of payoffs to create the buffer, thus assuring payments by $c$ would be ineffective at trying to change the home state's payoffs to make war more attractive than the status quo.}
+
+By contributing $\mu$ in each period, the patron supplies sufficient resources to the unrecognized state to ensure that the secessionist elite prefers the status quo to surrendering independence.
+
+\newpage
+Appendix
+
+{\bf Proof of Lemma 1} \\
+Recall that the upper case notation represents payoffs gross of investments by the outside players.
+Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions.
+Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{(1-\delta_s)}$.
+Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{(1-\delta_i)}$.
+The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$.
+Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta_i}$.
+
+In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2).
+The argument for player $g$ is symmetric.
+\hfill $\blacksquare$
+\\
+\\
+\noindent {\bf Proof of Lemma 2} \\
+Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
+If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
+
+Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
+}{1 -\de_p}-R_p^{RU}$.
+For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de_c} -R_c^{RU}$.
+Neither player will be willing to invest in its least preferred state.
+That is, $R_p^{RU} = R_c^{REC} =0$.
+So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de_c}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ}
+\right)}{1 -\de_p}$ each period toward maintaining it.
+
+Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g.
+counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de_c} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.
+
+Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de_p} > \frac{\beta+ R_c^{SQ}}{1 -\de_c}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
+Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
+Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\frac{-\alpha}{1 -\de_p}  > \frac{\beta}{1 -\de_c}$.
+\hfill $\blacksquare$
+\\
+\\
+\noindent {\bf Proof of Lemma 3} \\
+In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
+If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
+
+Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de_p} -R_p^{REC}$.
+For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de_c}-R_c^{REC}$.
+As noted above, $R_c^{REC} =0$.
+So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de_p}$ toward recognition.
+We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de_c}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$.
+
+Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
+counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $-\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.
+
+Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.
+
+Thus $\frac{-\nu}{1 -\de_c} > \frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
+A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
+\hfill $\blacksquare$
+\\
+\\
+\noindent {\bf Proof of Lemma 4} \\
+The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ} -R_p^{WAR}$.
+For player $c$ it is $\frac{\nu p_1 + \beta p_2}{1 -\de_c} + R_c^{SQ} -R_c^{WAR}$.
+As noted above, $R_c^{SQ} =0$.
+So $c$ is willing to invest up to $\frac{\nu p_1 + \beta p_2}{1 -\de_c}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
+
+$p$ is willing to invest up to $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+Thus we must only examine the case where this quantity is positive.
+
+We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}$ as this is most restrictive.
+In this only case of interest, we have $\frac{\alpha p_2}{1 -\de_p} +\frac{\lambda p_1}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0$.
+Note that the first term is negative while the rest are positive.
+
+Multiplying Condition 3 by $-p_{2s}$, we have $\frac{-\beta \cdot p_{2s}}{1-\de_c} \geq \frac{\alpha \cdot p_{2s}}{1-\de_p}$. Using this with Condition 4, we have
+\[
+  \textstyle -\left(\frac{\nu p_{1s} + \beta p_{2s}}{1 -\de_c}\right) = - \frac{\nu p_{1s}}{1 -\de_c} - \frac{\beta p_{2s}}{1 -\de_c} \geq \frac{\alpha p_{2s}}{1 -\de_p} +\frac{\lambda p_{1s}}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0 \ \forall p_{1s},p_{2s}
+\]
+
+Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude.
+Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place.
+\hfill $\blacksquare$
+\\
+
 \end{document}

### 2015-08-29 draft.tex
Fearon 1994 reference added
changed statement of individual discount factors to homogenized discount factor

@@
 
-Future payoffs are discounted with parameters $\delta_i$, where 1/$\delta_i$ is player $i$'s discount rate for $i \in (g, s, c, p)$, $0\leq\delta_i\leq1$.
-Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \delta_i^{n-1}$.
+Future payoffs are discounted with a common parameter $\delta$, where 1/$\delta$ is each player's discount rate and $0\leq\delta\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \delta^{n-1}$.
 
@@
 
+\item \hangindent=1cm Fearon, James D.  1994.  "Domestic Political Audiences and the Escalation of International Disputes."  \emph{The American Political Science Review} 88: 577-92.
+
 \item \hangindent=1cm Fearon, James D., and David D. Laitin. 2003. "Ethnicity, Insurgency, and Civil War." \emph {American Political Science Review} 97 (01): 75-90.

### 2015-08-31 draft.tex
modify lemmas 1 and 2 for homogenized delta

--- 2015-04-09 draft.tex
@@
 
-Future payoffs are discounted with parameters $\delta_i$, where 1/$\delta_i$ is player $i$'s discount rate for $i \in (g, s, c, p)$, $0\leq\delta_i\leq1$.
-Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \delta_i^{n-1}$.
+Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \de^{n-1}$.
 
@@
 
-\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta_i} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta_i} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
-
-\item \textit{$\frac{-\alpha}{1-\de_p} \geq \frac {\beta}{1-\delta_c}$: reunification is more important for the patron to avoid than for the international community to achieve.}
+\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
+
+\item \textit{$-\alpha \geq \beta$: reunification is more important for the patron to avoid than for the international community to achieve.}
 
@@
 Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions.
-Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{(1-\delta_s)}$.
-Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{(1-\delta_i)}$.
+Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{1-\delta}$.
+Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{1-\delta}$.
 The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$.
-Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta_i}$.
+Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta}$.
 
@@
 \noindent {\bf Proof of Lemma 2} \\
-Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
+In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
 If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
@@
 Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
-}{1 -\de_p}-R_p^{RU}$.
-For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de_c} -R_c^{RU}$.
+}{1 -\de}-R_p^{RU}$.
+For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de} -R_c^{RU}$.
 Neither player will be willing to invest in its least preferred state.
 That is, $R_p^{RU} = R_c^{REC} =0$.
-So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de_c}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ}
-\right)}{1 -\de_p}$ each period toward maintaining it.
+So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ}
+\right)}{1 -\de}$ each period toward maintaining it.
 
 Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g.
-counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de_c} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.
 
-Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de_p} > \frac{\beta+ R_c^{SQ}}{1 -\de_c}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
+Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
-Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\frac{-\alpha}{1 -\de_p}  > \frac{\beta}{1 -\de_c}$.
+Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $-\alpha  > \beta$.
 \hfill $\blacksquare$
@@
 \noindent {\bf Proof of Lemma 3} \\
-In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
+Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
 If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

### 2015-09-01 draft.tex
modify lemmas 3 and 4 for homogenized delta

@@
 
-\item  \textit{$\frac {-\nu\cdot p_{1s}}{1-\delta_c} \geq \frac{\lambda \cdot p_{1s} + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
+\item  \textit{$-\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
 
@@
 
-Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de_p} -R_p^{REC}$.
-For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de_c}-R_c^{REC}$.
+Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$.
+For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$.
 As noted above, $R_c^{REC} =0$.
-So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de_p}$ toward recognition.
-We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de_c}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$.
+So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
+We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.
 
@@
 
-Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.
-
-Thus $\frac{-\nu}{1 -\de_c} > \frac{\lambda + \mu}{1 -\de_p} + \frac{\beta}{1 -\de_c}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
+Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.
+
+Thus $\frac{-\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
 A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
@@
 \noindent {\bf Proof of Lemma 4} \\
-The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ} -R_p^{WAR}$.
-For player $c$ it is $\frac{\nu p_1 + \beta p_2}{1 -\de_c} + R_c^{SQ} -R_c^{WAR}$.
+The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_1 + \alpha p_2 }{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
+For player $c$ it is $\frac{\nu p_1 + \beta p_2}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
 As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{\nu p_1 + \beta p_2}{1 -\de_c}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
-
-$p$ is willing to invest up to $\frac{\lambda p_1 + \alpha p_2 }{1 -\de_p} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+So $c$ is willing to invest up to $\frac{\nu p_1 + \beta p_2}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
+
+$p$ is willing to invest up to $\frac{\lambda p_1 + \alpha p_2 }{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
 Thus we must only examine the case where this quantity is positive.
 
-We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}$ as this is most restrictive.
-In this only case of interest, we have $\frac{\alpha p_2}{1 -\de_p} +\frac{\lambda p_1}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0$.
+We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive.
+In this only case of interest, we have $\frac{\alpha p_2 + \lambda p_1 + \mu + \beta}{1 - \de}> 0$.
 Note that the first term is negative while the rest are positive.
 
-Multiplying Condition 3 by $-p_{2s}$, we have $\frac{-\beta \cdot p_{2s}}{1-\de_c} \geq \frac{\alpha \cdot p_{2s}}{1-\de_p}$. Using this with Condition 4, we have
+Multiplying Condition 3 by $-p_{2s}$, we have $-\beta \cdot p_{2s} \geq \alpha \cdot p_{2s}$.
+Using this with Condition 4, we have $- \nu p_{1s} - \beta p_{2s} \geq \alpha p_{2s} +\lambda p_{1s} + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
 \[
-  \textstyle -\left(\frac{\nu p_{1s} + \beta p_{2s}}{1 -\de_c}\right) = - \frac{\nu p_{1s}}{1 -\de_c} - \frac{\beta p_{2s}}{1 -\de_c} \geq \frac{\alpha p_{2s}}{1 -\de_p} +\frac{\lambda p_{1s}}{1 -\de_p} + \frac{\mu}{1-\de_p} + \frac{\beta}{1 - \de_c}> 0 \ \forall p_{1s},p_{2s}
+  \textstyle -\left(\frac{\nu p_{1s} + \beta p_{2s}}{1 -\de}\right) \geq \frac{\alpha p_{2s}+ \lambda p_{1s} + \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
 \]

### 2015-09-03 draft.tex
complete revisions to proposition 1 for homogenized discount factor

@@
 \noindent \bf \emph{Proposition 1:} \rm  \emph{There exists an equilibrium in which the outcome is perpetual unrecognized statehood.
-The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\delta_c}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}
+The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\de}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}
 
@@
 
-\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
 
@@
 Consider period 1.
-When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\delta_c}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
-That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\delta_c} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
-Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\delta_c} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
+When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
+That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
+Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
 
 In periods $n > 1$, there are two cases to consider.
-Either (a) the buffer created in period $n-1$ was precisely the necessary $\frac{\beta}{1-\delta_c}$, or (b) the buffer is larger than $\frac{\beta}{1-\delta_c}$.
-In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\delta_c}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\delta_c}$ provides lower welfare to player $p$.
+Either (a) the buffer created in period $n-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$.
+In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$.
 Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $n$.
 However, because in each period $q_{sn}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
-Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $n > 1$(Condition 6).
+Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $n > 1$ (Condition 6).
 
@@
 	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $n$ whenever this is affordable.
-That is, $p$ invests $R_{pn} = \max\left\{0,\frac{\beta}{1-\delta_c} -(q_{sn} - l_{sn})\right\}\leq B_{pn}$.
+That is, $p$ invests $R_{pn} = \max\left\{0,\frac{\beta}{1-\delta} -(q_{sn} - l_{sn})\right\}\leq B_{pn}$.
 If this inequality is violated, $p$ invests nothing.
-	\item Player $c$ invests $R_{cn} = \frac{\beta}{1-\delta_c} -(Q_{sn} - l_{sn})$ if $\frac{\beta}{1-\delta_c} > (Q_{sn} - l_{sn})$. Otherwise, it invests nothing. 
+	\item Player $c$ invests $R_{cn} = \frac{\beta}{1-\delta} -(Q_{sn} - l_{sn})$ if $\frac{\beta}{1-\delta} > (Q_{sn} - l_{sn})$. Otherwise, it invests nothing. 
 	\item Players $s$ and $g$ play Status Quo as long as it yields the highest continuation value, and play Cede or Fight if continuation values from either exceeds the status quo continuation value. 
@@
 
-Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de_c}$ (with a possible lump sum investment at $n=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
+Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $n=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
 \hfill $\blacksquare$

### 2015-09-03 draft.tex
complete all revisions (prop 2) for homogenized discount factor

@@
 
-Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de_p^m}{1-de_p} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de_p} + \frac{\de_p}{1-\de_p}\alpha (1-p_{1s}-p_{2s})$.
-For player $c$, it is $0 + \frac{\de_p^m}{1-de_p} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de_p} + \frac{\de_p}{1-\de_p}\beta (1-p_{1s}-p_{2s})$.
+Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} + \frac{\de}{1-\de}\alpha (1-p_{1s}-p_{2s})$.
+For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de} + \frac{\de}{1-\de}\beta (1-p_{1s}-p_{2s})$.
 Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
 \[
-  \textstyle \frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de_p} - \frac{\de_p \alpha}{1-\de_p} (p_{1s}+p_{2s}) + \alpha \sum_{i=1}^{m-1} \de_p^i > \frac{\nu p_{1s} + \beta p_{2s}}{1-\de_c} - \frac{\de_c \beta}{1-\de_c} (p_{1s}+p_{2s}) + \beta \sum_{i=1}^{m-1} \de_c^i
+  \textstyle \frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} - \frac{\de \alpha}{1-\de} (p_{1s}+p_{2s}) + \alpha \sum_{i=1}^{m-1} \de^i > \frac{\nu p_{1s} + \beta p_{2s}}{1-\de} - \frac{\de \beta}{1-\de} (p_{1s}+p_{2s}) + \beta \sum_{i=1}^{m-1} \de^i
 \]

### 2015-09-03 draft.tex
modify initial setup in section 2.2.1 to remove non-decisive war

@@
 We denote the payoffs of war as a lottery $\Omega$.\footnote{Because the unrecognized state already controls territory and the \emph{de facto} borders are armed, there is likely only a small advantage to be gained by attacking first for either side.
-Therefore, we argue it is not essential to differentiate between these war scenarios analytically.} This lottery determines one of three potential outcomes: outright victory by either of the players, which are both absorbing states, and an indecisive war where costs of war are borne but then the game returns to the status quo in which the game continues to the next period.
-Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
+Therefore, we argue it is not essential to differentiate between these war scenarios analytically.} This lottery determines whether the secessionists or government wins the war.
+For simplicity, we assume the outcome of the war is an absorbing state.\footnote{The model can easily be extended to encompass the possibility of an indecisive war.} Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
 Therefore, outright victory requires more than simply securing control of the disputed territory (which the secessionists have already done when the game begins) and requires the ability to impose the terms of settlement.
@@
 
-For probabilities $p_1$ of outright victory, $p_2$ of loss and $1-p_1-p_2$ of non-decisive war, player $i \in \{s, g\}$ in period $n$ with a fixed cost of war $\zeta_i$
-faces war lottery $ \omega_{in} \equiv (p_1 (W_{in}-\zeta_i), p_2( L_{in}-\zeta_i), (1-p_1-p_2)(Q_{in}-\zeta_i))$.
+For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $n$ with a fixed cost of war $\zeta_i$ faces war lottery $ \omega_{in} \equiv (p_1 (W_{in}-\zeta_i), \left(1-p_1\right)( L_{in}-\zeta_i))$.
 $\Omega_n \equiv (\omega_{gn}, \omega_{sn}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.} Players are assumed to approach war lotteries as expected values.

### 2015-09-03 draft.tex
begin revising proposition 1 for decisive war

@@
 
-\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta} \geq  -\zeta_i+\frac{L_{in}(p_2) +W_{in}(p_1)+Q_{in}(1-p_1-p_2)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
+\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta} \geq  -\zeta_i+\frac{W_{in}(p_1) + L_{in}(1-p_1)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
 
@@
 Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3).
-Although period $n$ may be reached because in period $n-1$ fighting resulted in a stalemate, both players ceded, or the status quo had been maintained, the strategic landscape in period $n$ is the same.
+Although period $n$ may be reached either because in period $n-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $n$ is the same.
 Thus all periods in which the players are able to move are in an equivalence class in which the only possible strategic difference is in the value of the state variables.
@@
 Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{1-\delta}$.
-The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, $L_s$ forever with probability $p_2$, and the status quo payoff forever with probability $1-p_1-p_2$.
-Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(p_2) +Q_{sn}(1-p_1-p_2)}{1-\delta}$.
+The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$.
+Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(1-p_1)}{1-\delta}$.
 

### 2015-09-04 draft.tex
revise lemma 4 for decisive war

@@
 \noindent {\bf Proof of Lemma 4} \\
-The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_1 + \alpha p_2 }{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
-For player $c$ it is $\frac{\nu p_1 + \beta p_2}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
+The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
+For player $c$ it is $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
 As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{\nu p_1 + \beta p_2}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
-
-$p$ is willing to invest up to $\frac{\lambda p_1 + \alpha p_2 }{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+So $c$ is willing to invest up to $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
+
+$p$ is willing to invest up to $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
 Thus we must only examine the case where this quantity is positive.
@@
 We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive.
-In this only case of interest, we have $\frac{\alpha p_2 + \lambda p_1 + \mu + \beta}{1 - \de}> 0$.
+In this only case of interest, we have $\frac{\alpha (1-p_{1s}) + \lambda p_{1s} + \mu + \beta}{1 - \de}> 0$.
 Note that the first term is negative while the rest are positive.
 
-Multiplying Condition 3 by $-p_{2s}$, we have $-\beta \cdot p_{2s} \geq \alpha \cdot p_{2s}$.
-Using this with Condition 4, we have $- \nu p_{1s} - \beta p_{2s} \geq \alpha p_{2s} +\lambda p_{1s} + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
+Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq \alpha \cdot (1-p_{1s})$.
+Using this with Condition 4, we have $- \nu p_{1s} - \beta (1-p_{1s}) \geq \alpha (1-p_{1s}) +\lambda p_{1s} + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
 \[
-  \textstyle -\left(\frac{\nu p_{1s} + \beta p_{2s}}{1 -\de}\right) \geq \frac{\alpha p_{2s}+ \lambda p_{1s} + \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
+  \textstyle -\left(\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\alpha (1-p_{1s})+ \lambda p_{1s} + \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
 \]

### 2015-09-07 draft.tex
revise proof of prop 2 for decisive war

@@
 
-Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} + \frac{\de}{1-\de}\alpha (1-p_{1s}-p_{2s})$.
-For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de} + \frac{\de}{1-\de}\beta (1-p_{1s}-p_{2s})$.
+Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de}$.
+For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de}$.
 Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
 \[
-  \textstyle \frac{\lambda p_{1s} + \alpha p_{2s}}{1-\de} - \frac{\de \alpha}{1-\de} (p_{1s}+p_{2s}) + \alpha \sum_{i=1}^{m-1} \de^i > \frac{\nu p_{1s} + \beta p_{2s}}{1-\de} - \frac{\de \beta}{1-\de} (p_{1s}+p_{2s}) + \beta \sum_{i=1}^{m-1} \de^i
+  \textstyle \frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de} - \frac{\de^m}{1-\de} \alpha > \frac{\nu p_{1s} + \beta (1-p_{1s})}{1-\de} - \beta \frac{\de^m}{1-\de} \alpha
 \]
@@
 Status Quo is no longer a long-run equilibrium strategy, but even without outside investments, it is possible that Fight could be a more attractive action than Cede.
-Degrading the status quo payoffs is not sufficient to make the continuation value from Cede the highest available.
-
-To see why, note that the sanctions only have a partial effect in reducing the Fight continuation value through the war lottery, while their entire impact is felt on the Status Quo continuation value.
-Thus the gap between the Status Quo and Fight continuation values narrows at the same time as the corresponding gap between the Status Quo and Cede continuation values.
-Although the former gap narrows more slowly, it is likely to be smaller at the outset so sanctions do not necessarily make Cede the most preferred action for player $s$.
-
-Therefore, in order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight.
+
+The status quo payoffs will degrade period by period in the absence of investments.
+Whether the secessionists choose Cede or Fight depends on which strategy yields the higher continuation value.
+In order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight.
 \hfill $\blacksquare$

### 2015-09-07 draft.tex
simplify statement of prop 2 for decisive war
change statment of alpha and nu parameters to make them positive

@@
 
-Player $p$'s payoffs in period $n$ are $U_{pn}= \alpha X+\lambda Y -R_{pn}$, while the payoffs for player $c$ are $U_{cn}= \beta X+\nu Y -R_{cn}$, with both payoffs denoted in currency units.
-
-The patron opposes reunification while player $c$ prefers it, so $\alpha$ is negative and $\beta$ is positive.
-The patron prefers recognized independence so $\lambda$ is positive.
-Player $c$ is averse to the creation of new states so $\nu$ is negative.
+Player $p$'s payoffs in period $n$ are $U_{pn}= -\alpha X+\lambda Y -R_{pn}$, while the payoffs for player $c$ are $U_{cn}= \beta X-\nu Y -R_{cn}$, with both payoffs denoted in currency units.
+Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
 
@@
 
-\item \textit{The continuation value from playing Status Quo must fall below that from playing Cede before it falls below that from playing Fight. }
+\item \textit{The secessionists' continuation value from playing Cede must be higher than their continuation value from playing Fight. }
 \end{enumerate}

### 2015-09-07 draft.tex
fix lemma 2 and condition 3 of prop 1 for alpha positive

@@
 
-\item \textit{$-\alpha \geq \beta$: reunification is more important for the patron to avoid than for the international community to achieve.}
+\item \textit{$\alpha \geq \beta$: reunification is more important for the patron to avoid than for the international community to achieve.}
 
@@
 In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
-If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= \alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
-
-Thus the difference in continuation values from switching from the status quo to reunification for player $p$ is $\frac{\alpha + R_p^{SQ} 
-}{1 -\de}-R_p^{RU}$.
-For player $c$ it is $\frac{\beta + R_c^{SQ}}{1 -\de} -R_c^{RU}$.
-Neither player will be willing to invest in its least preferred state.
-That is, $R_p^{RU} = R_c^{REC} =0$.
-So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de}$ toward reunification---or preventing the status quo---and $p$ is willing to invest $\frac{-\left(\alpha + R_p^{SQ}
-\right)}{1 -\de}$ each period toward maintaining it.
+If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= -\alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
+
+Since reunification is the patron's least preferred state, $R_p^{RU} =0$; that is, the patron is not willing to invest to provoke reunification.
+But the patron \textit{would} be willing to invest the amount it would lose from reunification to maintain the Status Quo: that is, $R_p^{SQ} = \frac{\alpha}{1 -\de}$.
+
+For player $c$ the difference in continuation values from switching from the status quo to reunification is $\frac{\beta + R_c^{SQ}}{1 -\de} -R_c^{RU}$.
+So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de}$ toward reunification---or preventing the status quo.
 
@@
 
-Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{-\alpha - R_p^{SQ}}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
+Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
-Given that $R_p^{SQ}$ is non-negative, a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $-\alpha  > \beta$.
+Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$.
 \hfill $\blacksquare$

### 2015-09-08 draft.tex
fix lemmas 3 and 4 for alpha and nu positive

@@
 
-\item  \textit{$-\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
+\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
 
@@
 Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
-If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= \nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
+If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
 
 Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$.
-For player $c$ it is $\frac{\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$.
-As noted above, $R_c^{REC} =0$.
-So $c$ is willing to invest up to $-\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
+For player $c$ it is $\frac{-\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$.
+$R_c^{REC} =0$ since the international community will not invest to create its least preferred outcome.
+So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
 We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.
@@
 Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $-\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.
@@
 
-Thus $\frac{-\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
+Thus $\frac{\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
 A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
@@
 \noindent {\bf Proof of Lemma 4} \\
-The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
-For player $c$ it is $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
+The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
+For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
 As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
-
-$p$ is willing to invest up to $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+So $c$ is willing to invest up to $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
+
+$p$ is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
 Thus we must only examine the case where this quantity is positive.
@@
 We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive.
-In this only case of interest, we have $\frac{\alpha (1-p_{1s}) + \lambda p_{1s} + \mu + \beta}{1 - \de}> 0$.
-Note that the first term is negative while the rest are positive.
-
-Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq \alpha \cdot (1-p_{1s})$.
-Using this with Condition 4, we have $- \nu p_{1s} - \beta (1-p_{1s}) \geq \alpha (1-p_{1s}) +\lambda p_{1s} + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
+In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}> 0$.
+
+Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$.
+Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
 \[
-  \textstyle -\left(\frac{\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\alpha (1-p_{1s})+ \lambda p_{1s} + \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
+  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
 \]

### 2015-09-08 draft.tex
fix proof of prop 2 for alpha and nu positive

@@
 Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$.
-Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s},p_{2s}$ and therefore also
+Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
 \[
-  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s},p_{2s}
+  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}
 \]
@@
 
-Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 + \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de}$.
-For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{\nu p_{1s} + \beta p_{2s}}{1-\de}$.
+Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de}$.
+For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_{1s} + \beta (1 - p_{1s})}{1-\de}$.
 Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
 \[
-  \textstyle \frac{\lambda p_{1s} + \alpha (1-p_{1s})}{1-\de} - \frac{\de^m}{1-\de} \alpha > \frac{\nu p_{1s} + \beta (1-p_{1s})}{1-\de} - \beta \frac{\de^m}{1-\de} \alpha
+  \textstyle \frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_{1s} + \beta (1-p_{1s})}{1-\de} -  \frac{\de^m}{1-\de} \beta
 \]

### 2015-09-10 draft.tex
change 'n' to 't' throughout section 2

@@
 The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
-There are an infinite number of periods $n=1,2,\ldots$, with play proceeding until an absorbing state is reached.
+There are an infinite number of periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
 
 Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
-This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{n=1}^{\infty} U_{in} \de^{n-1}$.
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_{it} \de^{t-1}$.
 
@@
 
-\item $p$ chooses a level of resource expenditure $R_{pn}\in \mathbb{R}^+$ in period $n$ and invests it to alter the payoffs of players $s$ and/or $g$.
-
-\item $c$ chooses a level of resource expenditure $R_{cn}\in \mathbb{R}^+$ in period $n$ and invests it to alter the payoffs of players $s$ and/or $g$.
+\item $p$ chooses a level of resource expenditure $R_{pt}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
+
+\item $c$ chooses a level of resource expenditure $R_{ct}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
 
@@
   $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	Fight & $\Omega_n$ &$\Omega_n$&$W_{gn}, L_{sn}$ \\ \hline
-	Status Quo& $\Omega_n$& $Q_{gn}, Q_{sn}$ &$W_{gn}, L_{sn}$ \\ \hline
-	Cede&$L_{gn}, W_{sn}$& $L_{gn}, W_{sn}$ &$Q_{gn}, Q_{sn}$ \\ \hline
+	Fight & $\Omega_t$ &$\Omega_t$&$W_{gt}, L_{st}$ \\ \hline
+	Status Quo& $\Omega_t$& $Q_{gt}, Q_{st}$ &$W_{gt}, L_{st}$ \\ \hline
+	Cede&$L_{gt}, W_{st}$& $L_{gt}, W_{st}$ &$Q_{gt}, Q_{st}$ \\ \hline
  \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
@@
 
-For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $n$ with a fixed cost of war $\zeta_i$ faces war lottery $ \omega_{in} \equiv (p_1 (W_{in}-\zeta_i), \left(1-p_1\right)( L_{in}-\zeta_i))$.
-$\Omega_n \equiv (\omega_{gn}, \omega_{sn}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.} Players are assumed to approach war lotteries as expected values.
+For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_{it} \equiv (p_1 (W_{it}-\zeta_i), \left(1-p_1\right)( L_{it}-\zeta_i))$.
+$\Omega_t \equiv (\omega_{gt}, \omega_{st}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.} Players are assumed to approach war lotteries as expected values.
 
@@
 As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
-Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_{s,n+1} = Q_{s,n} - \mu$.
+Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_{s,t+1} = Q_{s,t} - \mu$.
 Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.
 Other payoffs remain unchanged from period to period; they can only be altered by the investments of players $p$ or $c$, e.g.
-$L_{s,n} = l_{s,n} + R_{p,n} + R_{c,n}$ and $l_{s,n+1} = L_{s,n}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
+$L_{s,t} = l_{s,t} + R_{p,t} + R_{c,t}$ and $l_{s,t+1} = L_{s,t}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
 %who can expend resources ($R$) to increase a payoff at the rate $\gamma$.
@@
 At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_{pn}$ to increase the secessionists status quo payoffs, the beginning-of-period payoff $q_{sn}$ becomes end-of-period payoff $Q_{sn} = q_{sn} + R_{pn}$.
-We can write analogous expressions for $q_{gn}, \ l_{sn},\ l_{gn},\ w_{sn}$, and $w_{sn}$, and the international community can also make investments to increase whichever of these payoffs they like.
-$B_{pn}$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
+For example, if $p$ expends resources $R_{pt}$ to increase the secessionists status quo payoffs, the beginning-of-period payoff $q_{st}$ becomes end-of-period payoff $Q_{st} = q_{st} + R_{pt}$.
+We can write analogous expressions for $q_{gt}, \ l_{st},\ l_{gt},\ w_{st}$, and $w_{st}$, and the international community can also make investments to increase whichever of these payoffs they like.
+$B_{pt}$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
 
@@
 
-Player $p$'s payoffs in period $n$ are $U_{pn}= -\alpha X+\lambda Y -R_{pn}$, while the payoffs for player $c$ are $U_{cn}= \beta X-\nu Y -R_{cn}$, with both payoffs denoted in currency units.
+Player $p$'s payoffs in period $t$ are $U_{pt}= -\alpha X+\lambda Y -R_{pt}$, while the payoffs for player $c$ are $U_{ct}= \beta X-\nu Y -R_{ct}$, with both payoffs denoted in currency units.
 Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.

### 2015-09-10 draft.tex
change 'n' to 't' throughout section 3

@@
 \begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_{in} \geq L_{in} \ \forall n$: in each period, remaining in the status quo is better than ceding.}
-
-\item \emph{For both players $g$ and $s$, $\frac {Q_{in}}{1-\delta} \geq  -\zeta_i+\frac{W_{in}(p_1) + L_{in}(1-p_1)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
+\item \textit{For both players $g$ and $s$, $Q_{it} \geq L_{it} \ \forall t$: in each period, remaining in the status quo is better than ceding.}
+
+\item \emph{For both players $g$ and $s$, $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_1) + L_{it}(1-p_1)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
 
@@
 
-\item \textit{$B_{pn} \geq \mu, \ \forall n>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
-If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $n> 1$ compared to period $1$, (6) could be binding.}
+\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
+If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}
 
@@
 For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run.
-The inequality \emph{will} bind for some set of periods $n \geq 1$ because the secessionist payoffs from the status quo decrease over time.
+The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
 In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
@@
 Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3).
-Although period $n$ may be reached either because in period $n-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $n$ is the same.
+Although period $t$ may be reached either because in period $t-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $t$ is the same.
 Thus all periods in which the players are able to move are in an equivalence class in which the only possible strategic difference is in the value of the state variables.
@@
 
-To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{sn},q_{gn},w_{sn},w_{gn},l_{sn},$ and $l_{gn}$.\footnote{In the case of $w_{sn}$ and $w_{gn}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e.
+To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e.
 \emph{Recognition, Status Quo}, or \emph{Reunification}---they will target given that they will choose the most efficient way to alter the payoffs of players $s$ and $g$.
@@
 
-In periods $n > 1$, there are two cases to consider.
-Either (a) the buffer created in period $n-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$.
+In periods $t > 1$, there are two cases to consider.
+Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$.
 In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$.
-Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $n$.
-However, because in each period $q_{sn}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
-Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $n > 1$ (Condition 6).
+Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
+However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
+Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Condition 6).
 
@@
 \begin{itemize}
-	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $n$ whenever this is affordable.
-That is, $p$ invests $R_{pn} = \max\left\{0,\frac{\beta}{1-\delta} -(q_{sn} - l_{sn})\right\}\leq B_{pn}$.
+	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $t$ whenever this is affordable.
+That is, $p$ invests $R_{pt} = \max\left\{0,\frac{\beta}{1-\delta} -(q_{st} - l_{st})\right\}\leq B_{pt}$.
 If this inequality is violated, $p$ invests nothing.
-	\item Player $c$ invests $R_{cn} = \frac{\beta}{1-\delta} -(Q_{sn} - l_{sn})$ if $\frac{\beta}{1-\delta} > (Q_{sn} - l_{sn})$. Otherwise, it invests nothing. 
+	\item Player $c$ invests $R_{ct} = \frac{\beta}{1-\delta} -(Q_{st} - l_{st})$ if $\frac{\beta}{1-\delta} > (Q_{st} - l_{st})$. Otherwise, it invests nothing. 
 	\item Players $s$ and $g$ play Status Quo as long as it yields the highest continuation value, and play Cede or Fight if continuation values from either exceeds the status quo continuation value. 
@@
 
-Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $n=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
+Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $t=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
 \hfill $\blacksquare$

### 2015-09-10 draft.tex
change 'n' to 't' throughout appendix

@@
 The international community, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.
-sanctions are equivalent to $R_{cn} < 0$).
+sanctions are equivalent to $R_{ct} < 0$).
 Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.
@@
 Recall that the upper case notation represents payoffs gross of investments by the outside players.
-Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions.
-Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{1-\delta}$.
-Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{1-\delta}$.
+Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions.
+Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{st} \ \forall t$, so the continuation value is $\frac{Q_{st}}{1-\delta}$.
+Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{st}$ in each period, is $\frac{L_{st}}{1-\delta}$.
 The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$.
-Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(1-p_1)}{1-\delta}$.
+Thus the continuation value is $\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.
 
@@
 \noindent {\bf Proof of Lemma 2} \\
-In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
-If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= -\alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
+In the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$.
+If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pt}^{RU}= -\alpha -R_p^{RU}$ and $U_{ct}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
 
@@
 Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g.
-counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{st} - l_{st}$, where $Q_{st}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.
 
-Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest.
+Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{st} - l_{st}$ so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
@@
 \noindent {\bf Proof of Lemma 3} \\
-Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$.
-If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
+Again, in the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$.
+If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pt}^{REC}= \lambda -R_p^{REC}$ and $U_{ct}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
 
@@
 So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
-We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.
+We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $t=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.
 
 Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting.
+counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\nu \geq L_{gt} - q_{gt}$, where $L_{gt}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.
 
-Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.
+Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gt} - q_{gt}$.
 
@@
 \[
-  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}
+  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}.
 \]
@@
 {\bf Proof of Proposition 2} \\
-Suppose that sanctions are applied by player $c$ in period $n+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $n$.
-As the only alteration is a reduction in $Q_{sn}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed.
+Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$.
+As the only alteration is a reduction in $Q_{st}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed.
 As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

### 2015-09-11 draft.tex
rewrite description of status quo payoffs as state variable

@@
 If both $s$ and $g$ play Status Quo, then the status quo persists.
-The status quo payoffs for the secessionists are modeled as steadily declining due to the costs of non-recognition discussed in the introduction.
+The status quo payoffs for the secessionists are modeled as a state variable.
+The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
 As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
@@
 Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.
-Other payoffs remain unchanged from period to period; they can only be altered by the investments of players $p$ or $c$, e.g.
+
+Other payoffs are also modeled as state variables.
+However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
 $L_{s,t} = l_{s,t} + R_{p,t} + R_{c,t}$ and $l_{s,t+1} = L_{s,t}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
-%who can expend resources ($R$) to increase a payoff at the rate $\gamma$.
-For example, if $p$ expends resources to aid the secessionists, the payoff $Q_s$ becomes $Q_s+\gamma R_p$ in the round that the resources are expended.
 

### 2015-09-11 to_do_list.tex
Added to-dos for Ben based on draft response memo

@@
 
+Ben
+\begin{itemize}
+	\item R1: Finish \& Double Check Fearon reference
+		\item add appropriate language to memo if necessary
+
+	\item R1: Talk to Kristy about whether we want to offer a numerical example a la what R1 requests.
+
+	\item R1: Tone down our claims about being novel and unique
+
+	\item R1: Double check that we have fixed the typo identified by R1
+
+	\item R3: Read the literature referenced by R3 and add some cites
+
+	\item R3: Read Berg 2009 in particular and engage it at least briefly in the paper.
+
+\end{itemize}
+
 \end{document}

### 2015-09-12 draft.tex
fix typo in footnote about status quo payoffs as state variable, closes issue #1

@@
 The status quo payoffs for the secessionists are modeled as a state variable.
-The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
+The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists' status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
 As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.

### 2015-09-17 draft.tex
Added paragraph discussing Fearon 1994

--- 2015-09-11 draft.tex
@@
 \subsection{Unrecognized States in the Literature}
-
 %Political scientists care a great deal about both civil conflict and about state formation, but unrecognized states do not sit neatly in either literature.
@@
 
+While we do not model this directly, the stability of the status quo equilibrium may be further enhanced by the reluctance of the patron to withdraw support for an unrecognized state that was previously supported.
+Fearon (1994) and subsequent work on audience costs suggests that, past a certain threshold of escalation, it becomes difficult for leaders to back down from confrontation because they fear looking weak in front of domestic audiences.
+Leaders of patrons that have supported the secessionists in a previous period may risk losing office if they later back down -- even if it is the best interest of the patron state to do so.
+
 \subsection{Alternative Outcomes}

### 2015-09-17 draft.tex
Edited draft

@@
 
-We develop a novel, four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome.
+We develop a four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome.
 These outside actors are a patron, which prefers independence to reunification, and a fourth player we denote as player $c$ who represents one or more members of the international community.
@@
 While the importance of outside actors in civil conflict has been widely acknowledged in the empirical literature, it is rarely modeled formally.
-We introduce a unique four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome.
+We introduce a four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome.
 This allows us both to examine the means through which patron states sustain unrecognized statehood as a stable equilibrium, and to rigorously analyze the strategies available to other outside actors to pursue reunification.

### 2015-10-01 eqm_profiles.tex
add eqm_profiles.tex to house work on specifying full equilibrium strategy profiles

--- 2015-06-19 JLEO_eqm.tex
@@
 \begin{center}
-Unrecognized States -- JLEO Streamlined Model and Equilibrium
+Equilibrium Strategy Profiles (For JLEO Revision)
 \end{center}
@@
 \vskip.3in
-Kristy
-\begin{itemize}
-	\item Take indecisive war to decisive war
-	\item Change sign on payoff functions for $c$ and $p$
-	\item Homogenize discount factors
-	\item Decide on Reviewer 1's suggestion to recast Proposition 1 as a ``class of games'' instead of ``sufficient conditions''
-		\begin{itemize}
-			\item ``I understand why the set of sufficient conditions in this proposition is the key result.
-However, these mechanisms are essentially just assumptions on payoffs; and therefore I would be inclined to interpret these conditions as characterizations of “a class” of games, where this class would satisfy all the assumptions on payoffs (instead of saying sufficient conditions).
-\\
-				I would recommend clearing up Section 3.1 in terms of packaging the results: 
-					\begin{enumerate}
-						\item First, I would identify a class of games that satisfy the assumptions on payoffs (sufficient conditions (1) through (6) in Proposition 1). Denote this class as G for example. 
-						\item Then, formally define the “status quo” equilibrium – an equilibrium in which the equilibrium outcome is perpetual unrecognized statehood with all the actors’ strategies specified. 
-						\item Finally, I would state a much clearer and simpler proposition, e.g., For any game G, there exists a status quo equilibrium.
-...
-Then subsection 3.2 would basically be a discussion of this class of games.
-					\end{enumerate}
-		\end{itemize}
-	\item Clearly specify punishments / equilibrium path to satisfy editor
-		\begin{itemize}
-			\item ``Clear and detailed description of the noncooperative game''
-			\item ``I would also like to see the key logic laid out better in the text, and the full
-equilibrium strategy profiles noted so that it is clear how the players are rewarding and punishing each other on and off the equilibrium path''
-		\end{itemize}
-	\item Make sure equilibrium concept is clearly stated and defined
-		\begin{itemize}
-			\item ``it is important to provide formal definitions and to clearly describe the equilibrium concept being utilized.
-It seems to me that, as Referee 2 states, you are using the Markov perfect equilibrium concept.''
-		\end{itemize}
-\end{itemize}
 

### 2015-10-01 eqm_profiles.tex
add state variable details and dependence of each players' strategies on it
add period 1 strategies

@@
 \vskip.3in
+First, need to get timing clear for evolution of state variable. Take state variable to be $q_{st}$.
+\begin{itemize}
+	\item $q_{s1}$: beginning of the world (or, $Q_{s0} - \mu = q_{s1}$)
+	\item $Q_{s1} = q_{s1} + R_{p1} + R_{c1}$
+		\begin{itemize}
+			\item $q_{s2} = Q_{s1} - \mu$
+		\end{itemize}
+\end{itemize}
+
+Strategies
+\begin{itemize}
+	\item For patron: function of $q_{st}$
+	\item For c: function of $q_{st} + R_{pt}$
+	\item For s: function of $Q_{st}=q_{st} + R_{pt} + R_{ct}$
+	\item For g: not a function of $q_{st}$ at all
+\end{itemize}
+
+For Markov-perfect equilibrium, strategy profile must be dependent on state variable, $q_{s}$ only.
+
+\vskip.5in
+Period 1
+\begin{itemize}
+	\item Patron: $R_{p1} = \frac{\beta}{1-\de} - \left(q_{s1} - l_{s1}\right)$ to augment $q_{s1}$ if this is greater than 0. Else, $R_{p1} = 0$. Can write in max language.
+	\item International community: If $R_{p1} + q_{s1} - l_{s1} \geq \frac{\beta}{1-\de}$, $R_{c1} =0$. Otherwise $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right)$ to augment $l_{s1}$.
+	\item Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega{i1}$
+\end{itemize}
 

### 2015-10-01 eqm_profiles.tex
add notes on off-eqm path actions in first period

@@
 	\item Patron: $R_{p1} = \frac{\beta}{1-\de} - \left(q_{s1} - l_{s1}\right)$ to augment $q_{s1}$ if this is greater than 0. Else, $R_{p1} = 0$. Can write in max language.
-	\item International community: If $R_{p1} + q_{s1} - l_{s1} \geq \frac{\beta}{1-\de}$, $R_{c1} =0$. Otherwise $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right)$ to augment $l_{s1}$.
-	\item Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega{i1}$
+		\begin{itemize}
+			\item Patron might also want to invest to encourage recognition.
+Most efficient way to do this is to augment $L_{g1}$ (other options are to push either $g$ or $s$ to war, but probability of other outcome diminishes efficiency of investment, and assume payoff to war is lower to begin with).
+		\end{itemize}
+
+	\item International community: 
+		\begin{enumerate}
+			\item If $R_{p1} + q_{s1} - l_{s1} \geq \frac{\beta}{1-\de}$, $R_{c1} =0$
+			\item Otherwise $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right) + \ve$ to augment $l_{s1}$
+				\begin{itemize}
+					\item (SQ,Cede) is played and game ends
+				\end{itemize}
+
+		\end{enumerate}
+
+	\item Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega_{i1}$
+		\begin{itemize}
+			\item Game only continues if (SQ,SQ) or (Cede, Cede) was played
+		\end{itemize}
+
 \end{itemize}

### 2015-10-02 eqm_profiles.tex
reorganize to handle four ways that patron could invest

@@
 \vskip.5in
+Note that patron will never invest to augment $l_s$ or $w_g$. This leave four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$.
+\begin{itemize}
+	\item Also, Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega_{i1}$
+		\begin{itemize}
+			\item Game only continues if (SQ,SQ) or (Cede, Cede) was played
+		\end{itemize}
+\end{itemize}
+
 Period 1
-\begin{itemize}
+\begin{enumerate}
 	\item Patron: $R_{p1} = \frac{\beta}{1-\de} - \left(q_{s1} - l_{s1}\right)$ to augment $q_{s1}$ if this is greater than 0. Else, $R_{p1} = 0$. Can write in max language.
 		\begin{itemize}
-			\item Patron might also want to invest to encourage recognition.
-Most efficient way to do this is to augment $L_{g1}$ (other options are to push either $g$ or $s$ to war, but probability of other outcome diminishes efficiency of investment, and assume payoff to war is lower to begin with).
-		\end{itemize}
-
 			\item International community: 
@@
 						\end{itemize}
-
+				\end{enumerate}
+		\end{itemize}	
+	\item Patron might also want to invest to encourage recognition.
+Most efficient way to do this is to augment $L_{g1}$ (other options are to push either $g$ or $s$ to war, but probability of other outcome diminishes efficiency of investment, and assume payoff to war is lower to begin with).
 \end{enumerate}
 
-	\item Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega_{i1}$
-		\begin{itemize}
-			\item Game only continues if (SQ,SQ) or (Cede, Cede) was played
-		\end{itemize}
-
-\end{itemize}
-
 \end{document}

### 2015-10-02 eqm_profiles.tex
add rest of four cases for patron to invest

@@
 Most efficient way to do this is to augment $L_{g1}$ (other options are to push either $g$ or $s$ to war, but probability of other outcome diminishes efficiency of investment, and assume payoff to war is lower to begin with).
+	\item Patron is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+		\begin{itemize}
+			\item $c$ will counter up to this amount (NEED TO CHECK THAT THESE CALCULATONS ARE CORRECT GIVEN THAN INVESTMENT IN $w_s$ will be weighted down)
+		\end{itemize}
+	\item Patron could also invest in $g_q$ to counter something by $c$. But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out
 \end{enumerate}

### 2015-10-03 eqm_profiles.tex
clean up four cases for patron to invest

--- 2015-10-02 eqm_profiles.tex
@@
 		\end{itemize}	
-	\item Patron might also want to invest to encourage recognition.
-Most efficient way to do this is to augment $L_{g1}$ (other options are to push either $g$ or $s$ to war, but probability of other outcome diminishes efficiency of investment, and assume payoff to war is lower to begin with).
+	\item Patron might also want to invest to encourage recognition directly. How much to augment $l_{g1}$?	
+	\item Patron might also want to instigate fighting.
+If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}$ is positive, invest in $w_{s1}$ to change $\omega_{s1}$.
+Will not invest in $w_{s1}$ when the quantity is non-positive.
+		\begin{itemize}
+			\item $c$ will counter up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$
+		\end{itemize}
+	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
+But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
 \end{enumerate}

### 2015-10-03 eqm_profiles.tex
add defn of MPE and description of state variables

@@
 
-For Markov-perfect equilibrium, strategy profile must be dependent on state variable, $q_{s}$ only.
+\vskip.5in	
+According to Mailath and Samuelson (2006) p. 177.
+\begin{itemize}
+	\item The strategy profile $\sigma$ is a stationary Markov strategy if for any two ex post histories $\tilde{h}^t$ and $\tilde{h}^\tau$ (of equal or different lengths) terminating in the same state, $\sigma\left(\tilde{h}^t\right) = \sigma\left(\tilde{h}^\tau\right)$.
+	\item The strategy profile $\sigma$ is a stationary Markov equilibrium if $\sigma$ is a stationary Markov strategy profile and a subgame-perfect equilibrium.
+\end{itemize}
+
+\vskip.5in
+We have six state variables, so $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
+\begin{itemize}
+	\item Strategies are a function of all six state variables
+	\item Strategies for the patron and $c$ are also how much to invest in each of the six state variables. Some can be ruled out by preference assumptions:
+		\begin{itemize}
+			\item $c$ dislikes war, so will never invest in $w_s$ or $w_g$. It would also not want to make the government lose, so won't invest in $l_g$ either.
+			\item Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
+		\end{itemize}
+\end{itemize}
 

### 2015-10-14 draft.tex
add assumption that inside actors break ties in favor of status quo
Added 1 sentence and 2 footnotes regarding complexity of dependence on patron (for R3)

@@
 In practice we believe that in most cases there are many states that may act as player $c$ in our model, and often we observe groups of states like the OECD or the UN acting in this capacity.
+Modeling the international community as a single actor is a strong simplifying assumption, but by leaving the preferences of player $c$ quite general, the model can accommodate much of this diversity.\footnote{A variety of scholars, including Owtram (2011) and Coggins (2011) discuss the international relations of unrecognized states and variation across third-party states in their preferences with regard to unrecognized states.}
 
@@
 
-Analysis of the posited equilibrium proceeds most naturally by backward induction within each period. Thus we begin with the stage game between the government and secessionists.
+Analysis of the posited equilibrium proceeds most naturally by backward induction within each period.
+Thus we begin with the stage game between the government and secessionists.
+We assume that the government and secessionists choose the Status Quo strategy whenever indifferent between it and any other strategy.
 
@@
 The only unrecognized state currently in existence which has been able to survive without a patron is Somaliland.
-Somaliland has been able to survive as long as it has because of the extreme weakness of the home state (Somalia).
+Somaliland has been able to survive as long as it has because of the extreme weakness of the home state (Somalia).\footnote{Owtram (2011) argues that, over time, unrecognized states can become sufficiently "consolidated" as to no longer be dependent on a patron for support.
+However, no cases other than Somaliland support this claim empirically.}
 

### 2015-10-14 draft.tex
Added Owtram cite to biblio
Added summary of lit and 1 cite to biblio for R3

@@
 Pegg 1998; Kolst\o \ 2006; Geldenhuys 2009; Caspersen and Stansfield 2011; Caspersen 2013).
+Bakke (2011) summarizes four primary explanations for the persistence of unrecognized states, each of which plays a role in the theory we develop here.
+Fist, economic and political elites often benefit from the status quo (e.g.
+King 2001).
+Second, the secessionists have sufficient military capacity to make military reconquest costly (e.g.
+Kolst\o \ 2006).
+Third, secessionists often enjoy external support from a patron (e.g.
+Lynch 2002; Stanislawski 2008.
+Lastly, the governments of unrecognized states are often internally legitimate, enjoying support from residents of the unrecognized state (e.g.
+Lynch 2002, Kolst\o \ 2006).
+
  However, the literature continues to lack a clear general theory specifying the conditions under which this outcome emerges and persists.
@@
 
+\item \hangindent=1cm Bakke, Kristin M.
+2011 "After the war ends: violence and vialibity of post-Soviet unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
+\emph{Unrecognized States in the International System}.
+New York, NY: Routledge.
+
 \item \hangindent=1cm Bakke, Kristin M., John O'Loughlin, Gerard Toal, and Michael D.
@@
 \item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, settlement, and the shadow of the future." \emph{Journal of Economic Behavior $\&$ Organization} 105: 75-89.
+
+\item \hangindent=1cm Owtram, Francis.
+2011.
+"The foreign policies of unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
+\emph{Unrecognized States in the International System}.
+New York, NY: Routledge.
 

### 2015-10-15 draft.tex
Fixed 2 typos

@@
 Pegg 1998; Kolst\o \ 2006; Geldenhuys 2009; Caspersen and Stansfield 2011; Caspersen 2013).
-Bakke (2011) summarizes four primary explanations for the persistence of unrecognized states, each of which plays a role in the theory we develop here.
-Fist, economic and political elites often benefit from the status quo (e.g.
+Bakke (2011) summarizes four explanations for the persistence of unrecognized states, each of which plays a role in the theory we develop here.
+First, economic and political elites benefit from the status quo (e.g.
 King 2001).
@@
 Kolst\o \ 2006).
-Third, secessionists often enjoy external support from a patron (e.g.
+Third, secessionists enjoy external support from a patron (e.g.
 Lynch 2002; Stanislawski 2008.
-Lastly, the governments of unrecognized states are often internally legitimate, enjoying support from residents of the unrecognized state (e.g.
+Lastly, the governments of unrecognized states are internally legitimate, enjoying support from residents of the unrecognized state (e.g.
 Lynch 2002, Kolst\o \ 2006).

### 2015-10-15 to_do_list.tex
Updated Ben's to do list

--- 2015-04-09 to_do_list.tex
@@
 
+Ben
+\begin{itemize}
+	\item R1: Finish \& Double Check Fearon reference [DONE]
+		\item add appropriate language to memo if necessary [DONE]
+
+	\item R1: Talk to Kristy about whether we want to offer a numerical example a la what R1 requests.
+
+	\item R1: Tone down our claims about being novel and unique [DONE]
+
+	\item R1: Double check that we have fixed the typo identified by R1 [DONE]
+
+	\item R3: Read the literature referenced by R3 and add some cites [DONE]
+
+	\item R3: Read Berg 2009 in particular and engage it at least briefly in the paper. [DONE]
+
+\end{itemize}
+
 \end{document}

### 2015-10-16 eqm_profiles.tex
start reasoning about relevant regions of state variables
minor clarifications to eqm_profiles.tex

@@
 
-Strategies
+\vskip.2in
+Strategies (this is for the most interesting state variable, but could write it out for the other five)
 \begin{itemize}
@@
 	\item Strategies are a function of all six state variables
+		\begin{itemize}
+			\item We can restrict attention to relevant ranges of the state variables.
+CAN I CHANGE PROP 1 PARTS 1 AND 2 TO BE INITIAL VALUES OF STATE VARIABLES?
+If so, then these conditions and what is necessary in each period to make the game continue.
+		\end{itemize}
 	\item Strategies for the patron and $c$ are also how much to invest in each of the six state variables. Some can be ruled out by preference assumptions:
@@
 \vskip.5in
-Note that patron will never invest to augment $l_s$ or $w_g$. This leave four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$.
+This leaves four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$. Three in which $c$ may invest: $q_s$, $q_g$, $l_s$.
 \begin{itemize}
@@
 
+\vskip.2in
 Period 1

### 2015-10-18 eqm_profiles.tex
rewrite case 1 of eqm profile

@@
 \vskip.2in
+Assumptions
+\begin{itemize}
+	\item $Q_{s0} \geq L_{s0} \Rightarrow q_{s1} + \mu \geq l_{s1}$
+	\item $Q_{g0} \geq L_{g0} \Rightarrow q_{g1} \geq l_{g1}$
+\end{itemize}
+
+\vskip.2in
 Period 1
 \begin{enumerate}
-	\item Patron: $R_{p1} = \frac{\beta}{1-\de} - \left(q_{s1} - l_{s1}\right)$ to augment $q_{s1}$ if this is greater than 0. Else, $R_{p1} = 0$. Can write in max language.
+	\item Patron would like $Q_{s1} \geq L_{s1}$ to prevent $c$ from incentivizing secessionists to Cede
 		\begin{itemize}
-			\item International community: 
+			\item Equivalent to $q_{s1} + R_{p1} \geq l_{s1} + R_{c1}$ / $Q_{s0} + \mu + R_{p1} \geq L_{s0} + R_{c1}$ / $R_{p1} \geq R_{c1} - \left( Q_{s0} - \mu - L_{s0}\right)$
+				\begin{itemize}
+					\item Equivalent to $R_{c1} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{p1} \geq \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ allows the patron to ensure the original inequality
+					\item Initial assumption means $R_{p1} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
+					\item Since patron is willing to pay up to $R_{p1} = \frac{\alpha}{1-\de}$, when $\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$, it will invest $R_{p1} = \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ to augment $q_{s1}$ if this is greater than 0.
+Else, $R_{p1} = 0$.
 					  \begin{enumerate}
-					\item If $R_{p1} + q_{s1} - l_{s1} \geq \frac{\beta}{1-\de}$, $R_{c1} =0$
-					\item Otherwise $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right) + \ve$ to augment $l_{s1}$
+							\item If $R_{p1} + \left( Q_{s0} - \mu - L_{s0}\right) \geq \frac{\beta}{1-\de}$, $R_{c1} =0$. (SQ,SQ) is played and game continues.
+							\item If instead $R_{p1} + \left( Q_{s0} - \mu - L_{s0}\right) < \frac{\beta}{1-\de}$ (i.e.
+assumption doesn't hold), $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right) + \ve$ to augment $l_{s1}$.
+Note optimal $R_{p1}$ in this case is 0.
 								\begin{itemize}
@@
 						\end{enumerate}
+				\end{itemize}
 		\end{itemize}	

### 2015-10-18 eqm_profiles.tex
rewrite case 2 of eqm profile

@@
 				\begin{itemize}
-					\item Equivalent to $R_{c1} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{p1} \geq \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ allows the patron to ensure the original inequality
+					\item $R_{c1} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{p1} \geq \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ allows the patron to ensure the original inequality
 					\item Initial assumption means $R_{p1} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
@@
 		\end{itemize}	
-	\item Patron might also want to invest to encourage recognition directly. How much to augment $l_{g1}$?	
+	\item Patron would like $L_{g1} > Q_{g1}$ to achieve recognition directly. Must invest to augment $l_{g1}$
+		\begin{itemize}
+			\item Equivalent to $l_{g1} + R_{p1} > q_{g1} + R_{c1}$ / $L_{g0} + R_{p1} > Q_{g0} + R_{c1}$ / $R_{p1} > R_{c1} + \left( Q_{g0} - L_{g0}\right)$
+				\begin{itemize}
+					\item $R_{c1} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{p1} > \frac{\nu}{1-\de} + \left( Q_{g0} - L_{g0}\right)$ allows the patron to ensure the original inequality
+					\item Since patron is willing to pay up to $R_{p1} = \frac{\lambda +\mu + \beta}{1-\de}$, when $\frac{\lambda +\mu + \beta}{1-\de} \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{g1}$.
+$R_{c1} = 0$ as well.
+					\item If instead assumption 4 did not hold and $R_{p1} > \frac{\nu}{1-\de} + \left( Q_{g0} - L_{g0}\right)$, it will invest this amount to augment $l_{g1}$.
+Again, $R_{c1} = 0$.
+(Cede, SQ) is played and game ends.
+				\end{itemize}
+		\end{itemize}
 	\item Patron might also want to instigate fighting.

### 2015-10-18 eqm_profiles.tex
write case 3 of eqm profile; will have to re-write proof of lemma 4

@@
 		\end{itemize}
-	\item Patron might also want to instigate fighting.
+	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
 If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}$ is positive, invest in $w_{s1}$ to change $\omega_{s1}$.
-Will not invest in $w_{s1}$ when the quantity is non-positive.
+Will not invest when the quantity is non-positive.
+Inequality is $\omega_{s1} > Q_{s1}$
 		\begin{itemize}
-			\item $c$ will counter up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$
+			\item $\omega_{s1} = -\zeta_{s}(1-\de) + W_{s1}p_{1s} + L_{s1}(1-p_{1s}) > Q_{s1}$
+			\item $-\zeta_{s}(1-\de) + (W_{s0}+R_{p1})p_{1s} + L_{s0}(1-p_{1s}) > Q_{s0} - \mu + R_{c1}$
+			\item $R_{p1}p_{1s} > R_{c1} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)$
+			\item $R_{c1} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{p1}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)$, $R_{c1}=0$ and (SQ,Fight) is played in third stage.
+Outcome depends on war lottery.
+			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de}$
 		\end{itemize}

### 2015-10-20 eqm_profiles.tex
finish case 3
replicate period 1 profiles for period t

@@
 Outcome depends on war lottery.
-			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de}$
+			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
+				\begin{itemize}
+					\item So need
+						\begin{multline}
+							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)
+						\end{multline}
+				\end{itemize}
+		\end{itemize}
+	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
+But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
+\end{enumerate}
+
+\vskip.2in
+Now generalize profiles for all $t$
+\begin{enumerate}
+	\item Patron would like $Q_{st} \geq L_{st}$ to prevent $c$ from incentivizing secessionists to Cede
+		\begin{itemize}
+			\item Equivalent to $q_{st} + R_{pt} \geq l_{st} + R_{ct}$ / $Q_{s,t-1} + \mu + R_{pt} \geq L_{s,t-1} + R_{ct}$ / $R_{pt} \geq R_{ct} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$
+				\begin{itemize}
+					\item $R_{ct} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{pt} \geq \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ allows the patron to ensure the original inequality
+					\item Assumption 1 combined with equilibrium play (equivalently, if $L_{s,t-1} > Q_{s,t-1}$, the game would have ended before the start of period $t$) means $R_{pt} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
+					\item Since patron is willing to pay up to $R_{pt} = \frac{\alpha}{1-\de}$, when $\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$, it will invest $R_{pt} = \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ to augment $q_{st}$ if this is greater than 0.
+Else, $R_{pt} = 0$.
+					  \begin{enumerate}
+							\item If $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) \geq \frac{\beta}{1-\de}$, $R_{ct} =0$. (SQ,SQ) is played and game continues.
+							\item If instead $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) < \frac{\beta}{1-\de}$ (i.e.
+assumption 3 doesn't hold), $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$ where $\ve$ is small to augment $l_{st}$.
+Note optimal $R_{pt}$ in this case is 0.
+								\begin{itemize}
+									\item (SQ,Cede) is played and game ends
+								\end{itemize}
+						\end{enumerate}
+				\end{itemize}
+		\end{itemize}	
+	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$
+		\begin{itemize}
+			\item Equivalent to $l_{gt} + R_{pt} > q_{gt} + R_{ct}$ / $L_{g,t-1} + R_{pt} > Q_{g,t-1} + R_{ct}$ / $R_{pt} > R_{ct} + \left( Q_{g,t-1} - L_{g,t-1}\right)$
+				\begin{itemize}
+					\item $R_{ct} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ allows the patron to ensure the original inequality
+					\item Since patron is willing to pay up to $R_{pt} = \frac{\lambda +\mu + \beta}{1-\de} - \mu$, when $\frac{\lambda +\mu + \beta}{1-\de} -\mu \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{gt}$.
+$R_{ct} = 0$ as well.
+					\item If instead assumption 4 did not hold and $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, patron will invest this amount to augment $l_{gt}$.
+Again, $R_{ct} = 0$.
+(Cede, SQ) is played and game ends.
+				\end{itemize}
+		\end{itemize}
+	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
+If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$ is positive, invest in $w_{st}$ to change $\omega_{st}$.
+Will not invest when the quantity is non-positive.
+Inequality is $\omega_{st} > Q_{st}$
+		\begin{itemize}
+			\item $\omega_{st} = -\zeta_{s}(1-\de) + W_{st}p_{1s} + L_{st}(1-p_{1s}) > Q_{st}$
+			\item $-\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) > Q_{s,t-1} - \mu + R_{ct}$
+			\item $R_{pt}p_{1s} > R_{ct} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$
+			\item $R_{ct} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$, $R_{ct}=0$ and (SQ,Fight) is played in third stage.
+Outcome depends on war lottery.
+			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
+				\begin{itemize}
+					\item So need
+						\begin{multline}
+							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)
+						\end{multline}
+				\end{itemize}
 		\end{itemize}

### 2015-10-20 draft.tex
update assumptions in prop 1 in draft

@@
 \begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_{it} \geq L_{it} \ \forall t$: in each period, remaining in the status quo is better than ceding.}
-
-\item \emph{For both players $g$ and $s$, $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_1) + L_{it}(1-p_1)}{1-\delta} \ \forall n$: in each period, the expected outcome under war is worse than the status quo.}
-
-\item \textit{$\alpha \geq \beta$: reunification is more important for the patron to avoid than for the international community to achieve.}
+\item \textit{For both players $g$ and $s$, $Q_{i0} \geq L_{i0}$: remaining in the status quo is better than ceding at the beginning of the game.}
+
+\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_1) + L_{i0}(1-p_1)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+
+\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}
 
@@
 
-\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(Q_{s0} -\mu - L_{s0} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
 

### 2015-10-21 draft.tex
Shortened uncertainty section and merged with outside interactions.

@@
 \label{sec:ext}
-\subsection{Introducing Uncertainty}
-The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$.
+%\subsection{Introducing Uncertainty}
+%The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$.
 Payoffs as described in the stage game are those perceived by $g$ and $s$.
@@
 
-Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs.
+%Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs.
 After viewing these (uncertain) payoffs, $c$ and $p$ each invest some level of resources, altering the payoff structure before it is observed (accurately) by $s$ and $g$.
@@
 
-In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups.
+%In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups.
 This makes over-contribution by $c$ more likely than under-contribution by the patron.
@@
 
-In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources.
+%In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources.
 By adding some uncertainty about payoffs, we can account for the observed fact that $c$ sometimes expends resources unsuccessfully.
@@
 
-\subsection{Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
-The game that we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible.
+\subsection{Uncertainty and Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
+While we present a model with perfect information, there are some empirical cases in which it appears that various players have acted on mistaken beliefs about the strength or payoffs of other players.
+Modeling uncertainty falls outside the scope of this paper, but uncertainty does offer an additional plausible explanation for off-equilibrium-path behavior that we observe in empirical cases.
+For example, uncertainty regarding the payoff structure facing the secessionists may explain failed and costly efforts by the European Union and its members (i.e.
+player $c$) to promote a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.
+
+The game we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible.
 We do not model any direct exchange of resources or imposition of harm between $c$ and the patron, but we take the implications of these possible outside interactions seriously.
@@
 
-\section{Conclusions}
+\section{Conclusion}
 In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives.
 Current events, particularly, Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon.
-While the Ukrainian situation remains fluid, the status quo equilibrium we model is a plausible future.
+While the Ukrainian situation remains fluid, as of October 2015, the status quo equilibrium we model is a plausible future.
 The equilibrium could emerge as follows: with Russian support, the secessionists solidify \emph{de facto} control of parts of Eastern Ukraine, but risk of direct confrontation with the West is sufficiently high to prevent Russian annexation of Eastern Ukraine.\footnote{Annexation of Crimea was achieved at a reasonable cost to Russia.

### 2015-10-21 draft.tex
Update to-do list. Refer to 'dynamic four-player model' in the text.

@@
 Their persistence is both intellectually puzzling and normatively problematic, but unrecognized statehood can be a remarkably stable outcome, persisting for decades.
-Our four-player model reveals that unrecognized statehood emerges as an equilibrium outcome when a patron state is willing and able to persistently invest resources to sustain it.
+Our dynamic four-player model reveals that unrecognized statehood emerges as an equilibrium outcome when a patron state is willing and able to persistently invest resources to sustain it.
 %Counterintuitively, patrons will make these investments even when unrecognized statehood is not their most preferred outcome.
@@
 
-We develop a four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome.
+We develop a dynamic four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome.
 These outside actors are a patron, which prefers independence to reunification, and a fourth player we denote as player $c$ who represents one or more members of the international community.
@@
 While the importance of outside actors in civil conflict has been widely acknowledged in the empirical literature, it is rarely modeled formally.
-We introduce a four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome.
+We introduce a dynamic four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome.
 This allows us both to examine the means through which patron states sustain unrecognized statehood as a stable equilibrium, and to rigorously analyze the strategies available to other outside actors to pursue reunification.

### 2015-10-31 draft.tex
differentiate between inside players' winning probabilities in prop 1 assumptions

@@
 
-\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_1) + L_{i0}(1-p_1)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
 

### 2015-10-31 draft.tex
modify lemma 1 and proof for change to initial conditions

@@
 \begin{quotation}
-\noindent \textit{{\bf Lemma 1:} Conditions 1 and 2 are sufficient for status quo to be the outcome of Stage 3 in any period.} 
+\noindent \textit{{\bf Lemma 1:} $Q_{it} \geq L_{it}$ and $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_{1i}) + L_{it}(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for status quo to be the outcome of Stage 3 in period $t \ \forall t$.}
 \end{quotation}
@@
 The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$.
-Thus the continuation value is $\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.
-
-In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2).
-The argument for player $g$ is symmetric.
-\hfill $\blacksquare$
+Thus the continuation value is $-\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.
+
+Given that the continuation value from Status Quo is higher than both that from playing Cede and Fight, player $s$ will play Status Quo. The argument for player $g$ is symmetric. \hfill $\blacksquare$
 \\

### 2015-10-31 eqm_profiles.tex
copy outline of general strategy profiles to turn into prose
turn eqm path strategy part of strategy profile into prose

@@
 
+\newpage
+Prose version of equilibrium strategy profiles: \\
+
+In period $t$, the patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment the status quo payoffs of the secessionists, $q_{st}$.
+Player $c$ then chooses $R_{ct} =0$ so that playing status quo is a best response for both inside players.
+(SQ,SQ) is played in the third stage and the game continues to period $t+1$.
+
+\begin{enumerate}
+							\item If instead $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) < \frac{\beta}{1-\de}$ (i.e.
+assumption 3 doesn't hold), $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$ where $\ve$ is small to augment $l_{st}$.
+Note optimal $R_{pt}$ in this case is 0.
+								\begin{itemize}
+									\item (SQ,Cede) is played and game ends
+								\end{itemize}
+						\end{enumerate}
+
+\begin{enumerate}
+	\item Patron would like $Q_{st} \geq L_{st}$ to prevent $c$ from incentivizing secessionists to Cede
+
+				\begin{itemize}
+					\item Assumption 1 combined with equilibrium play (equivalently, if $L_{s,t-1} > Q_{s,t-1}$, the game would have ended before the start of period $t$) means $R_{pt} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
+				\end{itemize}
+
+	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$
+		\begin{itemize}
+			\item Equivalent to $l_{gt} + R_{pt} > q_{gt} + R_{ct}$ / $L_{g,t-1} + R_{pt} > Q_{g,t-1} + R_{ct}$ / $R_{pt} > R_{ct} + \left( Q_{g,t-1} - L_{g,t-1}\right)$
+				\begin{itemize}
+					\item $R_{ct} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ allows the patron to ensure the original inequality
+					\item Since patron is willing to pay up to $R_{pt} = \frac{\lambda +\mu + \beta}{1-\de} - \mu$, when $\frac{\lambda +\mu + \beta}{1-\de} -\mu \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{gt}$.
+$R_{ct} = 0$ as well.
+					\item If instead assumption 4 did not hold and $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, patron will invest this amount to augment $l_{gt}$.
+Again, $R_{ct} = 0$.
+(Cede, SQ) is played and game ends.
+				\end{itemize}
+		\end{itemize}
+	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
+If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$ is positive, invest in $w_{st}$ to change $\omega_{st}$.
+Will not invest when the quantity is non-positive.
+Inequality is $\omega_{st} > Q_{st}$
+		\begin{itemize}
+			\item $\omega_{st} = -\zeta_{s}(1-\de) + W_{st}p_{1s} + L_{st}(1-p_{1s}) > Q_{st}$
+			\item $-\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) > Q_{s,t-1} - \mu + R_{ct}$
+			\item $R_{pt}p_{1s} > R_{ct} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$
+			\item $R_{ct} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$, $R_{ct}=0$ and (SQ,Fight) is played in third stage.
+Outcome depends on war lottery.
+			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
+				\begin{itemize}
+					\item So need
+						\begin{multline}
+							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)
+						\end{multline}
+				\end{itemize}
+		\end{itemize}
+	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
+But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
+\end{enumerate}
+
 \end{document}

### 2015-10-31 eqm_profiles.tex
finish turning eqm path strategy profiles into prose

@@
 
-\begin{enumerate}
-							\item If instead $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) < \frac{\beta}{1-\de}$ (i.e.
-assumption 3 doesn't hold), $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$ where $\ve$ is small to augment $l_{st}$.
-Note optimal $R_{pt}$ in this case is 0.
-								\begin{itemize}
-									\item (SQ,Cede) is played and game ends
-								\end{itemize}
-						\end{enumerate}
+If instead $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ (i.e.
+assumption 3 doesn't hold), player $c$ will invest $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, to augment $l_{st}$.
+In this case, status quo remains a best response for the government but cede becomes the best response for the secessionists.
+The game ends with the secessionist territory being reunited.
 
 \begin{enumerate}	
-	\item Patron would like $Q_{st} \geq L_{st}$ to prevent $c$ from incentivizing secessionists to Cede
-
-				\begin{itemize}
-					\item Assumption 1 combined with equilibrium play (equivalently, if $L_{s,t-1} > Q_{s,t-1}$, the game would have ended before the start of period $t$) means $R_{pt} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
-				\end{itemize}
-
 	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$

### 2015-10-31 eqm_profiles.tex
turn lemma 3 strategy profile part into prose

@@
 
+On the equilibrium path, the patron invests to avoid the reunification outcome.
+That is, it invests in the secessionists' status quo payoffs to deter the international community from investing in the secessionists' payoffs from ceding.
 In period $t$, the patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment the status quo payoffs of the secessionists, $q_{st}$.
@@
 The game ends with the secessionist territory being reunited.
+This path is ruled out by Assumption 3 (see Lemma 2).
+
+The patron would prefer to invest to provoke the recognition outcome, which it can attempt to do directly by investing in the government's payoffs from ceding, $l_{gt}$.
+If the patron invests $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ will then choose $R_{ct} = 0$ and cede will be the best response for the government.
+Status quo remains the best response for the secessionists, so the game ends with the secessionists gaining recognition.
+This path is ruled out by Assumption 4 (see Lemma 3).
+
+As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, player $c$ will counter with $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$.
+This makes status quo a best response for the government as well as for the secessionists so that the status quo equilibrium is played and the game continues to period $t+1$.
+Note that the optimal investments in the government's payoffs zero.
 
 \begin{enumerate}		
-	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$
-		\begin{itemize}
-			\item Equivalent to $l_{gt} + R_{pt} > q_{gt} + R_{ct}$ / $L_{g,t-1} + R_{pt} > Q_{g,t-1} + R_{ct}$ / $R_{pt} > R_{ct} + \left( Q_{g,t-1} - L_{g,t-1}\right)$
-				\begin{itemize}
-					\item $R_{ct} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ allows the patron to ensure the original inequality
-					\item Since patron is willing to pay up to $R_{pt} = \frac{\lambda +\mu + \beta}{1-\de} - \mu$, when $\frac{\lambda +\mu + \beta}{1-\de} -\mu \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{gt}$.
-$R_{ct} = 0$ as well.
-					\item If instead assumption 4 did not hold and $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, patron will invest this amount to augment $l_{gt}$.
-Again, $R_{ct} = 0$.
-(Cede, SQ) is played and game ends.
-				\end{itemize}
-		\end{itemize}
 	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.

### 2015-11-01 eqm_profiles.tex
finish writing out strategy profiles into prose

@@
 
-\begin{enumerate}		
-	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
-If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$ is positive, invest in $w_{st}$ to change $\omega_{st}$.
-Will not invest when the quantity is non-positive.
-Inequality is $\omega_{st} > Q_{st}$
-		\begin{itemize}
-			\item $\omega_{st} = -\zeta_{s}(1-\de) + W_{st}p_{1s} + L_{st}(1-p_{1s}) > Q_{st}$
-			\item $-\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) > Q_{s,t-1} - \mu + R_{ct}$
-			\item $R_{pt}p_{1s} > R_{ct} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$
-			\item $R_{ct} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$, $R_{ct}=0$ and (SQ,Fight) is played in third stage.
-Outcome depends on war lottery.
-			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
-				\begin{itemize}
-					\item So need
-						\begin{multline}
-							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)
-						\end{multline}
-				\end{itemize}
-		\end{itemize}
-	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
-But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
-\end{enumerate}
+Instead of instigating recognition directly by improving the government's payoffs from ceding, the patron might want to instigate fighting, which would lead to recognition with probability $p_{1s}$.
+If the patron invests $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ to augment $w_{s,t}$, player $c$ will choose $R_{ct}=0$ and Fight becomes a best response for the secessionists so (SQ,Fight) is played in third stage.
+The game will end with an outcome that depends on the war lottery.
+This path is ruled out by a combination of Assumptions 3 and 4 (see Lemma 4).
+
+If the patron spends less than this amount, player $c$ will counter by investing $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$, the secessionists' status quo payoffs.
+This makes playing Status Quo a best response for both players so that the status quo equilibrium is played and the game continues to period $t+1$.
+Here again the optimal investment by both parties is zero.\footnote{Note that the patron could also invest in $q_{gt}$ to counter investment by player $c$ in the government's payoffs from war.
+But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.}
 

### 2015-11-05 to_do_list.tex
Update to-do list. Refer to 'dynamic four-player model' in the text.
revise to-do list for final big re-write of main section of draft

@@
 \begin{itemize}
-	\item Think through reviewer 2's suggestion about uncertainty
+	\item Recast Proposition 1 as a ``class of games'' instead of ``sufficient conditions'' (Reviewer 1)
 		\begin{itemize}
-			\item How would model actually work with asymmetric information?
-		\end{itemize}
-	\item Take indecisive war to decisive war
-	\item Change sign on payoff functions for $c$ and $p$
-	\item Homogenize discount factors
-	\item Decide on Reviewer 1's suggestion to recast Proposition 1 as a ``class of games'' instead of ``sufficient conditions''
-		\begin{itemize}
-			\item ``I understand why the set of sufficient conditions in this proposition is the key result.
-However, these mechanisms are essentially just assumptions on payoffs; and therefore I would be inclined to interpret these conditions as characterizations of “a class” of games, where this class would satisfy all the assumptions on payoffs (instead of saying sufficient conditions).
-\\
-				I would recommend clearing up Section 3.1 in terms of packaging the results: 
+			\item ``interpret these conditions as characterizations of “a class” of games, where this class would satisfy all the assumptions on payoffs (instead of saying sufficient conditions).
+			\item I would recommend clearing up Section 3.1 in terms of packaging the results: 
 					\begin{enumerate}
@@
 		\end{itemize}
-	\item State process governing status-quo payoffs clearly
-		\begin{itemize}
-			\item ``If I understand correctly, expenditure R by the patron is essentially a long-term investment that raises the status-quo payoff of the secessionist elite by R in every future period.
-I suggest stating it this way: The per-period status quo payoff is a state variable; its transition from one period to the next involves an automatic reduction by mu (for the reasons you discuss) and an increase by the amount of the patron's investment.
-		\end{itemize}
 	\item Make sure equilibrium concept is clearly stated and defined
@@
 		\end{itemize}
+	\item Think through reviewer 2's suggestion about uncertainty [DONE]
+		\begin{itemize}
+			\item How would model actually work with asymmetric information? [NOT GOING TO DO]
+		\end{itemize}
+	\item Take indecisive war to decisive war [DONE]
+	\item Change sign on payoff functions for $c$ and $p$ [DONE]
+	\item Homogenize discount factors [DONE]
+	\item State process governing status-quo payoffs clearly [DONE]
 \end{itemize}
@@
 
-	\item R1: Talk to Kristy about whether we want to offer a numerical example a la what R1 requests.
+	\item R1: Talk to Kristy about whether we want to offer a numerical example a la what R1 requests. [DONE]
 

### 2015-11-06 to_do_list.tex
simplify to-do list

@@
 					\begin{enumerate}
-						\item First, I would identify a class of games that satisfy the assumptions on payoffs (sufficient conditions (1) through (6) in Proposition 1). Denote this class as G for example. 
-						\item Then, formally define the “status quo” equilibrium – an equilibrium in which the equilibrium outcome is perpetual unrecognized statehood with all the actors’ strategies specified. 
-						\item Finally, I would state a much clearer and simpler proposition, e.g., For any game G, there exists a status quo equilibrium.
-...
-Then subsection 3.2 would basically be a discussion of this class of games.
+						\item Identify a class of games that satisfies the assumptions on payoffs (sufficient conditions (1) through (6) in Proposition 1). Denote this class as G for example. 
+						\item Formally define the “status quo” equilibrium -- an equilibrium in which the equilibrium outcome is perpetual unrecognized statehood with all the actors’ strategies specified. 
+						\item State a much clearer and simpler proposition, e.g., For any game G, there exists a status quo equilibrium.
+						\item Then subsection 3.2 would basically be a discussion of this class of games. 
 					\end{enumerate}
 		\end{itemize}
-	\item Clearly specify punishments / equilibrium path to satisfy editor
+	\item Clearly specify punishments / equilibrium path (editor)
 		\begin{itemize}
 			\item ``Clear and detailed description of the noncooperative game''
-			\item ``I would also like to see the key logic laid out better in the text, and the full
-equilibrium strategy profiles noted so that it is clear how the players are rewarding and punishing each other on and off the equilibrium path''
+			\item ``I would also like to see the key logic laid out better in the text, and the full equilibrium strategy profiles noted so that it is clear how the players are rewarding and punishing each other on and off the equilibrium path''
 		\end{itemize}

### 2015-11-06 draft.tex
define class of games that satisfy parameter restrictions

@@
 
+\newtheorem{definition}{Definition}
 \newcommand{\ta}{\theta}
@@
 
+We begin by listing and providing intuition for a set of restrictions on the preferences of the actors and their resources for which we can guarantee that unrecognized statehood is an equilibrium outcome of the repeated game.
+This set of restrictions identifies a class of games, $\mathcal{G}$:
+
+\begin{definition}
+\emph{The class $\mathcal{G}$ includes all those games for which the parameters satisfy the following restrictions:}
+
+\begin{enumerate}
+\item \textit{For both players $g$ and $s$, $Q_{i0} \geq L_{i0}$: remaining in the status quo is better than ceding at the beginning of the game.}
+
+\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+
+\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}
+
+\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
+
+\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(Q_{s0} -\mu - L_{s0} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+
+\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
+If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}
+
+\end{enumerate}
+\end{definition}
+
 \begin{quotation}
@@
 The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\de}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}
-
-\emph{The following are sufficient conditions for such an equilibrium: }
-
-\begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_{i0} \geq L_{i0}$: remaining in the status quo is better than ceding at the beginning of the game.}
-
-\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
-
-\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}
-
-\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
-
-\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(Q_{s0} -\mu - L_{s0} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
-
-\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
-If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}
-
-\end{enumerate}
 

### 2015-11-06 draft.tex
paste strategy profiles into eqm defn in draft

@@
 \newcommand{\de}{\delta}
+\newcommand{\ve}{\varepsilon}
 \linespread{1.05} % Palatino needs more leading (space between lines)
@@
 
+\begin{definition}
+ The Status Quo Equilibrium is a stationary Markov equilibrium in which the outcome is perpetual unrecognized statehood.
+
+On the equilibrium path, the patron invests to avoid the reunification outcome.
+That is, it invests in the secessionists' status quo payoffs to deter the international community from investing in the secessionists' payoffs from ceding.
+In period $t$, the patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment the status quo payoffs of the secessionists, $q_{st}$.
+Player $c$ then chooses $R_{ct} =0$ so that playing status quo is a best response for both inside players.
+(SQ,SQ) is played in the third stage and the game continues to period $t+1$.
+
+If instead $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ (i.e.
+assumption 3 doesn't hold), player $c$ will invest $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, to augment $l_{st}$.
+In this case, status quo remains a best response for the government but cede becomes the best response for the secessionists.
+The game ends with the secessionist territory being reunited.
+This path is ruled out by Assumption 3 (see Lemma 2).
+
+The patron would prefer to invest to provoke the recognition outcome, which it can attempt to do directly by investing in the government's payoffs from ceding, $l_{gt}$.
+If the patron invests $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ will then choose $R_{ct} = 0$ and cede will be the best response for the government.
+Status quo remains the best response for the secessionists, so the game ends with the secessionists gaining recognition.
+This path is ruled out by Assumption 4 (see Lemma 3).
+
+As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, player $c$ will counter with $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$.
+This makes status quo a best response for the government as well as for the secessionists so that the status quo equilibrium is played and the game continues to period $t+1$.
+Note that the optimal investments in the government's payoffs zero.
+
+Instead of instigating recognition directly by improving the government's payoffs from ceding, the patron might want to instigate fighting, which would lead to recognition with probability $p_{1s}$.
+If the patron invests $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ to augment $w_{s,t}$, player $c$ will choose $R_{ct}=0$ and Fight becomes a best response for the secessionists so (SQ,Fight) is played in third stage.
+The game will end with an outcome that depends on the war lottery.
+This path is ruled out by a combination of Assumptions 3 and 4 (see Lemma 4).
+
+If the patron spends less than this amount, player $c$ will counter by investing $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$, the secessionists' status quo payoffs.
+This makes playing Status Quo a best response for both players so that the status quo equilibrium is played and the game continues to period $t+1$.
+Here again the optimal investment by both parties is zero.\footnote{Note that the patron could also invest in $q_{gt}$ to counter investment by player $c$ in the government's payoffs from war.
+But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.}
+\end{definition}
+
 \begin{quotation}

### 2015-11-07 draft.tex
add defn of stationary markov eqm to draft

@@
 
-We provide a set of conditions under which perpetual unrecognized statehood is the outcome of a subgame-perfect equilibrium of this game.
+We provide a set of conditions under which perpetual unrecognized statehood is the outcome of a stationary Markov equilibrium of this game.
 %This requires that the patron is both willing and able to make investments sufficient to (a) offset the deterioration in the secessionists' status quo payoffs and (b) deter the international community from making investments that would induce a negotiated settlement.
@@
 
+Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006).
+That is, a strategy profile is a stationary Markov equilibrium if it is a stationary Markov strategy profile and a subgame-perfect equilibrium.
+In turn, a strategy profile is defined to be a stationary Markov strategy if any two ex post histories terminating in the same state play identically from the termination point forward.
+That is, stationary Markov strategies ignore all details of the history aside from the current state.
+
 \begin{definition}
@@
 
-There are many other potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
+\textbf{There are many other potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
 As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity.
-We will use the solution concept of subgame perfect equilibrium to show that, given the sufficient conditions in Proposition 1, at least one status quo equilibrium will exist.\\
+We will use the solution concept of stationary Markov equilibrium to show that, given the sufficient conditions in Proposition 1, at least one status quo equilibrium will exist.}\\
 
@@
 
+\item \hangindent=1cm Mailath, George J., and Larry Samuelson. 2006. "Repeated Games and Reputation." New York, NY: Oxford University Press.
+
 \item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, settlement, and the shadow of the future." \emph{Journal of Economic Behavior $\&$ Organization} 105: 75-89.

### 2015-11-07 draft.tex
add notes on state variables, rearrange text around MPE discussion

@@
 
-We begin by listing and providing intuition for a set of restrictions on the preferences of the actors and their resources for which we can guarantee that unrecognized statehood is an equilibrium outcome of the repeated game.
+We begin by listing and providing intuition for a set of restrictions on the preferences of the actors and their resources for which we can guarantee that unrecognized statehood is an equilibrium outcome.
 This set of restrictions identifies a class of games, $\mathcal{G}$:
@@
 \end{definition}
+
+There are many potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
+As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity.
+We will show that, given a game satisfying the parameter restrictions in Definition 1, at least one status quo equilibrium will exist.
 
@@
 
+\vskip.5in
+We have six state variables, so $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
+\begin{itemize}
+	\item Strategies for the patron and $c$ are also how much to invest in each of the six state variables. Some can be ruled out by preference assumptions:
+		\begin{itemize}
+			\item $c$ dislikes war, so will never invest in $w_s$ or $w_g$. It would also not want to make the government lose, so won't invest in $l_g$ either.
+			\item Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
+		\end{itemize}
+	\item This leaves four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$. Three in which $c$ may invest: $q_s$, $q_g$, $l_s$.
+\end{itemize}
+
+\vskip1in
 \begin{definition}
@@
 In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
-
-\textbf{There are many other potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
-As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity.
-We will use the solution concept of stationary Markov equilibrium to show that, given the sufficient conditions in Proposition 1, at least one status quo equilibrium will exist.}\\
 

### 2015-11-07 draft.tex
explain which state variables are possible for investment; rewrite defn of SQ eqm

@@
 
-\vskip.5in
-We have six state variables, so $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
-\begin{itemize}
-	\item Strategies for the patron and $c$ are also how much to invest in each of the six state variables. Some can be ruled out by preference assumptions:
-		\begin{itemize}
-			\item $c$ dislikes war, so will never invest in $w_s$ or $w_g$. It would also not want to make the government lose, so won't invest in $l_g$ either.
-			\item Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
-		\end{itemize}
-	\item This leaves four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$. Three in which $c$ may invest: $q_s$, $q_g$, $l_s$.
-\end{itemize}
-
-\vskip1in
+The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
+That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
+Strategies for the patron and the international community are how much to invest in each of the six state variables.
+However, positive investments in some of the state variables can be ruled out by preference assumptions.
+
+Player $c$ dislikes war and so will never invest in $w_s$ or $w_g$, which only serve to increase the likelihood that one of the inside actors chooses to fight.
+It would also not want to make the government lose and so won't invest in $l_g$.
+This leaves three state variables in which player $c$ might invest: $q_s$, $q_g$, $l_s$.
+
+Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
+The only reason the patron would invest in $q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $q_g$.
+This leaves four state variables in which the patron may invest: $q_s$, $l_g$ and $w_s$.
+
+In equilibrium, the outside actors' investments are paired: the patron investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
+Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding.
+Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
+In each case, the investments are sufficient to make the largest continuation payoffs in the game between the secessionists and the government those from playing status quo.
+
 \begin{definition}
 The Status Quo Equilibrium is a stationary Markov equilibrium in which the outcome is perpetual unrecognized statehood.
-
-On the equilibrium path, the patron invests to avoid the reunification outcome.
-That is, it invests in the secessionists' status quo payoffs to deter the international community from investing in the secessionists' payoffs from ceding.
-In period $t$, the patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment the status quo payoffs of the secessionists, $q_{st}$.
-Player $c$ then chooses $R_{ct} =0$ so that playing status quo is a best response for both inside players.
-(SQ,SQ) is played in the third stage and the game continues to period $t+1$.
-
-If instead $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ (i.e.
-assumption 3 doesn't hold), player $c$ will invest $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, to augment $l_{st}$.
-In this case, status quo remains a best response for the government but cede becomes the best response for the secessionists.
-The game ends with the secessionist territory being reunited.
-This path is ruled out by Assumption 3 (see Lemma 2).
-
-The patron would prefer to invest to provoke the recognition outcome, which it can attempt to do directly by investing in the government's payoffs from ceding, $l_{gt}$.
-If the patron invests $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ will then choose $R_{ct} = 0$ and cede will be the best response for the government.
-Status quo remains the best response for the secessionists, so the game ends with the secessionists gaining recognition.
-This path is ruled out by Assumption 4 (see Lemma 3).
-
-As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, player $c$ will counter with $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$.
-This makes status quo a best response for the government as well as for the secessionists so that the status quo equilibrium is played and the game continues to period $t+1$.
-Note that the optimal investments in the government's payoffs zero.
-
-Instead of instigating recognition directly by improving the government's payoffs from ceding, the patron might want to instigate fighting, which would lead to recognition with probability $p_{1s}$.
-If the patron invests $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ to augment $w_{s,t}$, player $c$ will choose $R_{ct}=0$ and Fight becomes a best response for the secessionists so (SQ,Fight) is played in third stage.
-The game will end with an outcome that depends on the war lottery.
-This path is ruled out by a combination of Assumptions 3 and 4 (see Lemma 4).
-
-If the patron spends less than this amount, player $c$ will counter by investing $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$, the secessionists' status quo payoffs.
-This makes playing Status Quo a best response for both players so that the status quo equilibrium is played and the game continues to period $t+1$.
-Here again the optimal investment by both parties is zero.\footnote{Note that the patron could also invest in $q_{gt}$ to counter investment by player $c$ in the government's payoffs from war.
-But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.}
+The strategy profiles are the following, organized according to the patron's investments in each state variable.
+In period $t$:
+
+\begin{enumerate}
+	\item The patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_{st}$. Player $c$ invests $R_{ct} =0$ in $l_{st}$.
+
+If $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, in $l_{st}$.
+Cede becomes the unique best response for the secessionists.
+
+	\item As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_{gt}$.
+Note that the outcome is that investments in $l_{gt}$ and $q_{gt}$ are zero.
+
+If $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = 0$. Cede becomes the unique best response for the government.
+
+	\item As long as $R_{pt} \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_{s,t}$, player $c$ invests $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$.
+Similar to case two, the outcome is that investments in $w_{st}$ and $q_{st}$ are zero.
+
+If $R_{pt} > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_{s,t}$, player $c$ will choose $R_{ct}=0$.
+Fight becomes the unique best response for the secessionists.
+\end{enumerate}
+The government and secessionists play their best responses in the stage game given the continuation values induced by the investments of the outside actors.
+Unless otherwise noted above, playing Status Quo is the best response for both inside actors.
 \end{definition}
+
+\textbf{Need text to give intuition or tell where I'm going to give intuition.}
 

### 2015-11-07 draft.tex
revise def of SQ eqm, minor grammatical revisions

@@
 In turn, a strategy profile is defined to be a stationary Markov strategy if any two ex post histories terminating in the same state play identically from the termination point forward.
-That is, stationary Markov strategies ignore all details of the history aside from the current state.
+In other words, stationary Markov strategies ignore all details of the history aside from the current state.
 
@@
 
-In equilibrium, the outside actors' investments are paired: the patron investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
+In equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
 Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding.
@@
 \begin{definition}
-The Status Quo Equilibrium is a stationary Markov equilibrium in which the outcome is perpetual unrecognized statehood.
-The strategy profiles are the following, organized according to the patron's investments in each state variable.
+A Status Quo Equilibrium is a stationary Markov equilibrium in which the outcome is perpetual unrecognized statehood.
+
+The strategies for the government and secessionists in this equilibrium are to play their best responses in the stage game given the continuation values induced by the investments of the outside actors.
+Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
+The strategies for the insides actors are the the following, organized according to the patron's investments in each state variable.
 In period $t$:
@@
 \end{enumerate}
-The government and secessionists play their best responses in the stage game given the continuation values induced by the investments of the outside actors.
-Unless otherwise noted above, playing Status Quo is the best response for both inside actors.
 \end{definition}

### 2015-11-07 draft.tex
complete description of game in sec:structure

@@
 \subsection{Timing and Structure of Interaction}
+\label{sec:structure}
 
@@
 There are an infinite number of periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
+This is a dynamic game because the fundamentals determining the payoffs of the inside actors are state variables whose values persist from period to period and in which the outside actors can make investments.
 
@@
 Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
-The strategies for the insides actors are the the following, organized according to the patron's investments in each state variable.
+The strategies for the insides actors are the following, organized according to the patron's investments in each state variable.
 In period $t$:

### 2015-11-07 draft.tex
add intuition for strategy profiles and move description of eqm path actions

@@
 
-\textbf{Need text to give intuition or tell where I'm going to give intuition.}
+To understand this equilibrium, notice that there are three possible ways to disturb the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order for the Status Quo Equilibrium to persist, each of these three disturbances must be deterred.
+
+Since the patron moves first, in the Status Quo Equilibrium, the only investment that takes place is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty.
+The international community's willingness to counteract investments by the patron toward the other two disturbances implies that there will be no investments in equilibrium in cases (2) and (3).
+
+If, however, off-path investments are ever made such that Status Quo does not yield the highest continuation value for one of the players, that player will play Cede or Fight and the game will end.
+
+Equilibrium actions are for the patron to maintain the status quo by investing $\mu$ each period once the difference in payoffs to the secessionists from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $t=1$ of up to that amount); for the international community to not invest and for both inside actors to play Status Quo each period.
 
@@
 
-Thus, given Conditions (1)-(6) hold, equilibrium strategies for each period in this status quo equilibrium are:
-\begin{itemize}
-	\item Player $p$ spends more than $c$ is willing to invest to ensure the incentives for the status quo outcome are in place in period $t$ whenever this is affordable.
-That is, $p$ invests $R_{pt} = \max\left\{0,\frac{\beta}{1-\delta} -(q_{st} - l_{st})\right\}\leq B_{pt}$.
-If this inequality is violated, $p$ invests nothing.
-	\item Player $c$ invests $R_{ct} = \frac{\beta}{1-\delta} -(Q_{st} - l_{st})$ if $\frac{\beta}{1-\delta} > (Q_{st} - l_{st})$. Otherwise, it invests nothing. 
-	\item Players $s$ and $g$ play Status Quo as long as it yields the highest continuation value, and play Cede or Fight if continuation values from either exceeds the status quo continuation value. 
-\end{itemize}
-
-Equilibrium actions are for $p$ to maintain the status quo by investing $\mu$ each period once the difference in payoffs to $s$ from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $t=1$ of up to that amount); for $c$ to not invest and for both $g$ and $s$ to play Status Quo each period.
-\hfill $\blacksquare$
-
 %Our status quo equilibrium requires that $p$ invests in every period enough to maintain $F^*$ by offsetting the $\mu$ decline in the secessionists' status quo payoffs.
@@
 By contributing $\mu$ in each period, the patron supplies sufficient resources to the unrecognized state to ensure that the secessionist elite prefers the status quo to surrendering independence.
+\hfill $\blacksquare$
 

### 2015-11-07 draft.tex
restate prop 1; change format of props to match defns

@@
 \newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
 \newcommand{\ta}{\theta}
@@
 
-\begin{quotation}
-\noindent \bf \emph{Proposition 1:} \rm  \emph{There exists an equilibrium in which the outcome is perpetual unrecognized statehood.
-The actions in this equilibrium are for $p$ to invest to create a buffer of $\frac{\beta}{1-\de}$ between the payoffs from ceding and the status quo in the first period and to invest up to $\mu$ each period thereafter to maintain the buffer; for $c$ to pay nothing; and for both $g$ and $s$ to play $Status$ $Quo$ each period.}
-
-\end{quotation}
+\begin{proposition}
+	For any game in the class of games $\mathcal{G}$, there exists at least one Status Quo Equilibrium.
+\end{proposition}
 
@@
 
-\begin{quotation}
-\noindent \bf \emph{Proposition 2:} \rm  \emph{Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
-In order for sanctions to lead to ceding by the secessionists, the following are required:}
+\begin{proposition}
+Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
+In order for sanctions to lead to ceding by the secessionists, the following are required:
 
@@
 
-\item \textit{The secessionists' continuation value from playing Cede must be higher than their continuation value from playing Fight. }
+\item \textit{The secessionists' continuation value from playing Cede must be higher than their continuation value from playing Fight.}
 \end{enumerate}
 
-\end{quotation}
+\end{proposition}
 
@@
 
-\begin{quotation}
-\noindent \bf \emph{Proposition 3:} \rm  \emph{Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
-The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.}
-
-\end{quotation}
+\begin{proposition}
+Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
+The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
+
+\end{proposition}
 

### 2015-11-08 draft.tex
change format of lemmas to match defns

@@
 \newtheorem{proposition}{Proposition}
+\newtheorem{lemma}{Lemma}
 \newcommand{\ta}{\theta}
@@
 
-These conditions for the status quo are sufficient but not necessary.
+The restrictions on the class of games in $\mathcal{G}$ are sufficient but not necessary.
 For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run.
@@
 
+\vskip.1in
 \noindent {\bf Proof of Proposition 1} \\
@@
 
-\begin{quotation}
-\noindent \textit{{\bf Lemma 1:} $Q_{it} \geq L_{it}$ and $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_{1i}) + L_{it}(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for status quo to be the outcome of Stage 3 in period $t \ \forall t$.}
-\end{quotation}
+\begin{lemma}
+$Q_{it} \geq L_{it}$ and $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_{1i}) + L_{it}(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
+\end{lemma}
 
@@
 
-\begin{quotation}
-\noindent \textit{{\bf Lemma 2:} When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.}
-\end{quotation}
+\begin{lemma}
+When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
+\end{lemma}
 
@@
 
-\begin{quotation}
-\noindent \textit{{\bf Lemma 3:} When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.}
-\end{quotation}
+\begin{lemma}
+When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
+\end{lemma}
 
@@
 
-\begin{quotation}
-\noindent \textit{{\bf Lemma 4:} Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.}
-\end{quotation}
+\begin{lemma}
+Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
+\end{lemma}
 

### 2015-11-08 draft.tex
replace 'condition' with 'restriction'

@@
 The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
-Strategies for the patron and the international community are how much to invest in each of the six state variables.
+That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
 However, positive investments in some of the state variables can be ruled out by preference assumptions.
@@
 The restrictions on the class of games in $\mathcal{G}$ are sufficient but not necessary.
-For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Condition (6) need not be met to maintain the status quo in the short run.
+For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run.
 The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
-In cases where condition (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
+In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
 
@@
 
-To begin, note that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} It is not necessary to focus on which investment vehicle is optimal for a government to choose; what is essential is to determine which outcomes---i.e.
-\emph{Recognition, Status Quo}, or \emph{Reunification}---they will target given that they will choose the most efficient way to alter the payoffs of players $s$ and $g$.
-
-%Given their preference orderings, some of these can be ruled out.
-Player $p$ will not invest to increase $s$'s payoffs from ceding ($l_{sn}$) or $g$'s payoffs from war when it could do so for $s$.
-It will also not invest in $q_{gn}$ when it can instead increase $l_{gn}$, leaving the list of rational investment choices for $p$ as $q_{sn},l_{gn}$, and $w_{sn}$.
-Symmetrically, $c$ will not invest in $q_{sn},L_{gn}$, or $w_{sn}$.
-Given the assumption that the international community follows a norm of encouraging peaceful resolution, it also will not invest in $w_{gn}$, leaving only $l_{sn}$ and $q_{gn}$ as rational investment options for player $c$.
+Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$.
+However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2.
+The remainder of the proof treats each set of matched investment decisions in turn.
 
@@
 \begin{lemma}
-When Condition 3 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
+When Restriction (3) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
 \end{lemma}
 
-\noindent Condition 3 provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
+\noindent Restriction (3) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
 To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}
@@
 Consider period 1.
-When Condition (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
+When Restriction (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
 That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
-Thus we need Condition (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
+Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
 
@@
 In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$.
-Conditions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
+Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
 However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
-Again, assuming Condition (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Condition 6).
+Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).
 
@@
 \begin{lemma}
-When Condition 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
+When Restriction 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
 \end{lemma}
@@
 This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved.
-The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Conditions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
+The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
 
@@
 \begin{lemma}
-Conditions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
+Restrictions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
 \end{lemma}
@@
 
-The conditions of Proposition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome.
+The restrictions in Definition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome.
 In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.
@@
 \begin{proposition}
-Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
 In order for sanctions to lead to ceding by the secessionists, the following are required:
@@
 
-If condition (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1.
-If conditions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of condition 3.
-Note here from condition (2) that sanctions can induce investment behavior by the patron that was ruled out under the conditions of Proposition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.
+If Restriction (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1.
+If Restrictions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Restriction 3.
+Note here from Restriction (2) that sanctions can induce investment behavior by the patron that was ruled out under the Restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.
 

### 2015-11-08 draft.tex
clean up main text of prop 1 proof

@@
 However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2.
-The remainder of the proof treats each set of matched investment decisions in turn.
+The remainder of the proof addresses each set of matched investment decisions in turn.
 
@@
 Consider period 1.
-When Restriction (4) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
+When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
 That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
-Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(q_{s1} - l_{s1} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
+Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(Q_{s0} - \mu - L_{s0} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
 
@@
 
-With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending for or against recognition of the secessionist state:
+With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending by the patron to help achieve recognition of the secessionist state:
 
 \begin{lemma}
-When Restriction 4 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
+When Restriction (4) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
 \end{lemma}
@@
 \begin{lemma}
-Restrictions 3 and 4 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
+Restrictions (3) and (4) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
 \end{lemma}

### 2015-11-08 draft.tex
make proof of prop 1 match defn 2

@@
 Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
-The strategies for the insides actors are the following, organized according to the patron's investments in each state variable.
+The strategies for the insides actors are the following, organized according to the patron's investments in each relevant state variable.
 In period $t$:
@@
 
-To understand this equilibrium, notice that there are three possible ways to disturb the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order for the Status Quo Equilibrium to persist, each of these three disturbances must be deterred.
+To understand the construction of this equilibrium, notice that there are three possible ways to disturb the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order for the Status Quo Equilibrium to persist, each of these three disturbances must be deterred.
 
@@
 Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3).
-Although period $t$ may be reached either because in period $t-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $t$ is the same.
-Thus all periods in which the players are able to move are in an equivalence class in which the only possible strategic difference is in the value of the state variables.
-
-Analysis of the posited equilibrium proceeds most naturally by backward induction within each period.
-Thus we begin with the stage game between the government and secessionists.
+Although period $t$ may be reached either because in period $t-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $t$ is the same given our restriction to stationary Markov strategies.
+
+To establish that the Status Quo Equilibrium of Definition 2 exists, we must show that the given strategies constitute a Nash equilibrium in each subgame within a period given the induced continuation values.\footnote{Subgame perfection within the one-shot game combined with the stationary Markovian property ensures subgame perfection in the whole game.} We begin with the stage game between the government and secessionists.
 We assume that the government and secessionists choose the Status Quo strategy whenever indifferent between it and any other strategy.
@@
 Lemma 1 establishes ranges for the payoffs for players $g$ and $s$---gross of investments by the outside players---in which the status quo outcome can occur.
-The rest of this section tackles the more difficult task of determining the incentives and actions of the outside actors---that is, the patron state and the international community---to impact those payoffs.
+
+We now turn to the incentives of the outside actors---that is, the patron state and the international community---in the subgames that begin in the first and second stages.
 
@@
 
-By contributing $\mu$ in each period, the patron supplies sufficient resources to the unrecognized state to ensure that the secessionist elite prefers the status quo to surrendering independence.
+Lemmas 2 through 4 establish that the only incentive compatible investment is that of the patron to deter the international community from inducing reunification, supported by the strategies described in cases (1)-(3) in Definition 2.
+This combined with Lemma 1 ensures the existence of the Status Quo Equilibrium described in Definition 2.
 \hfill $\blacksquare$

### 2015-11-09 draft.tex
update proofs of lemmas 2, 3 and part of 4

@@
 
-Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g.
-counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{st} - l_{st}$, where $Q_{st}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
-If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.
-
-Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{st} - l_{st}$ so that player $c$ will not find contesting the status quo to be in its interest.
+Player $c$ would be able to induce the secessionists to play Cede---e.g.
+counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
+If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.
+
+On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_{st} - L_{st}$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
-Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$.
+
+Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$.
+Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_{pt}$.
+
+Since the patron is willing to invest up to $\frac{\alpha}{1 -\de}$, $\frac{\beta}{1 -\de} + \mu \leq \frac{\alpha}{1 -\de}$ is a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome.
 \hfill $\blacksquare$
@@
 So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
-We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $t=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.
-
-Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\nu \geq L_{gt} - q_{gt}$, where $L_{gt}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
-If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.
-
-Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gt} - q_{gt}$.
-
-Thus $\frac{\nu}{1 -\de} > \frac{\lambda + \mu + \beta}{1 -\de}$ guarantees that player $c$ will always be willing to counter an investment by player $p$ toward achieving recognition, making $p$ unwilling to make that investment in the initial stage and making the response by player $c$ unnecessary.
-A slightly weaker version of this inequality is stated as condition 4 so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
+We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $t=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$.
+
+If $\frac{\nu}{1-\de} < L_{g,t-1} + R_{pt} - Q_{g,t-1}$, the patron's investment in $l_{gt}$ creates a large enough difference in $L_{st} - Q_{st}$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+
+On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
+counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_{pt} - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
+If this relationship is reversed, similar analysis and results hold.}
+
+Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_{pt}$.
+
+Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome.
+A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
 \hfill $\blacksquare$
@@
 
-We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de}$ as this is most restrictive.
-In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}> 0$.
-
-Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$.
-Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
+We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive.
+In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$.
+
+Multiplying Restriction (3) by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$.
+Using this with Restriction (4), we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
 \[

### 2015-11-10 draft.tex
finish rewrite of lemma 4 proof

@@
 As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).
-
-$p$ is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+So $c$ is willing to invest up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting.
+
+The patron is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
 Thus we must only examine the case where this quantity is positive.
@@
 We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive.
-In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$.
-
-Multiplying Restriction (3) by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$.
-Using this with Restriction (4), we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
-\[
-  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}.
-\]
-
-Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude.
-Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place.
+In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.
+
+If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_{st}\left(R_{pt}\right) - Q_{st}$, the patron's investment in $\omega_{st}$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+
+On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g.
+counter an addition to $\omega_{st}$ with an investment to $q_{st}$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_{st}\left(R_{pt}\right) - Q_{st}$.
+
+This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) - Q_{s,t-1} +\mu = \omega_{s,t-1} - Q_{s,t-1} +R_{pt} p_{1s} + \mu$.
+
+Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} \geq \omega_{s,t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_{pt} p_{1s}$.
+
+Since $R_{pt} \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
+
+Multiplying Restriction (3) by $-(1-p_{1s})$, we have $\frac{-\beta \cdot (1-p_{1s})}{1-\de} -\mu (1-p_{1s}) \geq \frac{-\alpha \cdot (1-p_{1s})}{1-\de}$.
+Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1-\de} -\mu \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$.
+Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu > p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$.
 \hfill $\blacksquare$

### 2015-11-10 draft.tex
revise section 4

@@
 
-If Restriction (1) fails, player $p$ will continue to invest to prevent reunification as in Proposition 1.
-If Restrictions (2) or (3) fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Restriction 3.
-Note here from Restriction (2) that sanctions can induce investment behavior by the patron that was ruled out under the Restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.
+If Condition 1 fails, player $p$ will continue to invest to prevent reunification as in Proposition 1.
+If Conditions 2 or 3 fail, sanctions will lead to fighting initiated by the secessionists---either supported by the patron, or without its support in the case of Condition 3.
+Note here from Condition 2 that sanctions can induce investment behavior by the patron that was ruled out under the restrictions of Definition 1: the goal of sanctions is to destabilize the Status Quo Equilibrium and they certainly can achieve that goal but there may be unintended consequences, most notably the initiation of war by the secessionists.
 
 Thus, even if we do not consider the sanctions to impose any direct costs on the home state (though in practice they likely do), it remains ambiguous whether sanctions will benefit the home state.
-Recall that in the status quo equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo.
+Recall that in the Status Quo Equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo.
 If sanctions induce the secessionists to play Fight, this is a worse outcome for the home state than if the status quo had been allowed to persist.
@@
 \begin{proposition}
-Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
@@
 The proof of Proposition 3 is immediate.
-Although under the conditions of Proposition 1 the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting.
+Although under the conditions of Proposition 1 (i.e.
+when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting.
 The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.
 
-Thus, propositions 2 and 3 imply that sanctions are both wealth destroying and violence increasing.
+Thus, Propositions 2 and 3 imply that sanctions are both wealth destroying and violence increasing.
 The sanctions destroy wealth directly by damaging the economy of the secessionist region and lowering the secessionists' payoffs from the status quo.

### 2015-11-10 draft.tex
minor proof-reading revisions to section 3

@@
 
-\item \emph{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+\item \textit{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
 
@@
 
-Player $c$ dislikes war and so will never invest in $w_s$ or $w_g$, which only serve to increase the likelihood that one of the inside actors chooses to fight.
+Player $c$ dislikes war and so will never invest in $w_s$ or $w_g$ since they increase the likelihood that one of the inside actors chooses to fight.
 It would also not want to make the government lose and so won't invest in $l_g$.
@@
 Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
-In each case, the investments are sufficient to make the largest continuation payoffs in the game between the secessionists and the government those from playing status quo.
+In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government those from playing status quo.
 
@@
 Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
-The strategies for the insides actors are the following, organized according to the patron's investments in each relevant state variable.
-In period $t$:
+
+The strategies for the insides actors are the following, organized according to the patron's investments in each relevant state variable. In period $t$:
 
@@
 
-	\item As long as $R_{pt} \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_{s,t}$, player $c$ invests $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$.
+	\item As long as $R_{pt} \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_{st}$, player $c$ invests $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$.
 Similar to case two, the outcome is that investments in $w_{st}$ and $q_{st}$ are zero.
 
-If $R_{pt} > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_{s,t}$, player $c$ will choose $R_{ct}=0$.
+If $R_{pt} > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_{st}$, player $c$ will choose $R_{ct}=0$.
 Fight becomes the unique best response for the secessionists.

### 2015-11-10 draft.tex
fix condition/restriction terminology throughout text

@@
 
-We provide a set of conditions under which perpetual unrecognized statehood is the outcome of a stationary Markov equilibrium of this game.
+We provide a set of parameter restrictions under which perpetual unrecognized statehood is the outcome of a stationary Markov equilibrium of this game.
 %This requires that the patron is both willing and able to make investments sufficient to (a) offset the deterioration in the secessionists' status quo payoffs and (b) deter the international community from making investments that would induce a negotiated settlement.
@@
 We next turn to presenting the model.
-In Section 3, we present a set of conditions under which unrecognized statehood is a stable, long-run equilibrium outcome.
+In Section 3, we present a set of parameter restrictions under which unrecognized statehood is a stable, long-run equilibrium outcome.
 In Section 4, we analyze the potential effects of sanctions on the resolution of territorial disputes while Section 5 explores extensions of the model.
@@
 
-\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, condition (5) is more likely to be binding than condition (6).
+\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, Restriction (5) is more likely to be binding than Restriction (6).
 If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}

### 2015-11-17 draft.tex
A few typos

@@
 Third, secessionists enjoy external support from a patron (e.g.
-Lynch 2002; Stanislawski 2008.
+Lynch 2002; Stanislawski 2008).
 Lastly, the governments of unrecognized states are internally legitimate, enjoying support from residents of the unrecognized state (e.g.
@@
 Although we are not aware of any formal models of unrecognized statehood there is a growing economics literature that employs game theoretic models of conflict.
-An early model of territorial disputes in which the divisibility of the contested territory is an important parameter is Grossman (2004).
+Grossman (2004) offers an early model of territorial disputes in which the divisibility of the contested territory is an important parameter.
 In Hirschleifer, Boldrin and Levin (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
@@
 
-A major innovation of our model is the introduction of international actors.
+One innovation of our model is the introduction of international actors.
 While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis 2000; Regan 2002; Balch-Lindsay, Enterline, and Joyce 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter 1997, 2002).
@@
 
-We construct a model with four players: the secessionist elite ($s$); the central government of the home state ($g$) from which $s$ is attempting to secede; and two other players, the patron ($p$) and the international community ($c$).
+We construct a model with four players: the secessionist elite ($s$), which seeks recognized independence; the central government of the home state ($g$) from which $s$ is attempting to secede, which seeks reunification; and two other players, the patron ($p$) and the international community ($c$).
 The latter two are outside actors---states, groups of states and/or individuals acting in concert---that have interests in the outcome of the attempted secession.
@@
 At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_{pt}$ to increase the secessionists status quo payoffs, the beginning-of-period payoff $q_{st}$ becomes end-of-period payoff $Q_{st} = q_{st} + R_{pt}$.
+For example, if $p$ expends resources $R_{pt}$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_{st}$ becomes end-of-period payoff $Q_{st} = q_{st} + R_{pt}$.
 We can write analogous expressions for $q_{gt}, \ l_{st},\ l_{gt},\ w_{st}$, and $w_{st}$, and the international community can also make investments to increase whichever of these payoffs they like.
@@
 Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
-In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government those from playing status quo.
+In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government are those from playing status quo.
 
@@
 In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists.
-The international community, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.
+Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.
 sanctions are equivalent to $R_{ct} < 0$).
@@
 In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives.
-Current events, particularly, Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon.
+Current events, particularly Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon.
 While the Ukrainian situation remains fluid, as of October 2015, the status quo equilibrium we model is a plausible future.

### 2015-11-19 draft.tex
add math specification in props 2 and 3, clarify assumption on war lottery

@@
 For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_{it} \equiv (p_1 (W_{it}-\zeta_i), \left(1-p_1\right)( L_{it}-\zeta_i))$.
-$\Omega_t \equiv (\omega_{gt}, \omega_{st}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.} Players are assumed to approach war lotteries as expected values.
+$\Omega_t \equiv (\omega_{gt}, \omega_{st}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.
+Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.
 
@@
 
-The strategies for the insides actors are the following, organized according to the patron's investments in each relevant state variable. In period $t$:
+The strategies for the inside actors are the following, organized according to the patron's investments in each relevant state variable. In period $t$:
 
@@
 \begin{proposition}
-Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e.
+$q_{s,t+1} = Q_{s,t} - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).
 In order for sanctions to lead to ceding by the secessionists, the following are required:
@@
 \begin{proposition}
-Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e.
+$q_{s,t+1} = Q_{s,t} - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
-
 \end{proposition}

### 2015-11-19 draft.tex
clarify outcomes vs. eq'a in alternative outcomes section

@@
 
-In the equilibrium under consideration here, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome.
+In the Status Quo equilibrium, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome.
 This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved.
@@
 If this inequality holds for just one player, then we have only the additional outcome in which that player chooses Fight and its opponent cedes.
-If both players strictly prefer Cede to Fight, then both the (Fight, Cede) and (Cede,Fight) outcomes are added to the status quo outcome.\footnote{In the case of any indifference, we get the relevant combination of (Fight, Fight), (Fight,Cede) and (Cede, Fight).}
+If both players strictly prefer Cede to Fight, then there exist equilibria where both (Fight,Cede) and (Cede,Fight) are outcomes of the stage game.\footnote{In the case of any indifference, we get the relevant combination of (Fight, Fight), (Fight,Cede) and (Cede, Fight).}
 

### 2015-11-19 draft.tex
move paragraph on sufficient but not necessary to alternative outcomes section

@@
 
-The restrictions on the class of games in $\mathcal{G}$ are sufficient but not necessary.
-For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run.
-The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
-In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
-
 \vskip.1in
@@
 
-The restrictions in Definition 1 \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome.
+First, note that the restrictions in Definition 1 are sufficient but not necessary.
+For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run.
+The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
+In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
+
+On the other hand, the restrictions in Definition 1 also \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome.
 In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.

### 2015-11-19 draft.tex
small typo fixes

@@
 The international system is characterized by a norm under which seceding entities are rarely recognized by foreign states unless they are first recognized by the home state.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states.
-In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).}%The higher the costs of secession, the greater the security of existing states.
-Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.
+In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).} Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.
 
@@
 We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely.
-Even in this circumstance, the status quo remains an equilibrium outcome.
+We will show that even in this circumstance the status quo remains an equilibrium outcome.
 
@@
 
-\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, Restriction (5) is more likely to be binding than Restriction (6).
-If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be binding.}
+\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
+If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}
 
@@
 
-In equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
+We now turn to defining and describing the Status Quo Equilibrium.
+We will see that, in this equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
 Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding.
@@
 
-To understand the construction of this equilibrium, notice that there are three possible ways to disturb the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order for the Status Quo Equilibrium to persist, each of these three disturbances must be deterred.
-
-Since the patron moves first, in the Status Quo Equilibrium, the only investment that takes place is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty.
+To understand the construction of this equilibrium, notice that there are three possible ways to disrupt the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order to establish that the Status Quo Equilibrium exists, it must be shown that each of these three deviations will be deterred.
+
+Since the patron moves first, the only investment that takes place in the Status Quo Equilibrium is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty.
 The international community's willingness to counteract investments by the patron toward the other two disturbances implies that there will be no investments in equilibrium in cases (2) and (3).

### 2016-03-02 draft.tex
Removed statement about work in progress from cover page
change time subscripts to superscripts through defn 1

@@
 \thispagestyle{empty}
-\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}
+%\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}
 
@@
 Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
-This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_{it} \de^{t-1}$.
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_i^t \de^{t-1}$.
 
@@
 
-\item $p$ chooses a level of resource expenditure $R_{pt}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
-
-\item $c$ chooses a level of resource expenditure $R_{ct}\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
+\item $p$ chooses a level of resource expenditure $R_p^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
+
+\item $c$ chooses a level of resource expenditure $R_c^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
 
@@
   $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	Fight & $\Omega_t$ &$\Omega_t$&$W_{gt}, L_{st}$ \\ \hline
-	Status Quo& $\Omega_t$& $Q_{gt}, Q_{st}$ &$W_{gt}, L_{st}$ \\ \hline
-	Cede&$L_{gt}, W_{st}$& $L_{gt}, W_{st}$ &$Q_{gt}, Q_{st}$ \\ \hline
+	Fight & $\Omega^t$ &$\Omega^t$&$W_g^t, L_s^t$ \\ \hline
+	Status Quo& $\Omega^t$& $Q_g^t, Q_s^t$ &$W_g^t, L_s^t$ \\ \hline
+	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$Q_g^t, Q_s^t$ \\ \hline
  \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
@@
 
-For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_{it} \equiv (p_1 (W_{it}-\zeta_i), \left(1-p_1\right)( L_{it}-\zeta_i))$.
-$\Omega_t \equiv (\omega_{gt}, \omega_{st}).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.
+For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p_1 (W_i^t-\zeta_i), \left(1-p_1\right)( L_i^t-\zeta_i))$.
+$\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.
 Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.
@@
 As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
-Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_{s,t+1} = Q_{s,t} - \mu$.
+Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_s^{t+1} = Q_s^t - \mu$.
 Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.
@@
 However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
-$L_{s,t} = l_{s,t} + R_{p,t} + R_{c,t}$ and $l_{s,t+1} = L_{s,t}$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
+$L_s^t = l_s^t + R_p^t + R_c^t$ and $l_s^{t+1} = L_s^t$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
 
@@
 At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_{pt}$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_{st}$ becomes end-of-period payoff $Q_{st} = q_{st} + R_{pt}$.
-We can write analogous expressions for $q_{gt}, \ l_{st},\ l_{gt},\ w_{st}$, and $w_{st}$, and the international community can also make investments to increase whichever of these payoffs they like.
-$B_{pt}$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
+For example, if $p$ expends resources $R_p^t$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_s^t$ becomes end-of-period payoff $Q_s^t = q_s^t + R_p^t$.
+We can write analogous expressions for $q_g^t, \ l_s^t,\ l_g^t,\ w_s^t$, and $w_g^t$, and the international community can also make investments to increase whichever of these payoffs they like.
+$B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
 
@@
 
-Player $p$'s payoffs in period $t$ are $U_{pt}= -\alpha X+\lambda Y -R_{pt}$, while the payoffs for player $c$ are $U_{ct}= \beta X-\nu Y -R_{ct}$, with both payoffs denoted in currency units.
+Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
 Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
@@
 \begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_{i0} \geq L_{i0}$: remaining in the status quo is better than ceding at the beginning of the game.}
-
-\item \textit{For both players $g$ and $s$, $\frac {Q_{i0}}{1-\delta} \geq  -\zeta_i+\frac{W_{i0}(p_{1i}) + L_{i0}(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}
+
+\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_{1i}) + L_i^0(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
 
@@
 
-\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(Q_{s0} -\mu - L_{s0} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
-
-\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
+\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+
+\item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
 If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}
@@
 The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_{st}$ and $w_{gt}$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
+That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_s^t$ and $w_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
 However, positive investments in some of the state variables can be ruled out by preference assumptions.
@@
 \begin{enumerate}
-	\item The patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_{st}$. Player $c$ invests $R_{ct} =0$ in $l_{st}$.
-
-If $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, in $l_{st}$.
+	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.
+
+If $R_p^t < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$.
 Cede becomes the unique best response for the secessionists.
 
-	\item As long as $R_{pt} \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = R_{pt} - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_{gt}$.
-Note that the outcome is that investments in $l_{gt}$ and $q_{gt}$ are zero.
-
-If $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_{gt}$, player $c$ invests $R_{ct} = 0$. Cede becomes the unique best response for the government.
-
-	\item As long as $R_{pt} \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_{st}$, player $c$ invests $R_{ct} = R_{pt}p_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_{st}$.
-Similar to case two, the outcome is that investments in $w_{st}$ and $q_{st}$ are zero.
-
-If $R_{pt} > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_{st}$, player $c$ will choose $R_{ct}=0$.
+	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_g^t$.
+Note that the outcome is that investments in $l_g^t$ and $q_g^t$ are zero.
+
+If $R_p^t > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
+
+	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_s^t$.
+Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.
+
+If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
 Fight becomes the unique best response for the secessionists.
@@
 \begin{lemma}
-$Q_{it} \geq L_{it}$ and $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_{1i}) + L_{it}(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
+$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_{1i}) + L_i^t(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
 \end{lemma}
@@
 
-Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_{st},q_{gt},w_{st},w_{gt},l_{st},$ and $l_{gt}$.
+Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_s^t,q_g^t,w_s^t,w_g^t,l_s^t,$ and $l_g^t$.
 However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2.
@@
 When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
-That is, in equilibrium, $R_{p1} = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
-Thus we need Restriction (5): $B_{p1} \geq\frac{\beta}{1-\de} - \left(Q_{s0} - \mu - L_{s0} \right)$, where $B_{p1}$ is the amount $p$ has available to spend on the conflict in period $1$.
+That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
+Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
 
@@
 Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
-However, because in each period $q_{st}$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
+However, because in each period $q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
 Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).
@@
 Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.
-sanctions are equivalent to $R_{ct} < 0$).
+sanctions are equivalent to $R_c^t < 0$).
 Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.
@@
 Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e.
-$q_{s,t+1} = Q_{s,t} - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).
+$q_s^{t+1} = Q_s^t - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).
 In order for sanctions to lead to ceding by the secessionists, the following are required:
@@
 Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e.
-$q_{s,t+1} = Q_{s,t} - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
+$q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
@@
 Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions.
-Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{st} \ \forall t$, so the continuation value is $\frac{Q_{st}}{1-\delta}$.
-Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{st}$ in each period, is $\frac{L_{st}}{1-\delta}$.
+Given the stationary equilibrium actions, playing Status Quo will lead to $Q_s^t \ \forall t$, so the continuation value is $\frac{Q_s^t}{1-\delta}$.
+Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$.
 The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$.
-Thus the continuation value is $-\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.
+Thus the continuation value is $-\zeta_s +\frac{W_s^t(p_1)+ L_s^t(1-p_1)}{1-\delta}$.
 
@@
 \noindent {\bf Proof of Lemma 2} \\
-In the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$.
-If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pt}^{RU}= -\alpha -R_p^{RU}$ and $U_{ct}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
+In the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$.
+If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_p^t^{RU}= -\alpha -R_p^{RU}$ and $U_c^t^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
 
@@
 Player $c$ would be able to induce the secessionists to play Cede---e.g.
-counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $q_s^t$ with an investment to $l_s^t$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.
 
-On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_{pt} - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_{st} - L_{st}$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
+On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_s^t - L_s^t$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
@@
 Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$.
-Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_{pt}$.
+Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.
 
@@
 \noindent {\bf Proof of Lemma 3} \\
-Again, in the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$.
-If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pt}^{REC}= \lambda -R_p^{REC}$ and $U_{ct}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
+Again, in the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$.
+If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_p^t^{REC}= \lambda -R_p^{REC}$ and $U_c^t^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
 
@@
 
-If $\frac{\nu}{1-\de} < L_{g,t-1} + R_{pt} - Q_{g,t-1}$, the patron's investment in $l_{gt}$ creates a large enough difference in $L_{st} - Q_{st}$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+If $\frac{\nu}{1-\de} < L_{g,t-1} + R_p^t - Q_{g,t-1}$, the patron's investment in $l_g^t$ creates a large enough difference in $L_s^t - Q_s^t$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
 
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_{pt} - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $l_g^t$ with an investment to $q_g^t$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_p^t - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.}
 
-Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_{pt}$.
+Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
 
@@
 
-If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_{st}\left(R_{pt}\right) - Q_{st}$, the patron's investment in $\omega_{st}$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
 
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g.
-counter an addition to $\omega_{st}$ with an investment to $q_{st}$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_{st}\left(R_{pt}\right) - Q_{st}$.
-
-This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) - Q_{s,t-1} +\mu = \omega_{s,t-1} - Q_{s,t-1} +R_{pt} p_{1s} + \mu$.
-
-Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} \geq \omega_{s,t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_{pt} p_{1s}$.
-
-Since $R_{pt} \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
+counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
+
+This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_{s,t-1}+R_p^t)p_{1s} + L_{s,t-1}(1-p_{1s}) - Q_{s,t-1} +\mu = \omega_{s,t-1} - Q_{s,t-1} +R_p^t p_{1s} + \mu$.
+
+Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} \geq \omega_{s,t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_p^t p_{1s}$.
+
+Since $R_p^t \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
 
@@
 Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$.
-As the only alteration is a reduction in $Q_{st}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed.
+As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed.
 As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

### 2016-03-02 draft.tex
change time subscripts to superscripts from defn 1 to end of appendix

@@
 \begin{enumerate}
-	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.
-
-If $R_p^t < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$.
+	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.
+
+If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$.
 Cede becomes the unique best response for the secessionists.
 
-	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_{g,t-1} - L_{g,t-1}\right)$ in $q_g^t$.
+	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ in $q_g^t$.
 Note that the outcome is that investments in $l_g^t$ and $q_g^t$ are zero.
 
-If $R_p^t > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
-
-	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_{s,t-1} + \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$ in $q_s^t$.
+If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
+
+	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)$ in $q_s^t$.
 Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.
 
-If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
+If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
 Fight becomes the unique best response for the secessionists.
@@
 Consider period 1.
-When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_{s1} \ (= q_{s1} + R_{s1})$ and $l_{s1}$ so that $c$ will not invest.
-That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_{s1} - l_{s1})$ as long as $p$ can afford to make this investment.
+When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1 \ (= q_s^1 + R_s^1)$ and $l_s^1$ so that $c$ will not invest.
+That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_s^1 - l_s^1)$ as long as $p$ can afford to make this investment.
 Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
@@
 \noindent {\bf Proof of Lemma 2} \\
-In the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$.
-If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_p^t^{RU}= -\alpha -R_p^{RU}$ and $U_c^t^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
+In the status quo, we have $U_p^{t,SQ}= -R_p^{SQ}$ and $U_c^{t,SQ}= -R_c^{SQ}$.
+If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_p^{t,RU}= -\alpha -R_p^{RU}$ and $U_c^{t,RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
 
@@
 Player $c$ would be able to induce the secessionists to play Cede---e.g.
-counter an addition to $q_s^t$ with an investment to $l_s^t$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $q_s^t$ with an investment to $l_s^t$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.
 
-On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_{s,t-1} -\mu + R_p^t - L_{s,t-1}$, the patron's investment creates a large enough difference in $Q_s^t - L_s^t$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
+On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$, the patron's investment creates a large enough difference in $Q_s^t - L_s^t$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
 
-Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$.
+Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$.
 Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.
@@
 \noindent {\bf Proof of Lemma 3} \\
-Again, in the status quo, we have $U_p^t^{SQ}= -R_p^{SQ}$ and $U_c^t^{SQ}= -R_c^{SQ}$.
-If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_p^t^{REC}= \lambda -R_p^{REC}$ and $U_c^t^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
+Again, in the status quo, we have $U_p^{t,SQ}= -R_p^{SQ}$ and $U_c^{t,SQ}= -R_c^{SQ}$.
+If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_p^{t,REC}= \lambda -R_p^{REC}$ and $U_c^{t,REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
 
@@
 
-If $\frac{\nu}{1-\de} < L_{g,t-1} + R_p^t - Q_{g,t-1}$, the patron's investment in $l_g^t$ creates a large enough difference in $L_s^t - Q_s^t$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+If $\frac{\nu}{1-\de} < L_g^{t-1} + R_p^t - Q_g^{t-1}$, the patron's investment in $l_g^t$ creates a large enough difference in $L_s^t - Q_s^t$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
 
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_g^t$ with an investment to $q_g^t$---as long as $\frac{\nu}{1-\de} \geq L_{g,t-1} + R_p^t - Q_{g,t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $l_g^t$ with an investment to $q_g^t$---as long as $\frac{\nu}{1-\de} \geq L_g^{t-1} + R_p^t - Q_g^{t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.}
 
-Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} > L_{s,t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
+Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
 
@@
 
-This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_{s,t-1}+R_p^t)p_{1s} + L_{s,t-1}(1-p_{1s}) - Q_{s,t-1} +\mu = \omega_{s,t-1} - Q_{s,t-1} +R_p^t p_{1s} + \mu$.
-
-Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} \geq \omega_{s,t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_p^t p_{1s}$.
+This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_{1s} + L_s^{t-1}(1-p_{1s}) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_{1s} + \mu$.
+
+Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_p^t p_{1s}$.
 

### 2016-03-02 draft.tex
simplify prob in war lotter from p1 to p

@@
 
-For probabilities $p_1$ of outright victory and $1-p_1$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p_1 (W_i^t-\zeta_i), \left(1-p_1\right)( L_i^t-\zeta_i))$.
+For probabilities $p$ of outright victory and $1-p$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p (W_i^t-\zeta_i), \left(1-p\right)( L_i^t-\zeta_i))$.
 $\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.
@@
 
-\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_{1i}) + L_i^0(1-p_{1i})}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
+\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
 
@@
 
-\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
+\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
 
@@
 
-	\item As long as $R_p^t \leq \frac{1}{p_{1s}} \left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_{1s} + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)$ in $q_s^t$.
+	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ in $q_s^t$.
 Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.
 
-If $R_p^t > \frac{1}{p_{1s}}\left[\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_{1s} + L_s^{t-1}(1-p_{1s})\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
+If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
 Fight becomes the unique best response for the secessionists.
@@
 \begin{lemma}
-$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_{1i}) + L_i^t(1-p_{1i})}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
+$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_i) + L_i^t(1-p_i)}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
 \end{lemma}
@@
 Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e.
-$q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_{1s}\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
+$q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
@@
 1998.
-"The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives EuropŽennes de Sociologie} 39 (01): 110-146.
+"The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives Europennes de Sociologie} 39 (01): 110-146.
 
@@
 Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$.
-The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$.
-Thus the continuation value is $-\zeta_s +\frac{W_s^t(p_1)+ L_s^t(1-p_1)}{1-\delta}$.
+The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p$, and $L_s$ forever with probability $1-p$.
+Thus the continuation value is $-\zeta_s +\frac{W_s^t(p)+ L_s^t(1-p)}{1-\delta}$.
 
@@
 \noindent {\bf Proof of Lemma 4} \\
-The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
-For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
+The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
+For player $c$ it is $\frac{-\nu p_s + \beta (1-p_s)}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
 As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting.
-
-The patron is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
+So $c$ is willing to invest up to $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}$ toward preventing fighting.
+
+The patron is willing to invest up to $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
 Thus we must only examine the case where this quantity is positive.
@@
 We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive.
-In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.
-
-If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+In this only case of interest, we have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.
+
+If $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
 
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g.
-counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
-
-This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_{1s} + L_s^{t-1}(1-p_{1s}) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_{1s} + \mu$.
-
-Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_p^t p_{1s}$.
-
-Since $R_p^t \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
-
-Multiplying Restriction (3) by $-(1-p_{1s})$, we have $\frac{-\beta \cdot (1-p_{1s})}{1-\de} -\mu (1-p_{1s}) \geq \frac{-\alpha \cdot (1-p_{1s})}{1-\de}$.
-Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1-\de} -\mu \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$.
-Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu > p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$.
+counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
+
+This can be expanded as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_s + L_s^{t-1}(1-p_s) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_s + \mu$.
+
+Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.
+
+Since $R_p^t \leq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  - \mu \geq p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} - p_s \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
+
+Multiplying Restriction (3) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$.
+Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
+Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
 \hfill $\blacksquare$
@@
 
-Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de}$.
-For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_{1s} + \beta (1 - p_{1s})}{1-\de}$.
+Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_s - \alpha (1-p_s)}{1-\de}$.
+For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_s + \beta (1 - p_s)}{1-\de}$.
 Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
 \[
-  \textstyle \frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_{1s} + \beta (1-p_{1s})}{1-\de} -  \frac{\de^m}{1-\de} \beta
+  \textstyle \frac{\lambda p_s - \alpha (1-p_s)}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_s + \beta (1-p_s)}{1-\de} -  \frac{\de^m}{1-\de} \beta
 \]

### 2016-03-02 draft.tex
remove lowercase state variables throughout sec. 2

@@
 As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
-Therefore, the payoffs from the status quo, in the absence of other actions, are given by $q_s^{t+1} = Q_s^t - \mu$.
-Lower case letters denote beginning-of-period quantities net of investments by outside actors, while upper case letters denote end-of-period quantities gross of such investments.
+Therefore, the payoffs from the status quo evolve according to the following law of motion: $Q_s^{t} = Q_s^{t-1} - \mu + R_p^t - R_c^t$.\footnote{Here we anticipate a bit the incentives of the patron and international community in the next section to sign the impact of contributions: the patron will want to increase the value of the status quo to the secessionists, while the international community could only desire to reduce this value.}
 
@@
 However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
-$L_s^t = l_s^t + R_p^t + R_c^t$ and $l_s^{t+1} = L_s^t$, likewise for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
+$L_s^t = L_s^{t-1} - R_p^t + R_c^t$ and similarly for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
 
@@
 At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_p^t$ to increase the secessionists' status quo payoffs, the beginning-of-period payoff $q_s^t$ becomes end-of-period payoff $Q_s^t = q_s^t + R_p^t$.
-We can write analogous expressions for $q_g^t, \ l_s^t,\ l_g^t,\ w_s^t$, and $w_g^t$, and the international community can also make investments to increase whichever of these payoffs they like.
+For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t$.
+We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like.
 $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

### 2016-03-02 draft.tex
remove lowercase state variables throughout body of paper

@@
 The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-That is, $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.\footnote{In the case of $w_s^t$ and $w_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
+That is, $s = \left(Q_s,Q_g,L_s,L_g,W_s,W_g \right)$.\footnote{In the case of $W_s^t$ and $W_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
 However, positive investments in some of the state variables can be ruled out by preference assumptions.
 
-Player $c$ dislikes war and so will never invest in $w_s$ or $w_g$ since they increase the likelihood that one of the inside actors chooses to fight.
-It would also not want to make the government lose and so won't invest in $l_g$.
-This leaves three state variables in which player $c$ might invest: $q_s$, $q_g$, $l_s$.
-
-Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
-The only reason the patron would invest in $q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $q_g$.
-This leaves four state variables in which the patron may invest: $q_s$, $l_g$ and $w_s$.
+Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight.
+It would also not want to make the government lose and so won't invest to augment $L_g$.
+This leaves three state variables in which player $c$ might invest: $Q_s$, $Q_g$, $L_s$.
+
+Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $W_g$ or $L_s$.
+The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$.
+This leaves four state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.
 
@@
 \begin{enumerate}
-	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $q_s^t$. Player $c$ invests $R_c^t =0$ in $l_s^t$.
-
-If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = l_s^t - \left(q_s^t + R_p^t \right) + \ve$, for $\ve$ small, in $l_s^t$.
+	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $Q_s^{t-1}$. Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.
+
+If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = L_s^{t-1} - \left(Q_s^{t-1} - \mu + R_p^t \right) + \ve$, for $\ve$ small, to augment $L_s^{t-1}$.
 Cede becomes the unique best response for the secessionists.
 
-	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ in $q_g^t$.
-Note that the outcome is that investments in $l_g^t$ and $q_g^t$ are zero.
-
-If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $l_g^t$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
-
-	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $w_s^t$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ in $q_s^t$.
-Similar to case two, the outcome is that investments in $w_s^t$ and $q_s^t$ are zero.
-
-If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $w_s^t$, player $c$ will choose $R_c^t=0$.
+	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $Q_g^{t-1}$.
+Note that the outcome is that investments to augment $L_g^{t-1}$ and $Q_g^{t-1}$ are zero.
+
+If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
+
+	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $W_s^{t-1}$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ to augment $Q_s^{t-1}$.
+Similar to case two, the outcome is that investments to augment $W_s^{t-1}$ and $Q_s^{t-1}$ are zero.
+
+If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $W_s^{t-1}$, player $c$ will choose $R_c^t=0$.
 Fight becomes the unique best response for the secessionists.
@@
 
-Recall that either outside actor could invest toward increasing any of the six state variables in a period: $q_s^t,q_g^t,w_s^t,w_g^t,l_s^t,$ and $l_g^t$.
+Recall that either outside actor could invest toward increasing any of the six state variables in a period.
 However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2.
@@
 Consider period 1.
-When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1 \ (= q_s^1 + R_s^1)$ and $l_s^1$ so that $c$ will not invest.
-That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(q_s^1 - l_s^1)$ as long as $p$ can afford to make this investment.
+When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest.
+That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment.
 Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
@@
 Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
-However, because in each period $q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
+However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
 Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).
@@
 Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e.
-$q_s^{t+1} = Q_s^t - \mu -\sigma$ where $\sigma$ is the reduction due to sanctions).
+$Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ where $\sigma$ is the reduction due to sanctions).
 In order for sanctions to lead to ceding by the secessionists, the following are required:
@@
 Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e.
-$q_s^{t+1} = Q_s^t - \mu -\sigma$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_{1g}\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
+$Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_g\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.

### 2016-03-03 draft.tex
put plainpages=false back and add pageanchor=false to overcome pageanchor error
remove lowercase state variables throughout appendix

@@
 %\usepackage{hypertex}
-\usepackage[pdftex, bookmarks, colorlinks=true, plainpages = false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
+\usepackage[pdftex, bookmarks, colorlinks=true, plainpages=false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
 \usepackage{amssymb}
@@
 \begin{document}
+\hypersetup{pageanchor=false}
 \begin{titlepage}
@@
 \end{titlepage}
+\hypersetup{pageanchor=true}
 \setcounter{page}{1}
@@
 Player $c$ would be able to induce the secessionists to play Cede---e.g.
-counter an addition to $q_s^t$ with an investment to $l_s^t$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $Q_s^{t-1}$ with an investment to $L_s^{t-1}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.
 
-On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$, the patron's investment creates a large enough difference in $Q_s^t - L_s^t$ (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
+On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$, the patron's investment creates a large enough difference in the secessionist's payoffs from the status quo versus ceding (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
 Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
@@
 
-If $\frac{\nu}{1-\de} < L_g^{t-1} + R_p^t - Q_g^{t-1}$, the patron's investment in $l_g^t$ creates a large enough difference in $L_s^t - Q_s^t$ (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
+If $\frac{\nu}{1-\de} < L_g^{t-1} + R_p^t - Q_g^{t-1}$, the patron's investment to augment $L_g^{t-1}$ creates a large enough difference in the secessionists' payoffs of ceding versus the status quo (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
 
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $l_g^t$ with an investment to $q_g^t$---as long as $\frac{\nu}{1-\de} \geq L_g^{t-1} + R_p^t - Q_g^{t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
+counter an addition to $L_g^{t-1}$ with an investment to augment $Q_g^{t-1}$---as long as $\frac{\nu}{1-\de} \geq L_g^{t-1} + R_p^t - Q_g^{t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
 If this relationship is reversed, similar analysis and results hold.}
@@
 On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g.
-counter an addition to $\omega_s^t$ with an investment to $q_s^t$---as long as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
+counter an addition to $\omega_s^t$ with an investment to augment $Q_s^{t-1}$---as long as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
 

### 2016-03-03 draft.tex
add \pdfoptionpdfminorversion=6 to preamble to get rid of warning about pdf version
move desc of state variables from sec:stage to sec:structure and rewrite, closes #2

@@
 \usepackage{graphicx,verbatim,scalefnt}
+\pdfoptionpdfminorversion=6 %gets rid of warning about pdf version
 
@@
 There are an infinite number of periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
-This is a dynamic game because the fundamentals determining the payoffs of the inside actors are state variables whose values persist from period to period and in which the outside actors can make investments.
+
+Importantly, the payoffs for what we term the `inside actors'---the secessionists and the home government---are modeled as state variables.
+The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.
+The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
+As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
+Therefore, the payoffs to the secessionists from the status quo evolve according to the law of motion $Q_s^{t} = Q_s^{t-1} - \mu + R_p^t - R_c^t$ where $R_p^t$ is the investment of the patron and $R_c^t$ is the investment of the international community in period $t$.\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions: the patron will want to increase the value of the status quo to the secessionists, while the international community could only desire to reduce this value.}
+
+The other payoffs for the inside actors are also modeled as state variables.
+However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
+$L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status.
+The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.
+
+Both because of the presence of absorbing states and these state variables whose values persist from period to period, this is a dynamic game rather than a standard repeated game.
 
@@
 \subsubsection{Stage Game Payoffs}
+\label{sec:stage}
 
@@
 If both $s$ and $g$ play Status Quo, then the status quo persists.
-The status quo payoffs for the secessionists are modeled as a state variable.
-The transition of this state variable from one period to the next involves an increase by the amount of the patron's investment and an automatic reduction by $\mu$.\footnote{Note that it is also possible for the international community to make an investment in the secessionists' status quo payoffs, but we will show that they have no incentive to do so.} The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
-As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
-Therefore, the payoffs from the status quo evolve according to the following law of motion: $Q_s^{t} = Q_s^{t-1} - \mu + R_p^t - R_c^t$.\footnote{Here we anticipate a bit the incentives of the patron and international community in the next section to sign the impact of contributions: the patron will want to increase the value of the status quo to the secessionists, while the international community could only desire to reduce this value.}
-
-Other payoffs are also modeled as state variables.
-However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
-$L_s^t = L_s^{t-1} - R_p^t + R_c^t$ and similarly for player $g$ and for the payoffs from winning for both players and the status quo for player $g$.
 

### 2016-03-03 draft.tex
small textual change in sec:structure

@@
 
-Both because of the presence of absorbing states and these state variables whose values persist from period to period, this is a dynamic game rather than a standard repeated game.
+Both because of the presence of absorbing states and these state variables whose values depend on their values in the previous period, this is a dynamic game rather than a standard repeated game.
 

### 2016-03-08 draft.tex
replace hardcoding of restriction numbering with hyperref coding

@@
 \begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}
-
-\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}
-
-\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}
-
-\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}
-
-\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}
+\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}\label{res:1}
+
+\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}
+
+\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}\label{res:3}
+
+\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}\label{res:4}
+
+\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}\label{res:5}
 
 \item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
-If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}
+If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}\label{res:6}
 
@@
 The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$.
-This leaves four state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.
+This leaves three state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.
 
@@
 \begin{lemma}
-When Restriction (3) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
+When Restriction (\ref{res:3}) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
 \end{lemma}
 
-\noindent Restriction (3) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
+\noindent Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
 To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}
@@
 Consider period 1.
-When Restriction (3) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest.
+When Restriction (\ref{res:3}) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest.
 That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment.
-Thus we need Restriction (5): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
+Thus we need Restriction (\ref{res:5}): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
 
@@
 In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$.
-Restrictions (3) and (5) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
+Restrictions (\ref{res:3}) and (\ref{res:5}) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
 However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
-Again, assuming Restriction (3) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction 6).
+Again, assuming Restriction (\ref{res:3}) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction \ref{res:6}).
 
@@
 \begin{lemma}
-When Restriction (4) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
+When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
 \end{lemma}
@@
 This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved.
-The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (5) and (6): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
+The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (\ref{res:5}) and (\ref{res:6}): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
 
@@
 \begin{lemma}
-Restrictions (3) and (4) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
+Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
 \end{lemma}
@@
 First, note that the restrictions in Definition 1 are sufficient but not necessary.
-For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (6) need not be met to maintain the status quo in the short run.
+For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (\ref{res:6}) need not be met to maintain the status quo in the short run.
 The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
-In cases where Restriction (6) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
+In cases where Restriction (\ref{res:6}) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
 
@@
 
-Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$.
+Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$.
 Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.
@@
 
-Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
+Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
 
 Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome.
-A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
+A slightly weaker version of this inequality is stated as Restriction (\ref{res:4}) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
 \hfill $\blacksquare$
@@
 
-Restriction (1) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.
+Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.
 
@@
 
-Multiplying Restriction (3) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$.
-Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
+Multiplying Restriction (\ref{res:3}) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$.
+Using this with Restriction (\ref{res:4}) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
 Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.

### 2016-03-08 draft.tex
add new restriction to cover case where s prefers war to ceding

@@
 
-\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq  -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}
+\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq \frac {\omega_i^0}{1-\delta} = -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}
+
+\item \textit{Either (a) $\frac{L_s^0}{1-\delta}  > \frac{\omega_s^0}{1-\delta}$ or (b) $0 > \frac{\lambda p_s - \alpha (1-p_s)}{1-\de} > -\mu$: either the secessionists prefer ceding to war or the patron's disutility from war is greater than the per-period cost of offsetting the deterioration in the secessionists' status quo payoffs.}\label{res:new}
 

### 2016-03-08 draft.tex
expand lemma 2, add new paragraph immediately after w/explanation

@@
 \begin{lemma}
-When Restriction (\ref{res:3}) of Definition 1 holds, the patron's willingness to invest to maintain the status quo is sufficient to deter the international community from intervening to encourage reunification.
+When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
 \end{lemma}
 
-\noindent Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
+\noindent Restriction (\ref{res:new}) guarantees one of two outcomes: either the patron is willing to invest in the secessionists' status quo payoffs to avoid them falling below the expected payoffs from war or the secessionists' payoffs from war remain below those from ceding so that it is only the patron's incentives concerning the ceding outcome with which we must be concerned.
+The following treats the latter case; a parallel argument using Restriction \ref{res:new}(b) can be made in the former case.
+
+Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
 To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}

### 2016-03-08 draft.tex
modify proof of lemma 2 to take account of war possibility

@@
 \noindent {\bf Proof of Lemma 2} \\
+Begin by assuming Restriction \ref{res:new}a holds.
 In the status quo, we have $U_p^{t,SQ}= -R_p^{SQ}$ and $U_c^{t,SQ}= -R_c^{SQ}$.
@@
 Since the patron is willing to invest up to $\frac{\alpha}{1 -\de}$, $\frac{\beta}{1 -\de} + \mu \leq \frac{\alpha}{1 -\de}$ is a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome.
+
+If Restriction \ref{res:new}a is violated, as the secessionists' status quo payoffs degrade from period to period, they will fall below $\omega_s^t$ before they fall below $L_s^t$.
+If no investment is made in $Q_s^{t-1}$, the secessionists will choose the war outcome.
+Restriction \ref{res:new}b guarantees that the patron will make the investment $R_p^{t'}  = \omega_s^{t-1} -\mu - Q_s^{t-1} \leq \mu$ to prevent the secessionists choosing war since this $R_p^{t'}$ is strictly less than the patron's expected loss from the war outcome.
+Note that the patron's actual investment will be the maximum of the investment required to prevent the war and ceding outcomes.
 \hfill $\blacksquare$

### 2016-03-08 draft.tex
add ctn values, remove BR language from defn 2; note Lem 4 not necessary when res:newb holds

@@
 
-\item \textit{Either (a) $\frac{L_s^0}{1-\delta}  > \frac{\omega_s^0}{1-\delta}$ or (b) $0 > \frac{\lambda p_s - \alpha (1-p_s)}{1-\de} > -\mu$: either the secessionists prefer ceding to war or the patron's disutility from war is greater than the per-period cost of offsetting the deterioration in the secessionists' status quo payoffs.}\label{res:new}
+\item \textit{Either (a) $\frac{L_s^0}{1-\delta}  > \frac{\omega_s^0}{1-\delta}$ or (b) $0 > -\mu > \frac{\lambda p_s - \alpha (1-p_s)}{1-\de}$: either the secessionists prefer ceding to war or the patron's disutility from war is greater than the per-period cost of offsetting the deterioration in the secessionists' status quo payoffs.}\label{res:new}
 
@@
 
-The strategies for the government and secessionists in this equilibrium are to play their best responses in the stage game given the continuation values induced by the investments of the outside actors.
-Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
+The strategies for the government and secessionists in this equilibrium are to play their best responses given the continuation values induced by the investments of the outside actors.
+The continuation value from choosing war is the expected future stream of payoffs from the two possible absorbing states net of the cost of war $\left(-\zeta_i +\frac{W_i^t(p_i)+ L_i^t(1-p_i)}{1-\delta}\right)$.
+Similarly, the continuation value from ceding is $\frac{L_i^t}{1-\de}$ while that for the status quo is $\sum_{s=t}^\infty \de^{t-s} Q_i^t$.
+Unless otherwise noted below, playing Status Quo is the dynamic best response for both inside actors.
 
@@
 \begin{enumerate}
-	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),0\right\}$ to augment $Q_s^{t-1}$. Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.
-
-If $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$ player $c$ invests $R_c^t = L_s^{t-1} - \left(Q_s^{t-1} - \mu + R_p^t \right) + \ve$, for $\ve$ small, to augment $L_s^{t-1}$.
-Cede becomes the unique best response for the secessionists.
+	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),\frac{\omega_s^{t-1}}{1-\de} +\mu - \frac{Q_s^{t-1}}{1-\de},0\right\}$ to augment $Q_s^{t-1}$.
+Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.
+
+If $\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right) > 0$ and $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$, player $c$ invests $R_c^t = L_s^{t-1} - \left(Q_s^{t-1} - \mu + R_p^t \right) + \ve$, for $\ve$ small, to augment $L_s^{t-1}$ so that the secessionists play Cede.
 
@@
 
-If $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = 0$. Cede becomes the unique best response for the government.
+If the patron invests $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ is deterred from investing in $Q_g^{t-1}$ ($R_c^t = 0$) so that the government plays Cede.
 
@@
 
-If $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $W_s^{t-1}$, player $c$ will choose $R_c^t=0$.
-Fight becomes the unique best response for the secessionists.
+If the patron invests $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $W_s^{t-1}$, player $c$ is deterred from investing in $Q_s^{t-1}$ ($R_c^t=0$) so that the secessionists play Fight.
 \end{enumerate}
@@
 One last possibility is ruled out by Lemma 4: that the patron would invest to encourage the secessionists to fight.
+Note that the patron has no incentive to make this investment if Restriction \ref{res:new}b holds.
 
@@
 If no investment is made in $Q_s^{t-1}$, the secessionists will choose the war outcome.
-Restriction \ref{res:new}b guarantees that the patron will make the investment $R_p^{t'}  = \omega_s^{t-1} -\mu - Q_s^{t-1} \leq \mu$ to prevent the secessionists choosing war since this $R_p^{t'}$ is strictly less than the patron's expected loss from the war outcome.
+Restriction \ref{res:new}b guarantees that the patron will make the investment $R_p^{t'}  = \frac{\omega_s^{t-1}}{1-\de} +\mu - \frac{Q_s^{t-1}}{1-\de} \leq \mu$ to prevent the secessionists choosing war since this $R_p^{t'}$ is strictly less than the patron's expected loss from the war outcome.
 Note that the patron's actual investment will be the maximum of the investment required to prevent the war and ceding outcomes.

### 2016-03-15 draft.tex
clarify wording on inside actors best response
adjust restriction 6 and item 1 in defn 2

@@
 
-\item \textit{$B_p^t \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
+\item \textit{$B_p^t \geq \mu, \ \forall t$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
 If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}\label{res:6}
@@
 Similarly, the continuation value from ceding is $\frac{L_i^t}{1-\de}$ while that for the status quo is $\sum_{s=t}^\infty \de^{t-s} Q_i^t$.
-Unless otherwise noted below, playing Status Quo is the dynamic best response for both inside actors.
+Unless otherwise noted below, playing Status Quo is the best response for both inside actors in terms of continuation values.
 
@@
 \begin{enumerate}
-	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),\frac{\omega_s^{t-1}}{1-\de} +\mu - \frac{Q_s^{t-1}}{1-\de},0\right\}$ to augment $Q_s^{t-1}$.
+	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),\omega_s^{t-1} +\mu - Q_s^{t-1},0\right\}$ to augment $Q_s^{t-1}$.
 Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.

### 2016-06-01 draft.tex
fix statement about infinite horizon
modify figure 1

@@
 
-The game is infinitely repeated and proceeds as follows.
+The dynamic game has an infinite horizon and proceeds as follows.
 At the beginning of each period, the payoffs for the secessionists from unrecognized statehood---to which we refer as the status quo---deteriorate by a fixed amount, representing the ongoing costs of non-recognition.
@@
 Grossman (2004) offers an early model of territorial disputes in which the divisibility of the contested territory is an important parameter.
-In Hirschleifer, Boldrin and Levin (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
+In Hirshleifer, Boldrin and Levin (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
 In our context, the contested territory is indivisible; this implies that peace would require very large concessions so that this pathway of using small compromises to avoid conflict is not available.
@@
 
-\subsection{Timing and Structure of Interaction}
+\subsection{Interaction within a Period}
 \label{sec:structure}

### 2016-06-01 draft.tex
add new, simplified exposition at start of section 2.2

@@
 The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
-There are an infinite number of periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
-
-Importantly, the payoffs for what we term the `inside actors'---the secessionists and the home government---are modeled as state variables.
+There are an infinite number of discrete periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
+
+Within each period, there is first an investment stage in which first the patron and then the international community each make decisions about how much to invest in altering the payoffs of the `inside actors'---the secessionists and the home government.
+Then, the inside actors play a ``conflict stage game'' in which they simultaneously choose whether to cede the issue of status, fight, or maintain the status quo.
+
+The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the per-period value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively.
+The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede).
+Otherwise, the game ends in an absorbing state as described below.
+
 The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.

### 2016-06-01 draft.tex
misc changes from editor, including fig. 2

@@
 \usepackage{pdfsync}
-\usepackage{url}
+\usepackage{url,bm}
 \usepackage{pst-all}
@@
 
-The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the per-period value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively.
+The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the evolving value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively.
 The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede).
@@
 The other payoffs for the inside actors are also modeled as state variables.
-However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, e.g.
-$L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status.
+However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, to be precise $L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status.
 The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.
 
-Both because of the presence of absorbing states and these state variables whose values depend on their values in the previous period, this is a dynamic game rather than a standard repeated game.
+Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.
 
 Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
-This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} U_i^t \de^{t-1}$.
-
-In each period, play proceeds as follows:
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
+
+In each period $t$, play proceeds as follows:
 
@@
 
-\item $p$ chooses a level of resource expenditure $R_p^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
-
-\item $c$ chooses a level of resource expenditure $R_c^t\in \mathbb{R}^+$ in period $t$ and invests it to alter the payoffs of players $s$ and/or $g$.
+\item $p$ chooses an investment level $R_p^t\in \mathbb{R}^+$.
+
+\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$.
 
@@
   $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	Fight & $\Omega^t$ &$\Omega^t$&$W_g^t, L_s^t$ \\ \hline
-	Status Quo& $\Omega^t$& $Q_g^t, Q_s^t$ &$W_g^t, L_s^t$ \\ \hline
-	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$Q_g^t, Q_s^t$ \\ \hline
+	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
+	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
+	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
  \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
@@
 
-If both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
-In this case neither player has demonstrated a willingness to give up more than the other.
-Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}
-
 There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously.
@@
 
+Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded.
 If both $s$ and $g$ play Status Quo, then the status quo persists.
+Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
+In this case neither player has demonstrated a willingness to give up more than the other.
+Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}
 
@@
 At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t$.
+For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t - R_c^t$.
 We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like.

### 2016-06-01 draft.tex
rearrange p/c and s/g payoffs to get them closer together

@@
 The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
-There are an infinite number of discrete periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.
-
-Within each period, there is first an investment stage in which first the patron and then the international community each make decisions about how much to invest in altering the payoffs of the `inside actors'---the secessionists and the home government.
-Then, the inside actors play a ``conflict stage game'' in which they simultaneously choose whether to cede the issue of status, fight, or maintain the status quo.
-
-The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the evolving value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively.
-The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede).
-Otherwise, the game ends in an absorbing state as described below.
-
-The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.
-The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
-As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
-Therefore, the payoffs to the secessionists from the status quo evolve according to the law of motion $Q_s^{t} = Q_s^{t-1} - \mu + R_p^t - R_c^t$ where $R_p^t$ is the investment of the patron and $R_c^t$ is the investment of the international community in period $t$.\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions: the patron will want to increase the value of the status quo to the secessionists, while the international community could only desire to reduce this value.}
-
-The other payoffs for the inside actors are also modeled as state variables.
-However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, to be precise $L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status.
-The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.
-
-Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.
-
-Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
-This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
-
-In each period $t$, play proceeds as follows:
+There are an infinite number of discrete periods $t=1,2,\ldots$.
+Play proceeds in each period $t$ as follows until an absorbing state is reached.
 
@@
 
-\item $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
+\item Conflict Stage Game: $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
 \end{enumerate}
 
-In order to fully describe the structure of the game, we first introduce the stage game payoffs.
+The payoffs at the end of a period are determined by these actions and six state variables.
+These state variables keep track of the value of the status quo ($Q_s^t$ and $Q_g^t$), losing ($L_s^t$ and $L_g^t$) and winning the issue of status ($W_s^t$ and $W_g^t$) for the secessionists and government respectively.
+These values evolve as the patron and player $c$ make investments and the status quo payoffs degrade over time.
+
+The game continues into the next period if the `inside players'---the secessionists and government---selected (status quo, status quo) or (cede, cede).
+Otherwise, the game ends in an absorbing state as described in Section~\ref{sec:stage}.
+Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.
+
+Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
+This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
+
+In order to fully describe the structure of the game, we next introduce the payoffs.
+
+\subsubsection{Payoffs for the Patron and International Community}
+
+At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game according to Equation~\ref{eqn:motion} below.
+$B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
+
+We assume player $c$ prefers peace to war.
+For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war.
+This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.
+
+We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
+$X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
+Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
+
+Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
+Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
+
+%The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources.
+As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.
+
+%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.
+If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.
+Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.
 
@@
 
-\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
+\noindent Stage game payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
 \vspace{2pt}
@@
 
+All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
+To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
+	\begin{equation}
+	\begin{aligned}
+		L_s^t = L_s^{t-1} - R_p^t + R_c^t \hskip.5in L_g^t = L_g^{t-1} + R_p^t - R_c^t \\
+		W_s^t = W_s^{t-1} + R_p^t - R_c^t \hskip.5in W_g^t = W_g^{t-1} - R_p^t + R_c^t \\
+		Q_g^t = Q_g^{t-1} - R_p^t + R_c^t \hskip.5in Q_s^t = Q_s^{t-1} - \mu + R_p^t - R_c^t
+		\label{eqn:motion}
+	\end{aligned}
+	\end{equation}
+
+The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.
+The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
+As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
+
+Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded.
+If both $s$ and $g$ play Status Quo, then the status quo persists.
+Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
+In this case neither player has demonstrated a willingness to give up more than the other.
+Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}
+
 If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e.
@@
 Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.
-
-Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded.
-If both $s$ and $g$ play Status Quo, then the status quo persists.
-Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
-In this case neither player has demonstrated a willingness to give up more than the other.
-Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}
-
-\subsubsection{Payoffs for the Patron and International Community}
-
-At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game.
-For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t - R_c^t$.
-We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like.
-$B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
-
-We assume player $c$ prefers peace to war.
-For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war.
-This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.
-
-We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
-$X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
-Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
-
-Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
-Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
-
-The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources.
-As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.
-
-%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.
-If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.
-Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.
 

### 2016-07-14 draft.tex
rename section 2.2
move common knowledge statement to end of sec 2.2

@@
 
-\subsection{Interaction within a Period}
+\subsection{Details of the Dynamic Game}
 \label{sec:structure}
@@
 
-The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
-There are an infinite number of discrete periods $t=1,2,\ldots$.
-Play proceeds in each period $t$ as follows until an absorbing state is reached.
+There are an infinite number of discrete periods $t=1,2,\ldots$. Play proceeds in each period $t$ as follows (and as shown in Figure 1) until an absorbing state is reached.
 
@@
 
-In order to fully describe the structure of the game, we next introduce the payoffs.
+The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
+We turn next to describing the payoffs.
 

### 2016-07-14 draft.tex
change titles of sections 2.2.1 and 2.2.2

@@
 
-\subsubsection{Payoffs for the Patron and International Community}
+\subsubsection{Within-period Payoffs for the Patron and International Community}
 
@@
 
-\subsubsection{Stage Game Payoffs}
+\subsubsection{Within-period Payoffs for the Secessionists and Government}
 \label{sec:stage}
 
-\noindent Stage game payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
+\noindent Payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
 \vspace{2pt}

### 2016-07-15 draft.tex
remove reduction by mu from list of period actions
revise rest of timeline and following paragraph; remove first paragraph of sec 2.2.1

@@
 \begin{enumerate} 
-\item The status quo payoffs for $s$ are reduced by $\mu$.
-
-\item $p$ chooses an investment level $R_p^t\in \mathbb{R}^+$.
-
-\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$.
-
-\item Conflict Stage Game: $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
+\item $p$ chooses an investment level $R_p^t\in \left[0,B_p^t\right]$ to influence the payoffs of $s$ and $g$, where $B_p^t$ is an exogenous limit on the finances of the patron.
+
+\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$ to influence the payoffs of $s$ and $g$.\footnote{We assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.}
+
+\item Conflict Stage Game: $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: Fight, Status Quo, Cede. 
 \end{enumerate}
@@
 These state variables keep track of the value of the status quo ($Q_s^t$ and $Q_g^t$), losing ($L_s^t$ and $L_g^t$) and winning the issue of status ($W_s^t$ and $W_g^t$) for the secessionists and government respectively.
-These values evolve as the patron and player $c$ make investments and the status quo payoffs degrade over time.
+The transitions of the state variables are described below.
 
@@
 \subsubsection{Within-period Payoffs for the Patron and International Community}
-
-At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game according to Equation~\ref{eqn:motion} below.
-$B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.
-
 We assume player $c$ prefers peace to war.

### 2016-07-15 draft.tex
footnote about initial conditions in section 2

@@
 
-All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
-To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
+All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.\footnote{We will see below that restrictions on the initial values of the state variables are essential for characterizing our main result, but in general we leave them unrestricted.} To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
 	\begin{equation}

### 2016-07-15 draft.tex
reverse sections 2.2.1 and 2.2.2 to make payoffs to outside actors clear
add footnote about outside actors' payoffs in absorbing states

@@
 
-\subsubsection{Within-period Payoffs for the Patron and International Community}
-We assume player $c$ prefers peace to war.
-For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war.
-This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.
-
-We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
-$X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
-Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
-
-Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
-Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
-
-%The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources.
-As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.
-
-%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.
-If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.
-Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.
-
 \subsubsection{Within-period Payoffs for the Secessionists and Government}
@@
 
+\subsubsection{Within-period Payoffs for the Patron and International Community}
+We assume player $c$ prefers peace to war.
+For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war.
+This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.
+
+We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
+$X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
+Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
+
+Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.\footnote{Note that, in equilibrium, after an absorbing state has been reached, neither outside player will make any investments.
+Thus the relevant per period payoff will be repeated with $R_p^i = R_c^i = 0$ for $i > t$ if an absorbing state is reached at time $t$.} Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
+
+%The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources.
+As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.
+
+%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.
+If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.
+Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.
+
 \subsubsection{The Low Payoffs from Ceding}

### 2016-07-15 draft.tex
change to 1.25 margins
switch to endnotes
make reference citations in text conform to JLEO style sheet

@@
 \usepackage{amsmath}
-\usepackage[hmargin=3.0cm,vmargin=3.0cm]{geometry}
+\usepackage[hmargin=3.25cm,vmargin=3.0cm]{geometry}
 %\usepackage{geometry}
@@
 \pdfoptionpdfminorversion=6 %gets rid of warning about pdf version
+
+\let\footnote=\endnote
 
@@
 Non-recognition locks unrecognized states out of a states-only club whose members enjoy benefits in terms of both security and economic integration.
-Unrecognized states do not benefit from the norm of territorial integrity that reduces the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Non-recognition also prevents entry into multilateral and bilateral trade agreements, dramatically reduces access to foreign aid, and excludes their territory from international legal frameworks, making it harder to secure foreign investment (Milhalkanian 2004; Caspersen 2013: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King 2001).
-However, as Moldova deepens ties with the EU, Transnistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu 2011).} %In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).
+Unrecognized states do not benefit from the norm of territorial integrity that reduces the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Non-recognition also prevents entry into multilateral and bilateral trade agreements, dramatically reduces access to foreign aid, and excludes their territory from international legal frameworks, making it harder to secure foreign investment (Milhalkanian, 2004; Caspersen, 2013: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King, 2001).
+However, as Moldova deepens ties with the EU, Transnistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu, 2011).} %In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).
 
@@
 The first literature to address unrecognized states directly was grounded in comparative politics, and a robust area-studies literature exists around each of the current cases of unrecognized statehood.\footnote{Lynch (2004), King (2001), Stanislawski (2008), and Bakke et al.
-(2013) provide notable treatments of the Former Soviet cases, and a pair of edited volumes by Bahcheli, Bartmann, and Srebrnik (2004) and Kingston and Spears (2004) each compile broader sets of case studies.} More recent literature has addressed wider ranges of cases and made important conceptual progress identifying patterns and commonalities across cases (e.g.
-Pegg 1998; Kolst\o \ 2006; Geldenhuys 2009; Caspersen and Stansfield 2011; Caspersen 2013).
+(2013) provide notable treatments of the Former Soviet cases, and a pair of edited volumes by Bahcheli et al.
+(2004) and Kingston and Spears (2004) each compile broader sets of case studies.} More recent literature has addressed wider ranges of cases and made important conceptual progress identifying patterns and commonalities across cases (e.g.
+Pegg, 1998; Kolst\o, \ 2006; Geldenhuys, 2009; Caspersen and Stansfield, 2011; Caspersen 2013).
 Bakke (2011) summarizes four explanations for the persistence of unrecognized states, each of which plays a role in the theory we develop here.
 First, economic and political elites benefit from the status quo (e.g.
-King 2001).
+King, 2001).
 Second, the secessionists have sufficient military capacity to make military reconquest costly (e.g.
-Kolst\o \ 2006).
+Kolst\o, \ 2006).
 Third, secessionists enjoy external support from a patron (e.g.
-Lynch 2002; Stanislawski 2008).
+Lynch, 2002; Stanislawski, 2008).
 Lastly, the governments of unrecognized states are internally legitimate, enjoying support from residents of the unrecognized state (e.g.
-Lynch 2002, Kolst\o \ 2006).
+Lynch, 2002, Kolst\o, \ 2006).
 
@@
 Grossman (2004) offers an early model of territorial disputes in which the divisibility of the contested territory is an important parameter.
-In Hirshleifer, Boldrin and Levin (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
+In Hirshleifer et al., (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
 In our context, the contested territory is indivisible; this implies that peace would require very large concessions so that this pathway of using small compromises to avoid conflict is not available.
@@
 
-Another strand of the literature explicitly models the bargaining process between potential adversaries, either with asymmetric (Yared 2010) or complete information (Jackson and Morelli 2007, Schwartz and Sonin 2008, Bevi\'{a} and Corch\'{o}n 2010, McBride and Skaperdas 2014).
-Acemoglu, Golosov, Tsyvinski and Yared (2012) develop a dynamic model of resource wars featuring limited commitment.
+Another strand of the literature explicitly models the bargaining process between potential adversaries, either with asymmetric (Yared, 2010) or complete information (Jackson and Morelli, 2007; Schwartz and Sonin, 2008; Bevi\'{a} and Corch\'{o}n, 2010; McBride and Skaperdas, 2014).
+Acemoglu et al.
+(2012) develop a dynamic model of resource wars featuring limited commitment.
 We abstract from questions of resource allocation and bargaining in order to focus on the dynamics created by the involvement of outside players: in the case of unrecognized states, the international community and patron states.
@@
 One innovation of our model is the introduction of international actors.
-While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis 2000; Regan 2002; Balch-Lindsay, Enterline, and Joyce 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter 1997, 2002).
+While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis, 2000; Regan 2002; Balch-Lindsay et al., 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter, 1997; 2002).
 
@@
 We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium.
-Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan, Gleditsch and Cunningham 2012), e.g.
+Patrons choose to contribute resources to secessionists for one or more of several reasons: 1) As an efficient mechanism for imposing costs on the home state (Salehyan et al., 2012), e.g.
 as Russia does to Georgia via South Ossetia and Abkhazia; 2) ethnic solidarity with the secessionists (e.g.
@@
 International norms against irredentism are strong, and the costs of annexing an unrecognized state appear to be prohibitively large in most cases (e.g.
-Zacher 2001).
+Zacher, 2001).
 Prior to Russia's annexation of Crimea, all post-WWII cases of annexation involved the annexation of colonial territory, rather than parts of the metropole.
@@
 Second, many of the payments that could be offered are not credible (e.g.
-Schultz 2010).
+Schultz, 2010).
 
 The difficulty of making credible payments in exchange for status is one clearly demonstrated in the civil war literature (e.g.
-Licklider 1995; Walter 1997, 2002; Fearon and Laitin 2007; Doyle and Sambanis 2006).
-Unrecognized states generally constitute ``sons of the soil" conflicts in which the central government cannot credibly commit to preserving the local demographic and political dominance of the secessionist elite once the disputed territory reverts to central government control (Weimer 1978; Fearon 2004).
+Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007).
+Unrecognized states generally constitute ``sons of the soil" conflicts in which the central government cannot credibly commit to preserving the local demographic and political dominance of the secessionist elite once the disputed territory reverts to central government control (Weimer, 1978; Fearon, 2004).
 While the central government might initially grant the secessionist elite a high level of autonomy in exchange for agreeing to reunification, the level of autonomy is likely to decrease over time, perhaps quite quickly.
@@
 \subsection{If the Patron Withdraws Support}
-Support by a foreign patron is, in almost all cases, necessary for the persistence of unrecognized statehood (Kolst\o \ 2006; Caspersen 2013).
+Support by a foreign patron is, in almost all cases, necessary for the persistence of unrecognized statehood (Kolst\o, \ 2006; Caspersen, 2013).
 When there is no patron, or when the patron withdraws its support, military reconquest by the home state is likely.
@@
 Empirically, we observe a number of cases in which $c$ applies pressure directly to encourage the patron to withdraw support from the unrecognized state.
-In an extreme example involving both sanctions and direct military confrontation, NATO coerced Serbia into, among other things, withdrawing its support from Republika Srpska and Republika Srpska Krajina, both of which were then reconquered militarily.\footnote{For an excellent discussion of the case of Republika Srpska, see Zahar 2004.} %In the final section of the paper, we consider direct coercion of the patron among the strategies available to player $c$.
+In an extreme example involving both sanctions and direct military confrontation, NATO coerced Serbia into, among other things, withdrawing its support from Republika Srpska and Republika Srpska Krajina, both of which were then reconquered militarily.\footnote{For an excellent discussion of the case of Republika Srpska, see Zahar, 2004.} %In the final section of the paper, we consider direct coercion of the patron among the strategies available to player $c$.
 
@@
 
-Serving as a third-party guarantor of autonomy rights is a way for player $c$ to potentially overcome problems of indivisibility and commitment and help the parties reach a credible compromise on status (Walter 2002).
+Serving as a third-party guarantor of autonomy rights is a way for player $c$ to potentially overcome problems of indivisibility and commitment and help the parties reach a credible compromise on status (Walter, 2002).
 However, this strategy is only tenable when the only impediment to settlement is the unenforceability of a bargain, and when $c$ is credible as an enforcer of that bargain.
@@
 
+\newpage
+
+\theendnotes
+
 \end{document}

### 2016-07-16 draft.tex
update references to JLEO style guide

@@
 
-\item \hangindent=1cm Acemoglu, Daron, Golosov, Mikhail, Tsyvinski, Aleh, and Pierre Yared. 2012. "A Dynamic Theory of Resource Wars." 
-\emph{The Quarterly Journal of Economics} 127: 283-331.
-
-\item \hangindent=1cm Bahcheli, Tozun, Barry Martmann, and Henry Srebrnik, eds. 2004 "De Facto States: The Quest for Sovereignty." New York, NY: Routledge.
+\item \hangindent=1cm Acemoglu, Daron, Mikhail Golosov, Aleh Tsyvinski, and Pierre Yared. 2012. "A Dynamic Theory of Resource Wars," 
+127 \emph{The Quarterly Journal of Economics} 283-331.
+
+\item \hangindent=1cm Bahcheli, Tozun, Barry Martmann, and Henry Srebrnik. 2004. \emph{De Facto States: The Quest for Sovereignty.} New York, NY: Routledge.
 
 \item \hangindent=1cm Bakke, Kristin M.
-2011 "After the war ends: violence and vialibity of post-Soviet unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
-\emph{Unrecognized States in the International System}.
+2011 "After the War Ends: Violence and Vialibity of Post-Soviet Unrecognized States," in Nina Caspersen and Gareth Stansfield, eds., \emph{Unrecognized States in the International System}.
 New York, NY: Routledge.
@@
 "Convincing State-Builders?
-Disaggregating Internal Legitimacy in Abkhazia."  \emph{International Studies Quarterly.}
+Disaggregating Internal Legitimacy in Abkhazia," 58 \emph{International Studies Quarterly} 591-607.
 
@@
 
-\item \hangindent=1cm Balch-Lindsay, Dylan, Andrew J. Enterline, and Kyle A. Joyce.  2008.  "Third-Party Intervention and the Civil War Process."  \emph{Journal of Peace Research} 45: 345-63.
-
-\item \hangindent=1cm Baliga, Sandeep, and Tomas Sjostrom. 2012. "The Strategy of Manipulating Conflict." \emph{The American Economic Review} 102 (6): 2897-2922.
-
-\item \hangindent=1cm Bevi\'{a}, Carmen and Luis C. Corch\'{o}n. 2010. ``Peace Agreements without Commitment.'' \emph{Games and Economic Behavior} 68 (2): 469-487.
-
-\item \hangindent=1cm Bueno de Mesquita, Ethan. 2013. ``Rebel tactics.'' \emph{Journal of Political Economy} 121 (2): 323-357.
-
-\item \hangindent=1cm Caspersen, Nina. 2013. "Unrecognized States: The Struggle for Sovereignty in the Modern International System." Malden, MA: Polity Press.
-
-\item \hangindent=1cm Caspersen, Nina, and Gareth Stansfield. 2011. "Unrecognized States in the International System." New York, NY: Routledge.
+\item \hangindent=1cm Balch-Lindsay, Dylan, Andrew J. Enterline, and Kyle A. Joyce. 2008. "Third-Party Intervention and the Civil War Process," 45 \emph{Journal of Peace Research} 345-63.
+
+\item \hangindent=1cm Baliga, Sandeep, and Tomas Sjostrom. 2012. "The Strategy of Manipulating Conflict," 102 \emph{The American Economic Review} 2897-2922.
+
+\item \hangindent=1cm Bevi\'{a}, Carmen and Luis C. Corch\'{o}n. 2010. ``Peace Agreements without Commitment,'' 68 \emph{Games and Economic Behavior} 469-87.
+
+\item \hangindent=1cm Bueno de Mesquita, Ethan. 2013. ``Rebel tactics,'' 121 \emph{Journal of Political Economy} 323-57.
+
+\item \hangindent=1cm Caspersen, Nina. 2013. \emph{Unrecognized States: The Struggle for Sovereignty in the Modern International System.} Malden, MA: Polity Press.
+
+\item \hangindent=1cm Caspersen, Nina, and Gareth Stansfield. 2011. \emph{Unrecognized States in the International System.} New York, NY: Routledge.
 
@@
 
-\item \hangindent=1cm Coggins, Bridget. 2011. "Friends in High Places: International Politics and the Emergence of States from Secessionism." \emph {International Organization} 65 (03): 433-467.
+\item \hangindent=1cm Coggins, Bridget. 2011. "Friends in High Places: International Politics and the Emergence of States from Secessionism," 65 \emph {International Organization} 433-67.
 
@@
 
-\item \hangindent=1cm Crawford, James. 2006. "The Creation of States in International Law." Oxford: Oxford University Press.
+\item \hangindent=1cm Crawford, James. 2006. \emph{The Creation of States in International Law.} Oxford: Oxford University Press.
 
@@
 
-\item \hangindent=1cm Doyle, Michael W., and Nicholas Sambanis. 2006. "Making War and Building Peace: United Nations Peace Operations." Princeton: Princeton University Press.
-
-\item \hangindent=1cm Elbadawi, Ibrahim, and Nicholas Sambanis. 2000. "External Interventions and the Duration of Civil Wars." 2433.
+\item \hangindent=1cm Doyle, Michael W., and Nicholas Sambanis. 2006. \emph{Making War and Building Peace: United Nations Peace Operations.} Princeton, NJ: Princeton University Press.
+
+\item \hangindent=1cm Elbadawi, Ibrahim, and Nicholas Sambanis. 2000. "External Interventions and the Duration of Civil Wars." World Bank Policy Research Working Paper 2433.
 
@@
 
-\item \hangindent=1cm Fearon, James D.  1994.  "Domestic Political Audiences and the Escalation of International Disputes."  \emph{The American Political Science Review} 88: 577-92.
-
-\item \hangindent=1cm Fearon, James D., and David D. Laitin. 2003. "Ethnicity, Insurgency, and Civil War." \emph {American Political Science Review} 97 (01): 75-90.
-
-\item \hangindent=1cm Fearon, James D. 2004. "Why Do Some Civil Wars Last So Much Longer than Others?" \emph {Journal of Peace Research} 41 (3): 275-301.
-
-\item \hangindent=1cm Geldenhuys, Deon. 2009. "Contested States in World Politics." Basingstoke: Palgrave McMillan.
+\item \hangindent=1cm Fearon, James D. 1994. "Domestic Political Audiences and the Escalation of International Disputes," 88 \emph{The American Political Science Review} 577-92.
+
+\item \hangindent=1cm Fearon, James D., and David D. Laitin. 2003. "Ethnicity, Insurgency, and Civil War," 97 \emph {American Political Science Review} 75-90.
+
+\item \hangindent=1cm Fearon, James D. 2004. "Why Do Some Civil Wars Last So Much Longer than Others?" 41 \emph {Journal of Peace Research} 275-301.
+
+\item \hangindent=1cm Geldenhuys, Deon. 2009. \emph{Contested States in World Politics.} Basingstoke: Palgrave McMillan.
 
@@
 
-\item \hangindent=1cm Grossman, Herschel I. 2004. ''Peace and war in territorial disputes.'' No. w10601. National Bureau of Economic Research.
-
-\item \hangindent=1cm Hirshleifer, Jack, Boldrin, Michele, and David K Levine. 2009. ''The slippery slope of concession.'' \emph{Economic Inquiry} 47 (2): 197-205.
-
-\item \hangindent=1cm Jackson, Matthew O., and Massimo Morelli. 2007. "Political bias and war." \emph{The American Economic Review} 97 (4): 1353-1373.
-
-\item \hangindent=1cm King, Charles. 2001. "The Benefits of Ethnic War: Understanding Eurasia's Unrecognized States." \emph{World Politics} 53 (4): 524-552.
-
-\item \hangindent=1cm Kingston, Paul, and Ian S. Spears, eds. 2004. "States-Within-States." New York, NY: Palgrave McMillan.
-
-\item \hangindent=1cm Kolst\o, P{\aa}l. 2006. "The Sustainability and Future of Unrecognized Quasi-States." \emph {Journal of Peace Research} 43 (6): 723-740.
-
-\item \hangindent=1cm Licklider, Roy. 1995. "The Consequences of Negotiated Settlements in Civil Wars, 1945-1993." \emph {American Political Science Review} (89): 681-90.
+\item \hangindent=1cm Grossman, Herschel I. 2004. ''Peace and war in territorial disputes.'' No. w10601. National Bureau of Economic Research. Working Paper
+
+\item \hangindent=1cm Hirshleifer, Jack, Michele Boldrin, and David K Levine. 2009. ''The slippery slope of concession,'' 47 \emph{Economic Inquiry} 197-205.
+
+\item \hangindent=1cm Jackson, Matthew O., and Massimo Morelli. 2007. "Political bias and war," 97 \emph{The American Economic Review} 1353-73.
+
+\item \hangindent=1cm King, Charles. 2001. "The Benefits of Ethnic War: Understanding Eurasia's Unrecognized States," 53 \emph{World Politics} 524-552.
+
+\item \hangindent=1cm Kingston, Paul, and Ian S. Spears. 2004. \emph{States-Within-States.} New York, NY: Palgrave McMillan.
+
+\item \hangindent=1cm Kolst\o, P{\aa}l. 2006. "The Sustainability and Future of Unrecognized Quasi-States," 43 \emph {Journal of Peace Research} 723-40.
+
+\item \hangindent=1cm Licklider, Roy. 1995. "The Consequences of Negotiated Settlements in Civil Wars, 1945-1993," 89 \emph {American Political Science Review} 681-90.
 
@@
 
-\item \hangindent=1cm Lynch, Dov. 2004. "Engaging Eurasia's Separatist States." Washington, DC: U.S. Institute of Peace.
-
-\item \hangindent=1cm Mailath, George J., and Larry Samuelson. 2006. "Repeated Games and Reputation." New York, NY: Oxford University Press.
-
-\item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, settlement, and the shadow of the future." \emph{Journal of Economic Behavior $\&$ Organization} 105: 75-89.
+\item \hangindent=1cm Lynch, Dov. 2004. \emph{Engaging Eurasia's Separatist States.} Washington, DC: U.S. Institute of Peace.
+
+\item \hangindent=1cm Mailath, George J., and Larry Samuelson. 2006. \emph{Repeated Games and Reputation.} New York, NY: Oxford University Press.
+
+\item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, Settlement, and the Shadow of the Future," 105 \emph{Journal of Economic Behavior $\&$ Organization} 75-89.
 
@@
 2011.
-"The foreign policies of unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
-\emph{Unrecognized States in the International System}.
+"The Foreign Policies of Unrecognized States," in: Nina Caspersen and Gareth Stansfield, eds., \emph{Unrecognized States in the International System}.
 New York, NY: Routledge.
 
-\item \hangindent=1cm Pegg, Scott. 1998. "International society and the de facto state." New York: Ashgate.
+\item \hangindent=1cm Pegg, Scott. 1998. \emph{International Society and the de Facto State.} New York, NY: Ashgate.
 
@@
 
-\item \hangindent=1cm Regan, P.M. 2002. "Civil wars and foreign powers: Outside intervention in intrastate conflict." Ann Arbor: University of Michigan Press.
-
-\item \hangindent=1cm Roeder, Philip G. 2007. "Where nation-states come from: institutional change in the age of nationalism." Princeton: Princeton University Press.
-
-\item \hangindent=1cm Salehyan, Idean, Kristian Skrede Gleditsch, and David E.
-Cunningham.
-2011.
-"Explaining External Support for Insurgent Groups." \emph {International Organization} 65 (04): 709-744.
-
-\item \hangindent=1cm Schwarz, Michael, and Konstantin Sonin. 2008. "A theory of brinkmanship, conflicts, and commitments." \emph{Journal of Law, Economics, and Organization} 24 (1): 163-183.
-
-\item \hangindent=1cm Schultz, Kenneth A.
-2010.
-"The Enforcement Problem in Coercive Bargaining: Interstate Conflict over Rebel Support in Civil Wars." \emph {International Organization} 64 (02): 281-312.
+\item \hangindent=1cm Regan, P.M. 2002. \emph{Civil Wars and Foreign Powers: Outside Intervention in Intrastate Conflict.} Ann Arbor, MI: University of Michigan Press.
+
+\item \hangindent=1cm Roeder, Philip G. 2007. \emph{Where Nation-states Come From: Institutional Change in the Age of Nationalism.} Princeton, NJ: Princeton University Press.
+
+\item \hangindent=1cm Salehyan, Idean, Kristian Skrede Gleditsch, and David E. Cunningham. 2011. "Explaining External Support for Insurgent Groups," 65 \emph {International Organization} 709-44.
+
+\item \hangindent=1cm Schwarz, Michael, and Konstantin Sonin. 2008. "A Theory of Brinkmanship, Conflicts, and Commitments." 24 \emph{Journal of Law, Economics, and Organization} 163-183.
+
+\item \hangindent=1cm Schultz, Kenneth A. 2010. "The Enforcement Problem in Coercive Bargaining: Interstate Conflict over Rebel Support in Civil Wars," 64 \emph{International Organization} 281-312.
 
@@
 2008.
-"Para-States, Quasi-States, and Black Spots: Perhaps Not States, But Not `Ungoverned Territories,' Either." \emph {International Studies Review} 10 (2): 366-396.
+"Para-States, Quasi-States, and Black Spots: Perhaps Not States, But Not `Ungoverned Territories,' Either," 10 \emph{International Studies Review} 366-396.
 
@@
 
-\item \hangindent=1cm van Houten, Pieter.
-1998.
-"The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives Europennes de Sociologie} 39 (01): 110-146.
-
-\item \hangindent=1cm Walter, Barbara F. 1997. "The Critical Barrier to Civil War Settlement." \emph {International Organization} 51 (03): 335-364.
-
-\item \hangindent=1cm Walter, Barbara F.. 2002. "Committing to Peace: The successful settlement of civil wars." Princeton: Princeton University Press.
+\item \hangindent=1cm van Houten, Pieter. 1998. "The Role of a Minority's Reference State in Ethnic Relations." 39 \emph {European Journal of Sociology / Archives Europennes de Sociologie} 110-146.
+
+\item \hangindent=1cm Walter, Barbara F. 1997. "The Critical Barrier to Civil War Settlement," 51 \emph {International Organization} 335-64.
+
+\item \hangindent=1cm ----. 2002. \emph{Committing to Peace: The Successful Settlement of Civil Wars.} Princeton, NJ: Princeton University Press.
 
@@
 
-\item \hangindent=1cm Yared, Pierre. 2010. ``A Dynamic Theory of War and Peace.'' \emph{Journal of Economic Theory} 145: 1921-1950.
-
-\item \hangindent=1cm Zacher, Mark W. 2001. "The Territorial Integrity Norm: International Boundaries and the Use of Force." \emph {International Organization} 55 (02): 215-250.
+\item \hangindent=1cm Yared, Pierre. 2010. ``A Dynamic Theory of War and Peace,'' 145 \emph{Journal of Economic Theory} 1921-50.
+
+\item \hangindent=1cm Zacher, Mark W. 2001. "The Territorial Integrity Norm: International Boundaries and the Use of Force," 55 \emph {International Organization} 215-50.
 

### 2016-09-17 midwest_04_12_15.tex
add midwest_04_12_15.tex, last presentation of the project

\documentclass{beamer} 
%\documentclass[handout]{beamer} 
\usetheme{Ilmenau}
\usepackage{graphicx,verbatim,hyperref}
\usepackage{textpos}

\usecolortheme{beaver}
\useinnertheme{default}
\setbeamertemplate{itemize item}[triangle]
\setbeamertemplate{itemize subitem}[triangle]
\setbeamertemplate{itemize subsubitem}[circle]
\setbeamertemplate{enumerate items}[default]
\setbeamertemplate{blocks}[upper=block head,rounded]
\setbeamercolor{item}{fg=black}
\usefonttheme{serif} %should allow ccfonts to take effect

\usepackage{cite}
\usepackage{times, verbatim,xcolor,bm}
\usepackage{amsbsy,amssymb, amsmath, amsthm}
\usepackage{booktabs}
%David miller's fonts
	\usepackage[T1]{fontenc}
	\usepackage[boldsans]{ccfonts}
	\usepackage[euler-hat-accent]{eulervm}

\newcommand{\al}{\alpha}
\newcommand{\expect}{\mathbb{E}}
\newcommand{\Bt}{B(\bm{\tau^a})}
\newcommand{\bta}{\bm{\tau^a}}
\newcommand{\btn}{\bm{\tau^{tw}}}
\newcommand{\ga}{\gamma}
\newcommand{\ve}{\varepsilon}
\newcommand{\ta}{\theta}
\newcommand{\de}{\delta}
\newcommand{\ov}{\overline}

\newenvironment{changemargin}[2]{% 
  \begin{list}{}{% 
    \setlength{\topsep}{0pt}% 
    \setlength{\leftmargin}{#1}% 
    \setlength{\rightmargin}{#2}% 
    \setlength{\listparindent}{\parindent}% 
    \setlength{\itemindent}{\parindent}% 
    \setlength{\parsep}{\parskip}% 
  }% 
  \item[]}{\end{list}}

\begin{document}
\title[Unrecognized States: A Theory of Self-Determination and Foreign Influence\hspace{1.25in}\insertframenumber/\inserttotalframenumber]{Unrecognized States: A Theory of Self-Determination and Foreign Influence}
\author[Kristy Buzard, Benjamin Graham and Ben Horne]{Kristy Buzard, Benjamin Graham and Ben Horne\\ kbuzard@syr.edu}
\date{April 12, 2015}
\maketitle
%\insertpresentationendpage removed b/c of appendix

\section{Overview}
\subsection{Preview}
\begin{frame}
\frametitle{Unrecognized States}
Unrecognized states (URS): control and govern territory, seek recognition
\pause
\begin{itemize}[<+->]
  \item Six current URS more than 20 yrs old (+ Eastern Ukraine)
	\item No explanation in the literature for how this can be a stable outcome
\end{itemize}
\end{frame}

\begin{frame}{What we do}

\pause
We demonstrate unrecognized statehood can be a ``status quo'' outcome
\pause
\begin{itemize}[<+->]
	\item (SPN) Equilibrium in a repeated game
	\item Four players
		\begin{itemize}[<+->]
			\item Home government
			\item Secessionist elite
			\item Patron state
			\item International community
		\end{itemize}
	\item State variable: Status Quo (SQ) payoffs for secessionists
\end{itemize}

\end{frame}

\begin{frame}{The General Idea}

\pause
In each period: secessionists and gov't each choose (simultaneously) among $\left\{\text{Fight, Status Quo, Cede}\right\}$
\pause
		\begin{itemize}
			\item We need SQ to be a stage game best response for both secessionists and gov't
		\end{itemize}

\pause
\vskip.2in
We add some more realistic elements:
\pause
		\begin{enumerate}[<+->]
			\item Unrecognized status reduces SQ payoffs of secessionists each period
			\item Patron and int'l community can make investments in both actors' payoffs
		\end{enumerate}

\end{frame}

\begin{frame}{Outline of Talk}
\pause
\begin{enumerate}[<+->]
	\item Model
	\item Status quo equilibrium
	\item What we can say about sanctions
	\item Conclusion
\end{enumerate}
\end{frame}

\section{Model}
\subsection{Economic and Political Structure}
\begin{frame}{Timeline}
\pause
\begin{figure}
\includegraphics[width=\linewidth,height=\textheight,keepaspectratio]{Timeline2.pdf}
\caption{Timeline}
\end{figure}
\end{frame}

\subsection{The Players}

\begin{frame}{Outside Actors}
	\pause
International community ($c$): desires reunification \\
\pause
Patron ($p$): desires recognized independence
\pause
\begin{itemize}	
	\item Each has preferences that are aligned with one of the home state actors
	\pause
	\item Patron may most prefer SQ; our assumption stacks the deck against us
\end{itemize}

\pause
\vskip.2in
$p$'s per period payoffs: $U_{pt} = -\alpha X + \lambda Y - R_{pt}$ \\
$c$'s per period payoffs: $U_{ct} = \beta X - \nu Y - R_{ct}$
\begin{itemize}
	\item $X=1$ if the secessionists rejoin the home state \\
	\item $Y=1$ if the home state recognizes the secessionists 
\end{itemize}
\end{frame}

\begin{frame}{Home State Actors}
	\pause
Central Government of Home State ($g$): desires reunification \\
\pause
Secessionists ($s$): desire recognized independence
\pause
\begin{itemize}	
	\item Central issue of contention: recognized independence vs. reunification.
\end{itemize}

\pause
\vskip.2in
Assumptions:
\begin{itemize}[<+->]
	\item Issue of status is indivisible, highly valued by both sides
	\item Insufficient credible side payments for easy settlement
\end{itemize}
\end{frame}

\begin{frame}{Stage Game between Home Gov't $\&$ Secessionists}
\pause
	\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $\omega_{gt}, \omega_{st}$ &$\omega_{gt}, \omega_{st}$&$W_{gt}, L_{st}$ \\ \hline
	Status Quo& $\omega_{gt}, \omega_{st}$& $Q_{gt}, Q_{st}$ &$W_{gt}, L_{st}$ \\ \hline
	Cede&$L_{gt}, W_{st}$& $L_{gt}, W_{st}$ &$Q_{gt}, Q_{st}$ \\ \hline
 \multicolumn{4}{c} {Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
where $\omega_{it} \equiv p_1 \left(W_{it}-\zeta_i\right) + \left(1-p_1\right)\left( L_{it}-\zeta_i\right)$
\begin{itemize}
	\item fixed cost of war $\zeta_i$
	\item probabilities $p_1$ of victory, $1 - p_1$ of loss 
		\begin{itemize}
			\item can also handle non-decisive war
		\end{itemize}
\end{itemize}
and $Q_{st} \equiv Q_{s,t-1} -\mu + R_{pt} - R_{ct}$
\end{frame}

\section{Status Quo Equilibrium}
\subsection{}

\begin{frame}{Main Result}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Proposition 1}
  There exists an equilibrium in which the outcome is perpetual unrecognized statehood. The actions in this equilibrium are 
	\begin{enumerate}
		\item $p$ invests to create a buffer of $\frac{\beta}{1-\delta}$ between the payoffs from ceding and the status quo in the first period;
		\item $p$ invests up to $\mu$ each period thereafter to maintain the buffer;
		\item $c$ invests nothing;
		\item both $g$ and $s$ play $Status$ $Quo$ each period.
	\end{enumerate}

\end{beamerboxesrounded}
\end{frame}

\begin{frame}{Proposition 1 Continued}
\pause
  The following are sufficient conditions for the equilibrium:
	\begin{enumerate}[<+->]
		\item \textit{For both $g$ and $s$, $Q_{it} \geq L_{it} \ \forall t$.}

		\item \emph{For both $g$ and $s$, $\frac {Q_{it}}{1-\delta} \geq  -\zeta_i+\frac{W_{it}(p_1) + L_{it}(1-p_1)}{1-\delta} \ \forall t$.}

\item \textit{$\alpha \geq \beta$: reunification is more important for $p$ to avoid than for $c$ to achieve.}

\item  \textit{$\nu\cdot p_{1s} \geq \lambda \cdot p_{1s} + \mu + \beta$: recognition of $s$ is more important for $c$ to avoid than for $p$ to achieve.}

\item  \textit{$B_{p1} \geq\frac{\beta}{1-\delta} - \left(q_{s1} - l_{s1} \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}

\item \textit{$B_{pt} \geq \mu, \ \forall t>1$: the patron can afford to pay to maintain the status quo.}
	\end{enumerate}
\end{frame}

\section{Sanctions}
\begin{frame}{Sanctions (I)}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Proposition 2}
  Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s Status Quo payoffs. In order for sanctions to lead to ceding by the secessionists, the following are required:
	\pause
	\begin{enumerate}[<+->]
\item The patron must be unable or find it not worthwhile to invest the additional amount now required to maintain the status quo.

\item The patron must be unable or find it not worthwhile to invest to instigate fighting by the secessionists.

\item The continuation value from SQ must fall below that from Cede before it falls below that from Fight.
\end{enumerate}

\end{beamerboxesrounded}

\end{frame}

\begin{frame}{Sanctions (II)}
\begin{beamerboxesrounded}[upper=palette tertiary, shadow=true]{Proposition 3}
  Assume the conditions of \emph{Proposition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities. \\
	\pause
	The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
\end{beamerboxesrounded}

\end{frame}

\section{Conclusion}
\subsection{}
\begin{frame}{Conclusion}
\begin{itemize}[<+->]
	\item We present a unified four-player framework for analyzing unrecognized statehood and its alternatives
	\item We show unrecognized statehood can emerge in equilibrium even when it is a terrible outcome for all involved
		\begin{itemize}
			\item Unfortunately, it is not difficult to imagine Eastern Ukraine on our list of URS in 20 years
		\end{itemize}
	\item The model demonstrates why sanctions may be counterproductive and suggests that positive inducements may be more productive
\end{itemize}

\end{frame}

\end{document}

### 2016-09-21 presentation.tex
initial changes to presentation for PARCC conferencce

--- 2016-09-17 midwest_04_12_15.tex
@@
-\documentclass{beamer} 
-%\documentclass[handout]{beamer} 
+%\documentclass{beamer} 
+\documentclass[handout]{beamer} 
 \usetheme{Ilmenau}
@@
 \author[Kristy Buzard, Benjamin Graham and Ben Horne]{Kristy Buzard, Benjamin Graham and Ben Horne\\ kbuzard@syr.edu}
-\date{April 12, 2015}
+\date{September 23, 2016}
 \maketitle
@@
 		\begin{itemize}
-			\item We need SQ to be a stage game best response for both secessionists and gov't
+			\item We need Status Quo to be a stage game best response for both secessionists and gov't
 		\end{itemize}
@@
 		\begin{enumerate}[<+->]
-			\item Unrecognized status reduces SQ payoffs of secessionists each period
+			\item Unrecognized status reduces Status Quo payoffs of secessionists each period
 			\item Patron and int'l community can make investments in both actors' payoffs

### 2017-04-17 teaching.tex
create file for explanation of version for undergraduate teaching

--- 2015-10-31 eqm_profiles.tex
@@
 \linespread{1.1}
+
+\usepackage[pdftex,
+bookmarks=true,
+bookmarksnumbered=false,
+pdfview=fitH,
+bookmarksopen=true]{hyperref}
 
@@
 
+\newtheorem{proposition}{Proposition}
 \newcommand{\ve}{\varepsilon}
@@
 \newcommand{\expect}{\mathbb{E}}
+\newcommand{\Bt}{B(\bm{\tau^a})}
+\newcommand{\bta}{\bm{\tau^a}}
+\newcommand{\btn}{\bm{\tau^n}}
+\newcommand{\btw}{\bm{\tau^{tw}}}
+\newcommand{\ga}{\gamma}
+\newcommand{\Ga}{\Gamma}
 \newcommand{\de}{\delta}
@@
 \begin{center}
-Equilibrium Strategy Profiles (For JLEO Revision)
+  Unrecognized States: A Theory of Self-Determination and Foreign Influence \\
+	By Kristy Buzard, Benjamin A.T. Graham and Ben Horne
 \end{center}
@@
 \vskip.3in
-First, need to get timing clear for evolution of state variable. Take state variable to be $q_{st}$.
+\section{Motivation}
+
+\vskip1in
+\section{Players and Strategy Spaces}
+\begin{enumerate}
+	\item Home government (G) chooses $S_G \in \left\{SQ,C\right\}$ where SQ means 'Status Quo' of unrecognized statehood and C means 'Cede' the issue of status and recognize the secessionists as an independent state.
+	\item Secessionists (S) choose $S_S \in \left\{SQ,C\right\}$, same as for $G$ except when they Cede, they rejoin the home state.
+	\item Patron state (P) chooses $p \in [0,\infty)$ to invest in the secessionists status quo payoffs
+	\item International community (C) chooses $c \in [0,\infty)$ to invest in the secessionists payoffs from rejoining the home state
+\end{enumerate}
+
+\vskip1in
+\section{Payoffs}
+The payoffs after (SQ,SQ) are -p,-c,
+
+\vskip1in
+\section{Model}
+N
 \begin{itemize}
-	\item $q_{s1}$: beginning of the world (or, $Q_{s0} - \mu = q_{s1}$)
-	\item $Q_{s1} = q_{s1} + R_{p1} + R_{c1}$
-		\begin{itemize}
-			\item $q_{s2} = Q_{s1} - \mu$
-		\end{itemize}
+	\item 
 \end{itemize}
 
-\vskip.2in
-Strategies (this is for the most interesting state variable, but could write it out for the other five)
-\begin{itemize}
-	\item For patron: function of $q_{st}$
-	\item For c: function of $q_{st} + R_{pt}$
-	\item For s: function of $Q_{st}=q_{st} + R_{pt} + R_{ct}$
-	\item For g: not a function of $q_{st}$ at all
-\end{itemize}
-
-\vskip.5in	
-According to Mailath and Samuelson (2006) p. 177.
-\begin{itemize}
-	\item The strategy profile $\sigma$ is a stationary Markov strategy if for any two ex post histories $\tilde{h}^t$ and $\tilde{h}^\tau$ (of equal or different lengths) terminating in the same state, $\sigma\left(\tilde{h}^t\right) = \sigma\left(\tilde{h}^\tau\right)$.
-	\item The strategy profile $\sigma$ is a stationary Markov equilibrium if $\sigma$ is a stationary Markov strategy profile and a subgame-perfect equilibrium.
-\end{itemize}
-
-\vskip.5in
-We have six state variables, so $s = \left(q_s,q_g,l_s,l_g,w_s,w_g \right)$.
-\begin{itemize}
-	\item Strategies are a function of all six state variables
-		\begin{itemize}
-			\item We can restrict attention to relevant ranges of the state variables.
-CAN I CHANGE PROP 1 PARTS 1 AND 2 TO BE INITIAL VALUES OF STATE VARIABLES?
-If so, then these conditions and what is necessary in each period to make the game continue.
-		\end{itemize}
-	\item Strategies for the patron and $c$ are also how much to invest in each of the six state variables. Some can be ruled out by preference assumptions:
-		\begin{itemize}
-			\item $c$ dislikes war, so will never invest in $w_s$ or $w_g$. It would also not want to make the government lose, so won't invest in $l_g$ either.
-			\item Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $w_g$ or $l_s$.
-		\end{itemize}
-\end{itemize}
-
-\vskip.5in
-This leaves four state varibles in which the patron may invest: $q_s$, $q_g$, $l_g$ and $w_s$. Three in which $c$ may invest: $q_s$, $q_g$, $l_s$.
-\begin{itemize}
-	\item Also, Gov't / Secessionists: Choose unilateral, simultaneous best responses depending on magnitudes of $Q_{i1}$, $L_{i1}$ and $\omega_{i1}$
-		\begin{itemize}
-			\item Game only continues if (SQ,SQ) or (Cede, Cede) was played
-		\end{itemize}
-\end{itemize}
-
-\vskip.2in
-Assumptions
-\begin{itemize}
-	\item $Q_{s0} \geq L_{s0} \Rightarrow q_{s1} + \mu \geq l_{s1}$
-	\item $Q_{g0} \geq L_{g0} \Rightarrow q_{g1} \geq l_{g1}$
-\end{itemize}
-
-\vskip.2in
-Period 1
-\begin{enumerate}
-	\item Patron would like $Q_{s1} \geq L_{s1}$ to prevent $c$ from incentivizing secessionists to Cede
-		\begin{itemize}
-			\item Equivalent to $q_{s1} + R_{p1} \geq l_{s1} + R_{c1}$ / $Q_{s0} + \mu + R_{p1} \geq L_{s0} + R_{c1}$ / $R_{p1} \geq R_{c1} - \left( Q_{s0} - \mu - L_{s0}\right)$
-				\begin{itemize}
-					\item $R_{c1} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{p1} \geq \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ allows the patron to ensure the original inequality
-					\item Initial assumption means $R_{p1} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
-					\item Since patron is willing to pay up to $R_{p1} = \frac{\alpha}{1-\de}$, when $\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$, it will invest $R_{p1} = \frac{\beta}{1-\de} - \left( Q_{s0} - \mu - L_{s0}\right)$ to augment $q_{s1}$ if this is greater than 0.
-Else, $R_{p1} = 0$.
-					  \begin{enumerate}
-							\item If $R_{p1} + \left( Q_{s0} - \mu - L_{s0}\right) \geq \frac{\beta}{1-\de}$, $R_{c1} =0$. (SQ,SQ) is played and game continues.
-							\item If instead $R_{p1} + \left( Q_{s0} - \mu - L_{s0}\right) < \frac{\beta}{1-\de}$ (i.e.
-assumption doesn't hold), $R_{c1} = l_{s1} - \left(q_{s1} + R_{p1} \right) + \ve$ to augment $l_{s1}$.
-Note optimal $R_{p1}$ in this case is 0.
-								\begin{itemize}
-									\item (SQ,Cede) is played and game ends
-								\end{itemize}
-						\end{enumerate}
-				\end{itemize}
-		\end{itemize}	
-	\item Patron would like $L_{g1} > Q_{g1}$ to achieve recognition directly. Must invest to augment $l_{g1}$
-		\begin{itemize}
-			\item Equivalent to $l_{g1} + R_{p1} > q_{g1} + R_{c1}$ / $L_{g0} + R_{p1} > Q_{g0} + R_{c1}$ / $R_{p1} > R_{c1} + \left( Q_{g0} - L_{g0}\right)$
-				\begin{itemize}
-					\item $R_{c1} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{p1} > \frac{\nu}{1-\de} + \left( Q_{g0} - L_{g0}\right)$ allows the patron to ensure the original inequality
-					\item Since patron is willing to pay up to $R_{p1} = \frac{\lambda +\mu + \beta}{1-\de}$, when $\frac{\lambda +\mu + \beta}{1-\de} \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{g1}$.
-$R_{c1} = 0$ as well.
-					\item If instead assumption 4 did not hold and $R_{p1} > \frac{\nu}{1-\de} + \left( Q_{g0} - L_{g0}\right)$, it will invest this amount to augment $l_{g1}$.
-Again, $R_{c1} = 0$.
-(Cede, SQ) is played and game ends.
-				\end{itemize}
-		\end{itemize}
-	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
-If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de}$ is positive, invest in $w_{s1}$ to change $\omega_{s1}$.
-Will not invest when the quantity is non-positive.
-Inequality is $\omega_{s1} > Q_{s1}$
-		\begin{itemize}
-			\item $\omega_{s1} = -\zeta_{s}(1-\de) + W_{s1}p_{1s} + L_{s1}(1-p_{1s}) > Q_{s1}$
-			\item $-\zeta_{s}(1-\de) + (W_{s0}+R_{p1})p_{1s} + L_{s0}(1-p_{1s}) > Q_{s0} - \mu + R_{c1}$
-			\item $R_{p1}p_{1s} > R_{c1} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)$
-			\item $R_{c1} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{p1}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)$, $R_{c1}=0$ and (SQ,Fight) is played in third stage.
-Outcome depends on war lottery.
-			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
-				\begin{itemize}
-					\item So need
-						\begin{multline}
-							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s0} - \left(-\zeta_{s}(1-\de) + W_{s0}p_{1s} + L_{s0}(1-p_{1s})\right)
-						\end{multline}
-				\end{itemize}
-		\end{itemize}
-	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
-But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
-\end{enumerate}
-
-\vskip.2in
-Now generalize profiles for all $t$
-\begin{enumerate}
-	\item Patron would like $Q_{st} \geq L_{st}$ to prevent $c$ from incentivizing secessionists to Cede
-		\begin{itemize}
-			\item Equivalent to $q_{st} + R_{pt} \geq l_{st} + R_{ct}$ / $Q_{s,t-1} + \mu + R_{pt} \geq L_{s,t-1} + R_{ct}$ / $R_{pt} \geq R_{ct} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$
-				\begin{itemize}
-					\item $R_{ct} \leq \frac{\beta}{1-\de} \Rightarrow$ $R_{pt} \geq \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ allows the patron to ensure the original inequality
-					\item Assumption 1 combined with equilibrium play (equivalently, if $L_{s,t-1} > Q_{s,t-1}$, the game would have ended before the start of period $t$) means $R_{pt} \geq \frac{\beta}{1-\de} + \mu$ is a tighter condition
-					\item Since patron is willing to pay up to $R_{pt} = \frac{\alpha}{1-\de}$, when $\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$, it will invest $R_{pt} = \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ to augment $q_{st}$ if this is greater than 0.
-Else, $R_{pt} = 0$.
-					  \begin{enumerate}
-							\item If $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) \geq \frac{\beta}{1-\de}$, $R_{ct} =0$. (SQ,SQ) is played and game continues.
-							\item If instead $R_{pt} + \left( Q_{s,t-1} - \mu - L_{s,t-1}\right) < \frac{\beta}{1-\de}$ (i.e.
-assumption 3 doesn't hold), $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$ where $\ve$ is small to augment $l_{st}$.
-Note optimal $R_{pt}$ in this case is 0.
-								\begin{itemize}
-									\item (SQ,Cede) is played and game ends
-								\end{itemize}
-						\end{enumerate}
-				\end{itemize}
-		\end{itemize}	
-	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$
-		\begin{itemize}
-			\item Equivalent to $l_{gt} + R_{pt} > q_{gt} + R_{ct}$ / $L_{g,t-1} + R_{pt} > Q_{g,t-1} + R_{ct}$ / $R_{pt} > R_{ct} + \left( Q_{g,t-1} - L_{g,t-1}\right)$
-				\begin{itemize}
-					\item $R_{ct} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ allows the patron to ensure the original inequality
-					\item Since patron is willing to pay up to $R_{pt} = \frac{\lambda +\mu + \beta}{1-\de} - \mu$, when $\frac{\lambda +\mu + \beta}{1-\de} -\mu \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{gt}$.
-$R_{ct} = 0$ as well.
-					\item If instead assumption 4 did not hold and $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, patron will invest this amount to augment $l_{gt}$.
-Again, $R_{ct} = 0$.
-(Cede, SQ) is played and game ends.
-				\end{itemize}
-		\end{itemize}
-	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
-If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$ is positive, invest in $w_{st}$ to change $\omega_{st}$.
-Will not invest when the quantity is non-positive.
-Inequality is $\omega_{st} > Q_{st}$
-		\begin{itemize}
-			\item $\omega_{st} = -\zeta_{s}(1-\de) + W_{st}p_{1s} + L_{st}(1-p_{1s}) > Q_{st}$
-			\item $-\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) > Q_{s,t-1} - \mu + R_{ct}$
-			\item $R_{pt}p_{1s} > R_{ct} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$
-			\item $R_{ct} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$, $R_{ct}=0$ and (SQ,Fight) is played in third stage.
-Outcome depends on war lottery.
-			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
-				\begin{itemize}
-					\item So need
-						\begin{multline}
-							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)
-						\end{multline}
-				\end{itemize}
-		\end{itemize}
-	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
-But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
-\end{enumerate}
-
-\newpage
-Prose version of equilibrium strategy profiles: \\
-
-In period $t$, the patron invests $R_{pt} = \max\left\{\frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right),0\right\}$ to augment the status quo payoffs of the secessionists, $q_{st}$.
-Player $c$ then chooses $R_{ct} =0$ so that playing status quo is a best response for both inside players.
-(SQ,SQ) is played in the third stage and the game continues to period $t+1$.
-
-If instead $R_{pt} < \frac{\beta}{1-\de} - \left( Q_{s,t-1} - \mu - L_{s,t-1}\right)$ (i.e.
-assumption 3 doesn't hold), player $c$ will invest $R_{ct} = l_{st} - \left(q_{st} + R_{pt} \right) + \ve$, for $\ve$ small, to augment $l_{st}$.
-In this case, status quo remains a best response for the government but cede becomes the best response for the secessionists.
-The game ends with the secessionist territory being reunited.
-
-\begin{enumerate}	
-	\item Patron would like $L_{gt} > Q_{gt}$ to achieve recognition directly. Must invest to augment $l_{gt}$
-		\begin{itemize}
-			\item Equivalent to $l_{gt} + R_{pt} > q_{gt} + R_{ct}$ / $L_{g,t-1} + R_{pt} > Q_{g,t-1} + R_{ct}$ / $R_{pt} > R_{ct} + \left( Q_{g,t-1} - L_{g,t-1}\right)$
-				\begin{itemize}
-					\item $R_{ct} \leq \frac{\nu}{1-\de} \Rightarrow$ $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$ allows the patron to ensure the original inequality
-					\item Since patron is willing to pay up to $R_{pt} = \frac{\lambda +\mu + \beta}{1-\de} - \mu$, when $\frac{\lambda +\mu + \beta}{1-\de} -\mu \leq \frac{\nu}{1-\de}$, patron will invest 0 in $l_{gt}$.
-$R_{ct} = 0$ as well.
-					\item If instead assumption 4 did not hold and $R_{pt} > \frac{\nu}{1-\de} + \left( Q_{g,t-1} - L_{g,t-1}\right)$, patron will invest this amount to augment $l_{gt}$.
-Again, $R_{ct} = 0$.
-(Cede, SQ) is played and game ends.
-				\end{itemize}
-		\end{itemize}
-	\item Patron might want to instigate fighting to have a $p_{1s}$ probability of achieving recognition indirectly.
-If $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$ is positive, invest in $w_{st}$ to change $\omega_{st}$.
-Will not invest when the quantity is non-positive.
-Inequality is $\omega_{st} > Q_{st}$
-		\begin{itemize}
-			\item $\omega_{st} = -\zeta_{s}(1-\de) + W_{st}p_{1s} + L_{st}(1-p_{1s}) > Q_{st}$
-			\item $-\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) > Q_{s,t-1} - \mu + R_{ct}$
-			\item $R_{pt}p_{1s} > R_{ct} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$
-			\item $R_{ct} \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$, so if $R_{pt}p_{1s} > \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)$, $R_{ct}=0$ and (SQ,Fight) is played in third stage.
-Outcome depends on war lottery.
-			\item Patron will spend at most $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu$
-				\begin{itemize}
-					\item So need
-						\begin{multline}
-							\textstyle p_{1s}\left[\frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu\right] < \frac{\lambda p_{1s} - \alpha (1-p_{1s}) +\mu + \beta}{1 -\de} - \mu \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \\ \leq \frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu + Q_{s,t-1} - \left(-\zeta_{s}(1-\de) + W_{s,t-1}p_{1s} + L_{s,t-1}(1-p_{1s})\right)
-						\end{multline}
-				\end{itemize}
-		\end{itemize}
-	\item Patron could also invest in $q_g$ to counter investment by $c$ in government's payoffs.
-But this would only been needed if $c$ had incentive to invest in $g$ playing Fight, and we've ruled that out.
-\end{enumerate}
-
 \end{document}

### 2017-04-17 teaching.tex
revise for 3-player game with S preferring SQ

@@
 
-\vskip1in
+\vskip.5in
 \section{Players and Strategy Spaces}
@@
 	\item Patron state (P) chooses $p \in [0,\infty)$ to invest in the secessionists status quo payoffs
-	\item International community (C) chooses $c \in [0,\infty)$ to invest in the secessionists payoffs from rejoining the home state
+%	\item International community (C) chooses $c \in [0,\infty)$ to invest in the secessionists payoffs from rejoining the home state
 \end{enumerate}
 
-\vskip1in
+\vskip.5in
 \section{Payoffs}
-The payoffs after (SQ,SQ) are -p,-c,
+Let, for instance, ($SQ,C$) mean that $G$ plays $SQ$ and $S$ plays $C$. \\
 
-\vskip1in
+Let us consider a first case where the payoffs for the secessionists are higher in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+This, for instance, may be the case very early after taking control of the territory before their economic situation has had a chance to deteriorate.
+Then 
+\begin{itemize}
+	\item The payoffs after ($SQ,SQ$) are $-p,3,2+p$
+	\item The payoffs after ($SQ,C$) are $-10-p,5,0$
+	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
+	\item The payoffs after ($C,C$) are $-p,3,2$
+\end{itemize}
+
+\vskip.5in
 \section{Model}

### 2017-04-26 teaching.tex
finish first, simplest version of teaching model

@@
 \vskip.3in
-\section{Motivation}
+\section{3-player version, Status Quo better than Reunification}
+Let us consider a first case where the payoffs for the secessionists are higher in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+This, for instance, may be the case very early after taking control of the territory before their economic situation has had a chance to deteriorate.
 
-\vskip.5in
-\section{Players and Strategy Spaces}
+\subsection{Players and Strategy Spaces}
 \begin{enumerate}
@@
 
+\vskip.3in
+\subsection{Timing and Information}
+The patron makes its investment. This investment is seen by all other players. Then the home government and secessions move simultaneously.
+
 \vskip.5in
-\section{Payoffs}
-Let, for instance, ($SQ,C$) mean that $G$ plays $SQ$ and $S$ plays $C$. \\
-
-Let us consider a first case where the payoffs for the secessionists are higher in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
-This, for instance, may be the case very early after taking control of the territory before their economic situation has had a chance to deteriorate.
-Then 
+\subsection{Payoffs}
+Let, for instance, ($SQ,C$) mean that $G$ plays $SQ$ and $S$ plays $C$. Then 
 \begin{itemize}
 	\item The payoffs after ($SQ,SQ$) are $-p,3,2+p$
-	\item The payoffs after ($SQ,C$) are $-10-p,5,0$
+	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
 	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
@@
 \vskip.5in
-\section{Model}
-N
+\subsection{Analysis}
+We proceed by backward induction.
 \begin{itemize}
-	\item 
+	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+		\begin{itemize}
+			\item Therefore the Nash Equilibrium (NE) of the subgame is determined by $p$: 
+				\begin{itemize}
+					\item If $2+p \geq 1$ (i.e. $p \geq -1$), the NE is ($SQ,SQ$).
+					\item If $2+p \leq 1$ (i.e. $p \leq -1$), the NE is ($SQ,C$).
+					\item Recall that the smallest value of $p$ that the Patron can choose is $p=0$.
+					\item Since $p\leq-1$ is not possible, ($SQ,C$) is not possible.
+					\item Therefore ($SQ,SQ$) will be the equilibrium.
 				\end{itemize}
+		\end{itemize}
+	\item The Patron's decision: Since ($SQ,SQ$) will be the equilibrium no matter what, the Patron will get a payoff of $-p$.
+		\begin{itemize}
+			\item Since the payoff is $-p$, the Patron maximizes its payoffs by choosing the smallest possible $p$, which is $0$. 
+		\end{itemize}
+\end{itemize}
+Therefore the Subgame Perfect Nash Equilibrium is $p=0$, ($SQ,SQ$) regardless of the value of $p$.
 

### 2017-04-26 teaching.tex
finish second, simiple version of teaching model

@@
 
+\vskip.5in
+\section{3-player version, Reunification better than Status Quo}
+Let us now consider a case where the payoffs for the secessionists are \underline{lower} in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+We believe this will be the case a few months or years after they take control of the territory and unrecognized status has caused their economic situation to deteriorate.
+
+\subsection{Players and Strategy Spaces}
+Same as in Section 1.
+
+\vskip.3in
+\subsection{Timing and Information}
+Same as in Section 1.
+
+\vskip.5in
+\subsection{Payoffs}
+\begin{itemize}
+	\item The payoffs after ($SQ,SQ$) are $-p,3,\bm{0}+p$
+	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
+	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
+	\item The payoffs after ($C,C$) are $-p,3,\bm{0}$
+\end{itemize}
+The only change(s) from Section 1 is that the secessionist's payoff from ($SQ,SQ$) has been reduced, and the payoffs from ($C,C$) along with it.
+
+\vskip.5in
+\subsection{Analysis}
+We proceed by backward induction.
+\begin{itemize}
+	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+		\begin{itemize}
+			\item Therefore the Nash Equilibrium (NE) of the subgame is determined by $p$: 
+				\begin{itemize}
+					\item If $0+p \geq 1$ (i.e. $p \geq 1$), the NE is ($SQ,SQ$).
+					\item If $0+p \leq 1$ (i.e. $p \leq 1$), the NE is ($SQ,C$).
+				\end{itemize}
+		\end{itemize}
+	\item The Patron's decision: if ($SQ,SQ$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
+If ($SQ,C$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
+		\begin{itemize}
+			\item Since ($SQ,C$) is the worse outcome for the Patron, the patron does not want to invest at all ($p=0$) if ($SQ,C$) is going to be the outcome. 
+				\begin{itemize}
+					\item In this case, the Patron chooses $p=0$ and the Patron's total payoff is $-10$.
+				\end{itemize}
+			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of the $p \geq 1$.
+Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$, which is $p=1$.
+				\begin{itemize}
+					\item In this case, the Patron chooses $p=1$ and the Patron's total payoff is $-1$.
+				\end{itemize}
+			\item Since $-1 > -10$, the Patron will choose to invest $p=1$.
+		\end{itemize}
+\end{itemize}
+Therefore the Subgame Perfect Nash Equilibrium is $p=1$, and ($SQ,SQ$) as long as $p \geq 1$, ($SQ,C$) if $p \leq 1$.
+\begin{itemize}
+	\item $p=1$, ($SQ,SQ$) is an equilibrium outcome.
+	\item $p=1$, ($SQ,C$) is also an equilibrium outcome. You can rule this one out by specifying that the Secessionists choose $SQ$ whenever they are indifferent.
+\end{itemize}
+
 \end{document}

### 2017-04-26 teaching.tex
start 4-player case

@@
 
+\vskip.5in
+\section{4-player version, Reunification better than Status Quo}
+Now we look at the full version of the model including the international community, and continue to assume that the payoffs for the secessionists are lower in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+
+\subsection{Players and Strategy Spaces}
+Same as in Section 1.
+
+\vskip.3in
+\subsection{Timing and Information}
+Same as in Section 1.
+
+\vskip.5in
+\subsection{Payoffs}
+\begin{itemize}
+	\item The payoffs after ($SQ,SQ$) are $-p,3,\bm{0}+p$
+	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
+	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
+	\item The payoffs after ($C,C$) are $-p,3,\bm{0}$
+\end{itemize}
+The only change(s) from Section 1 is that the secessionist's payoff from ($SQ,SQ$) has been reduced, and the payoffs from ($C,C$) along with it.
+
+\vskip.5in
+\subsection{Analysis}
+We proceed by backward induction.
+\begin{itemize}
+	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+		\begin{itemize}
+			\item Therefore the Nash Equilibrium (NE) of the subgame is determined by $p$: 
+				\begin{itemize}
+					\item If $0+p \geq 1$ (i.e. $p \geq 1$), the NE is ($SQ,SQ$).
+					\item If $0+p \leq 1$ (i.e. $p \leq 1$), the NE is ($SQ,C$).
+				\end{itemize}
+		\end{itemize}
+	\item The Patron's decision: if ($SQ,SQ$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
+If ($SQ,C$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
+		\begin{itemize}
+			\item Since ($SQ,C$) is the worse outcome for the Patron, the patron does not want to invest at all ($p=0$) if ($SQ,C$) is going to be the outcome. 
+				\begin{itemize}
+					\item In this case, the Patron chooses $p=0$ and the Patron's total payoff is $-10$.
+				\end{itemize}
+			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of the $p \geq 1$.
+Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$, which is $p=1$.
+				\begin{itemize}
+					\item In this case, the Patron chooses $p=1$ and the Patron's total payoff is $-1$.
+				\end{itemize}
+			\item Since $-1 > -10$, the Patron will choose to invest $p=1$.
+		\end{itemize}
+\end{itemize}
+Therefore the Subgame Perfect Nash Equilibrium is $p=1$, and ($SQ,SQ$) as long as $p \geq 1$, ($SQ,C$) if $p \leq 1$.
+\begin{itemize}
+	\item $p=1$, ($SQ,SQ$) is an equilibrium outcome.
+	\item $p=1$, ($SQ,C$) is also an equilibrium outcome. You can rule this one out by specifying that the Secessionists choose $SQ$ whenever they are indifferent.
+\end{itemize}
+
 \end{document}

### 2017-04-27 teaching.tex
modify all by analysis for 4-player game

@@
 \subsection{Players and Strategy Spaces}
-Same as in Section 1.
+We add the international community:
+\begin{enumerate}
+	\item Home government (G) chooses $S_G \in \left\{SQ,C\right\}$ where SQ means 'Status Quo' of unrecognized statehood and C means 'Cede' the issue of status and recognize the secessionists as an independent state.
+	\item Secessionists (S) choose $S_S \in \left\{SQ,C\right\}$, same as for $G$ except when they Cede, they rejoin the home state.
+	\item Patron state (P) chooses $p \in [0,\infty)$ to invest in the secessionists status quo payoffs
+	\item \textbf{International community (C) chooses} $c \in [0,\infty)$ \textbf{ to invest in the secessionists payoffs from rejoining the home state}
+\end{enumerate}
 
@@
 \subsection{Timing and Information}
-Same as in Section 1.
+The patron makes its investment.
+This investment is seen by all other players.
+\textbf{Then the international community makes its investment.
+This investment is seen by all other players.} Then the home government and secessions move simultaneously.
 
@@
 \subsection{Payoffs}
+We add payoffs for the international community, and the international community's investment in the secessionists payoffs from ceding.
 \begin{itemize}
-	\item The payoffs after ($SQ,SQ$) are $-p,3,\bm{0}+p$
-	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
-	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
-	\item The payoffs after ($C,C$) are $-p,3,\bm{0}$
+	\item The payoffs after ($SQ,SQ$) are $-p,\bm{-c},3,0+p$
+	\item The payoffs after ($SQ,C$) are $-10-p,\bm{5-c},5,1\bm{+c}$
+	\item The payoffs after ($C,SQ$) are $3-p,\bm{-7-c},0,5+p$
+	\item The payoffs after ($C,C$) are $-p,\bm{-c},3,0\bm{+c}$
 \end{itemize}
-The only change(s) from Section 1 is that the secessionist's payoff from ($SQ,SQ$) has been reduced, and the payoffs from ($C,C$) along with it.
 

### 2017-04-28 teaching.tex
finish 4-player game analysis for teaching version

@@
 				\end{itemize}
-			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of the $p \geq 1$.
+			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of $p \geq 1$.
 Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$, which is $p=1$.
@@
 				\begin{itemize}
-					\item If $0+p \geq 1$ (i.e. $p \geq 1$), the NE is ($SQ,SQ$).
-					\item If $0+p \leq 1$ (i.e. $p \leq 1$), the NE is ($SQ,C$).
+					\item If $0+p \geq 1+c$ (i.e. $p \geq 1+c$), the NE is ($SQ,SQ$).
+					\item If $0+p \leq 1+c$ (i.e. $p \leq 1+c$), the NE is ($SQ,C$).
 				\end{itemize}
+		\end{itemize}
+	\item Player $C$'s decision: if ($SQ,SQ$) is the equilibrium in the subgame, Player $C$ will get a payoff of $-c$.
+If ($SQ,C$) is the equilibrium in the subgame, Player $C$ will get a payoff of $5-c$.
+		\begin{itemize}
+			\item Since ($SQ,SQ$) is the worse outcome for Player $C$, Player $C$ does not want to invest at all ($c=0$) if ($SQ,SQ$) is going to be the outcome. 
+				\begin{itemize}
+					\item In this case, the Player $C$ chooses $c=0$ and Player $C$'s total payoff is $0$.
+				\end{itemize}
+			\item Player $C$ prefers ($SQ,C$), but to make this happen Player $C$ must make an investment of the $c \geq p - 1$.
+Because Player $C$'s payoff is $5-c$, Player $C$ maximizes its payoffs in the case where ($SQ,C$) is the outcome by choosing the smallest $c$, which is $c=p-1$.
+				\begin{itemize}
+					\item In this case, Player $C$'s investment of $c=p-1$ makes its total payoff is $5-c = 5 - (p -1) = 6-p$.
+				\end{itemize}
+			\item If $6 - p \leq 0$ (where $0$ is Player $C$'s payoff from ($SQ,SQ$)), Player $C$ will choose to invest $c = p-1$ and the outcome will be ($SQ,C$).
+			\item If $6 - p > 0$ (i.e. $6 > p$), Player $C$ will choose to invest $c = 0$ and the outcome will be ($SQ,SQ$).
 		\end{itemize}
@@
 				\end{itemize}
-			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of the $p \geq 1$.
-Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$, which is $p=1$.
+			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of $p \geq 6$.
+Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$ that leads to ($SQ,SQ$), which is $p=6$.
 				\begin{itemize}
-					\item In this case, the Patron chooses $p=1$ and the Patron's total payoff is $-1$.
+					\item In this case, the Patron chooses $p=6$ and the Patron's total payoff is $-6$.
 				\end{itemize}
-			\item Since $-1 > -10$, the Patron will choose to invest $p=1$.
+			\item Since $-6 > -10$, the Patron will choose to invest $p=6$.
 		\end{itemize}
 \end{itemize}
-Therefore the Subgame Perfect Nash Equilibrium is $p=1$, and ($SQ,SQ$) as long as $p \geq 1$, ($SQ,C$) if $p \leq 1$.
+Therefore the Subgame Perfect Nash Equilibrium is $p=6$, $c = 0$ if $p \geq 6$, $c = p-1$ if $p < 6$ and ($SQ,SQ$) as long as $p \geq c+1$, ($SQ,C$) if $p \leq c+ 1$.
 \begin{itemize}
-	\item $p=1$, ($SQ,SQ$) is an equilibrium outcome.
-	\item $p=1$, ($SQ,C$) is also an equilibrium outcome. You can rule this one out by specifying that the Secessionists choose $SQ$ whenever they are indifferent.
+	\item $p=6$, $c = 0$, ($SQ,SQ$) is an equilibrium outcome.
 \end{itemize}

### 2017-06-30 editedvolume.tex
add .tex file from Ben w/rough draft of version for Transforming Conflict book project

\documentclass[12pt,letterpaper, notitlepage]{article}
\usepackage{epsf,graphicx,psfrag}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage[reqno]{amsmath}
\usepackage{endnotes}
\usepackage{natbib}
\usepackage{amsfonts}
\usepackage{amssymb,enumerate}
\usepackage{dcolumn}
\usepackage{verbatim}
%\usepackage{hypertex}
\usepackage[dvips, bookmarks, colorlinks=true, plainpages = false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[hmargin=3.0cm,vmargin=3.0cm]{geometry}
%\usepackage{geometry}
\usepackage{nicefrac}
\usepackage{pdfsync}
\usepackage{url}
\usepackage{pst-all}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{graphicx,verbatim,scalefnt}

\newcommand{\ta}{\theta}
\newcommand{\Ta}{\Theta}
\newcommand{\ga}{\gamma}
\newcommand{\un}{\underline}
\newcommand{\ov}{\overline}
\linespread{1.05} % Palatino needs more leading (space between lines)
\title{Unrecognized States:  Theory and Cases}

\date{\today}
\begin{document}
\begin{titlepage}
\maketitle
\thispagestyle{empty}
%\center{Word Count: 12,842\\ This count includes title page, abstract, footnotes, and bibliography, but excludes the technical appendix, which does not need to be included in the print version of the article.}

\begin{abstract}
\noindent  Unrecognized states destabilize the international system and impoverish their residents. Thus, unrecognized statehood is a profoundly undesirable outcome, and yet it is often a  stable equilibrium. Game theoretic modeling has proven effective at clarifying the strategic logic that sustains unrecognized states, and offers insight into possible paths to resolution. In this chapter we draw on these insights, and illustrate them with discussion individual cases. The game theory on which we draw analyzes not only the actions of unrecognized states and the home states from which they are attempting to secede, but also the patrons that support these unrecognized states and the actors in the international community who work to induce peaceful settlement. In this piece we focus particularly on evaluating the policy options available to peace and development-seeking actors in the international community as they work to resolve these stalemated conflicts.
\end{abstract}

\end{titlepage}
\setcounter{page}{1}

\doublespacing

\section*{Introduction}

Unrecognized states destabilize the international system and impoverish their residents. Because their territorial borders are contested, the threat of violent conflict is ever present. Facing sanctions, unable to sign trade agreements or receive most foreign aid, their residents are isolated and impoverished. Thus, unrecognized statehood is a profoundly undesirable outcome, and yet it is often a  stable equilibrium. In addition to Somaliland and the Turkish Republic of Northern Cyrpus, many of the unrecognized states that emerged when the Soviet Union collapsed still enjoy \emph{de facto} independence a quarter century later.

1. Unrecognized states are destabilizing to the international system, bad for their residents, and yet often very long in tenure.
2. Game theory helps us understand the mechanisms by which they are sustained, and the ways in which they can be transformed
3. Engagement with the details of actual cases informs us to what actual paths to resolution of these disputes look like, and what roadblocks stand in the way

[Include an updated version of Table 1 from the JLEO article. Ben's RA will update this]

\section*{Self Determination and Foreign Influence}
[Summary of our theory in approx 1500 words]

\section*{How Unrecognized Statehood Ends}

Unrecognized Statehood exists as a halfway point between recognized independent statehood and reunification with the home state. Unrecognized statehood ends when either recognized statehood is achieved, or reunification occurs.  In the following sections we discuss how these transitions occur -- both through negotiated settlement and as a result of decisive military victory by one side or the other.

\subsection*{Reunification via Military Reconquest}

Most attempted secessions end in military defeat before territorial control is ever achieved (Fazal and Griffith 2008). Unrecognized states are thus a relatively elite set of secessionist movements, those that are unusually militarily powerful relative to the home state. However, even among secessionist movements that succeed in maintaining territorial control for a minimum of two years, the most common form of resolution remains military reconquest by the home state. When unrecognized states return to war with the home state, it is almost always the home state that initiates and the home state that is victorious (we discuss military victory by the secessionists in a later section).

In the case of most prolonged stalemates, a patron provides enough military assistance to the secessionists to make military reconquest by the patron prohibitively costly.  The 11 cases of military reconquest in Table 1 occur in cases with no patron or cases in which the patron withdraws or reduces its support.

The cases with no patron are fairly straightforward. For example, Chechnya achieved its \emph{de facto} independence immediately after the fall of the Soviet Union when Russia was very weak.   As Russia strengthened, there was no patron support to offset the relative decline in the Chechens' military capabilities. Over time, Russia's military advantage grew and in 1999 the Russian government invaded and reconquered Chechnya.

It is worth exploring, however, the reasons why a patron might support a secessionist group during its initial rebellion and then withdraw support at a later date. Patrons' strategic interests in the unrecognized state vary from patron to patron, and both budget constraints and salience of interest vary over time. For example, domestic political concerns (primarily ethnic solidarity with the secessionists) induced a modest level of Indian support for the Tamil Tigers in Sri Lanka 1983-1987.  These domestic political concerns were eventually outweighed by broader strategic security concerns and a desire for regional stability. In 1987 the Indian government signed a peace accord with Sri Lanka (the home state) and largely withdrew their support from the Tamil secessionists, even sending in peacekeepers that later clashed with the secessionists militarily.\footnote{Singer 1992}

The patron's decision to withdraw support for the secessionists is sometimes motivated by interactions between the patron and the international community, which may bribe or coerce the Patron to abandon its support. In an extreme example involving both sanctions and direct military confrontation, the United States and other members of the North Atlantic Treaty Organization (NATO) coerced Serbia into withdrawing its support from Republika Srpska and Republika Srpska Krajina, both of which had secured \emph{de facto} independence after the collapse of Yugoslavia.\footnote{For an excellent discussion of the case of Republika Srpska, see Zahar 2004.}

\subsection*{Negotiated Reunification}

Just as military reconquest becomes more likely when Patron support is withdrawn or declines, so too does negotiated reunification. Negotiated agreements are struck when the patron does not contribute sufficiently to prevent the secessionists from preferring ceding to the status quo, and when a deal is available that both sides prefer to war. Since WWII, four peacefully negotiated reunifications have occurred.\footnote{We limit our discussion here to entities that had existed in a period of stalemate prior to reaching a settlement -- i.e. those that had maintained territorial control for at least two years.} Secessionists in Ajara, Bouganville, and Gagauzia have opted to rejoin the home state.  In all four of the cases of negotiated settlement, the observed outcomes seem to match the model well: the payoffs to the secessionist elite from ceding have been low, and the payoffs to the central government high.

In Ajara, where the level of patron (Russian) support was quite low, the choice facing the secessionist elite was between agreeing to reunify with Georgia or facing military defeat. In Bouganville, which separated from Papua New Guinea, secessionists lacked not only a patron, but also a clear preference for secession -- demands for secession had emerged only late in a struggle that began as an effort to stop a mining operation.\footnote{Ghai and Regan 2006}  Here the value of status to the secessionists was actually quite low, and they were willing to surrender it in exchange for relatively small side payments.

However, despite past failures, or theory suggests that a sufficiently motivated patron can induce negotiated settlement if it so chooses. The means through which the international community might induce negotiated settlement are discussed in detail in the section on policy implications. It is notable, however, that we do not expect any future cases of peacefully negotiated independence.  While negotiated reunification is the preferred outcome of the international community, and they may be willing spend to achieve it, recognized statehood is generally not the preferred outcome of the patron or the international community. Our analysis suggests that the most likely path to recognized statehood is, and will remain, military defeat of the home state.

The difficulty of making credible payments in exchange for status is one clearly demonstrated in the civil war literature.\footnote{e.g. Licklider 1995; Walter 1997, 2002; Fearon and Laitin 2007; Doyle and Sambanis 2006} Unrecognized states generally constitute ``sons of the soil" conflicts in which the central government cannot credibly commit to preserving the local demographic and political dominance of the secessionist elite once the disputed territory reverts to central government control.\footnote{Weimer 1978; Fearon 2004}  While the central government might initially grant the secessionist elite a high level of autonomy in exchange for agreeing to reunification, the level of autonomy is likely to decrease over time, perhaps quite quickly.  Reference to the cases of Abkhazia and Gagauzia are informative here.

At the time of secession, ethnic Akbhaz made up a minority of the population of Abkhazia,\footnote{Cornell 2001; Wooleh 2006} but they now [DOUBLE CHECK WHAT THE CURRENT SITUATION IS] control almost all political posts in the \textit{de facto} government of the region. In 2004, the basket of payments offered by the Georgians in exchange for reunification included a provision guaranteeing that ethnic Abkhaz would retain a majority in the regional parliament, even if the return of internally displaced persons (IDPs) once again placed ethnic Abkhaz in a minority demographic position in the region. The promise, however, was not very meaningful.  First, even if the promise were upheld, it would still mean a step back from the total dominance the ethnic Abkhaz currently enjoy in the region.  Second, if Georgian IDPs returned, they may demand and receive a more equitable system of representation.  These concerns are not abstract; this type of reneging has already occurred in cases that did reach settlement.

Gagauzia achieved de facto independence at the time of the Soviet Union's collapse, but agreed to rejoin Moldova in 1994 as an autonomous region.  While Gagauzia was granted substantial autonomy under the Moldovan Law on the Special Legal Status of Gagauzia, when the governor of Gagauzia, Dmitrii Croiter, moved to assert these powers in 1999, the Moldovan government balked.  By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government.  The Moldovan government jailed a number of other Gagauz politicians, and while Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the de facto level of autonomy has been limited by continued by central government meddling in less-than-free regional elections.\footnote{Protsyk (2010) provides an account of the "salami tactics" by which Moldovan authorities have gradually reclaimed powers originally granted to the regional government.} The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.\footnote{Roper (2002) argues that secessionists in Transnistria are wary of negotiated reunification precisely because of the creeping re-centralization they have observed in Gagauzia.}

\subsection*{Recognition via Secessionist Military Victory}
While the path to independent statehood via secession is an extremely narrow one, recognition does sometimes occur. It has occurred primarily in cases where the secessionists (often supported by a patron) are so strong militarily that they not only achieve territorial control in the initial conflict, but also threaten the home state government outside the unrecognized state.  Bangladesh and Eritrea both secured recognition as part of the peace agreement ending the war of secession.

[DISCUSS BANGLADESH AND ERITREA IN DETAIL]

\section*{Negotiated Recognition}
No unrecognized state has yet managed to gain recognition from the home state when recognition or a referendum was not agreed to as a condition of ending the initial war of secession. Wars that have reignited after a period of unrecognized statehood have always either resulted in reunification or left the status quo intact.  However, if an unrecognized state were to gain an outright military victory over the home state at any time, this does represent a plausible path to recognition. Once unrecognized statehood has emerged as an equilibrium, however, the path to recognition is narrower still.

DISCUSS KOSOVO AND SOUTH SUDAN HERE.

\section*{Policy Implications: Options for The International Community}
 In general, the international community has preferences for reunification over independence, for resolution over the status quo, and for peace instead of war.  In this section we consider  four means through which the international community might pursue these ends: sanctions against the secessionist region, direct incentives provided to the secessionists in exchange for ceding, enforcement of concessions offered by the home state, and direct coercion of the patron.

The intended effect of sanctions is to make the status quo less appealing vis-\`{a}-vis ceding. However any sanctions that increase the secessionists' hostility toward reunification will also increase the range of conditions under which war will be chosen.  Sanctions can have this effect if they reduce the secessionists' quality of life under the status quo %(via imposition of economic costs)
and reduce the quality of the deal secessionists expect to get if they opt for negotiated resettlement. As the peaceful options become worse, war becomes relatively more attractive.
 Compounding this, sanctions that reduce the secessionists' military capabilities (and thus reduce the secessionists' expected payoffs from war) also have the effect of making military reconquest easier for the home state, making it more likely that the home state will attack. In either case, the range of conditions under which war will be initiated becomes broader.\footnote{In most cases, the military position of the home state is stronger than that of the secessionists, so a further tip in the balance of military power toward the home state is more likely to induce war than a similar change in favor of the secessionists.}

%However, if $\delta_s$ is high or the payoffs from war exceed the lowered payoffs from ceding for any other reason (such as animosity of the secessionist public toward the sanctions-imposing home state government), the effect of sanctions will be to induce war rather than negotiated settlement.

There is a better way. If the international community tries to promote settlement by supplementing the payoffs from unification, they are able to induce negotiated settlement without simultaneously increasing the risk of war.  This can be done either through promises of benefits to the unrecognized state provided directly by the international community, like aid, or by a commitment from the international community to serve as a third-party guarantor of side payments promised by the ceding side. In the case of contingent promises of aid, the calculation is relatively straightforward: 1) the promise of aid must be credibly contingent on negotiated settlement, and 2) the aid offered must be valued more highly than the concessions required to reach an agreement. It is the second condition that is most problematic. Because both sides place such a high value on status (independence vs. reunification), even large amounts of aid are likely to be valued less than the concessions necessary to reach an agreement.

Serving as a third-party guarantor of autonomy rights is a way for the international community to potentially overcome problems of indivisibility and commitment and help the parties reach a credible compromise on status\footnote{Walter 2002} However, this strategy is only tenable when the only impediment to settlement is the unenforcability of a bargain, and when the international community is credible as an enforcer of that bargain.

%Credible enforcement of future autonomy rights can be viewed either as increasing the value of available side payments or as making the central issue of contention divisible. In either view, a range of previously untenable agreements are made possible.

In Southern Sudan, the international community invested substantial resources to help negotiate a settlement and to ensure that the Sudanese government government both allowed the promised a referendum and respected its results. While the international community acted in Southern Sudan to enforce independence, not autonomy, it has shown itself capable of enforcing difficult concessions by the home state government. This bodes well for the future credibility of the international community as a third-party enforcer.  However, the role of the international community in enforcing other past agreements might give secessionists pause. For example, a referendum on independence in Western Sahara, which the UN ruled to be necessary more than thirty years ago, has never come to pass.\footnote{For a thorough analysis of the Western Sahara case, see Zunes and Mundy (2010).}  Nonetheless, it is possible for the international community to invest resources to enforce agreements, allowing for negotiated settlements that would otherwise be impossible to reach.

To show that it is possible for the international community to enforce the terms of negotiated agreements at a reasonable cost is not sufficient to imply that such an outcome is likely. The political will necessary to achieve success in Southern Sudan was motivated largely by the magnitude of the atrocities that accompanied the war of secession, and enforcement was made credible, in part, due to the weakness of Sudan relative to the international community. Enforcing the terms of an agreement between Russia and Georgia, for example, would be more difficult.

It is also possible for the international community to affect the payoffs of the patron through interactions in other games outside of our model. Such actions would manifest themselves within the model as reductions in the patron's willingness to pay to sustain the status quo. If the patron is unwilling to pay to sustain the status quo, the war payoffs and status quo payoffs of the secessionists will decline over time, eventually leading to either war or negotiated settlement. Under these conditions, the within-game costs to the international community of inducing negotiated reunification also fall.

In this section we have argued that successful intervention by the international community is possible. The key, however, is motivation: the international community is capable of inducing peaceful settlement when it is willing to invest the resources necessary. However, strong preferences of secessionists against reunification and the opposing intervention of the patron make the costs of such interventions prohibitively high in most cases.  Unrecognized statehood is a stable equilibrium because the international community is unwilling to invest sufficient resources to outspend the patron and induce its preferred outcome.

\section*{Conclusion}

\end{document}

### 2017-06-30 editedvolume.tex
remove dvips from hyperref options to quiet error
add questions from editors that need to be addressed in chapter

@@
 %\usepackage{hypertex}
-\usepackage[dvips, bookmarks, colorlinks=true, plainpages = false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
+\usepackage[bookmarks, colorlinks=true, plainpages = false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
 \usepackage{amssymb}
@@
 [Include an updated version of Table 1 from the JLEO article. Ben's RA will update this]
+
+From Kristy, 6/30/17:\\
+Note new requirement from editors, accompanied by increase in word limit to 7,000 (we would probably have done this anyway):
+To make the book  more coherent, we are asking that your chapter address the following questions:
+\begin{enumerate}
+	\item How does your work relate to a key change or  new challenge- change in environment, your field, way of thinking or intervening?
+	\item Why was the past approach/thinking incomplete/not adequate?
+	\item How does the new approach/thinking/framing in your chapter advance the field?
+\end{enumerate}
 

### 2017-06-30 editedvolume.tex
fix minor typos in text; add notes from initial read

@@
 
-However, despite past failures, or theory suggests that a sufficiently motivated patron can induce negotiated settlement if it so chooses.
+However, despite past failures, our theory suggests that a sufficiently motivated patron can induce negotiated settlement if it so chooses.
 The means through which the international community might induce negotiated settlement are discussed in detail in the section on policy implications.
@@
 By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government.
-The Moldovan government jailed a number of other Gagauz politicians, and while Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the de facto level of autonomy has been limited by continued by central government meddling in less-than-free regional elections.\footnote{Protsyk (2010) provides an account of the "salami tactics" by which Moldovan authorities have gradually reclaimed powers originally granted to the regional government.} The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.\footnote{Roper (2002) argues that secessionists in Transnistria are wary of negotiated reunification precisely because of the creeping re-centralization they have observed in Gagauzia.}
+The Moldovan government jailed a number of other Gagauz politicians, and while Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the de facto level of autonomy has been limited by continued central government meddling in less-than-free regional elections.\footnote{Protsyk (2010) provides an account of the "salami tactics" by which Moldovan authorities have gradually reclaimed powers originally granted to the regional government.} The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.\footnote{Roper (2002) argues that secessionists in Transnistria are wary of negotiated reunification precisely because of the creeping re-centralization they have observed in Gagauzia.}
 
@@
 
-In Southern Sudan, the international community invested substantial resources to help negotiate a settlement and to ensure that the Sudanese government government both allowed the promised a referendum and respected its results.
+In Southern Sudan, the international community invested substantial resources to help negotiate a settlement and to ensure that the Sudanese government both allowed the promised a referendum and respected its results.
 While the international community acted in Southern Sudan to enforce independence, not autonomy, it has shown itself capable of enforcing difficult concessions by the home state government.

### 2017-06-30 theory.tex
create theory.tex containing all theory/appendix matter; include in editedvolume.tex

--- 2016-07-15 draft.tex
@@
-\documentclass[11pt,letterpaper, notitlepage]{article}
-\usepackage{epsf,graphicx,psfrag}
-\usepackage{natbib}
-\usepackage{setspace}
-\usepackage{fullpage}
-\usepackage[reqno]{amsmath}
-\usepackage{endnotes}
-\usepackage{natbib}
-\usepackage{amsfonts}
-\usepackage{amssymb,enumerate}
-\usepackage{dcolumn}
-\usepackage{verbatim}
-%\usepackage{hypertex}
-\usepackage[pdftex, bookmarks, colorlinks=true, plainpages=false, citecolor = blue, urlcolor = blue, filecolor = blue, breaklinks]{hyperref}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage[hmargin=3.25cm,vmargin=3.0cm]{geometry}
-%\usepackage{geometry}
-\usepackage{nicefrac}
-\usepackage{pdfsync}
-\usepackage{url,bm}
-\usepackage{pst-all}
-\usepackage{pdflscape}
-\usepackage{rotating}
-\usepackage[T1]{fontenc}
-\usepackage[sc]{mathpazo}
-\usepackage{graphicx,verbatim,scalefnt}
-\pdfoptionpdfminorversion=6 %gets rid of warning about pdf version
-
-\let\footnote=\endnote
-
-\newtheorem{definition}{Definition}
-\newtheorem{proposition}{Proposition}
-\newtheorem{lemma}{Lemma}
-\newcommand{\ta}{\theta}
-\newcommand{\Ta}{\Theta}
-\newcommand{\ga}{\gamma}
-\newcommand{\un}{\underline}
-\newcommand{\ov}{\overline}
-\newcommand{\de}{\delta}
-\newcommand{\ve}{\varepsilon}
-\linespread{1.05} % Palatino needs more leading (space between lines)
-\title{Unrecognized States: A Theory of Self-Determination and Foreign Influence}
-
-\author{Kristy Buzard\\
-Syracuse University\\
-\\
-Benjamin A.T. Graham \\
-University of Southern California\\
-\\
-Ben Horne\\ 
-University of California, San Diego}
-
-\date{\today}
-\begin{document}
-\hypersetup{pageanchor=false}
-\begin{titlepage}
-\maketitle
-\thispagestyle{empty}
-%\center{This article is a working paper and is subject to future revision during the peer-review process. The authors welcome any comments, and may be reached by e-mail at benjamin.a.graham@usc.edu.}
-
-\begin{abstract}
-\noindent Unrecognized states are characterized by stagnant or crumbling economies and political instability, often serve as havens for illicit trade, and challenge the territorial sovereignty of recognized states.
-Their persistence is both intellectually puzzling and normatively problematic, but unrecognized statehood can be a remarkably stable outcome, persisting for decades.
-Our dynamic four-player model reveals that unrecognized statehood emerges as an equilibrium outcome when a patron state is willing and able to persistently invest resources to sustain it.
-%Counterintuitively, patrons will make these investments even when unrecognized statehood is not their most preferred outcome.
-We assess options available to actors in the international community who seek to impose their preferred outcomes in these disputes and find that, although sanctions are the most frequently employed, they can often lead to renewed conflict instead of the intended resolution.
-\end{abstract}
-
-\end{titlepage}
-\hypersetup{pageanchor=true}
-\setcounter{page}{1}
-
-\doublespacing
-\section{Introduction}
-South Ossetia is an archetypical unrecognized state -- characteristic of those regions of the world in which non-state actors control territory and govern populations.
-From 1990 to 1992, Ossetian rebels fought a successful secessionist civil war against the Georgian government that ended with a ceasefire and left the rebels in \textit{de facto} control of much of the region of South Ossetia, which sits along Georgia's northern border with Russia.
-In the 22 years since the ceasefire was signed, South Ossetia has functioned as an unrecognized state, governing its own affairs but unrecognized by foreign nations.\footnote{South Ossetia, along with Abkhazia, was recognized by Russia and Nicaragua in 2008, Venezuela and Nauru in 2009, and Tuvalu in 2011.} The Georgian government maintains its claim to the territory of South Ossetia, while the South Ossetians continue to seek international recognition of their independence.
-In 2004, the Georgian government began intermittent efforts to close trade with the separatist region, and in 2008, following escalating provocations from the Ossetian side, Georgia attempted to reclaim the territory by military force.
-Russian troops acting in support of the Ossetians quickly crushed the would-be reconquest, and the status quo was restored.%\footnote{Good sources on the history of the conflict include Lynch 2004, Asmus 2010, and a series of reports by the International Crisis Group.}
-
-The international system is characterized by a norm under which seceding entities are rarely recognized by foreign states unless they are first recognized by the home state.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states.
-In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm makes sense in a context where almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby reducing the incentives for secessionist movements.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize." For the definitive international legal treatment of recognition, see Crawford (2006).} Thus, unrecognized states emerge when seceding entities gain \emph{de facto} territorial control through successful wars of secession, but are not militarily strong enough to force recognition of their independence from the home state from which they have seceded.
-
-Persistent unrecognized statehood is both intellectually puzzling and normatively undesirable.
-It is an extremely costly outcome for the secessionists, the home states from which they are attempting to secede, and the international community more broadly.
-Nonetheless, unrecognized statehood has been a recurring phenomenon since WWII, with half a dozen current unrecognized states having existed for more than 20 years.\footnote{Abkhazia, Ngorno-Karabakh, Transnistria, Somaliland, South Ossetia, and the Turkish Republic of Northern Cyprus.} Much of the existing literature treats unrecognized statehood as either a transient phenomenon or simply as a failure to reach recognized statehood.
-In contrast, we argue that unrecognized statehood is an important outcome in its own right, and one that is potentially extremely stable over time.
-A major contribution of this paper is the demonstration of unrecognized statehood as a long-run equilibrium outcome of the strategic interaction between not just the secessionists and the home government, but also the international community and states or non-state actors who act as patrons of the secessionists.
-We show that unrecognized statehood can be a stable outcome despite its persistently high costs to all actors involved.
-
-The costs to the secessionists are arguably the highest.
-Non-recognition locks unrecognized states out of a states-only club whose members enjoy benefits in terms of both security and economic integration.
-Unrecognized states do not benefit from the norm of territorial integrity that reduces the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Non-recognition also prevents entry into multilateral and bilateral trade agreements, dramatically reduces access to foreign aid, and excludes their territory from international legal frameworks, making it harder to secure foreign investment (Milhalkanian, 2004; Caspersen, 2013: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King, 2001).
-However, as Moldova deepens ties with the EU, Transnistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu, 2011).} %In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).
-
-The home state bears costs of non-settlement as well, although they are not as high as those facing the unrecognized state.
-Instability in the border region and the diversion of military resources to monitor the \textit{de facto} border are costly, as is maintenance of economic sanctions against the unrecognized state.
-In Azerbaijan, where an oil boom sent GDP soaring in the mid-2000s, the unresolved secessionist conflict in Nagorno-Karabakh led a massive military buildup to take precedence over other government spending.\footnote{Military spending increased 51\% in 2004-2005, and went up another 82\% in 2006 (International Crisis Group 2007)} Foreign investors are more wary of investing in states that do not effectively control their own territory, and there are diplomatic costs as well.
-For Georgia, failure to resolve its outstanding secessionist conflicts has hindered progress toward NATO accession.
-
-%Unrecognized statehood is a thorn in the side of the state system.
-Many unrecognized states are havens for smuggling and trafficking illicit goods, they are prone to violent conflict, and their very existence challenges the norm of territorial sovereignty (King 2001; Stanislawski 2008).
-Like most unrecognized states, South Ossetia pays high costs for non-recognition.
-Its political survival is tenuous, its licit economy is in shambles after years of isolation and instability, and yet, if the history of similar entities are any guide, its chances of eventual recognition are slim.
-Stalemate imposes similar, though less severe, costs on home states (e.g.
-Georgia) as well.
-
-%\subsection{Non-Recognition and Its Costs}
-%Since WWII, the path to recognized statehood has been narrow, and securing territorial control is not sufficient to gain recognition.
-The international community has been exceedingly clear that a national right of self-determination applies only to cases of decolonization: it does not extend to cases of secession.
-In 1970, UN secretary general U Thant said, ``The United Nations has never accepted and does not accept and I do not believe will ever accept the principle of secession of a part of its member states."\footnote{Pegg 1998: 8} In cases where the home state's sovereignty is disputed or where the home state government commits mass atrocity crimes, some states may choose to grant recognition, but there remains a strong norm of home-state veto.That is, secessionist entities are rarely recognized by foreign states unless they are first recognized by the home state from which they are attempting to secede.
-This has lead to the existence of an odd group of state-like entities that control territory and govern population but are not recognized as states and cannot participate in the international system.
-We refer to these entities as unrecognized states, though, variously specified, they are also referred to states-within-states\footnote{Kingston and Spears 2004} \textit{de facto} states,\footnote{Pegg 1998, Lynch 2004} or almost-states.\footnote{Stanislawski 2008}
-
-%Since WWII, the path to recognized statehood has been narrow, and securing territorial control is not sufficient to gain recognition.
-Seceding entities are rarely recognized by foreign states unless they are first recognized by the home state from which they are attempting to secede.\footnote{This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states.
-In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.} This norm of home state veto emerges because almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, thereby making secession less likely.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize."} The higher the costs of secession, the greater the security of existing states.
-
-%Unrecognized statehood is a particularly compelling outcome from secession in part because while the secessionists have succeeded in securing control of (most of) the territory they aspire to rule they are not recognized by most other states, trapping them in a costly limbo.
-
-%Unrecognized states are subject to high costs of non-recognition generated by an international system extremely hostile to non-state territorial units.
-The modern state system is based on institutions designed to facilitate peaceful and economically beneficial relations between like actors, and on the empowerment and legitimation of states through mutual recognition (Spruyt 1996).
-The norm challenged most directly by unrecognized states is that which, in the post-WWII era, has become most fundamental to it -- the norm of territorial sovereignty.
-Almost all states have reason to fear the emergence of secessionist movements within their own borders, and a state system that places seceding entities at an extreme disadvantage lowers the expected benefits of secession, and thereby makes secession less likely.\footnote{Coggins (2011: 451) notes that, "The more acute the domestic threat, the more the reticence to recognize."} The higher the costs of secession, the greater the security of existing states.
-
-%The most direct manifestation of this systemic hostility to secession is the norm of states refusing to recognize a seceding entity unless the home state recognizes it first.\footnote{For a thorough international legal discussion of the issue, see Crawford 2006.} This norm is not absolute: in some cases, like the People's Republic of China, the sovereignty of the home state government (over Taiwan) has never been recognized by some states.
-In other cases, like Kosovo, the commission of mass atrocities by the home state government may supersede its sovereignty.
-However, the norm is strong and gives recognition by the home state its significance: without recognition by the home state, the post-WWII norm is for other states is to withhold recognition as well.\footnote{Caspersen (2012) traces the history of collective non-recognition, which begins after WWI and solidifies after WWII.
-In terms of the model presented below, the payoff from territorial control before WWI was equivalent to the payoff from recognized statehood.
-Players did not need to participate in the game described in the model.
-We limit our discussion to the post-WWII era, when the norm of home state veto is fully developed.}
-
-%Abkhazia, Ngorno-Karabakh, Transnistria, Somaliland, South Ossetia, and the Turkish Republic of Northern Cyprus have all controlled territory for over a decade without gaining recognition by more than four other states.
-
-%Non-recogition locks unrecognized states out of a states-only club whose members enjoy benefits in terms of both security and economic integration.
-For recognized states, the norm of territorial integrity lowers the cost of territorial defense by increasing the chance that foreign powers will intervene against, or at least sanction, an invader.\footnote{See Zacher (2001) on the norm of territorial integrity.} Recognition also allows entry into multilateral and bilateral trade agreements, dramatically bolsters access to foreign aid, and incorporates territory into international legal frameworks that make it easier to secure foreign investment (Milhalkanian (2004); Caspersen 2012: 40-45).\footnote{Transnistria represents a case where the gains from smuggling, particularly for the elite, may have outweighed these other economic costs in the 1990s (King 2001).
-However, as Moldova deepens ties with the EU, Transinistria's lack of access to EU markets and EU aid becomes a much larger relative cost to bear (Tudoroiu 2011).} In their most direct form -- loans and foreign aid -- these benefits are referred to as ``rents to sovereignty" (Collier and Hoeffler 2005).
-
-%Most unrecognized states are low income countries, and the costs of non-recognition grow over time as their economies further decay.
-Because the resources of unrecognized state governments are primarily focused on security, unrecognized states are generally characterized by a lack of public investment in infrastructure and education.
-They also face severe brain drain and a lack of private investment caused by the absence of security and restrictions on trade.
-
-%Non-settlement has costs for the home state as well, but they are simply not as high as those facing the unrecognized state.
-Instability in the border region and the diversion of military resources to monitor the \textit{de facto} border are costly, as is maintenance of economic sanctions against the unrecognized state.
-In Azerbaijan, where an oil boom sent GDP soaring in the mid-2000s, the unresolved secessionist conflict in Nagorno-Karabakh lead a massive military buildup to take precedence over other government spending.\footnote{Military spending increased 51\% in 2004-2005, and went up another 82\% in 2006 (International Crisis Group 2007)} Foreign investors may be more wary of investing in states that do not effectively control their own territory, and there are diplomatic costs as well.
-For Georgia, failure to resolve its outstanding secessionist conflicts has hindered progress toward NATO accession.
-
-The purpose of this paper is to identify the mechanisms that sustain unrecognized statehood as a stable equilibrium and to evaluate potential strategies through which actors may preserve the stability of this equilibrium, or disrupt it by inducing war or negotiated settlement.
-While there are many game theoretic treatments of civil conflict, we believe ours is the first formal model in which unrecognized statehood is addressed as an outcome.
-
-We develop a dynamic four-player model that focuses on the core conflict between the home state and the secessionist elite, but also incorporates the actions of outside states with interests in the outcome.
-These outside actors are a patron, which prefers independence to reunification, and a fourth player we denote as player $c$ who represents one or more members of the international community.
-Player $c$ is assumed to prefer reunification to independence.
-
-The dynamic game has an infinite horizon and proceeds as follows.
-At the beginning of each period, the payoffs for the secessionists from unrecognized statehood---to which we refer as the status quo---deteriorate by a fixed amount, representing the ongoing costs of non-recognition.
-Next, the patron and international community can, in turn, make investments to augment the payoffs of the home state government and the secessionists.
-Finally, the home government and secessionists play a simultaneous stage game in which each decides whether to initiate conflict, maintain the status quo of unrecognized statehood, or cede the issue of status.
-
-We provide a set of parameter restrictions under which perpetual unrecognized statehood is the outcome of a stationary Markov equilibrium of this game.
-%This requires that the patron is both willing and able to make investments sufficient to (a) offset the deterioration in the secessionists' status quo payoffs and (b) deter the international community from making investments that would induce a negotiated settlement.
-That is, it must be that the prospect of reunification is more important for the patron to avoid than for the international community to achieve.
-Additionally, in order for the status quo to persist, preferences must be such that recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve; otherwise, the international community will not be able to deter the patron from making investments to instigate recognition.
-Somewhat counterintuitively, in this equilibrium a patron will expend resources every period if necessary to sustain the status quo of unrecognized statehood even when the patron would prefer that the secessionists achieve recognized independence.
-This occurs when the international community's preferences against recognition are sufficiently strong---as is almost always likely to the case---to make it prohibitively expensive for the patron to facilitate recognition in the face of counteractivity by the international community.
-
-In the end, the status quo outcome is observed when the patron---who faces a much tighter budget constraint---outspends an international community with access to a much larger pool of resources.
-This occurs because each outside actor's willingness to pay to achieve its most preferred outcome is outstripped by the other's desire to avoid its least desired outcome.
-We are left with an unresolved conflict with no clear winner, and even if there is no official loser, all parties pay significant costs.
-
-We go on to analyze a tool that has been one of the most frequently employed by the international community in these types of territorial disputes: economic and military sanctions.
-We show that, although the goal of such sanctions is presumably to encourage secessionists to cede the issue of status, their most pronounced effect is to increase the range of conditions under which war occurs.
-%Instead, the patron may compensate for the additional cost of sanctions through increased investments or the secessionists may turn to fighting instead of ceding if the gap between the payoffs from the status quo and reunification are too large.
-Alternatively, if the secessionists' military capabilities are sufficiently weakened, the home state government may be emboldened to initiate fighting itself.
-%Thus while sanctions can be effective in only a limited range of cases, they can counterproductively lead to increased conflict in others. 
-In contrast, we argue that members of the international community (player $c$) are capable of intervening through other means to overcome the influence of the patron and induce settlement if they are sufficiently motivated to do so.
-Policy options include the enforcement of bargains and the granting of positive inducements for settlement, rather than direct pressure on the unrecognized state via sanctions.
-
-\subsection{The Empirical Landscape}
-
-The contributions of this paper are primarily theoretical, but it is useful to introduce the empirical landscape that informs our model.
-%We reference individual cases throughout to illustrate our theory and to justify its assumptions.
-Table 1 catalogs the post-WWII cases of unrecognized statehood, separated according to their current status.
-We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state.\footnote{This excludes, for example, the territories governed by FARC in Colombia because FARC does not seek recognized statehood and includes cases, like South Ossetia, where recognition is obtained by only a handful of other states.
-We also exclude entities, like the Slovak Republic, whose secession was not opposed by the state from which it seceded.
-We acknowledge that other scholars might differ with us with regard to the inclusion or exclusion of specific cases.
-We assert that our model is a useful tool for analyzing each of the cases in this table; it may be useful in explaining some of the ambiguous cases we exclude as well.} All six current unrecognized states have now been unrecognized for over twenty years.
-
-%\pagebreak
-\begin{center}
-\includegraphics[width=6.25in]{Table_1_082214.pdf}
-\end{center}
-
-Of those cases in the table, all but Moheli controlled territory and govern(ed) populations for at least two years.
-These cases represent the most successful cases of attempted secession in the Post-WWII era, and yet eventual military defeat by the home state is still the modal form of resolution.
-Recognition by the home state is similarly rare, occurring in only three cases and never except as a direct result of concessions won on the battlefield.
-In cases where recognition by the home state or the right to a referendum on independence is not secured as part of the initial peace agreement, it has not historically been forthcoming.
-Only four cases of negotiated reunification are observed: secessionists who are strong enough to secure and retain territorial control are rarely willing to surrender their independence at the bargaining table, even though the chances of eventual recognition are vanishingly slim.
-Thus, the number of long-running, costly stalemates has been substantial, most of them eventually ending in military reconquest by the home state.
-
-\subsection{Unrecognized States in the Literature}
-%Political scientists care a great deal about both civil conflict and about state formation, but unrecognized states do not sit neatly in either literature.
-The literature on civil war is focused primarily on war onset (e.g.
-Fearon and Laitin 2003; Hegre and Sambanis 2006), war intensity and duration (e.g.
-Collier and Hoeffler 2004; Cunningham 2006), and the durability of post-conflict peace (e.g.
-Hartzell, Hoddie and Rothchild 2001).
-Unrecognized statehood does not fit neatly into these areas of study because, while unrecognized states begin, and often end, through violent conflict, periods of unrecognized statehood generally contain little, if any, fighting.
-%Unrecognized statehood often represents a relatively stable end to civil war, but the cessation of fighting is not coterminous with resolution of the conflict in any meaningful way.
-We argue that persistent unrecognized statehood is not a successful resolution to secessionist civil war; it is a costly and normatively bad outcome in its own right, and it needs to be analyzed as such.
-%Unrecognized states also fall outside most treatments of state formation, because most unrecognized states never achieve recognition (or have not achieved it yet).\footnote{For example, Roeder (2007) treats unrecognized states simply as failures to gain recognized statehood, not as outcomes to be analyzed in their own right.}
-
-The political science literature on civil war is focused primarily on war onset, %\footnote{e.g. Fearon and Laitin 2003, Hegre and Sambanis 2006}
-war intensity and duration, %\footnote{e.g. Collier and Hoeffler 2004; Cunningham 2006} 
-and the durability of post-conflict peace. %\footnote{e.g. Hartzell, Hoddie and Rothchild 2001} 
-Unrecognized statehood does not fit neatly into these areas of study because, while unrecognized states begin, and often end, through violent conflict, periods of unrecognized statehood generally contain little, if any, fighting.
-Unrecognized states also fall outside most treatments of state formation, because most unrecognized states never achieve recognition (or have not achieved it yet).\footnote{For example, Roeder (2007) treats unrecognized states simply as failures to gain recognized statehood, not as outcomes to be analyzed in their own right.}
-
-The first literature to address unrecognized states directly was grounded in comparative politics, and a robust area-studies literature exists around each of the current cases of unrecognized statehood.\footnote{Lynch (2004), King (2001), Stanislawski (2008), and Bakke et al.
-(2013) provide notable treatments of the Former Soviet cases, and a pair of edited volumes by Bahcheli et al.
-(2004) and Kingston and Spears (2004) each compile broader sets of case studies.} More recent literature has addressed wider ranges of cases and made important conceptual progress identifying patterns and commonalities across cases (e.g.
-Pegg, 1998; Kolst\o, \ 2006; Geldenhuys, 2009; Caspersen and Stansfield, 2011; Caspersen 2013).
-Bakke (2011) summarizes four explanations for the persistence of unrecognized states, each of which plays a role in the theory we develop here.
-First, economic and political elites benefit from the status quo (e.g.
-King, 2001).
-Second, the secessionists have sufficient military capacity to make military reconquest costly (e.g.
-Kolst\o, \ 2006).
-Third, secessionists enjoy external support from a patron (e.g.
-Lynch, 2002; Stanislawski, 2008).
-Lastly, the governments of unrecognized states are internally legitimate, enjoying support from residents of the unrecognized state (e.g.
-Lynch, 2002, Kolst\o, \ 2006).
-
-However, the literature continues to lack a clear general theory specifying the conditions under which this outcome emerges and persists.
-
-One of the major contributions of this article is to provide a unified analytic framework for understanding the mechanisms sustaining unrecognized statehood as a stable equilibrium.
-By modeling unrecognized statehood formally, we move away from a case-by-case treatment toward development of a rigorous general theory.
-The model allows us to assess the conditions under which unrecognized statehood persists, and those under which war and negotiated settlement occur.
-It also allows us to evaluate various strategies available to actors, particularly states and coalitions of states who want to facilitate a peaceful and permanent resolution.
-
-Although we are not aware of any formal models of unrecognized statehood there is a growing economics literature that employs game theoretic models of conflict.
-Grossman (2004) offers an early model of territorial disputes in which the divisibility of the contested territory is an important parameter.
-In Hirshleifer et al., (2009), conflict can be avoided when there is common knowledge, a common rate of time preference, and players can make a series of small concessions.
-In our context, the contested territory is indivisible; this implies that peace would require very large concessions so that this pathway of using small compromises to avoid conflict is not available.
-
-In contrast to Bueno de Mesquita (2013), which exemplifies a class of literature that examines how rebel groups make decisions about resistance strategies, we begin our analysis at the point where initial rebellion has already been successful to the point of gaining \textit{de facto} control of the disputed territory.
-
-Another strand of the literature explicitly models the bargaining process between potential adversaries, either with asymmetric (Yared, 2010) or complete information (Jackson and Morelli, 2007; Schwartz and Sonin, 2008; Bevi\'{a} and Corch\'{o}n, 2010; McBride and Skaperdas, 2014).
-Acemoglu et al.
-(2012) develop a dynamic model of resource wars featuring limited commitment.
-We abstract from questions of resource allocation and bargaining in order to focus on the dynamics created by the involvement of outside players: in the case of unrecognized states, the international community and patron states.
-
-We next turn to presenting the model.
-In Section 3, we present a set of parameter restrictions under which unrecognized statehood is a stable, long-run equilibrium outcome.
-In Section 4, we analyze the potential effects of sanctions on the resolution of territorial disputes while Section 5 explores extensions of the model.
-Section 6 evaluates policy implications of the model, assessing various strategies available to outside actors that seek to induce their preferred outcomes.
-Section 7 concludes.
-
 \section{A Model of Unrecognized Statehood} 
@@
 
+\textbf{<<COMP: Place Figure 1 about here>>}
+
 \subsection{Details of the Dynamic Game}
@@
 There are an infinite number of discrete periods $t=1,2,\ldots$. Play proceeds in each period $t$ as follows (and as shown in Figure 1) until an absorbing state is reached.
-
-\begin{figure}
-\includegraphics{Timeline2.pdf}
-\caption{Timeline}
-\end{figure}
 
@@
 \vspace{2pt}
-\begin{center}
-\begin{tabular}{|  c| c | c | c |} 
-\hline
-  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
-	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
-	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
- \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
-\end{tabular}
-\end{center}
+\textbf{<<COMP: Place Figure 2 about here>>}
 \vspace{2.5pt}
@@
 
-%The Tamil war of independence began in 1983, with India initially serving as a patron of the secessionists.
-India withdrew its support following a failed peacekeeping intervention in 1989, after which the secessionists relied on economic support from the Tamil diaspora -- i.e.
-ethnic Tamils living abroad.
-
-%An empirical investigation of post-WWII cases provides illustration of these propositions.
-We observe in practice that, when the home state imposes sanctions on the secessionists, it is most often the home state that initiates the return to violence.
-Table X provides 13 cases in which a period of unrecognized statehood (\emph{de facto} territorial control by the secessionists) was ended through decisive military victory by one side or the other.
-In five of these cases, sanctions by the home state were present against the secessionists.
-In four of the cases the home state initiated the fighting; in one case the secessionists initiated; in six cases either the initial conflict through which the secessionists achieved \emph{de facto} territorial control never actually ceased; and in two cases
-
-%weakening the unrecognized state militarily is equivalent to strengthening the home government militarily
-
 %\vspace{3pt}
@@
 A flow payment of carrots, even backed by the promises of a ``neutral" third party, may not be credible in the eyes of the secessionists, which could explain the frequent resort to sticks.
-
-%As we consider tools such as sanctions aimed at disrupting the status quo equilibrium, it is important to note that not every force exerted on the situation will lead to a change in strategy.
-Only if a knife-edge condition exists or if large enough investments are made or sanctions applied to overcome a buffer will the perturbations of payoffs lead to a change in the actions of the decision makers.
-Some situations very much favor the status quo because of relatively high status quo payoffs compared to the alternatives.
-These situations are hard to perturb even if significant pressures are exerted from the outside players.
-The stability of the game can be amplified by strong preferences of both $p$ and $c$ which lead them to contribute rather than allow their least-preferred outcome.
-
-\section{Extensions}
-\label{sec:ext}
-%\subsection{Introducing Uncertainty}
-%The assumption of perfect information may be somewhat unrealistic, and here we explore adjusting the model to accommodate some uncertainty on the part of $c$ and $p$.
-Payoffs as described in the stage game are those perceived by $g$ and $s$.
-If these payoff values are not precisely known by $c$ and $p$, enough uncertainty may be present in order for both those players to contribute resources in equilibrium.
-
-%Here, $c$ and $p$ observe a random draw of the stage game payoffs, with each payoff drawn independently and with a mean matching the original stage game payoffs.
-After viewing these (uncertain) payoffs, $c$ and $p$ each invest some level of resources, altering the payoff structure before it is observed (accurately) by $s$ and $g$.
-Based on this altered payoff structure, $c$ and $g$ choose their strategies.
-Uncertainty can lead to outcomes where either $c$ or $p$ invests too much, wastes resources, or makes a more severe misstep, such as $p$ investing too little and allowing reunification to occur.
-
-%In practice we expect that uncertainty is lower for the patron than for $c$ because the patron is close to, and intimately involved in, the conflict and therefore may have a better grasp of the two players' payoffs than do other outside states or groups.
-This makes over-contribution by $c$ more likely than under-contribution by the patron.
-%Of particular note here are the resources the European Union and its members have devoted toward promoting a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.
-
-%In the equilibria constructed in Section~\ref{sec:main}, full information on all sides implies that one of the parties would, in equilibrium, not give any resources.
-By adding some uncertainty about payoffs, we can account for the observed fact that $c$ sometimes expends resources unsuccessfully.
-This type of spending can also be explained as non-strategic spending -- i.e.
-spending aimed at goals other than promoting settlement, like pure humanitarianism.
-
-%\subsection{Partial Recognition}
-%The norm of home state veto gives recognition by the home state its significance: recognition by the home state is the core demand of the secessionists in our model.
-In cases where the secessionists can gain recognition from a large number of states without first gaining recognition by the home state, the status quo is a less costly outcome for the secessionists.
-This increases the patience of the secessionists.
-In the model, if the payoffs to the secessionist leaders for the status quo rise, the equilibrium decision by the secessionist elite to play "status quo" becomes more stable.
-Less support from a patron is needed and deterioration of conditions for the unrecognized state does not immediately disrupt the status quo equilibrium.
-Because the home state is aware of this payoff change, it may realize that holding out is less likely to be fruitful -- reabsorption of the secessionists by the home state is less likely to occur through the secessionists ceding.
-Additionally, if the international community's preferences shift toward protection of the secessionists following mass atrocity crimes, expected payoffs from a military conflict would decrease for the home state.
-It is possible that the status quo payoffs are lowered for the home state as well, due to lack of support for its cause by the international community.
-If these dynamics lead to a sufficiently low home state payoff for both status quo and war, the the home state will cede.
-
-%With the official adoption of the Responsibility to Protect doctrine by the United Nations and the precedent of Kosovo, home states are put on notice that the commission of mass atrocity crimes against the residents of seceding entities may lead to recognition of that entity by other countries.\footnote{For a discussion of the sanction theory of recognition, see Berlin (2009).} However, the norm of home state veto remains strong, and so long as further atrocities are not committed, the Responsibility to Protect doctrine is unlikely to affect recognition of the unrecognized states already in existence.
-
-\subsection{Uncertainty and Outside Interactions Between the Patron and \texorpdfstring{$c$}{c}}
-While we present a model with perfect information, there are some empirical cases in which it appears that various players have acted on mistaken beliefs about the strength or payoffs of other players.
-Modeling uncertainty falls outside the scope of this paper, but uncertainty does offer an additional plausible explanation for off-equilibrium-path behavior that we observe in empirical cases.
-For example, uncertainty regarding the payoff structure facing the secessionists may explain failed and costly efforts by the European Union and its members (i.e.
-player $c$) to promote a negotiated reunification of the Turkish Republic of Northern Cyprus into Cyprus.
-
-The game we model, with the control of the unrecognized state at stake, is only one of several strategic games in which the patron and $c$ may be interacting at any given time, and linkages between games are possible.
-We do not model any direct exchange of resources or imposition of harm between $c$ and the patron, but we take the implications of these possible outside interactions seriously.
-The willingness and ability of either party to contribute resources within the game we model may be affected by their interaction with one another in other contexts.
-As we will discuss in greater detail below, in a number of cases a patron has withdrawn support for an unrecognized state in response to pressure exerted by actors in the international community ($c$) in other venues.
-
-\subsection{If the Patron Withdraws Support}
-Support by a foreign patron is, in almost all cases, necessary for the persistence of unrecognized statehood (Kolst\o, \ 2006; Caspersen, 2013).
-When there is no patron, or when the patron withdraws its support, military reconquest by the home state is likely.
-The only unrecognized state currently in existence which has been able to survive without a patron is Somaliland.
-Somaliland has been able to survive as long as it has because of the extreme weakness of the home state (Somalia).\footnote{Owtram (2011) argues that, over time, unrecognized states can become sufficiently "consolidated" as to no longer be dependent on a patron for support.
-However, no cases other than Somaliland support this claim empirically.}
-
-%In the model, $g$ or $s$ has an incentive to fight rather than stay in the status quo if the expected discounted stream of payoffs is greater by fighting.
-Specifically, if condition (2) for the status quo equilibrium is not met, then one player will prefer war to both ceding and the status quo.
-In the case of most prolonged stalemates, strategic spending by the patron averts these outcomes.
-The patron provides military assistance to the secessionists at such a level that $g$ does not prefer to initiate conflict, and provides sufficient economic and humanitarian assistance to prevent $s$ from preferring war to a continuation of the status quo.
-However, if there is no patron or if the patron is not sufficiently interested, war may become a more attractive outcome than the status quo for at least one of the parties.
-
-%The cases where there is no patron, such as in Chechnya, this is relatively easy to explain. 
-The case of Chechnya is quite typical of unrecognized states without patrons.
-Chechnya fought and won its initial war of secession at a time when the home state (Russia) was in political and economic disarray following the collapse of the Soviet Union.
-As Russia gradually recovered and strengthened, Chechnya had no patron support to offset the relative decline in its military capabilities.
-Over time the war lottery became progressively more skewed in favor of Russian victory, the payoffs to ceding for the Chechen secessionists remained extremely low, and the Russian government invaded and reconquered Chechnya in 1999.
-
-It is worth exploring, however, the reasons why a patron might support a secessionist group during its initial rebellion and then withdraw support at a later date.
-Patrons' strategic interests in the unrecognized state vary from patron to patron, and both budget constraints and salience of interest vary over time.
-Returning to the case of Tamil Eelam, domestic political concerns (primarily Tamil ethnic solidarity) induced a modest level of Indian support for the secessionists  from 1983-1987.
-These domestic political concerns were eventually outweighed by broader strategic security concerns and a desire for regional stability.
-In 1987 the Indian government signed a peace accord with Sri Lanka (the home state) and largely withdrew their support from the secessionists.
-
-%, even sending in peacekeepers that later clashed with the secessionists militarily (Singer 1992).
-The Tamil diaspora funded the secessionist military until 2009, when this support proved insufficient and the territory was reconquered by the Sri Lankan government.\footnote{On a smaller scale, the Isaaq diaspora has also fulfilled some of the roles of the patron in Somaliland (Galipo 2011).}
-
-As noted in the section on outside interactions between the patron and player $c$, the patron's decision to withdraw support for the secessionists is sometimes motivated by interactions between the patron and $c$ that we do not model directly.
-Empirically, we observe a number of cases in which $c$ applies pressure directly to encourage the patron to withdraw support from the unrecognized state.
-In an extreme example involving both sanctions and direct military confrontation, NATO coerced Serbia into, among other things, withdrawing its support from Republika Srpska and Republika Srpska Krajina, both of which were then reconquered militarily.\footnote{For an excellent discussion of the case of Republika Srpska, see Zahar, 2004.} %In the final section of the paper, we consider direct coercion of the patron among the strategies available to player $c$.
-
-\section{Policy Implications: Options for Player \texorpdfstring{$c$}{c}}
-
-Sanctions imposed by the home state, often co-enforced by player $c$, aim to disrupt the status quo equilibrium and force the secessionists to cede.
-However, as Section \ref{sec:sanctions} demonstrates, sanctions also increase the risk of war, something player $c$ prefers to avoid.
-
-Fortunately, there is a better way.
-If instead of enforcing sanctions, player $c$ tries to promote settlement by supplementing the payoffs from unification, it is able to induce negotiated settlement without simultaneously increasing the risk of war.
-This can be done either through promises of benefits to the unrecognized state provided directly by $c$, like aid, or by a commitment from $c$ to serve as a third-party guarantor of side payments promised by the ceding side: both directly increase $L_i$ with the goal of raising it above $Q_i$ for the player who will cede in the settlement.
-In the case of contingent promises of aid, the calculation is relatively straightforward: 1) the promise of aid must be credibly contingent on negotiated settlement, and 2) the aid offered must be valued more highly than the concessions required to reach an agreement.
-It is the second condition that is most problematic.
-Because both sides place such a high value on status (independence vs.
-reunification), even large amounts of aid are likely to be valued less than the concessions necessary to reach an agreement.
-
-Serving as a third-party guarantor of autonomy rights is a way for player $c$ to potentially overcome problems of indivisibility and commitment and help the parties reach a credible compromise on status (Walter, 2002).
-However, this strategy is only tenable when the only impediment to settlement is the unenforceability of a bargain, and when $c$ is credible as an enforcer of that bargain.
-
-%Credible enforcement of future autonomy rights can be viewed either as increasing the value of available side payments or as making the central issue of contention divisible.
-In either view, a range of previously untenable agreements are made possible.
-
-In Southern Sudan, third-party actors, including the UN, invested substantial resources to help negotiate a settlement to the initial war of secession and to ensure that the Sudanese government both allowed the promised referendum and respected its results.
-While the UN and others invested resources in Southern Sudan to enforce independence, not autonomy, they have demonstrated that outside actors are capable of enforcing difficult concessions by the home state government.
-This bodes well for the future credibility of organizations like the UN as third-party enforcers.
-However, the role of outside actors in enforcing other past agreements might give secessionists pause.
-For example, a referendum on independence in Western Sahara, which the UN ruled to be necessary more than thirty years ago, has never come to pass.\footnote{For a thorough analysis of the Western Sahara case, see Zunes and Mundy (2010).} Nonetheless, it is possible for an outside state or coalition to invest resources to enforce agreements, allowing for negotiated settlements that would otherwise be impossible to reach.
-
-To show that it is possible for an actor like player $c$ to enforce the terms of negotiated agreements at a reasonable cost is not sufficient to imply that such an outcome is likely.
-The political will necessary to achieve success in Southern Sudan was motivated largely by the magnitude of the atrocities that accompanied the war of secession, and enforcement was made credible, in part, due to the weakness of Sudan relative to the outside states involved.
-Enforcing the terms of an agreement between Georgia, South Ossetia, and Russia, for example, would be more difficult.
-
-It is also possible for actors like $c$ to affect the payoffs of the patron through interactions in other games outside of our model.
-Such actions would manifest themselves within the model as reductions in the patron's willingness to pay to sustain the status quo.
-If the patron is unwilling to pay to sustain the status quo, the status quo payoffs of the secessionists will decline over time, eventually leading to either war or negotiated settlement.
-Under these conditions, the within-game costs to $c$ of inducing negotiated reunification also fall.
-
-In this section we have argued that outside states are capable of imposing their preferred outcome, including peaceful reunification.
-The key, however, is motivation: actors like player $c$ are capable of inducing peaceful reunification when they are willing to invest the resources necessary.
-However, strong preferences of secessionists against reunification and the opposing intervention of the patron make the costs of such interventions prohibitively high in most cases.
-Unrecognized statehood is a stable equilibrium because, while there are many actors in the international community that share the preferences we ascribe to $c$, they are usually unwilling to invest sufficient resources to outspend the patron and induce their preferred outcome.
-
-\section{Conclusion}
-In this paper we establish unrecognized states as an outcome of interest for social scientists and provide a unified framework for analyzing that outcome and its alternatives.
-Current events, particularly Russia's military and economic support of secessionist rebels in Eastern Ukraine, suggest that the phenomenon of unrecognized statehood will not fade from relevance soon.
-While the Ukrainian situation remains fluid, as of October 2015, the status quo equilibrium we model is a plausible future.
-The equilibrium could emerge as follows: with Russian support, the secessionists solidify \emph{de facto} control of parts of Eastern Ukraine, but risk of direct confrontation with the West is sufficiently high to prevent Russian annexation of Eastern Ukraine.\footnote{Annexation of Crimea was achieved at a reasonable cost to Russia.
-The costs of annexation Eastern Ukraine would likely be substantially higher.} Further Russian economic aid keeps the status quo payoffs of the secessionists sufficiently high that they decline negotiated reunification.
-If necessary, contributions to the home state by U.S.
-and the EU (player $c$) strengthen the home state enough to prevent the secessionists from overthrowing the Ukrainian government and forcing recognition.
-In twenty years, Eastern Ukraine may look a good deal like South Ossetia does today.
-
-While the importance of outside actors in civil conflict has been widely acknowledged in the empirical literature, it is rarely modeled formally.
-We introduce a dynamic four-player model that captures the core strategic interactions of the secessionist elite and the home state central government, as well as the interventions of outside actors with interests in the outcome.
-This allows us both to examine the means through which patron states sustain unrecognized statehood as a stable equilibrium, and to rigorously analyze the strategies available to other outside actors to pursue reunification.
-It is not always in the interests of outside states to bear the costs of inducing their preferred outcome, but we identify the mechanisms through which this is possible and the thresholds that must be overcome.
-
-In the model we present, we show that unrecognized statehood can emerge as an equilibrium outcome even when it is a terrible outcome for all players involved.
-The patron, even when it prefers outright independence, is willing to bear costs in every period to uphold an outcome that is far short of its ideal, and which imposes high costs on others as well.
-The international community, though wealthier than the patron, does does not outspend the patron to force reunification because the patron's desire to avoid reunification is stronger than the international community's desire to achieve it.
-
-By providing economic and other aid to the unrecognized state, the patron keeps the secessionist elite's payoffs from the status quo high enough to prevent a negotiated reunification.
-The stability of this equilibrium is abetted by the indivisible nature of independence and the difficulty of enforcing autonomy as a condition of reunification.
-In cases where there is no patron or where the patron eventually becomes unwilling to continue its support, the result has almost always been violent reconquest by the home state.
-
-Our model also suggests, however, that the historical pattern of costly stalemate followed by violent resolution is not the only possible path.
-We show that the stabilizing effect of the patron can be overcome by a sufficiently motivated international community.
-While the most frequently employed means through which outside actors attempt to induce settlement---sanctions---also increases the risk of war, we show that it is possible for outside actors to induce their preferred outcome without running this risk.
-In particular, we suggest that they can provide positive inducements for settlement and serve as a third-party guarantor of negotiated settlements in which unrecognized states rejoin the home state as autonomous regions.
-It is often not the lack of available means that prevents outside actors from inducing their preferred outcome, but rather the lack of will.
-
-\pagebreak
-
-\section{References}\label{A:references}
-\hangindent=1cm
-
-\begin{enumerate}[1.]
-
-\item \hangindent=1cm Acemoglu, Daron, Golosov, Mikhail, Tsyvinski, Aleh, and Pierre Yared. 2012. "A Dynamic Theory of Resource Wars." 
-\emph{The Quarterly Journal of Economics} 127: 283-331.
-
-\item \hangindent=1cm Bahcheli, Tozun, Barry Martmann, and Henry Srebrnik, eds. 2004 "De Facto States: The Quest for Sovereignty." New York, NY: Routledge.
-
-\item \hangindent=1cm Bakke, Kristin M.
-2011 "After the war ends: violence and vialibity of post-Soviet unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
-\emph{Unrecognized States in the International System}.
-New York, NY: Routledge.
-
-\item \hangindent=1cm Bakke, Kristin M., John O'Loughlin, Gerard Toal, and Michael D.
-Ward.
-2013.
-"Convincing State-Builders?
-Disaggregating Internal Legitimacy in Abkhazia."  \emph{International Studies Quarterly.}
-
-%\item \hangindent=1cm Balch-Lindsay, Dylan, and Andrew J. Enterline. 2000. "Killing Time: The World Politics of Civil War Duration, 1820-1992." \emph {International Studies Quarterly} 44 (4):615-642.
-
-\item \hangindent=1cm Balch-Lindsay, Dylan, Andrew J. Enterline, and Kyle A. Joyce.  2008.  "Third-Party Intervention and the Civil War Process."  \emph{Journal of Peace Research} 45: 345-63.
-
-\item \hangindent=1cm Baliga, Sandeep, and Tomas Sjostrom. 2012. "The Strategy of Manipulating Conflict." \emph{The American Economic Review} 102 (6): 2897-2922.
-
-\item \hangindent=1cm Bevi\'{a}, Carmen and Luis C. Corch\'{o}n. 2010. ``Peace Agreements without Commitment.'' \emph{Games and Economic Behavior} 68 (2): 469-487.
-
-\item \hangindent=1cm Bueno de Mesquita, Ethan. 2013. ``Rebel tactics.'' \emph{Journal of Political Economy} 121 (2): 323-357.
-
-\item \hangindent=1cm Caspersen, Nina. 2013. "Unrecognized States: The Struggle for Sovereignty in the Modern International System." Malden, MA: Polity Press.
-
-\item \hangindent=1cm Caspersen, Nina, and Gareth Stansfield. 2011. "Unrecognized States in the International System." New York, NY: Routledge.
-
-%\item \hangindent=1cm Christin, Thomas, and Simon Hug. 2012. "Federalism, the Geographic Location of Groups, and Conflict." \emph {Conflict Management and Peace Science} 29 (1): 93-122.
-
-\item \hangindent=1cm Coggins, Bridget. 2011. "Friends in High Places: International Politics and the Emergence of States from Secessionism." \emph {International Organization} 65 (03): 433-467.
-
-%\item \hangindent=1cm Collier, Paul, and Anke Hoeffler. 2004. "Greed and Grievance in Civil War." \emph {Oxford Economic Papers} 56 (4): 563-595.
-
-%\item \hangindent=1cm Collier, Paul and Anke Hoeffler. 2005. "Resource Rents, Governance, and Conflict." \emph {Journal of Conflict Resolution} 49 (4): 625-633.
-
-\item \hangindent=1cm Crawford, James. 2006. "The Creation of States in International Law." Oxford: Oxford University Press.
-
-%\item \hangindent=1cm Cunningham, David E. 2006. "Veto Players and Civil War Duration." \emph {American Journal of Political Science} 50 (4): 875-892.
-
-\item \hangindent=1cm Doyle, Michael W., and Nicholas Sambanis. 2006. "Making War and Building Peace: United Nations Peace Operations." Princeton: Princeton University Press.
-
-\item \hangindent=1cm Elbadawi, Ibrahim, and Nicholas Sambanis. 2000. "External Interventions and the Duration of Civil Wars." 2433.
-
-%\item \hangindent=1cm Fazal, Tanisha. 2007. "State Death: The Politics and Geography of Conquest, Occupation, and Annexation." Princeton: Princeton University Press.
-
-%\item \hangindent=1cm Fearon, James D. 1995. "Rationalist explanations for war." \emph {International Organization} 49 (03): 379-414.
-
-\item \hangindent=1cm Fearon, James D.  1994.  "Domestic Political Audiences and the Escalation of International Disputes."  \emph{The American Political Science Review} 88: 577-92.
-
-\item \hangindent=1cm Fearon, James D., and David D. Laitin. 2003. "Ethnicity, Insurgency, and Civil War." \emph {American Political Science Review} 97 (01): 75-90.
-
-\item \hangindent=1cm Fearon, James D. 2004. "Why Do Some Civil Wars Last So Much Longer than Others?" \emph {Journal of Peace Research} 41 (3): 275-301.
-
-\item \hangindent=1cm Geldenhuys, Deon. 2009. "Contested States in World Politics." Basingstoke: Palgrave McMillan.
-
-%\item \hangindent=1cm Ghai, Yash, and Anthony Regan. 2006. "Unitary State, Devolution, Autonomy, Aecession: State Building and Nation Building in Bouganville, Papua New Guinea." 95 (386).
-
-%\item \hangindent=1cm Hartzell, Caroline, Matthew Hoddie, and Donald Rothchild.
-2001.
-"Stabilizing the peace after civil war: An investigation of some key variables." \emph {International Organization} 55 (01): 183-208.
-
-%\item \hangindent=1cm Hegre, Havard, and Nicholas Sambanis. 2006. "Sensitivity Analysis of Empirical Results on Civil War Onset." \emph {Journal of Peace Research} 50 (4): 508-535.
-
-\item \hangindent=1cm Grossman, Herschel I. 2004. ''Peace and war in territorial disputes.'' No. w10601. National Bureau of Economic Research.
-
-\item \hangindent=1cm Hirshleifer, Jack, Boldrin, Michele, and David K Levine. 2009. ''The slippery slope of concession.'' \emph{Economic Inquiry} 47 (2): 197-205.
-
-\item \hangindent=1cm Jackson, Matthew O., and Massimo Morelli. 2007. "Political bias and war." \emph{The American Economic Review} 97 (4): 1353-1373.
-
-\item \hangindent=1cm King, Charles. 2001. "The Benefits of Ethnic War: Understanding Eurasia's Unrecognized States." \emph{World Politics} 53 (4): 524-552.
-
-\item \hangindent=1cm Kingston, Paul, and Ian S. Spears, eds. 2004. "States-Within-States." New York, NY: Palgrave McMillan.
-
-\item \hangindent=1cm Kolst\o, P{\aa}l. 2006. "The Sustainability and Future of Unrecognized Quasi-States." \emph {Journal of Peace Research} 43 (6): 723-740.
-
-\item \hangindent=1cm Licklider, Roy. 1995. "The Consequences of Negotiated Settlements in Civil Wars, 1945-1993." \emph {American Political Science Review} (89): 681-90.
-
-%\item \hangindent=1cm Levy, Jack S. 1987. "Declining Power and the Preventive Motivation for War." \emph{World Politics} 40: 82-107.
-
-\item \hangindent=1cm Lynch, Dov. 2004. "Engaging Eurasia's Separatist States." Washington, DC: U.S. Institute of Peace.
-
-\item \hangindent=1cm Mailath, George J., and Larry Samuelson. 2006. "Repeated Games and Reputation." New York, NY: Oxford University Press.
-
-\item \hangindent=1cm McBride, Michael, and Stergios Skaperdas. 2014. "Conflict, settlement, and the shadow of the future." \emph{Journal of Economic Behavior $\&$ Organization} 105: 75-89.
-
-\item \hangindent=1cm Owtram, Francis.
-2011.
-"The foreign policies of unrecognized states." in: Caspersen, Nina, and Gareth Stansfield.
-\emph{Unrecognized States in the International System}.
-New York, NY: Routledge.
-
-\item \hangindent=1cm Pegg, Scott. 1998. "International society and the de facto state." New York: Ashgate.
-
-%\item \hangindent=1cm Powell, Robert. 1999a. "The Modeling Enterprise and Security Studies." \emph {International Security} 24 (2): 97-106.
-
-%\item \hangindent=1cm Powell, Robert. 1999b. \emph{In the Shadow of Power: States and Strategies in International Politics.} Princeton: Princeton University Press.
-
-\item \hangindent=1cm Regan, P.M. 2002. "Civil wars and foreign powers: Outside intervention in intrastate conflict." Ann Arbor: University of Michigan Press.
-
-\item \hangindent=1cm Roeder, Philip G. 2007. "Where nation-states come from: institutional change in the age of nationalism." Princeton: Princeton University Press.
-
-\item \hangindent=1cm Salehyan, Idean, Kristian Skrede Gleditsch, and David E.
-Cunningham.
-2011.
-"Explaining External Support for Insurgent Groups." \emph {International Organization} 65 (04): 709-744.
-
-\item \hangindent=1cm Schwarz, Michael, and Konstantin Sonin. 2008. "A theory of brinkmanship, conflicts, and commitments." \emph{Journal of Law, Economics, and Organization} 24 (1): 163-183.
-
-\item \hangindent=1cm Schultz, Kenneth A.
-2010.
-"The Enforcement Problem in Coercive Bargaining: Interstate Conflict over Rebel Support in Civil Wars." \emph {International Organization} 64 (02): 281-312.
-
-%\item \hangindent=1cm Singer, Marshall R. 1992. "Sri Lanka's Tamil Sinhalese Ethnic Conflict: Alternative Solutions." \emph {Asian Survey} 32 (8): 712-722.
-
-%\item \hangindent=1cm Spears, Ian S.
-2004.
-"States-Within-States: An Introduction to Their Empirical Attributes." In States-Within-States, eds.
-Paul Kingston and Ian S.
-Spears.
-New York: Palgrave McMillan.
-
-%\item \hangindent=1cm Spruyt, Hendrik. 1996. "The Sovereign State and its Competitors." Princeton, NJ: Princeton University Press.
-
-\item \hangindent=1cm Stanislawski, Bartosz H.
-2008.
-"Para-States, Quasi-States, and Black Spots: Perhaps Not States, But Not `Ungoverned Territories,' Either." \emph {International Studies Review} 10 (2): 366-396.
-
-%\item \hangindent=1cm Tir, Jaroslav. 2006. "Domestic-Level Territorial Disputes: Conflict Management via Secession." \emph {Conflict Management and Peace Science} 23 (4): 309-328.
-
-\item \hangindent=1cm van Houten, Pieter.
-1998.
-"The role of a minority's reference state in ethnic relations." \emph {European Journal of Sociology / Archives Europennes de Sociologie} 39 (01): 110-146.
-
-\item \hangindent=1cm Walter, Barbara F. 1997. "The Critical Barrier to Civil War Settlement." \emph {International Organization} 51 (03): 335-364.
-
-\item \hangindent=1cm Walter, Barbara F.. 2002. "Committing to Peace: The successful settlement of civil wars." Princeton: Princeton University Press.
-
-%\item \hangindent=1cm Walter, Barbara F. 2006. "Building Reputation: Why Governments Fight Some Separatists but Not Others." \emph {American Journal of Political Science} 50 (2): 313-330.
-
-\item \hangindent=1cm Yared, Pierre. 2010. ``A Dynamic Theory of War and Peace.'' \emph{Journal of Economic Theory} 145: 1921-1950.
-
-\item \hangindent=1cm Zacher, Mark W. 2001. "The Territorial Integrity Norm: International Boundaries and the Use of Force." \emph {International Organization} 55 (02): 215-250.
-
-\end{enumerate}
 
@@
 
+%\newpage
+
+%\theendnotes
+
 \newpage
 
-\theendnotes
+\begin{figure}
+\includegraphics{Timeline2.pdf}
+\caption{Timeline}
+\end{figure}
 
-\end{document}
+\newpage
+
+\begin{center}
+\begin{tabular}{|  c| c | c | c |} 
+\hline
+  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
+	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
+	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
+	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
+ \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
+\end{tabular}
+\end{center}
+
+\newpage
+\begin{center}
+\includegraphics[width=6.25in]{Table_1_082214.pdf}
+\end{center}

### 2017-07-01 theory.tex
cut down 'model' and 'the players' subsections

@@
 reunification, is both difficult to divide and highly valued by both sides.
-The secessionists seek recognized statehood, the home state seeks reunification, and these demands do not vary over time.
 The side payments that can be offered in exchange for the opponent's surrender of the independence/reunification issue are sharply limited by the absence of large concessions that can credibly be made (Walter 1997, 2002; Schultz 2010).
 
-One innovation of our model is the introduction of international actors.
-While the role of outside actors in determining the duration and outcome of civil conflicts is well documented (e.g., Elbadawi and Sambanis, 2000; Regan 2002; Balch-Lindsay et al., 2008), the role of these actors has not been incorporated into the modeling of these conflicts.\footnote{One exception is van Houten (1998), who models the patron state ("reference state'') as a player in ethnic conflicts but otherwise takes an approach quite different from ours.} This is true even of work that addresses the role of outside actors as potential third-party enforcers (Walter, 1997; 2002).
-
-The model presented here allows us to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}
+Because our model incorporates the incentives and actions of international actors, it is about to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}
 
@@
 
-We construct a model with four players: the secessionist elite ($s$), which seeks recognized independence; the central government of the home state ($g$) from which $s$ is attempting to secede, which seeks reunification; and two other players, the patron ($p$) and the international community ($c$).
-The latter two are outside actors---states, groups of states and/or individuals acting in concert---that have interests in the outcome of the attempted secession.
+We construct a model with four players: the secessionist elite ($s$), which seeks recognized independence; the central government of the home state ($g$) from which $s$ is attempting to secede, which seeks reunification; and two outside actors:the patron ($p$) and the international community ($c$).
 
 Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders.
-As discussed earlier, most states in the international system prefer any given secessionist conflict to be resolved by reunification.
-In practice we believe that in most cases there are many states that may act as player $c$ in our model, and often we observe groups of states like the OECD or the UN acting in this capacity.
-Modeling the international community as a single actor is a strong simplifying assumption, but by leaving the preferences of player $c$ quite general, the model can accommodate much of this diversity.\footnote{A variety of scholars, including Owtram (2011) and Coggins (2011) discuss the international relations of unrecognized states and variation across third-party states in their preferences with regard to unrecognized states.}
+In practice we often observe groups of states like the OECD or the UN acting in this capacity.
 
@@
 Turkey's support of the Turkish Republic of Northern Cyprus); 3) hope of eventual annexation of the disputed territory (e.g.
-Armenia's support of Nagorno-Karabakh).\footnote{While annexation is appealing to many patrons (and some unrecognized states), annexation is not an outcome we model.
-International norms against irredentism are strong, and the costs of annexing an unrecognized state appear to be prohibitively large in most cases (e.g.
-Zacher, 2001).
-Prior to Russia's annexation of Crimea, all post-WWII cases of annexation involved the annexation of colonial territory, rather than parts of the metropole.
-Israel's occupation of the Golan Heights and the West Bank are possible exceptions here.}
-
-We acknowledge that there may exist patrons whose most-preferred outcome is the status quo.
-This naturally makes a status quo equilibrium easier to achieve, as there is an actor who strictly prefers the status quo and can expend resources to make it more likely.
-We choose to examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely.
+Armenia's support of Nagorno-Karabakh).
+Although there may exist patrons whose most-preferred outcome is the status quo, we examine the case where the patron's most preferred outcome is independence because this is the condition under which the status quo is least likely.
 We will show that even in this circumstance the status quo remains an equilibrium outcome.

### 2017-07-01 theory.tex
first pass at cutting down model section

@@
 
-The game begins at a status quo in which the secessionist elite controls at least some of the disputed territory, but cannot gain international recognition unless the central government cedes its claim to the territory.
+The game begins at a status quo in which the secessionist elite controls at least some of the disputed territory but cannot gain international recognition unless the central government cedes its claim to the territory.
 This condition is archetypical of cases in which a militarily successful war of secession ends in a ceasefire.
@@
 
-\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$ to influence the payoffs of $s$ and $g$.\footnote{We assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.}
+\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$ to influence the payoffs of $s$ and $g$.
 
@@
 Otherwise, the game ends in an absorbing state as described in Section~\ref{sec:stage}.
-Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.
 
-Future payoffs are discounted with a common parameter $\de$, where 1/$\de$ is each player's discount rate and $0\leq\de\leq1$.\footnote{All of the results that follow go through essentially unchanged in the general case where the players potentially discount at different rates.
-This simplifying assumption is made solely to highlight the essential mechanisms of the model.} Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
+Future payoffs are discounted with a common parameter $\de \in [0,1]$.
+Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
 
-The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur.
-We turn next to describing the payoffs.
+The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.
 
@@
 
-All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.\footnote{We will see below that restrictions on the initial values of the state variables are essential for characterizing our main result, but in general we leave them unrestricted.} To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
+All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
+To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
 	\begin{equation}
@@
 The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
-As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.
 
@@
 reunification) in exchange for some set of (relatively small) payments from the opposing player.
-For example, if the secessionists cede, the secessionist region is reunified with the home state and the payoffs are $L_s$ for the secessionists and $W_g$ for the home state.
 Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.
@@
 There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously.
-We denote the payoffs of war as a lottery $\Omega$.\footnote{Because the unrecognized state already controls territory and the \emph{de facto} borders are armed, there is likely only a small advantage to be gained by attacking first for either side.
-Therefore, we argue it is not essential to differentiate between these war scenarios analytically.} This lottery determines whether the secessionists or government wins the war.
-For simplicity, we assume the outcome of the war is an absorbing state.\footnote{The model can easily be extended to encompass the possibility of an indecisive war.} Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
-Therefore, outright victory requires more than simply securing control of the disputed territory (which the secessionists have already done when the game begins) and requires the ability to impose the terms of settlement.
-In practice, this likely involves the capture of the home state capital and/or the overthrow of the home state government.
+We denote the payoffs of war as a lottery $\Omega$.
+This lottery determines whether the secessionists or government wins the war.
+For simplicity, we assume the outcome of the war is an absorbing state.
+Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
 
 For probabilities $p$ of outright victory and $1-p$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p (W_i^t-\zeta_i), \left(1-p\right)( L_i^t-\zeta_i))$.
-$\Omega^t \equiv (\omega_g^t, \omega_s^t).$\footnote{Baseline costs of war are fixed ($\zeta_i$) but additional costs of war based on war's result are captured in $W_i$ and $L_i$.
-Note that the structure of the war lottery along with the assumption that the outside actors influence the payoffs only and not the probabilities of winning imply that any way in which the patron influences the payoffs from war can be counteracted by the international community through its own investments.} Players are assumed to approach war lotteries as expected values.
+$\Omega^t \equiv (\omega_g^t, \omega_s^t).$ Players are assumed to approach war lotteries as expected values.
+
+We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
+We assert that the payoffs for the party that cedes the issue of status (independence vs.
+reunification) are consistently low.
+This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).
 
@@
 We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
-$X$ is a binary variable representing reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
-Y represents recognition by the home state of the secessionists' independence: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
+$X$ represents reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
+Y represents recognition: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
 
-Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.\footnote{Note that, in equilibrium, after an absorbing state has been reached, neither outside player will make any investments.
-Thus the relevant per period payoff will be repeated with $R_p^i = R_c^i = 0$ for $i > t$ if an absorbing state is reached at time $t$.} Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
-
-%The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources.
-As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.
-
-%While in many cases both the central government and the de facto government of the unrecognized states may be relatively autocratic, leaders still rely on some level of public support to remain in power.
-If the leaders' actions deviate too far from the public's preferred course, they may be removed from office or forced to bear the costs of additional repression of regime opponents.
-Therefore, both the central government and the secessionist elite have incentives to take the public's opinions into account, and the preferences of the public are thus internalized in the payoffs expressed in the model for $c$ and $g$.
-
-\subsubsection{The Low Payoffs from Ceding}
-Aside from the central assumption that the government of the home state most prefers reunification and the secessionists most prefer independence accompanied by international recognition, we will for the most part leave the model general enough to incorporate a range of preferences and capabilities for players $g$ and $s$.
-%Our results will specify parameter ranges for these preferences over which outcomes of interest---such as perpetual unrecognized statehood---can be supported in equilibrium.
-We do, however, assert that the payoffs for the party that cedes the issue of status (independence vs.
-reunification) are consistently low.
-This reflects a combination of two factors.
-First, the issue of status is indivisible\footnote{Either the unrecognized state has sovereignty over its territory and is co-equal with the home state, or the secessionist region (and its government) are subordinate to the central government.} and highly valued by each side, making its surrender undesirable.
-Various forms of ethno-nationalism often motivate secession, and the values attached to independence by secessionists (and to reunification by citizens of the home state) are generally large relative to the values placed on economic prosperity and other goals.
-Second, many of the payments that could be offered are not credible (e.g.
-Schultz, 2010).
-
-The difficulty of making credible payments in exchange for status is one clearly demonstrated in the civil war literature (e.g.
-Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007).
-Unrecognized states generally constitute ``sons of the soil" conflicts in which the central government cannot credibly commit to preserving the local demographic and political dominance of the secessionist elite once the disputed territory reverts to central government control (Weimer, 1978; Fearon, 2004).
-While the central government might initially grant the secessionist elite a high level of autonomy in exchange for agreeing to reunification, the level of autonomy is likely to decrease over time, perhaps quite quickly.
-The payments that can be offered by the unrecognized state to the central government are similarly small or unenforceable: once recognition is achieved, the secessionists have little reason to uphold any prior commitments.
-
-%At the time of secession, ethnic Akbhaz made up a minority of the population of Abkhazia, but by the late 1990s they controlled almost all political posts in the \textit{de facto} government of the region (Cornell 2001; Wooleh 2006).
-In 2004, the basket of payments offered by the Georgians in exchange for reunification included a provision guaranteeing that ethnic Abkhaz would retain a majority in the regional parliament, even if the return of internally displaced persons (IDPs) once again placed ethnic Abkhaz in a minority demographic position in the region.
-The promise, however, was not very meaningful.
-First, even if the promise were upheld, it would still mean a step back from the total dominance the ethnic Abkhaz currently enjoy in the region.
-Second, if Georgian IDPs returned, they may demand and receive a more equitable system of representation.
-These concerns are not abstract; this type of reneging has already occurred in cases that did reach settlement.
-
-The case of Gagauzia illustrates the commitment problem nicely.
-Gagauzia achieved \emph{de facto} independence at the time of the Soviet Union's collapse, but agreed to rejoin Moldova in 1994 as an autonomous region.
-While Gagauzia was granted substantial autonomy under the Moldovan Law on the Special Legal Status of Gagauzia, when the governor of Gagauzia, Dmitrii Croiter, moved to assert these powers in 1999, the Moldovan government balked.
-By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government.
-The Moldovan government jailed a number of other Gagauz politicians, and while \emph{de jure} Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the \emph{de facto} level of autonomy has been limited by continued central government meddling in less-than-free regional elections.
-The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.
-%WE CAN CUT THE GAGAUZ EXAMPLE IF NECESSARY -- BG July 2014
-
-%The payments that can be offered by the unrecognized state to the central government are similarly small or unenforceable.
-To return to the Abkhaz example: Under a scenario in which Georgia recognizes Abkhazia as an independent state, the ethnically Georgian region of Gali, currently under Abkhaz control, would likely rejoin Georgia, Russian troops would be (at least temporarily) expelled, and compensation might be paid to displaced Georgians, but other side payments are difficult to picture.
-Abkhazia might promise Georgia privileged access to Abkhaz ports, or promise to keep Russian troops out of its territory permanently, but once recognition is granted, any such promises could easily be reneged upon.
-
-%Furthermore, while relinquishing territorial claims relieves the home state of a persistent source of instability, conceivably supplementing $g$'s payoffs from ceding, it also would increase the probability of future attempts at secession by other regions (Walter 2006).
-The lack of enforceable side payments and the ambiguous impact on stability make ceding the issue of status a low-payoff option for $g$, just as it is for $s$.
-
-%Among the means through which $p$ or $c$ can spend resources to alter the payoffs to $g$ and $s$ is to enforce bargains and guarantee future concessions by either party, raising the payoffs from ceding.
-For example, if both the home state and the secessionists prefer reunification with autonomy rights to both war and continued stalemate, $c$ can spend resources to enforce an agreement in which the secessionists are promised specific autonomy rights and $c$ agrees to ensure that these rights are later retained.
-Should the home state later attempt to revoke the promised autonomy, $c$ can levee sanctions or employ other coercive measures to deter this action.
-The promises and pitfalls of this approach are discussed in the section on policy implications of the model.
+Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
+Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
 

### 2017-07-01 theory.tex
remove appendix

@@
 By contrast, reunification through negotiated settlement has been very rare.
-We will explore potential explanations for this seeming inability of the international community to achieve its aims in Sections~\ref{sec:alt} and \ref{sec:ext}.
+We will explore potential explanations for this seeming inability of the international community to achieve its aims in Section~\ref{sec:alt}.
 First, we turn to the outcome that we find the most puzzling: perpetual unrecognized statehood.
@@
 Our central result shows that unrecognized statehood can be an equilibrium outcome capable of being sustained in perpetuity by fully rational, perfectly informed actors.
-This is true even when there is no actor that prefers unrecognized statehood as a first-best outcome.
 
@@
 
-The proofs of the lemmas are in the Appendix.
+The proofs of the lemmas can be found in Buzard, Graham and Horne (2017).
 
@@
 
-The proof of Proposition 2 is in the Appendix.
+The proof of Proposition 2 can be found in Buzard, Graham and Horne (2017).
 
@@
 
-\pagebreak
-\appendix
-\section{Appendix}
-
-{\bf Proof of Lemma 1} \\
-Recall that the upper case notation represents payoffs gross of investments by the outside players.
-Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions.
-Given the stationary equilibrium actions, playing Status Quo will lead to $Q_s^t \ \forall t$, so the continuation value is $\frac{Q_s^t}{1-\delta}$.
-Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_s^t$ in each period, is $\frac{L_s^t}{1-\delta}$.
-The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p$, and $L_s$ forever with probability $1-p$.
-Thus the continuation value is $-\zeta_s +\frac{W_s^t(p)+ L_s^t(1-p)}{1-\delta}$.
-
-Given that the continuation value from Status Quo is higher than both that from playing Cede and Fight, player $s$ will play Status Quo. The argument for player $g$ is symmetric. \hfill $\blacksquare$
-\\
-\\
-\noindent {\bf Proof of Lemma 2} \\
-Begin by assuming Restriction \ref{res:new}a holds.
-In the status quo, we have $U_p^{t,SQ}= -R_p^{SQ}$ and $U_c^{t,SQ}= -R_c^{SQ}$.
-If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_p^{t,RU}= -\alpha -R_p^{RU}$ and $U_c^{t,RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.
-
-Since reunification is the patron's least preferred state, $R_p^{RU} =0$; that is, the patron is not willing to invest to provoke reunification.
-But the patron \textit{would} be willing to invest the amount it would lose from reunification to maintain the Status Quo: that is, $R_p^{SQ} = \frac{\alpha}{1 -\de}$.
-
-For player $c$ the difference in continuation values from switching from the status quo to reunification is $\frac{\beta + R_c^{SQ}}{1 -\de} -R_c^{RU}$.
-So $c$ is willing to invest up to $\frac{\beta+ R_c^{SQ}}{1 -\de}$ toward reunification---or preventing the status quo.
-
-Player $c$ would be able to induce the secessionists to play Cede---e.g.
-counter an addition to $Q_s^{t-1}$ with an investment to $L_s^{t-1}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} > Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$.\footnote{Here we assume that the payoff relationship at $t$ is such that ceding is better than fighting.
-If this relationship is reversed, similar analysis and results hold.} For this to work, the investment required of player $c$ must be no more than the gain player $c$ receives from investing.
-
-On the other hand, if $\frac{\beta + R_c^{SQ}}{1 -\de} \leq Q_s^{t-1} -\mu + R_p^t - L_s^{t-1}$, the patron's investment creates a large enough difference in the secessionist's payoffs from the status quo versus ceding (net of player $c$'s investment) so that player $c$ will not find contesting the status quo to be in its interest.
-Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$.
-
-Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$.
-Combining this with $R_c^{SQ} =0$, a tighter condition for preventing player $c$ from inducing reunification is $\frac{\beta}{1 -\de} + \mu \leq R_p^t$.
-
-Since the patron is willing to invest up to $\frac{\alpha}{1 -\de}$, $\frac{\beta}{1 -\de} + \mu \leq \frac{\alpha}{1 -\de}$ is a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome.
-
-If Restriction \ref{res:new}a is violated, as the secessionists' status quo payoffs degrade from period to period, they will fall below $\omega_s^t$ before they fall below $L_s^t$.
-If no investment is made in $Q_s^{t-1}$, the secessionists will choose the war outcome.
-Restriction \ref{res:new}b guarantees that the patron will make the investment $R_p^{t'}  = \frac{\omega_s^{t-1}}{1-\de} +\mu - \frac{Q_s^{t-1}}{1-\de} \leq \mu$ to prevent the secessionists choosing war since this $R_p^{t'}$ is strictly less than the patron's expected loss from the war outcome.
-Note that the patron's actual investment will be the maximum of the investment required to prevent the war and ceding outcomes.
-\hfill $\blacksquare$
-\\
-\\
-\noindent {\bf Proof of Lemma 3} \\
-Again, in the status quo, we have $U_p^{t,SQ}= -R_p^{SQ}$ and $U_c^{t,SQ}= -R_c^{SQ}$.
-If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_p^{t,REC}= \lambda -R_p^{REC}$ and $U_c^{t,REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.
-
-Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$.
-For player $c$ it is $\frac{-\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$.
-$R_c^{REC} =0$ since the international community will not invest to create its least preferred outcome.
-So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition.
-We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $t=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$.
-
-If $\frac{\nu}{1-\de} < L_g^{t-1} + R_p^t - Q_g^{t-1}$, the patron's investment to augment $L_g^{t-1}$ creates a large enough difference in the secessionists' payoffs of ceding versus the status quo (net of player $c$'s investment) so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
-
-On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g.
-counter an addition to $L_g^{t-1}$ with an investment to augment $Q_g^{t-1}$---as long as $\frac{\nu}{1-\de} \geq L_g^{t-1} + R_p^t - Q_g^{t-1}$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting.
-If this relationship is reversed, similar analysis and results hold.}
-
-Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} > L_s^{t-1}$ so that a tighter condition for preventing player $p$ from inducing recognition is $\frac{\nu}{1 -\de} \geq R_p^t$.
-
-Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome.
-A slightly weaker version of this inequality is stated as Restriction (\ref{res:4}) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4.
-\hfill $\blacksquare$
-\\
-\\
-\noindent {\bf Proof of Lemma 4} \\
-The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ} -R_p^{WAR}$.
-For player $c$ it is $\frac{-\nu p_s + \beta (1-p_s)}{1 -\de} + R_c^{SQ} -R_c^{WAR}$.
-As noted above, $R_c^{SQ} =0$.
-So $c$ is willing to invest up to $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}$ toward preventing fighting.
-
-The patron is willing to invest up to $\frac{\lambda p_s - \alpha (1-p_s)}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive.
-Thus we must only examine the case where this quantity is positive.
-
-We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive.
-In this only case of interest, we have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.
-
-If $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  < \omega_s^t\left(R_p^t\right) - Q_s^t$, the patron's investment in $\omega_s^t$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.
-
-On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g.
-counter an addition to $\omega_s^t$ with an investment to augment $Q_s^{t-1}$---as long as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq \omega_s^t\left(R_p^t\right) - Q_s^t$.
-
-This can be expanded as $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_s^{t-1}+R_p^t)p_s + L_s^{t-1}(1-p_s) - Q_s^{t-1} +\mu = \omega_s^{t-1} - Q_s^{t-1} +R_p^t p_s + \mu$.
-
-Restriction (\ref{res:1}) combined with equilibrium play implies that  $Q_s^{t-1} \geq \omega_s^{t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu \geq R_p^t p_s$.
-
-Since $R_p^t \leq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_s - \beta (1-p_s)}{1 -\de}  - \mu \geq p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1 - \de} - p_s \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.
-
-Multiplying Restriction (\ref{res:3}) by $-(1-p_s)$, we have $\frac{-\beta \cdot (1-p_s)}{1-\de} -\mu (1-p_s) \geq \frac{-\alpha \cdot (1-p_s)}{1-\de}$.
-Using this with Restriction (\ref{res:4}) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_s - \beta (1-p_s)}{1-\de} -\mu \geq \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
-Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu > p_s \frac{\lambda p_s - \alpha (1-p_s) + \mu + \beta}{1-\de} -p_s \mu$.
-\hfill $\blacksquare$
-\\
-\\
-{\bf Proof of Proposition 2} \\
-Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$.
-As the only alteration is a reduction in $Q_s^t$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed.
-As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.
-
-Therefore if sanctions are to disrupt the Status Quo equilibrium and potentially lead to ceding by the secessionists, it must be that the patron is either not able or not willing to make the investment required to maintain the status quo under player $c$'s sanctions (Condition 1).
-
-Assume then that Condition 1 holds so that the patron makes no investments in the Status Quo payoffs of player $s$.
-We also need the patron to withhold investments toward encouraging the secessionists to fight.
-Lemma 4 of Proposition 1 is not sufficient to guarantee this behavior since no investments are being made in the Status Quo.
-In addition, the reference outcome changes: players $p$ and $c$ must now compare their payoffs in the war scenario to the reunification outcome.
-To be precise, the least stringent condition is $m$ periods of Status Quo followed by the Reunification outcome, where $m$ is the maximal number of periods that it takes for the sanctions to degrade the status quo payoffs below the level of the payoffs from ceding for player $s$.
-
-Player $p$'s continuation value when the status quo is played for $m-1$ periods before ceding is $0 - \frac{\de^m}{1-\de} \alpha$, whereas the continuation value from war is $\frac{\lambda p_s - \alpha (1-p_s)}{1-\de}$.
-For player $c$, it is $0 + \frac{\de^m}{1-\de} \beta$ versus $\frac{-\nu p_s + \beta (1 - p_s)}{1-\de}$.
-Thus the condition for $p$'s willingness to instigate fighting to outweigh $c$'s willingness to deter fighting is
-\[
-  \textstyle \frac{\lambda p_s - \alpha (1-p_s)}{1-\de} + \frac{\de^m}{1-\de} \alpha > \frac{-\nu p_s + \beta (1-p_s)}{1-\de} -  \frac{\de^m}{1-\de} \beta
-\]
-
-If this condition holds, then there will be no investments following sanctions.
-But it is still not guaranteed that the secessionists will choose to cede.
-Status Quo is no longer a long-run equilibrium strategy, but even without outside investments, it is possible that Fight could be a more attractive action than Cede.
-
-The status quo payoffs will degrade period by period in the absence of investments.
-Whether the secessionists choose Cede or Fight depends on which strategy yields the higher continuation value.
-In order to avoid conflict, sanctions must cause player $s$'s continuation value from playing Status Quo to fall below its continuation value from playing Cede before it falls below the continuation value from playing Fight.
-\hfill $\blacksquare$
-
-%Several possibilities exist.
-First, the sanctions could degrade the payoffs to ceding more than they degrade the status quo payoffs.
-If $S_L > \mu + S_Q$ where $S_L$ and $S_Q$ are the per-period reduction in the payoffs to ceding and maintaining the Status Quo, respectively, both payoffs will degrade until eventually player $s$ will choose \emph{Fight}.
-This occurs because the continuation value from \emph{Cede} will remain below that from playing \emph{Status Quo}, but the sanctions only have a partial effect in reducing the \emph{Fight} continuation value through the war lottery, while their entire impact is felt on the \emph{Status Quo} continuation value.
-The gap between the Status Quo and Fight continuation values that was positive without sanctions will eventually become negative, leading to conflict.
-
-%Thus we must have $\mu + S_Q > S_L$ if there is any chance that the secessionists will choose to cede.
-
-%\newpage
-
-%\theendnotes
-
 \newpage

### 2017-07-01 theory.tex
first pass through 'Explaining outcomes' section

@@
 For example, while a drought in the unrecognized state might decrease the secessionist elite's payoffs from the status quo and increase their need for international trade and assistance, additional humanitarian and economic assistance from the patron can offset the effects of the shock and preserve the status quo.
-Likewise, if the home state gains military strength, altering the probabilities in the war lottery, the patron can offset these changes by providing arms or otherwise investing in the defenses of the unrecognized state.
-See further discussion along these lines in Section~\ref{sec:sanctions}.
-
-While we do not model this directly, the stability of the status quo equilibrium may be further enhanced by the reluctance of the patron to withdraw support for an unrecognized state that was previously supported.
-Fearon (1994) and subsequent work on audience costs suggests that, past a certain threshold of escalation, it becomes difficult for leaders to back down from confrontation because they fear looking weak in front of domestic audiences.
-Leaders of patrons that have supported the secessionists in a previous period may risk losing office if they later back down -- even if it is the best interest of the patron state to do so.
 
@@
 In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.
-
-If the payoffs from Fight are strictly greater than the payoffs from Cede for both players, then (Fight,Fight) will be the only additional equilibrium outcome.
-If this inequality holds for just one player, then we have only the additional outcome in which that player chooses Fight and its opponent cedes.
-If both players strictly prefer Cede to Fight, then there exist equilibria where both (Fight,Cede) and (Cede,Fight) are outcomes of the stage game.\footnote{In the case of any indifference, we get the relevant combination of (Fight, Fight), (Fight,Cede) and (Cede, Fight).}
+The equilibria may involve (Fight,Fight), (Fight,Cede) and/or (Cede,Fight) as outcomes of the stage game.
 
@@
 Second, most of the types of outcomes that we observe in the post-WWII era are consistent with the set of parameters outlined in Proposition 1 that support the status quo outcome.
-In the next section, we turn to the use of sanctions by player $c$ and the home state to attempt to force reunification.
 

### 2017-07-01 theory.tex
first pass at cutting down section on sanctions

@@
 Let us begin with the simplest case, in which the sanctions affect only the secessionists' status quo payoffs, as when the imposition of sanctions has a negative impact on the economy of the unrecognized state.
-In the event that the secessionists cede or are defeated militarily, we presume the sanctions would be immediately lifted -- the home state would not want to sanction itself.
-In the event the secessionists achieve military victory, we presume they are able to force the home state to lift the sanctions.\footnote{Recall that our definition of military victory includes the ability to dictate the terms of settlement.} Sanctions are not expected to affect the payoffs to either ceding or military victory and neither should the cost of fighting itself be negatively impacted.
-
-Thus, sanctions serve only two purposes: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede.
+We assume sanctions have only two effects: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede.
 If the patron wishes to maintain the status quo, its per-period investment must increase to compensate for the additional degradation of the status quo payoffs caused by the sanctions.
-All of this implies that the effect of sanctions on the unrecognized state's choice is not unambiguous.
+In this case, the effect of sanctions on the unrecognized state's choice is ambiguous.
 Proposition 2 lists necessary conditions for the existence of an equilibrium in which ``ceding'' is the equilibrium outcome once sanctions are introduced.
-
-%Unfortunately, while the intended effect of sanctions is to reduce the unrecognized state's payoffs from the status quo, sanctions may also reduce the secessionists' payoffs from ceding.
-When the home state collaborates with $c$ to enforce sanctions and impose economic suffering on the residents of the secessionist region, this can have the unintended consequence of increasing the hostility of the secessionists toward reunification.
 
@@
 
-This logic is well illustrated by the case of Tamil Eelam, a territory in Northern Sri Lanka that existed as an unrecognized state from 1987-2009.
-Throughout the conflict, the Sri Lankan government used sanctions and blockades to reduce the secessionists' status quo payoffs and degrade their military capabilities.
-However, the effects of these sanctions were offset by two patrons: India, during the early stages of the conflict, and the Tamil diaspora throughout.
-The disruption of the status quo equilibrium began in 2006, when the United States, European Union, Canada, and India formally designated the leading secessionist organization, the Liberation Tigers of Tamil Eelam (LTTE), as terrorists.\footnote{Other states then followed suit, including Australia in 2008.} This designation greatly strengthened the sanctions against the secessionists and led to a sharp decline in both the quality of life in Tamil Eelam and the military capabilities of the LTTE.
-In January 2008, the Sri Lankan government abrogated the existing ceasefire agreement and in 2009 it launched a full-scale military offensive that ended with a decisive victory over the LTTE and reunification of Tamil Eelam and Sri Lanka.
-While there were many factors at play, the strengthening of sanctions increased the home state's probability of victory and thereby played a role in its decision to initiate a return to war.
-
-%\vspace{3pt}
-%\begin{center}
-%\begin{tabular}{|  c| c | c | c |}
-% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-%	Cede& $$& $$ &$(-),(\downarrow)$ \\ \hline
-%	Status Quo& $(\uparrow), (\downarrow)$& $(-),(\downarrow)$&$(-), (\downarrow)$ \\ \hline
-%	Fight & $(\uparrow), (\downarrow)$ &$(\uparrow), (\downarrow)$&$(-),(\downarrow)$\\ \hline
-% \multicolumn{4}{c} {\emph{Figure 5: Economic Sanctions by the Home State and Player $c$}}
-%\end{tabular}
-%\end{center}
-%\vspace{5pt}
-
-Despite this potential for perverse effects, sanctions are a common tool of outside actors ($c$), more common than aid and other positive inducements.
-A flow payment of carrots, even backed by the promises of a ``neutral" third party, may not be credible in the eyes of the secessionists, which could explain the frequent resort to sticks.
-
 \newpage

### 2017-07-03 theory.tex
strip math out of Def 1 and a few other misc locations as proof of concept

@@
 \begin{enumerate} 
-\item $p$ chooses an investment level $R_p^t\in \left[0,B_p^t\right]$ to influence the payoffs of $s$ and $g$, where $B_p^t$ is an exogenous limit on the finances of the patron.
+\item $p$ chooses an investment level to influence the payoffs of $s$ and $g$.
 
-\item $c$ chooses an investment level $R_c^t\in \mathbb{R}^+$ to influence the payoffs of $s$ and $g$.
+\item $c$ chooses an investment level to influence the payoffs of $s$ and $g$.
 
@@
 The payoffs at the end of a period are determined by these actions and six state variables.
-These state variables keep track of the value of the status quo ($Q_s^t$ and $Q_g^t$), losing ($L_s^t$ and $L_g^t$) and winning the issue of status ($W_s^t$ and $W_g^t$) for the secessionists and government respectively.
-The transitions of the state variables are described below.
+These state variables keep track of the value of the status quo, losing and winning the issue of status for the secessionists and government respectively.
 
@@
 \begin{definition}
-\emph{The class $\mathcal{G}$ includes all those games for which the parameters satisfy the following restrictions:}
+\emph{The class $\mathcal{G}$ includes all those games for which following restrictions are satisfied:}
 
 \begin{enumerate}
-\item \textit{For both players $g$ and $s$, $Q_i^0 \geq L_i^0$: remaining in the status quo is better than ceding at the beginning of the game.}\label{res:1}
+\item \textit{For both players $g$ and $s$, remaining in the status quo is better than ceding at the beginning of the game.}\label{res:1}
 
-\item \textit{For both players $g$ and $s$, $\frac {Q_i^0}{1-\delta} \geq \frac {\omega_i^0}{1-\delta} = -\zeta_i+\frac{W_i^0(p_i) + L_i^0(1-p_i)}{1-\delta}$: the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}
+\item \textit{For both players $g$ and $s$, the expected outcome under war is worse than the status quo at the beginning of the game.}\label{res:2}
 
-\item \textit{Either (a) $\frac{L_s^0}{1-\delta}  > \frac{\omega_s^0}{1-\delta}$ or (b) $0 > -\mu > \frac{\lambda p_s - \alpha (1-p_s)}{1-\de}$: either the secessionists prefer ceding to war or the patron's disutility from war is greater than the per-period cost of offsetting the deterioration in the secessionists' status quo payoffs.}\label{res:new}
+\item \textit{Either the secessionists prefer ceding to war or the patron's disutility from war is greater than the per-period cost of offsetting the deterioration in the secessionists' status quo payoffs.}\label{res:new}
 
-\item \textit{$\frac{\alpha}{1-\de} \geq \frac{\beta}{1-\de} + \mu$: reunification is more important for the patron to avoid than for the international community to achieve.}\label{res:3}
+\item \textit{Reunification is more important for the patron to avoid than for the international community to achieve.}\label{res:3}
 
-\item  \textit{$\nu\cdot p_s \geq \lambda \cdot p_s + \mu + \beta$: recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}\label{res:4}
+\item  \textit{Recognition of the secessionist state is more important for the international community to avoid than for the patron to achieve.}\label{res:4}
 
-\item  \textit{$B_p^1 \geq\frac{\beta}{1-\delta} - \left(Q_s^0 -\mu - L_s^0 \right)$: the patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}\label{res:5}
+\item  \textit{The patron can afford to deter player $c$ from inducing reunification at the beginning of the game.}\label{res:5}
 
-\item \textit{$B_p^t \geq \mu, \ \forall t$: the patron can afford to pay to maintain the status quo.}\footnote{Depending on parameters, one might think that Restriction (5) is more likely to be binding than Restriction (6).
-If there is great variance in budget between periods for the patron, such as a greatly reduced budget in some period $t> 1$ compared to period $1$, (6) could be the binding constraint.}\label{res:6}
+\item \textit{The patron can afford to pay to maintain the status quo.}\label{res:6}
 
@@
 Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006).
-That is, a strategy profile is a stationary Markov equilibrium if it is a stationary Markov strategy profile and a subgame-perfect equilibrium.
-In turn, a strategy profile is defined to be a stationary Markov strategy if any two ex post histories terminating in the same state play identically from the termination point forward.
-In other words, stationary Markov strategies ignore all details of the history aside from the current state.
+Strategies in a stationary Markov equilibrium ignore all details of the history aside from the current state.
 
 The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-That is, $s = \left(Q_s,Q_g,L_s,L_g,W_s,W_g \right)$.\footnote{In the case of $W_s^t$ and $W_g^t$, it is more intuitive to imagine investments increasing the probability of winning a war; because the lottery is additively separable and the ``Win'' outcome is always preferred when it occurs independently in the stage game, this more convenient modeling choice is inconsequential.} Strategies for the patron and the international community are how much to invest in each of the six state variables.
-However, positive investments in some of the state variables can be ruled out by preference assumptions.
+That is, $s = \left(Q_s,Q_g,L_s,L_g,W_s,W_g \right)$.
 

### 2017-07-03 theory.tex
remove proof of proposition 1

@@
 
-\subsection{Analysis of the ``Status Quo'' Equilibrium} 
+\subsection{The ``Status Quo'' Equilibrium} 
 \label{sec:mainproof}
@@
 
-There are many potential equilibrium outcomes of this game, including, under the right parameters, immediate ceding by either party as well as fighting (see Section~\ref{sec:alt}).
-As we are interested particularly in the outcome of long-term unrecognized statehood, here we focus on the question of the existence of an equilibrium that leads to this outcome in perpetuity.
-We will show that, given a game satisfying the parameter restrictions in Definition 1, at least one status quo equilibrium will exist.
-
+We show that at least one status quo equilibrium exists for any game satisfying the restrictions in Definition 1.
 Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006).
@@
 
+\begin{itemize}
+	\item Reunification: When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
+	\item Recognition: When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
+	\item Fight Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
+		\begin{itemize}
+			\item Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
+All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.
+		\end{itemize}
+\end{itemize}
+
 If, however, off-path investments are ever made such that Status Quo does not yield the highest continuation value for one of the players, that player will play Cede or Fight and the game will end.
@@
 
-\vskip.1in
-\noindent {\bf Proof of Proposition 1} \\
-Recall that each period is composed of three action stages: player $p$'s investment decision (stage 1), player $c$'s investment decision (stage 2), and the simultaneous game between players $g$ and $s$ (stage 3).
-Although period $t$ may be reached either because in period $t-1$ both players ceded or the status quo had been maintained, the strategic landscape in period $t$ is the same given our restriction to stationary Markov strategies.
-
-To establish that the Status Quo Equilibrium of Definition 2 exists, we must show that the given strategies constitute a Nash equilibrium in each subgame within a period given the induced continuation values.\footnote{Subgame perfection within the one-shot game combined with the stationary Markovian property ensures subgame perfection in the whole game.} We begin with the stage game between the government and secessionists.
-We assume that the government and secessionists choose the Status Quo strategy whenever indifferent between it and any other strategy.
-
-\begin{lemma}
-$Q_i^t \geq L_i^t$ and $\frac {Q_i^t}{1-\delta} \geq  -\zeta_i+\frac{W_i^t(p_i) + L_i^t(1-p_i)}{1-\delta}$ for $i \in \left\{g,s\right\}$ are sufficient for (Status Quo,Status Quo) to be the outcome of Stage 3 in period $t \ \forall t$.
-\end{lemma}
-
-The proofs of the lemmas can be found in Buzard, Graham and Horne (2017).
-
-Lemma 1 establishes ranges for the payoffs for players $g$ and $s$---gross of investments by the outside players---in which the status quo outcome can occur.
-
-We now turn to the incentives of the outside actors---that is, the patron state and the international community---in the subgames that begin in the first and second stages.
-
-Recall that either outside actor could invest toward increasing any of the six state variables in a period.
-However, each outside actor only considers investments in three of the state variables as treated in the three cases in Definition 2.
-The remainder of the proof addresses each set of matched investment decisions in turn.
-
-Lemma 2 addresses potential efforts by the international community to influence the outcome toward \emph{Reunification}:
-
-\begin{lemma}
-When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
-\end{lemma}
-
-\noindent Restriction (\ref{res:new}) guarantees one of two outcomes: either the patron is willing to invest in the secessionists' status quo payoffs to avoid them falling below the expected payoffs from war or the secessionists' payoffs from war remain below those from ceding so that it is only the patron's incentives concerning the ceding outcome with which we must be concerned.
-The following treats the latter case; a parallel argument using Restriction \ref{res:new}(b) can be made in the former case.
-
-Restriction (\ref{res:3}) provides a bound on the amount that player $p$ must be \emph{willing} to invest each period in order to prevent player $c$ from contesting the status quo outcome.
-To complete the equilibrium construction, we must determine the utility maximizing investments by $p$ and conditions to ensure that it is \emph{able} to make those investments.\footnote{Efforts by the patron to create the conditions for the \emph{Recognition} outcome are addressed in Lemma 3 once the equilibrium investments are established.}
-
-Consider period 1.
-When Restriction (\ref{res:3}) holds, the patron will want to invest just enough to create a buffer of $\frac{\beta}{1-\de}$ between $Q_s^1$ and $L_s^1$ so that $c$ will not invest.
-That is, in equilibrium, $R_p^1 = \frac{\beta}{1-\de} -(Q_s^0 - \mu - L_s^0)$ as long as $p$ can afford to make this investment.
-Thus we need Restriction (\ref{res:5}): $B_p^1 \geq\frac{\beta}{1-\de} - \left(Q_s^0 - \mu - L_s^0 \right)$, where $B_p^1$ is the amount $p$ has available to spend on the conflict in period $1$.
-
-In periods $t > 1$, there are two cases to consider.
-Either (a) the buffer created in period $t-1$ was precisely the necessary $\frac{\beta}{1-\de}$, or (b) the buffer is larger than $\frac{\beta}{1-\de}$.
-In case (a), the patron must spend exactly $\mu$ to offset the degradation in the status quo payoffs and re-establish the buffer of $\frac{\beta}{1-\de}$.\footnote{A third case in which the buffer is smaller than $\frac{\beta}{1-\de}$ provides lower welfare to player $p$.
-Restrictions (\ref{res:3}) and (\ref{res:5}) ensure that player $p$ can avoid this case.} In case (b), the patron can spend less than $\mu$ in period $t$.
-However, because in each period $Q_s^t$ degrades by $\mu$, eventually the buffer will be reached and we will be returned to case (a).
-Again, assuming Restriction (\ref{res:3}) holds, the patron will want to make this investment if its budget allows, and so a sufficient condition is that $p$'s budget is at least as large as $\mu$ in every period $t > 1$ (Restriction \ref{res:6}).
-
-With the equilibrium status quo investments determined, we can proceed to Lemma 3, which rules out spending by the patron to help achieve recognition of the secessionist state:
-
-\begin{lemma}
-When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
-\end{lemma}
-
-In the Status Quo equilibrium, the patron will invest to maintain an outcome that is not its most preferred, but it will not invest to achieve its most preferred outcome.
-This behavior may appear counterintuitive, but we frequently observe patron states whose preferred outcomes are recognized independence for the secessionists who  nonetheless contribute resources to sustain a status-quo outcome that is costly to all involved.
-The patron does not attempt to contribute sufficient resources to force recognition by the home state because doing so would induce offsetting expenditures by the international community to prevent this outcome.\footnote{We do not provide conditions on the budget of player $c$ similar to Restrictions (\ref{res:5}) and (\ref{res:6}): since the size of the international community relative to any particular country is large, it can be assumed that a budget constraint does not bind.}
-
-One last possibility is ruled out by Lemma 4: that the patron would invest to encourage the secessionists to fight.
-Note that the patron has no incentive to make this investment if Restriction \ref{res:new}b holds.
-
-\begin{lemma}
-Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
-\end{lemma}
-
-Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
-All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.
-
-%Our status quo equilibrium requires that $p$ invests in every period enough to maintain $F^*$ by offsetting the $\mu$ decline in the secessionists' status quo payoffs.
-The total per-period equilibrium investment $R_{pn}$, a flow payment, in the long run is thus $\mu$ per period in this steady state.
-In a status quo equilibrium $c$ need not invest at all since stage game payoffs of $g$ do not deteriorate.\footnote{ In the absence of the assumption in the model setup above that player $c$ does not want war, the patron would need to retain a second buffer against $c$ funding war.
-The expected value of war for the home state would need to be maintained at a level lower than the status quo payoff by a buffer of $F*=\frac{\beta}{1-\delta_c}$.
-If this buffer did not exist in the first round of the game, the patron would fund the alteration of payoffs to create the buffer, thus assuring payments by $c$ would be ineffective at trying to change the home state's payoffs to make war more attractive than the status quo.}
-
-Lemmas 2 through 4 establish that the only incentive compatible investment is that of the patron to deter the international community from inducing reunification, supported by the strategies described in cases (1)-(3) in Definition 2.
-This combined with Lemma 1 ensures the existence of the Status Quo Equilibrium described in Definition 2.
-\hfill $\blacksquare$
+The proof of Proposition 1 is available in Buzard, Graham and Horne (2017).
 

### 2017-07-03 theory.tex
complete major overhaul of analysis section

@@
 
-\section{Explaining Outcomes of Secessionist Conflicts} 
+\section{Explaining Outcomes of Secessionist Conflicts: The ``Status Quo'' Equilibrium} 
 \label{sec:main}
-%The game has three outcomes that we will consider, each of which can be characterized by a class of equilibria.
-The outcome of greatest interest is both players choosing the status quo in perpetuity; we will also examine the two absorbing states, reunification and recognition by the home state (i.e.
-recognized statehood).\footnote{Other classes of equilibria are possible, for example it is theoretically possible for the players to agree to a lottery between the absorbing states which would yield a higher expected payoff than the status quo.
-However, in practice there is a credibility issue: the losing party does not have an incentive to cede if they lose the lottery.
-Such strategies are, in any case, not analyzed here.}
 
@@
 By contrast, reunification through negotiated settlement has been very rare.
-We will explore potential explanations for this seeming inability of the international community to achieve its aims in Section~\ref{sec:alt}.
-First, we turn to the outcome that we find the most puzzling: perpetual unrecognized statehood.
+In this section, we use a game theoretic model to explain the outcome that we find the most puzzling: perpetual unrecognized statehood.
 
-\subsection{The ``Status Quo'' Equilibrium} 
-\label{sec:mainproof}
 Unrecognized states are frequently viewed as temporary phenomena or as non-equilibrium outcomes attributable to players' misperceptions of the strategic situation, or their fundamental irrationality.
@@
 
-We begin by listing and providing intuition for a set of restrictions on the preferences of the actors and their resources for which we can guarantee that unrecognized statehood is an equilibrium outcome.
-This set of restrictions identifies a class of games, $\mathcal{G}$:
+We begin by listing a set of restrictions on the preferences of the actors and their resources for which we can guarantee that unrecognized statehood is an equilibrium outcome.
 
@@
 
-We show that at least one status quo equilibrium exists for any game satisfying the restrictions in Definition 1.
-Our concept of equilibrium will be stationary Markov equilibrium as defined in Mailath and Samuelson (2006).
-Strategies in a stationary Markov equilibrium ignore all details of the history aside from the current state.
+We can show that at least one status quo equilibrium exists for any game satisfying the restrictions in Definition 1.
+Our concept of equilibrium is stationary Markov equilibrium in which strategies ignore all details of the history aside from the current state.
 
 The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-That is, $s = \left(Q_s,Q_g,L_s,L_g,W_s,W_g \right)$.
-
 Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight.
-It would also not want to make the government lose and so won't invest to augment $L_g$.
-This leaves three state variables in which player $c$ might invest: $Q_s$, $Q_g$, $L_s$.
-
-Because the patron's preferences are aligned with the secessionists and against the government, it never invests in $W_g$ or $L_s$.
-The only reason the patron would invest in $Q_g$ is to counter an investment by player $c$ in the government's payoffs from war; since player $c$ will not make such an investment, we can also rule out investments by the patron in $Q_g$.
-This leaves three state variables in which the patron may invest: $Q_s$, $L_g$ and $W_s$.
-
-We now turn to defining and describing the Status Quo Equilibrium.
-We will see that, in this equilibrium, the outside actors' investments are paired: the patron's investments in the secessionists' status quo payoffs deter player $c$'s investments in the secessionists' payoffs from ceding.
-Likewise, investments by player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding.
-Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
-In each case, the potential  investments are such that the largest continuation payoffs in the game between the secessionists and the government are those from playing status quo.
+It would also not invest in the government's payoffs from ceding.
+The patron will never invest in the government's payoffs from winning or the secessionists' payoffs from ceding, and it will not invest in the government's status quo payoffs because player $c$ will not invest in the government's payoffs from war so there is no need to counter such an investment.
+This leaves three state variables in which each outside actor might invest, which we address while defining the Status Quo Equilibrium.
 
@@
 The strategies for the government and secessionists in this equilibrium are to play their best responses given the continuation values induced by the investments of the outside actors.
-The continuation value from choosing war is the expected future stream of payoffs from the two possible absorbing states net of the cost of war $\left(-\zeta_i +\frac{W_i^t(p_i)+ L_i^t(1-p_i)}{1-\delta}\right)$.
-Similarly, the continuation value from ceding is $\frac{L_i^t}{1-\de}$ while that for the status quo is $\sum_{s=t}^\infty \de^{t-s} Q_i^t$.
 Unless otherwise noted below, playing Status Quo is the best response for both inside actors in terms of continuation values.
 
-The strategies for the inside actors are the following, organized according to the patron's investments in each relevant state variable. In period $t$:
+The strategies for the outside actors in period $t$ are:
 
 \begin{enumerate}
-	\item The patron invests $R_p^t = \max\left\{\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right),\omega_s^{t-1} +\mu - Q_s^{t-1},0\right\}$ to augment $Q_s^{t-1}$.
-Player $c$ invests $R_c^t =0$ to augment $L_s^{t-1}$.
+	\item The patron invests enough in the secessionists Status Quo payoffs to deter Player $c$ from investing in the secessionists' payoffs from Ceding.
+Otherwise, Player $c$ invests enough to induce the secessionists to play Cede.
 
-If $\frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right) > 0$ and $R_p^t < \frac{\beta}{1-\de} - \left( Q_s^{t-1} - \mu - L_s^{t-1}\right)$, player $c$ invests $R_c^t = L_s^{t-1} - \left(Q_s^{t-1} - \mu + R_p^t \right) + \ve$, for $\ve$ small, to augment $L_s^{t-1}$ so that the secessionists play Cede.
+	\item Player $c$ is prepared to counter the largest investment the patron is willing to make in the government's payoffs from ceding by making a counter-investment in the government's status quo payoffs.
+If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the government would play Cede.
 
-	\item As long as $R_p^t \leq \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ invests $R_c^t = R_p^t - \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $Q_g^{t-1}$.
-Note that the outcome is that investments to augment $L_g^{t-1}$ and $Q_g^{t-1}$ are zero.
-
-If the patron invests $R_p^t > \frac{\nu}{1-\de} + \left( Q_g^{t-1} - L_g^{t-1}\right)$ to augment $L_g^{t-1}$, player $c$ is deterred from investing in $Q_g^{t-1}$ ($R_c^t = 0$) so that the government plays Cede.
-
-	\item As long as $R_p^t \leq \frac{1}{p_s} \left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right]$ to augment $W_s^{t-1}$, player $c$ invests $R_c^t = R_p^tp_s + \mu - Q_s^{t-1} + \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)$ to augment $Q_s^{t-1}$.
-Similar to case two, the outcome is that investments to augment $W_s^{t-1}$ and $Q_s^{t-1}$ are zero.
-
-If the patron invests $R_p^t > \frac{1}{p_s}\left[\frac{\nu p_s - \beta (1-p_s)}{1 -\de} - \mu + Q_s^{t-1} - \left(-\zeta_{s}(1-\de) + W_s^{t-1}p_s + L_s^{t-1}(1-p_s)\right)\right] $ to augment $W_s^{t-1}$, player $c$ is deterred from investing in $Q_s^{t-1}$ ($R_c^t=0$) so that the secessionists play Fight.
+	\item Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
+If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the secessionists would play Fight.
 \end{enumerate}
@@
 Since the patron moves first, the only investment that takes place in the Status Quo Equilibrium is the patron's investment in the status quo payoffs of the secessionists to deter the international community from provoking the secessionists to cede the issue of sovereignty.
-The international community's willingness to counteract investments by the patron toward the other two disturbances implies that there will be no investments in equilibrium in cases (2) and (3).
+This requires that Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold.
+The patron must also have sufficient resources as per Restrictions (\ref{res:5}) and (\ref{res:6}).
 
-\begin{itemize}
-	\item Reunification: When Restrictions (\ref{res:new}) and (\ref{res:3}) of Definition 1 hold, the patron's willingness to invest to maintain the status quo is sufficient to both deter the international community from intervening to encourage reunification and to avoid the secessionists choosing to go to war.
-	\item Recognition: When Restriction (\ref{res:4}) of Definition 1 holds, the international community's willingness to invest to avoid recognition of the secessionist state is sufficient to deter the patron from investing to achieve recognition.
-	\item Fight Restrictions (\ref{res:3}) and (\ref{res:4}) of Definition 1 ensure that the international community's willingness to invest to discourage new conflict is sufficient to deter the patron from investing to instigate such fighting.
-		\begin{itemize}
-			\item Note that this result depends on an implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
-All other conflict scenarios are ruled out by the international community's assumed preference to avoid conflict.
-		\end{itemize}
-\end{itemize}
+The international community's willingness to counteract investments by the patron toward the other two disturbances (i.e., Restriction (\ref{res:4})) implies that there will be no investments in equilibrium in cases (2) and (3).
+Case (3) also requires Restrictions (\ref{res:3}) of Definition 1 and the implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
 
@@
 
-Equilibrium actions are for the patron to maintain the status quo by investing $\mu$ each period once the difference in payoffs to the secessionists from playing Status Quo and Cede reaches $\frac{\beta}{1-\de}$ (with a possible lump sum investment at $t=1$ of up to that amount); for the international community to not invest and for both inside actors to play Status Quo each period.
+Equilibrium actions are for the patron to maintain the status quo by investing enough to overcome the deterioration in the secessionists' status quo payoffs; for the international community to not invest and for both inside actors to play Status Quo each period.
 
@@
 
-The proof of Proposition 1 is available in Buzard, Graham and Horne (2017).
+The full proof of Proposition 1 is available in Buzard, Graham and Horne (2017).
 
@@
 \label{sec:sanctions}
-%It is useful to look more specifically at how spending by the patron and player $c$ can alter the game's payoffs, and potentially its outcomes.
-By analyzing comparative statics in the normal form game, we can evaluate the different strategies through which these players pursue their desired outcomes and the conditions under which they might be successful.
 
-%Let us first consider what occurs if the patron supplies the secessionists with weapons or other military support, changing the expected payoffs from war.
-This would make war more appealing to the secessionist elite and less appealing to the central government, deterring attempts at reconquest.
-At a certain point, if the central government (and $c$) do not invest in the home state military to counteract this support, due either to their preferences or budget constraint, the expected payoff of war for the secessionists can surpass the status quo payoff, and the secessionists would have the incentives to fight:\footnote{Arrows indicate the direction of change in payoffs due to the outside action.
-A blank cell or "-" indicates no change.}
-
-%\begin{center}
-%\begin{tabular}{|  c| c | c | c |}
-% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-%	Cede& $$& $$ &$$ \\ \hline
-%	Status Quo& $(\downarrow),(\uparrow)$& $$&$$ \\ \hline
-%	Fight & $(\downarrow),(\uparrow)$ &$(\downarrow),(\uparrow)$&$$\\ \hline
-% \multicolumn{4}{c} {\textit{Figure 2: Additional Military Support to the Secessionists}}
-%\end{tabular}
-%\end{center}
-%\vspace{5pt}
-
-%Alternatively, the patron state may supply humanitarian support (such as providing passports to citizens of the unrecognized state or funding schools), making the status quo more appealing and stable:
-
-%\begin{center}
-%\begin{tabular}{|  c| c | c | c |}
-% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	%Cede& $$& $$ &$(-),(\uparrow)$ \\ \hline
-%	Status Quo& $$& $(-),(\uparrow)$&$$ \\ \hline
-	%Fight & $$ &$$&$$\\ \hline
-% \multicolumn{4}{c} { \emph{Figure 3: The Effects of Humanitarian Assistance to the Secessionists}}
-%\end{tabular}
-%\end{center}
-%\vspace{5pt}
-
-%Similarly, $c$ can contribute resources to make ceding more likely by either party.
-However, because $c$ generally prefers reunification to independence, these resources are most likely to be committed to encouraging ceding by the secessionist elite (reunification), instead of ceding by the home state (recognition).
-
-%Changes in payoffs can be made either by carrots or by sticks.
-In one option $c$ provides the secessionist elite with positive inducements, like aid, in exchange for rejoining the home state.
-$c$ may also expend resources to make payments by the home state, such as various autonomy rights, more credible.
-The effect is the same: for a price, $c$ can increase the secessionists' payoffs from ceding.
-
-%\begin{center}
-%\begin{tabular}{|  c| c | c | c |}
-% \hline  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-%	Cede& $$& $$ &$$ \\ \hline
-%	Status Quo& $$& $$&$(-),(\uparrow)$ \\ \hline
-%	Fight & $$ &$$&$(-),(\uparrow)$\\ \hline
-% \multicolumn{4}{c} {\emph{Figure 4: Positive Inducements From Player $c$}}
-%\end{tabular}
-%\end{center}
-%\vspace{5pt}
-
-In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists.
+In Section~\ref{sec:main}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists.
 Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.

### 2017-07-03 theory.tex
take out payoff notation for outside actors, state variable notation

@@
 All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
-To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
-	\begin{equation}
-	\begin{aligned}
-		L_s^t = L_s^{t-1} - R_p^t + R_c^t \hskip.5in L_g^t = L_g^{t-1} + R_p^t - R_c^t \\
-		W_s^t = W_s^{t-1} + R_p^t - R_c^t \hskip.5in W_g^t = W_g^{t-1} - R_p^t + R_c^t \\
-		Q_g^t = Q_g^{t-1} - R_p^t + R_c^t \hskip.5in Q_s^t = Q_s^{t-1} - \mu + R_p^t - R_c^t
-		\label{eqn:motion}
-	\end{aligned}
-	\end{equation}
-
 The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.
@@
 In this case neither player has demonstrated a willingness to give up more than the other.
-Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.%\footnote{This assumption is appropriate as precise simultaneity is an artifact of discrete time modeling: it is not a phenomenon we observe in the real world.}
+Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.
 
@@
 
-We use two binary variables to express the preferences among the three outcomes (status quo, reunification, recognition by the home state) for both $p$ and $c$.
-$X$ represents reunification: $X=0$ in the status quo and $X=1$ if the secessionists rejoin the home state.
-Y represents recognition: $Y=1$ if the home state recognizes the secessionists as independent, $Y=0$ otherwise.
-
-Player $p$'s payoffs in period $t$ are $U_p^t= -\alpha X+\lambda Y -R_p^t$, while the payoffs for player $c$ are $U_c^t= \beta X-\nu Y -R_c^t$, with both payoffs denoted in currency units.
-Taking $\alpha, \ \beta, \lambda,$ and $\nu$ to be positive, these payoffs represent the idea that the patron opposes reunification while player $c$ prefers it, and that the patron prefers recognized independence while player $c$ is averse to the creation of new states.
+The patron opposes reunification while player $c$ prefers it, and the patron prefers recognized independence while player $c$ is averse to the creation of new states.
 
@@
 The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
-In cases where Restriction (\ref{res:6}) is not met, we can have unrecognized statehood for some time, but it is not a long-run equilibrium outcome.
 
@@
 First, it indicates that there may be an important role for external actors to play in coordinating expectations about which equilibrium will be played, and in the absence of such coordination, equilibrium switching from the status quo equilibrium to one of the other outcomes is possible.
-Second, most of the types of outcomes that we observe in the post-WWII era are consistent with the set of parameters outlined in Proposition 1 that support the status quo outcome.
+Second, most of the types of outcomes that we observe in the post-WWII era are consistent with the set of restrictions outlined in Proposition 1 that support the status quo outcome.
 

### 2017-07-04 theory.tex
remove table for inside players' payoffs

@@
 
-\noindent Payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
-\vspace{2pt}
-\textbf{<<COMP: Place Figure 2 about here>>}
-\vspace{2.5pt}
+We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
+We assert that the payoffs for the party that cedes the issue of status (independence vs.
+reunification) are consistently low.
+This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).
 
@@
 
-Only two outcomes of the stage game do not lead to absorbing states, and the payoffs for these are bolded.
+Only two outcomes of the stage game do not lead to absorbing states.
 If both $s$ and $g$ play Status Quo, then the status quo persists.
@@
 
-If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e.
-the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2).
+If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e., the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2).
 We interpret one player ceding as that player ceding the issue of status (independence vs.
@@
 This lottery determines whether the secessionists or government wins the war.
-For simplicity, we assume the outcome of the war is an absorbing state.
+We assume the outcome of the war is an absorbing state and that players approach war lotteries as expected values.
 Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
-
-For probabilities $p$ of outright victory and $1-p$ of loss, player $i \in \{s, g\}$ in period $t$ with a fixed cost of war $\zeta_i$ faces war lottery $\omega_i^t \equiv (p (W_i^t-\zeta_i), \left(1-p\right)( L_i^t-\zeta_i))$.
-$\Omega^t \equiv (\omega_g^t, \omega_s^t).$ Players are assumed to approach war lotteries as expected values.
-
-We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
-We assert that the payoffs for the party that cedes the issue of status (independence vs.
-reunification) are consistently low.
-This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).
 
@@
 In Section~\ref{sec:main}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists.
-Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e.
-sanctions are equivalent to $R_c^t < 0$).
+Player $c$, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo.
 Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.
@@
 \begin{proposition}
-Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo (i.e.
-$Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ where $\sigma$ is the reduction due to sanctions).
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect only player $s$'s payoffs to maintaining the Status Quo.
 In order for sanctions to lead to ceding by the secessionists, the following are required:
@@
 \item \textit{The patron must either be unable or find that it is not worthwhile to invest to instigate fighting by the secessionists.}
-
-%\item \textit{The secessionists' continuation values from playing \emph{Cede} must be higher than their continuation values from playing \emph{Fight}.}
-
-%\item \textit{The sanctions must overcome the gap between the \emph{Status Quo} and \emph{Cede} continuation values faster than the gap between the \emph{Status Quo} and \emph{Fight} continuation values closes.}
 
@@
 
-%The situations where condition 2 or 3 fails and the secessionists initiate war follow the same general logic as political science models in the power transitions literature that predict pre-emptive war as the response to the increasing relative capabilities of a rival (Levy 1997; Powell 1999b).
-The expected payoffs from war are not increasing, but the alternatives are getting worse.
-The parallels with preemptive war become more direct if we increase the realism of our assumptions and allow sanctions to have a negative effect not only on the economy (the Status Quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).
-
 Moving beyond this simple case, we can add realism by allowing sanctions to have a negative effect not only on the economy (the status quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).
@@
 \begin{proposition}
-Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities (i.e.
-$Q_s^t = Q_s^{t-1} - \mu -\sigma +R_p^t - R_c^t$ \emph{and} $p_s\left(\sigma\right)$ $\left(p_g\left(\sigma\right)\right)$ is decreasing (increasing) in $\sigma$).
+Assume the restrictions of \emph{Definition 1} hold in the absence of sanctions and that sanctions affect both player $s$'s status quo payoffs and its military capabilities.
 The parameter space over which a war will be initiated by the home state is increasing in the magnitude of the sanctions' impact on the secessionists' military capabilities.
@@
 The proof of Proposition 3 is immediate.
-Although under the conditions of Proposition 1 (i.e.
-when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting.
+Although under the conditions of Proposition 1 (i.e., when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting.
 The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.
@@
 \end{figure}
-
-\newpage
-
-\begin{center}
-\begin{tabular}{|  c| c | c | c |} 
-\hline
-  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
-	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
-	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
-	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
- \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
-\end{tabular}
-\end{center}
-
-\newpage
-\begin{center}
-\includegraphics[width=6.25in]{Table_1_082214.pdf}
-\end{center}

### 2017-07-04 theory.tex
first pass at cutting down sanctions section

@@
 
-\textbf{<<COMP: Place Figure 1 about here>>}
+\begin{figure}
+\includegraphics{Timeline2.pdf}
+\caption{Timeline}
+\end{figure}
 
@@
 Let us begin with the simplest case, in which the sanctions affect only the secessionists' status quo payoffs, as when the imposition of sanctions has a negative impact on the economy of the unrecognized state.
-We assume sanctions have only two effects: to narrow the difference between the payoffs from Status Quo and Fight, and to narrow the difference between the payoffs from Status Quo and Cede.
-If the patron wishes to maintain the status quo, its per-period investment must increase to compensate for the additional degradation of the status quo payoffs caused by the sanctions.
 In this case, the effect of sanctions on the unrecognized state's choice is ambiguous.
-Proposition 2 lists necessary conditions for the existence of an equilibrium in which ``ceding'' is the equilibrium outcome once sanctions are introduced.
 
@@
 
-Thus, even if we do not consider the sanctions to impose any direct costs on the home state (though in practice they likely do), it remains ambiguous whether sanctions will benefit the home state.
-Recall that in the Status Quo Equilibrium, the home state's expected returns from war are lower than from a continuation of the status quo.
-If sanctions induce the secessionists to play Fight, this is a worse outcome for the home state than if the status quo had been allowed to persist.
-
-Moving beyond this simple case, we can add realism by allowing sanctions to have a negative effect not only on the economy (the status quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).
+We can add realism by allowing sanctions to have a negative effect not only on the economy (the status quo payoffs) but also on the military capabilities of the secessionists (the expected payoffs from war).
 This is an important extension because one motivation for sanctions is often precisely that -- to weaken the military capability of the secessionists.
 
-In the model, this is represented as reducing (increasing) the secessionists' probability of victory (loss) in the war lottery.
+In the model, this is represented as reducing the secessionists' probability of victory in the war lottery.
 This should serve to increase the range of parameters over which the conditions of Proposition 2 hold.
 However, at the same time, the home government experiences changes of the same magnitude and opposite sign in its war lottery, increasing its payoffs from playing Fight.
-The effect of sanctions on the home state's strategic considerations is clear cut:
 
@@
 
-The proof of Proposition 3 is immediate.
-Although under the conditions of Proposition 1 (i.e., when the restriction of Definition 1 hold and the Status Quo Equilibrium is played) the home state's continuation value from maintaining the status quo is higher than from initiating conflict in the absence of sanctions, when those sanctions degrade the secessionists' military capabilities they increase the chances that the home state would prevail in a conflict, thus increasing the home government's continuation value from fighting.
-The stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.
+It is immediate that the stronger is the impact of sanctions on the secessionists military, the stronger is the effect on the home government's value of fighting and the greater is the range of parameters over which this change in payoffs will lead to a change in behavior.
 
@@
 Conversely, if the sanctions degrade the secessionists military capabilities sufficiently, it induces the home state to fight.
-
-\newpage
-
-\begin{figure}
-\includegraphics{Timeline2.pdf}
-\caption{Timeline}
-\end{figure}

### 2017-07-04 theory.tex
collapse section 1 and combine where possible

@@
 We model a dispute over a piece of territory that is controlled by a secessionist group and also claimed by a home state.
-The central issue of contention, independence vs.
-reunification, is both difficult to divide and highly valued by both sides.
-The side payments that can be offered in exchange for the opponent's surrender of the independence/reunification issue are sharply limited by the absence of large concessions that can credibly be made (Walter 1997, 2002; Schultz 2010).
-
-Because our model incorporates the incentives and actions of international actors, it is about to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.\footnote{For another example of the strategic manipulation of decision-makers into (and out of) conflict see Baliga and Sjostrom (2012).}
+Because our model incorporates the incentives and actions of international actors, it is about to both articulate the mechanisms that create these persistent stalemates and to assess the consequences, intended and otherwise, of outside actors' attempts to foster their desired outcome.
 
@@
 In practice we often observe groups of states like the OECD or the UN acting in this capacity.
+We also assume player $c$ prefers peace to war; this implies that player $c$ will not fund a military buildup that it expects will induce war.
 
-Player $p$ most prefers recognized independence, aligning its interests with the secessionists.
+In contrast, Player $p$ most prefers recognized independence and opposed independence, aligning its interests with the secessionists.
 We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium.
@@
 
-\begin{figure}
-\includegraphics{Timeline2.pdf}
-\caption{Timeline}
-\end{figure}
+We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
+We assert that the payoffs for the party that cedes the issue of status (independence vs.
+reunification) are consistently low.
+This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).
 
@@
 
+\begin{figure}
+\includegraphics{Timeline2.pdf}
+\caption{Timeline}
+\end{figure}
+
 \begin{enumerate} 
@@
 
-The payoffs at the end of a period are determined by these actions and six state variables.
-These state variables keep track of the value of the status quo, losing and winning the issue of status for the secessionists and government respectively.
+The payoffs at the end of a period are determined by these actions and the values of state variables that keep track of the value of the status quo, losing and winning the issue of status for the secessionists and government respectively.
 
-The game continues into the next period if the `inside players'---the secessionists and government---selected (status quo, status quo) or (cede, cede).
-Otherwise, the game ends in an absorbing state as described in Section~\ref{sec:stage}.
+All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
+The status quo payoffs for the secessionists are automatically reduced by $\mu$ each period.
+The steady reduction by $\mu$ represents the costs of non-recognition.
+
+Only two outcomes of the stage game do not lead to absorbing states.
+If both $s$ and $g$ play Status Quo, then the status quo persists.
+Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
+In this case neither player has demonstrated a willingness to give up more than the other.
+
+If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game ends with payoffs in every subsequent period given by the corresponding payoffs in the stage game.
+Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.
+
+There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously.
+We use a lottery to determine whether the secessionists or government wins the war.
+Victory would, among other things, allow an unrecognized state to force recognition by the home state government.
 
@@
 The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.
-
-\subsubsection{Within-period Payoffs for the Secessionists and Government}
-\label{sec:stage}
-
-We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
-We assert that the payoffs for the party that cedes the issue of status (independence vs.
-reunification) are consistently low.
-This reflects the fact that the issue of status is indivisible and highly valued by each side and that many of the payments that could be offered are not credible (Licklider, 1995; Walter 1997, 2002; Doyle and Sambanis, 2006; Fearon and Laitin, 2007; Schultz, 2010).
-
-All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
-The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$.
-The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction.
-
-Only two outcomes of the stage game do not lead to absorbing states.
-If both $s$ and $g$ play Status Quo, then the status quo persists.
-Likewise, if both states simultaneously play Cede, we assume that both renege immediately and that the status quo is preserved for that period.
-In this case neither player has demonstrated a willingness to give up more than the other.
-Therefore, payoffs for both players ceding simultaneously are identical to the status quo payoffs.
-
-If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game enters an absorbing state (i.e., the game ends), with payoffs in every subsequent period given by the corresponding entry in the stage game (Figure 2).
-We interpret one player ceding as that player ceding the issue of status (independence vs.
-reunification) in exchange for some set of (relatively small) payments from the opposing player.
-Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.
-
-There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously.
-We denote the payoffs of war as a lottery $\Omega$.
-This lottery determines whether the secessionists or government wins the war.
-We assume the outcome of the war is an absorbing state and that players approach war lotteries as expected values.
-Outright victory would, among other things, allow an unrecognized state to force recognition by the home state government.
-
-\subsubsection{Within-period Payoffs for the Patron and International Community}
-We assume player $c$ prefers peace to war.
-For simplicity, we limit our modeling of $c$'s preference for peace to the assumption that $c$ will not choose to fund a military buildup that it expects will induce war.
-This assumption is not necessary for the basic results to hold; however, it justifies our decision not to address military support of armed reconquest by the home state as a deliberate strategy by $c$ to achieve reunification.
-
-The patron opposes reunification while player $c$ prefers it, and the patron prefers recognized independence while player $c$ is averse to the creation of new states.
 

### 2017-07-04 theory.tex
tighten up section 1

@@
 
-Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders.
-In practice we often observe groups of states like the OECD or the UN acting in this capacity.
-We also assume player $c$ prefers peace to war; this implies that player $c$ will not fund a military buildup that it expects will induce war.
+Player $c$ prefers reunification to recognized independence---a preference that is common to most states, and especially among those that fear the prospect of secessionist movements within their own borders.\footnote{In practice we often observe groups of states like the OECD or the UN acting in this capacity.} We also assume player $c$ prefers peace to war; this implies that player $c$ will not fund a military buildup that it expects will induce war.
 
-In contrast, Player $p$ most prefers recognized independence and opposed independence, aligning its interests with the secessionists.
+In contrast, Player $p$ most prefers recognized independence and opposes independence, aligning its interests with the secessionists.
 We refer to $p$ as the patron because $p$ contributes resources to the unrecognized state in the status quo equilibrium.
@@
 All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment.
-The status quo payoffs for the secessionists are automatically reduced by $\mu$ each period.
-The steady reduction by $\mu$ represents the costs of non-recognition.
+The status quo payoffs for the secessionists are automatically reduced by $\mu$ each period, reflecting the costs of non-recognition.
 
-Only two outcomes of the stage game do not lead to absorbing states.
 If both $s$ and $g$ play Status Quo, then the status quo persists.
@@
 In this case neither player has demonstrated a willingness to give up more than the other.
+These are the only outcomes of the stage game that do not lead to absorbing states.
 
-If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game ends with payoffs in every subsequent period given by the corresponding payoffs in the stage game.
-Therefore, if one player agrees to cede while the other player chooses to remain in the status quo or fight, the result is a negotiated settlement benefiting the player who did not cede.
+If either $s$ or $g$ plays Cede while the other plays Fight or Status Quo, the game ends with payoffs in every subsequent period given by the corresponding payoffs in the stage game---i.e., the result is a negotiated settlement benefiting the player who did not cede.
 
-There are three ways to end up in war: either of the parties may attack first, or both may attack simultaneously.
+If either of the parties attacks first or both attack simultaneously, the result is war.
 We use a lottery to determine whether the secessionists or government wins the war.
-Victory would, among other things, allow an unrecognized state to force recognition by the home state government.
+The victor is able to force recognition/reunification.
 
-Future payoffs are discounted with a common parameter $\de \in [0,1]$.
-Therefore payoffs for the entire game for player $i\in \{s, g, p, c\} $ can be expressed by the discounted stream of payments $\Sigma_{t=1}^{\infty} \de^{t-1} U_i^t $ where $U_i^t$ is player $i$'s payoff in period $t$.
-
-The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.
+Future payoffs are discounted with a common parameter $\de \in [0,1]$. The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players.
 

### 2017-07-04 theory.tex
tightening through sec 2 and 3

@@
 
-We do not restrict the preferences and capabilities for players $g$ and $s$, with one exception.
+We do not make any assumptions about the preferences and capabilities for players $g$ and $s$, with one exception.
 We assert that the payoffs for the party that cedes the issue of status (independence vs.
@@
 
-\section{Explaining Outcomes of Secessionist Conflicts: The ``Status Quo'' Equilibrium} 
+\section{Explaining the ``Status Quo'' Equilibrium} 
 \label{sec:main}
 
-Despite the preferences of the international community for peace, the most common outcome of secessionist conflicts in the post-WWII period has been reunification with the home state via outright military reconquest.
+%Despite the preferences of the international community for peace, the most common outcome of secessionist conflicts in the post-WWII period has been reunification with the home state via outright military reconquest.
 By contrast, reunification through negotiated settlement has been very rare.
@@
 
-The state is a vector of six variables comprised of the payoffs from the status quo, ceding, and winning the issue of status for each of the two inside actors.
-Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight.
+Of the six state variables, Player $c$ dislikes war and so will never invest in either state variable associated with winning since they increase the likelihood that one of the inside actors chooses to fight.
 It would also not invest in the government's payoffs from ceding.
@@
 The strategies for the government and secessionists in this equilibrium are to play their best responses given the continuation values induced by the investments of the outside actors.
-Unless otherwise noted below, playing Status Quo is the best response for both inside actors in terms of continuation values.
+Unless otherwise noted below, playing Status Quo is the best response for both inside actors.
 
-The strategies for the outside actors in period $t$ are:
+The strategies for the outside actors in each period are:
 
 \begin{enumerate}
-	\item The patron invests enough in the secessionists Status Quo payoffs to deter Player $c$ from investing in the secessionists' payoffs from Ceding.
+	\item The patron invests enough in the secessionists Status Quo payoffs to deter Player $c$ from investing in the secessionists' payoffs from ceding.
 Otherwise, Player $c$ invests enough to induce the secessionists to play Cede.
 
-	\item Player $c$ is prepared to counter the largest investment the patron is willing to make in the government's payoffs from ceding by making a counter-investment in the government's status quo payoffs.
+	\item Potential investments by Player $c$ in the government's status quo payoffs deter the patron from investing in the government's payoffs from ceding.
 If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the government would play Cede.
 
-	\item Investments by player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
+	\item Potential investments by Player $c$ in the secessionists' status quo payoffs deter the patron from investing in the secessionists' payoffs from winning the conflict via fighting.
 If the patron were to make an investment larger than its willingness to pay, Player $c$ would not counter and the secessionists would play Fight.
@@
 
-To understand the construction of this equilibrium, notice that there are three possible ways to disrupt the Status Quo: (1) the international community can provoke the secessionists to Cede, (2) the patron can provoke the government to cede, or (3) the patron can provoke the secessionists to fight.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.} In order to establish that the Status Quo Equilibrium exists, it must be shown that each of these three deviations will be deterred.
+Equilibrium actions are for the patron to maintain the status quo by investing enough to overcome the deterioration in the secessionists' status quo payoffs; for the international community to not invest and for both inside actors to play Status Quo each period.
+
+In order to establish that the Status Quo Equilibrium exists, we must show that each of three possible deviations will be deterred: (1) the secessionists provoked to Cede the international community, (2) the government provoked to cede by the patron, or (3) the secessionists provoked to fight by the patron.\footnote{The international community might also provoke the government to fight, but it is assumed to avoid conflict.}
 
@@
 The international community's willingness to counteract investments by the patron toward the other two disturbances (i.e., Restriction (\ref{res:4})) implies that there will be no investments in equilibrium in cases (2) and (3).
-Case (3) also requires Restrictions (\ref{res:3}) of Definition 1 and the implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
+Case (3) also requires Restriction (\ref{res:3}) of Definition 1 and the implicit assumption that the patron is not able to skew the odds of the secessionists winning the conflict in a way that cannot be nullified by the international community.
 
 If, however, off-path investments are ever made such that Status Quo does not yield the highest continuation value for one of the players, that player will play Cede or Fight and the game will end.
-
-Equilibrium actions are for the patron to maintain the status quo by investing enough to overcome the deterioration in the secessionists' status quo payoffs; for the international community to not invest and for both inside actors to play Status Quo each period.
 
@@
 
-First, note that the restrictions in Definition 1 are sufficient but not necessary.
-For example, if the status quo initially has a much higher long term payoff than the next best alternative for the secessionists, Restriction (\ref{res:6}) need not be met to maintain the status quo in the short run.
-The inequality \emph{will} bind for some set of periods $t \geq 1$ because the secessionist payoffs from the status quo decrease over time.
-
-On the other hand, the restrictions in Definition 1 also \emph{do not} provide for a unique equilibrium, or even a unique equilibrium outcome.
-In fact, at least one additional equilibrium outcome always coexists with the status quo outcome.
-The equilibria may involve (Fight,Fight), (Fight,Cede) and/or (Cede,Fight) as outcomes of the stage game.
+The restrictions in Definition 1 do not provide for a unique equilibrium, or even a unique equilibrium outcome.
+At least one additional equilibrium outcome always coexists with the status quo outcome: at least one of (Fight,Fight), (Fight,Cede) and (Cede,Fight) will always be an equilibrium outcome of the stage game.
 

### 2020-09-28 teaching.tex
add intro and citations to teaching document

@@
 \end{center}
+
+\section{Unrecognized states in context (i.e. the strategic situation)}
+Unrecognized states destabilize the international system and impoverish their residents.
+Because the borders of these territories are contested, the threat of violent conflict is ever present.
+Unrecognized states are unable to sign trade agreements or receive most foreign aid, and most face economic sanctions, a combination that leaves their residents isolated and impoverished.
+Thus, unrecognized statehood is a profoundly undesirable outcome, and yet it is nonetheless often a stable equilibrium.
+The long-standing unrecognized states of Somaliland and the Turkish Republic of Northern Cyprus, as well as many of the unrecognized states that emerged when the Soviet Union collapsed, still enjoy de facto independence a quarter century later.
+Unrecognized states are thus quintessential intractable conflicts: difficult but not impossible to resolve (e.g.
+Crocker, Hampson, and Aall 2004; 2005).\\
+
+Because unrecognized statehood is such a clearly unfavorable outcome for the two primary parties involved – the unrecognized state itself and the home state from which it is attempting to secede – much of the existing work has either treated unrecognized states as an aberration or a temporary phenomenon, or, conversely, as utterly intractable conflicts rooted in the irrationality of one or more actors.
+We use game theoretic analysis to challenge these understandings.
+We argue that unrecognized statehood can, in fact, be a stable equilibrium outcome, and that it can be sustained even when all the players involved are perfectly informed and behaving rationally.
+Game theory provides value in this context by laying bare the mechanisms by which unrecognized statehood is sustained as a stable equilibrium outcome, and thus illuminating the ways in which these intractable conflicts can be peacefully transformed.
+\\
+
+We analyze not only the actions of unrecognized states and the home states from which they are attempting to secede, but also the patrons that support these unrecognized states and the actors in the international community who work to induce peaceful settlement.
+By formalizing the incentives of these third-party actors and the actions available to them, we are able to analyze the conditions under which their conduct can lead to peaceful resolution, and when it can lead to war or continued stalemate (Buzard, Graham, and Horne 2017).
+\\
+
+We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state.
+The newest unrecognized states were formed when the Soviet Union collapsed in the early 1990s.
+In the case of more recent cases of militarily successful secession, such as South Sudan, unrecognized statehood has been avoided.
+This suggests that, if the stalemates sustaining the six current unrecognized states (Abkhazia, Nagorno-Karabakh, Somaliland, South Ossetia, Transnistria, and the Turkish Republic of Northern Cyprus) can be successfully resolved, a world without unrecognized states is possible (Buzard, Graham and Horne 2019).\\
 
@@
 
+\section{Works Cited}
+
+Buzard, Kristy, Benjamin AT Graham, and Ben Horne.
+2019.
+Unrecognized States: Theory and Cases and Policy Implications, in Overcoming Intractable Conflicts: New Approaches to Constructive Transformations, Miriam Elman, Catherine Gerard, Galia Golan, and Louis Kriesberg, eds.
+London: Rowman $\&$ Littlefield.\\
+
+Buzard, Kristy, Benjamin AT Graham, and Ben Horne.
+2017.
+"Unrecognized States: A Theory of Self-Determination and Foreign Influence." The Journal of Law, Economics, and Organization 33 no.
+3: 578-611.\\
+
+Crocker, Chester A., Fen Osler Hampson, and Pamela Aall. 2004. Taming intractable conflicts: Mediation in the Hardest Cases. Washington, DC, United States Institute of Peace Press. \\
+
+Crocker, Chester A., Fen Osler Hampson, and Pamela Aall, Ed. 2005. Grasping the Nettle: Analyzing Cases of Intractable Conflicts. Washington, D.C., United States Institute of Peace Press.\\
+
 \end{document}

### 2020-10-31 teaching.tex
Add explanation for payoffs, tiebreak rule, +p only for status quo

@@
 \section{3-player version, Status Quo better than Reunification}
-Let us consider a first case where the payoffs for the secessionists are higher in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+Let us consider a first case where the payoffs for the secessionists are higher in the unrecognized state (`Status Quo') than they are if they cede and rejoin the home state.
 This, for instance, may be the case very early after taking control of the territory before their economic situation has had a chance to deteriorate.
@@
 \begin{enumerate}
-	\item Home government (G) chooses $S_G \in \left\{SQ,C\right\}$ where SQ means 'Status Quo' of unrecognized statehood and C means 'Cede' the issue of status and recognize the secessionists as an independent state.
-	\item Secessionists (S) choose $S_S \in \left\{SQ,C\right\}$, same as for $G$ except when they Cede, they rejoin the home state.
+	\item Home government (G) chooses $S_G \in \left\{Status Quo,Recognize\right\}$ where the `Status Quo' is unrecognized statehood and `Recognize' means the government cedes the issue of status and recognize the secessionists as an independent state.
+	\item Secessionists (S) choose $S_S \in \left\{Status Quo,Rejoin\right\}$, where `Status Quo' is remaining as an unrecognized state and `Rejoin' means they cede the issue of status and rejoin the home state.
 	\item Patron state (P) chooses $p \in [0,\infty)$ to invest in the secessionists status quo payoffs
@@
 \subsection{Payoffs}
-Let, for instance, ($SQ,C$) mean that $G$ plays $SQ$ and $S$ plays $C$. Then 
+Let, for instance, ($Status Quo,Rejoin$) mean that $G$ plays $Status Quo$ and $S$ plays $Rejoin$. Then 
 \begin{itemize}
-	\item The payoffs after ($SQ,SQ$) are $-p,3,2+p$
-	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
-	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
-	\item The payoffs after ($C,C$) are $-p,3,2$
+	\item The payoffs after ($StatusQuo,StatusQuo$) are $0-p,3,2+p$
+		\begin{itemize}
+			\item The patron gets neither its favorite or its least favorite outcome.
+We normalize the patron's payoff from the political outcome to be zero in this case.
+But the patron pays a cost of whatever amount $p$ it transfers to the secessionists.
+			\item Likewise, the government and the secessionists see this outcome as being between their favorite and least favorite (they just have different favorites...).
+We normalize the status quo to 3 for the government and 2 for the secessionists.
+However, the secessionists also receive a transfer of $p$ from the patron.
+		\end{itemize}
+	\item The payoffs after ($StatusQuo,Rejoin$) are $-10-p,5,1$
+		\begin{itemize}
+			\item The patron's worst outcome is when the secessionists rejoin the home state, so the patron gets a large negative payoff from the political outcome of -10.
+			\item The government likes this outcome better than the status quo, so gets 5 instead of 3.
+			\item The secessionists dislike this outcome compared to being unrecognized, so get 1 instead of 2.
+		\end{itemize}
+	\item The payoffs after ($Recognize,StatusQuo$) are $3-p,0,5$
+		\begin{itemize}
+			\item This is the patron's favorite outcome, but the patron likes this outcome less than it dislikes reunification, so we make the patron's payoff positive but not as large as the magnitude of dislike for reunification of 10.
+			\item The government dislikes this outcome compared to having an unrecognized state, so gets 0 instead of 3.
+			\item The secessionists like this outcome better than the status quo, so gets 5 instead of 2.
+		\end{itemize}
+	\item The payoffs after ($Recognize,Rejoin$) are $-p,3,2$
+		\begin{itemize}
+			\item This is a strange case, because it's not obvious what should happen if both give in at the same time.
+We assume that, when this happens, nothing changes---that is, we stay in the status quo.
+But the secessionists don't get the $p$ because they didn't do what the patron wanted them to do.
+		\end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+	\item In the simultaneous game between $G$ and $S$, $Recognize$ is dominated by $StatusQuo$ for player $G$.
 		\begin{itemize}
@@
 				\begin{itemize}
-					\item If $2+p \geq 1$ (i.e. $p \geq -1$), the NE is ($SQ,SQ$).
-					\item If $2+p \leq 1$ (i.e. $p \leq -1$), the NE is ($SQ,C$).
+					\item For simplicity, we will assume that the secessionists stay in the Status Quo when they're indifferent (i.e., they must be strictly worse off from being in the Status Quo in order to make a change).
+					\item If $2+p \geq 1$ (i.e. $p \geq -1$), the NE is ($Status Quo,Status Quo$).
+					\item If $2+p < 1$ (i.e. $p < -1$), the NE is ($Status Quo,Rejoin$).
 					\item Recall that the smallest value of $p$ that the Patron can choose is $p=0$.
-					\item Since $p\leq-1$ is not possible, ($SQ,C$) is not possible.
-					\item Therefore ($SQ,SQ$) will be the equilibrium.
+					\item Since $p <-1$ is not possible, ($Status Quo,Rejoin$) is not possible.
+					\item Therefore ($Status Quo,Status Quo$) will be the equilibrium.
 				\end{itemize}
 		\end{itemize}
-	\item The Patron's decision: Since ($SQ,SQ$) will be the equilibrium no matter what, the Patron will get a payoff of $-p$.
+	\item The Patron's decision: Since ($Status Quo,Status Quo$) will be the equilibrium no matter what, the Patron will get a payoff of $-p$.
 		\begin{itemize}
@@
 \end{itemize}
-Therefore the Subgame Perfect Nash Equilibrium is $p=0$, ($SQ,SQ$) regardless of the value of $p$.
+Therefore the Subgame Perfect Nash Equilibrium is $p=0$, ($Status Quo,Status Quo$) regardless of the value of $p$.
 
@@
 \section{3-player version, Reunification better than Status Quo}
-Let us now consider a case where the payoffs for the secessionists are \underline{lower} in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+Let us now consider a case where the payoffs for the secessionists are \underline{lower} in the unrecognized state ('Status Quo') than they are if they rejoin the home state.
 We believe this will be the case a few months or years after they take control of the territory and unrecognized status has caused their economic situation to deteriorate.
@@
 \begin{itemize}
-	\item The payoffs after ($SQ,SQ$) are $-p,3,\bm{0}+p$
-	\item The payoffs after ($SQ,C$) are $-10-p,5,1$
-	\item The payoffs after ($C,SQ$) are $3-p,0,5+p$
-	\item The payoffs after ($C,C$) are $-p,3,\bm{0}$
+	\item The payoffs after ($Status Quo,Status Quo$) are $-p,3,\bm{0}+p$
+	\item The payoffs after ($Status Quo,Rejoin$) are $-10-p,5,1$
+	\item The payoffs after ($Rejoin,Status Quo$) are $3-p,0,5$
+	\item The payoffs after ($Rejoin,Rejoin$) are $-p,3,\bm{0}$
 \end{itemize}
-The only change(s) from Section 1 is that the secessionist's payoff from ($SQ,SQ$) has been reduced, and the payoffs from ($C,C$) along with it.
+The only change(s) from Section 1 is that the secessionist's payoff from ($Status Quo,Status Quo$) has been reduced, and the payoffs from ($Rejoin,Rejoin$) along with it.
 
@@
 \begin{itemize}
-	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+	\item In the simultaneous game between $G$ and $S$, $Recognize$ is dominated by $Status Quo$ for player $G$.
 		\begin{itemize}
@@
 				\begin{itemize}
-					\item If $0+p \geq 1$ (i.e. $p \geq 1$), the NE is ($SQ,SQ$).
-					\item If $0+p \leq 1$ (i.e. $p \leq 1$), the NE is ($SQ,C$).
+					\item Again, assume that the secessionists stay in the Status Quo when they're indifferent.
+					\item If $0+p \geq 1$ (i.e. $p \geq 1$), the NE is ($Status Quo,Status Quo$).
+					\item If $0+p < 1$ (i.e. $p < 1$), the NE is ($Status Quo,Rejoin$).
 				\end{itemize}
 		\end{itemize}
-	\item The Patron's decision: if ($SQ,SQ$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
-If ($SQ,C$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
+	\item The Patron's decision: if ($Status Quo,Status Quo$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
+If ($Status Quo,Rejoin$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
 		\begin{itemize}
-			\item Since ($SQ,C$) is the worse outcome for the Patron, the patron does not want to invest at all ($p=0$) if ($SQ,C$) is going to be the outcome. 
+			\item Since ($Status Quo,Rejoin$) is the worse outcome for the Patron, the patron does not want to invest at all ($p=0$) if ($Status Quo,Rejoin$) is going to be the outcome. 
 				\begin{itemize}
@@
 				\end{itemize}
-			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of $p \geq 1$.
-Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$, which is $p=1$.
+			\item The Patron prefers ($Status Quo,Status Quo$), but to make this happen the Patron must make an investment of $p \geq 1$.
+Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($Status Quo,Status Quo$) by choosing the smallest $p$, which is $p=1$.
 				\begin{itemize}
@@
 \end{itemize}
-Therefore the Subgame Perfect Nash Equilibrium is $p=1$, and ($SQ,SQ$) as long as $p \geq 1$, ($SQ,C$) if $p \leq 1$.
+Therefore the Subgame Perfect Nash Equilibrium is $p=1$, and ($Status Quo,Status Quo$) as long as $p \geq 1$, ($Status Quo,Rejoin$) if $p < 1$.
 \begin{itemize}
-	\item $p=1$, ($SQ,SQ$) is an equilibrium outcome.
-	\item $p=1$, ($SQ,C$) is also an equilibrium outcome. You can rule this one out by specifying that the Secessionists choose $SQ$ whenever they are indifferent.
+	\item $p=1$, ($Status Quo,Status Quo$) is the equilibrium outcome.
 \end{itemize}
@@
 \section{4-player version, Reunification better than Status Quo}
-Now we look at the full version of the model including the international community, and continue to assume that the payoffs for the secessionists are lower in the unrecognized state ('Status Quo') than they are if they cede and rejoin the home state.
+Now we look at the full version of the model including the international community, and continue to assume that the payoffs for the secessionists are lower in the unrecognized state (`Status Quo') than they are if they cede and rejoin the home state.
 
@@
 \begin{enumerate}
-	\item Home government (G) chooses $S_G \in \left\{SQ,C\right\}$ where SQ means 'Status Quo' of unrecognized statehood and C means 'Cede' the issue of status and recognize the secessionists as an independent state.
-	\item Secessionists (S) choose $S_S \in \left\{SQ,C\right\}$, same as for $G$ except when they Cede, they rejoin the home state.
+	\item Home government (G) chooses $S_G \in \left\{Status Quo,Recognize\right\}$ where Status Quo means staying in unrecognized statehood and `Recognize' means to cede the issue of status and recognize the secessionists as an independent state.
+	\item Secessionists (S) choose $S_S \in \left\{Status Quo,Rejoin\right\}$, where `Status Quo' is remaining as an unrecognized state and `Rejoin' means they cede the issue of status and rejoin the home state.
 	\item Patron state (P) chooses $p \in [0,\infty)$ to invest in the secessionists status quo payoffs
@@
 \subsection{Payoffs}
-We add payoffs for the international community, and the international community's investment in the secessionists payoffs from ceding.
+We add payoffs for the international community in the second position in the payoff vector since the international community moves second.
+We also add the international community's investment in the secessionists payoffs from ceding.
+The international community's payoff from the status quo is normalized to 0.
+The international community most prefers reunification (the secessionists rejoin, +5) and is strongly opposed to the secessionists being recognized (-7).
+Here, note that the international community likes its preferred outcome MORE than the patron does, but it dislikes its least preferred outcome LESS than the patron does.
+This last comparison is crucial for obtaining the outcome we get.
 \begin{itemize}
-	\item The payoffs after ($SQ,SQ$) are $-p,\bm{-c},3,0+p$
-	\item The payoffs after ($SQ,C$) are $-10-p,\bm{5-c},5,1\bm{+c}$
-	\item The payoffs after ($C,SQ$) are $3-p,\bm{-7-c},0,5+p$
-	\item The payoffs after ($C,C$) are $-p,\bm{-c},3,0\bm{+c}$
+	\item The payoffs after ($Status Quo,Status Quo$) are $-p,\bm{-c},3,0+p$
+	\item The payoffs after ($Status Quo,Rejoin$) are $-10-p,\bm{5-c},5,1\bm{+c}$
+	\item The payoffs after ($Recognize,Status Quo$) are $3-p,\bm{-7-c},0,5$
+	\item The payoffs after ($Recognize,Rejoin$) are $-p,\bm{-c},3,0$
 \end{itemize}
@@
 \begin{itemize}
-	\item In the simultaneous game between $G$ and $S$, $C$ is dominated by $SQ$ for player $G$.
+	\item In the simultaneous game between $G$ and $S$, $Recognize$ is dominated by $Status Quo$ for player $G$.
 		\begin{itemize}
-			\item Therefore the Nash Equilibrium (NE) of the subgame is determined by $p$: 
+			\item Therefore the Nash Equilibrium (NE) of the subgame is determined by $p$ and $c$: 
 				\begin{itemize}
-					\item If $0+p \geq 1+c$ (i.e. $p \geq 1+c$), the NE is ($SQ,SQ$).
-					\item If $0+p \leq 1+c$ (i.e. $p \leq 1+c$), the NE is ($SQ,C$).
+					\item If $0+p \geq 1+c$ (i.e. $p \geq 1+c$), the NE is ($Status Quo,Status Quo$).
+					\item If $0+p < 1+c$ (i.e. $p < 1+c$), the NE is ($Status Quo,Rejoin$).
 				\end{itemize}
 		\end{itemize}
-	\item Player $C$'s decision: if ($SQ,SQ$) is the equilibrium in the subgame, Player $C$ will get a payoff of $-c$.
-If ($SQ,C$) is the equilibrium in the subgame, Player $C$ will get a payoff of $5-c$.
+	\item Player $C$'s decision: if ($Status Quo,Status Quo$) is the equilibrium in the subgame, Player $C$ will get a payoff of $-c$.
+If ($Status Quo,Rejoin$) is the equilibrium in the subgame, Player $C$ will get a payoff of $5-c$.
 		\begin{itemize}
-			\item Since ($SQ,SQ$) is the worse outcome for Player $C$, Player $C$ does not want to invest at all ($c=0$) if ($SQ,SQ$) is going to be the outcome. 
+			\item Since ($Status Quo,Status Quo$) is worse for Player $C$, Player $C$ does not want to invest at all ($c=0$) if ($Status Quo,Status Quo$) is going to be the outcome. 
 				\begin{itemize}
@@
 				\end{itemize}
-			\item Player $C$ prefers ($SQ,C$), but to make this happen Player $C$ must make an investment of the $c \geq p - 1$.
-Because Player $C$'s payoff is $5-c$, Player $C$ maximizes its payoffs in the case where ($SQ,C$) is the outcome by choosing the smallest $c$, which is $c=p-1$.
+			\item Player $C$ prefers ($Status Quo,Rejoin$), but to make this happen Player $C$ must make an investment of the $c \geq p - 1$.
+Because Player $C$'s payoff is $5-c$, Player $C$ maximizes its payoffs in the case where the outcome is ($Status Quo,Rejoin$) by choosing the smallest $c$, which is $c=p-1$.
 				\begin{itemize}
@@
 				\end{itemize}
-			\item If $6 - p \leq 0$ (where $0$ is Player $C$'s payoff from ($SQ,SQ$)), Player $C$ will choose to invest $c = p-1$ and the outcome will be ($SQ,C$).
-			\item If $6 - p > 0$ (i.e. $6 > p$), Player $C$ will choose to invest $c = 0$ and the outcome will be ($SQ,SQ$).
+			\item If $6 - p \leq 0$ (where $0$ is Player $C$'s payoff from ($Status Quo,Status Quo$)), Player $C$ will choose to invest $c = p-1$ and the outcome will be ($Status Quo,Rejoin$).
+			\item If $6 - p > 0$ (i.e. $6 > p$), Player $C$ will choose to invest $c = 0$ and the outcome will be ($Status Quo,Status Quo$).
 		\end{itemize}
-	\item The Patron's decision: if ($SQ,SQ$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
-If ($SQ,C$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
+	\item The Patron's decision: if ($Status Quo,Status Quo$) is the equilibrium in the subgame, the Patron will get a payoff of $-p$.
+If ($Status Quo,Rejoin$) is the equilibrium in the subgame, the Patron will get a payoff of $-10-p$.
 		\begin{itemize}
-			\item Since ($SQ,C$) is the worse outcome for the Patron, the patron does not want to invest at all ($p=0$) if ($SQ,C$) is going to be the outcome. 
+			\item Since ($Status Quo,Rejoin$) is worse for the Patron, the patron does not want to invest at all ($p=0$) if ($Status Quo,Rejoin$) is going to be the outcome. 
 				\begin{itemize}
@@
 				\end{itemize}
-			\item The Patron prefers ($SQ,SQ$), but to make this happen the Patron must make an investment of $p \geq 6$.
-Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($SQ,SQ$) by choosing the smallest $p$ that leads to ($SQ,SQ$), which is $p=6$.
+			\item The Patron prefers ($Status Quo,Status Quo$), but to make this happen the Patron must make an investment of $p \geq 6$.
+Because the Patron's payoff is $-p$, the Patron maximizes its payoffs in this case where it induces ($Status Quo,Status Quo$) by choosing the smallest $p$ that leads to ($Status Quo,Status Quo$), which is $p=6$.
 				\begin{itemize}
@@
 \end{itemize}
-Therefore the Subgame Perfect Nash Equilibrium is $p=6$, $c = 0$ if $p \geq 6$, $c = p-1$ if $p < 6$ and ($SQ,SQ$) as long as $p \geq c+1$, ($SQ,C$) if $p \leq c+ 1$.
+Therefore the Subgame Perfect Nash Equilibrium is $p=6$, $c = 0$ if $p \geq 6$, $c = p-1$ if $p < 6$ and ($Status Quo,Status Quo$) as long as $p \geq c+1$, ($Status Quo,Rejoin$) if $p \leq c+ 1$.
 \begin{itemize}
-	\item $p=6$, $c = 0$, ($SQ,SQ$) is an equilibrium outcome.
+	\item $p=6$, $c = 0$, ($Status Quo,Status Quo$) is the equilibrium outcome.
 \end{itemize}
+
+\section{Conclusion}
+We see from the comparison between the outcomes in Sections 2 and 3 that the patron only gets involved if that involvement will change the outcome--if its investments make a difference for the secessionists choice.
+The patron will pay enough to maintain the status quo as long as the payment is small enough (recall the cost has to be less than 10 in the analysis in Section 3.4).
+\\
+
+This logic carries over to the version of the game that includes the international community.
+The patron is willing to pay up to 10 (the difference between its payoffs from the political status in the status quo and rejoin outcomes).
+As long as the international community is willing to invest less to avoid its least preferred outcome than the patron is willing to invest to avoid its least preferred outcome, the patron will invest, the international community will not invest, and the status quo will prevail.
 

### 2022-02-26 teaching.tex
update and add detail to teaching example

@@
 Unrecognized states are unable to sign trade agreements or receive most foreign aid, and most face economic sanctions, a combination that leaves their residents isolated and impoverished.
-Thus, unrecognized statehood is a profoundly undesirable outcome, and yet it is nonetheless often a stable equilibrium.
-The long-standing unrecognized states of Somaliland and the Turkish Republic of Northern Cyprus, as well as many of the unrecognized states that emerged when the Soviet Union collapsed, still enjoy de facto independence a quarter century later.
+Thus, \textbf{unrecognized statehood is a profoundly undesirable outcome}, and yet it is nonetheless often a stable equilibrium.
+\textbf{Somaliland and the Turkish Republic of Northern Cyprus are long-standing unrecognized states}, and many of the unrecognized states that emerged when the Soviet Union collapsed still enjoy de facto independence a quarter century later.
 Unrecognized states are thus quintessential intractable conflicts: difficult but not impossible to resolve (e.g.
@@
 We argue that unrecognized statehood can, in fact, be a stable equilibrium outcome, and that it can be sustained even when all the players involved are perfectly informed and behaving rationally.
-Game theory provides value in this context by laying bare the mechanisms by which unrecognized statehood is sustained as a stable equilibrium outcome, and thus illuminating the ways in which these intractable conflicts can be peacefully transformed.
+\textbf{Game theory provides value in this context by laying bare the mechanisms by which unrecognized statehood is sustained as a stable equilibrium outcome, and thus illuminating the ways in which these intractable conflicts can be peacefully transformed.} \\
+
+\textbf{We analyze not only the actions of unrecognized states and the home states from which they are attempting to secede, but also the patrons that support these unrecognized states and the actors in the international community who work to induce peaceful settlement.} By formalizing the incentives of these third-party actors and the actions available to them, \textbf{we are able to analyze the conditions under which their conduct can lead to peaceful resolution, and when it can lead to war or continued stalemate} (Buzard, Graham, and Horne 2017).
 \\
 
-We analyze not only the actions of unrecognized states and the home states from which they are attempting to secede, but also the patrons that support these unrecognized states and the actors in the international community who work to induce peaceful settlement.
-By formalizing the incentives of these third-party actors and the actions available to them, we are able to analyze the conditions under which their conduct can lead to peaceful resolution, and when it can lead to war or continued stalemate (Buzard, Graham, and Horne 2017).
-\\
-
-We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state.
-The newest unrecognized states were formed when the Soviet Union collapsed in the early 1990s.
-In the case of more recent cases of militarily successful secession, such as South Sudan, unrecognized statehood has been avoided.
-This suggests that, if the stalemates sustaining the six current unrecognized states (Abkhazia, Nagorno-Karabakh, Somaliland, South Ossetia, Transnistria, and the Turkish Republic of Northern Cyprus) can be successfully resolved, a world without unrecognized states is possible (Buzard, Graham and Horne 2019).\\
+\textbf{We define unrecognized states as territories in which a non-state actor controls territory, governs a population, and seeks but does not receive broad recognition as an independent state.} In the most recent militarily-successful secession---South Sudan---unrecognized statehood has been avoided.
+\textbf{This suggests that, if the stalemates sustaining the eight current unrecognized states (Abkhazia, Donetsk, Luhansk, Nagorno-Karabakh, Somaliland, South Ossetia, Transnistria, and the Turkish Republic of Northern Cyprus) can be successfully resolved, a world without unrecognized states is possible (Buzard, Graham and Horne 2019).
+On the other hand, as the current conflict in Ukraine demonstrates, the ongoing existence of unrecognized states can be profoundly destabilizing (Sommerlad, 2022).} \\
 
@@
 \subsection{Timing and Information}
-The patron makes its investment. This investment is seen by all other players. Then the home government and secessions move simultaneously.
+The patron makes its investment. This investment is seen by all other players. Then the home government and secessionists move simultaneously.
 
@@
 \end{itemize}
+
+\vskip.5in
+\subsection{Solution Concept}
+This game has a dynamic component, as the government and the secessionists move after seeing the patron's choice.
+Further, we assume that the players have symmetric information about payoffs; that is, all the players know the payoffs associated with each outcome.
+We will also assume that the players know each other well enough to be able to correctly predict each other's behavior; this is likely to be the case when the same leaders have been in place for a while so that they have interacted repeatedly.
+As this is a high-stakes `game,' all the players have strong incentives to seek out information and think carefully about what the other players will do before choosing the strategy that will maximize their expected payoffs, and this is common knowledge amongst the players.
+Finally, no commitment devices are available to enforce players' promises, so the players cannot make commitments that are incredible.
+\\
+
+These assumptions (the game is dynamic, information in symmetric, and the actors have common knowledge of sequential rationality and can correctly predict each other's behavior) imply that Subgame Perfect Nash Equilibrium (SPNE) is the appropriate solution concept for this strategic situation (Watson p.
+45 and p.
+189).
 
@@
 
+Sommerlad, J.
+2022.
+``Ukraine crisis: What are the Donetsk and Luhansk People’s Republics?.'' The Independent.
+February 25, 2022.
+https://www.independent.co.uk/news/world/europe/ukraine-donetsk-luhansk-region-russia-war-b2023842.html.
+Accessed February 26, 2022.
+
 \end{document}

### 2022-04-13 teaching.tex
improve argumentation for soln concept

@@
 
-\vskip.3in
-\subsection{Timing and Information}
-The patron makes its investment. This investment is seen by all other players. Then the home government and secessionists move simultaneously.
-
-\vskip.5in
 \subsection{Payoffs}
@@
 
-\vskip.5in
+\subsection{Timing and Information}
+The patron makes its investment. This investment is seen by all other players. Then the home government and secessionists move simultaneously.\\
+
+We assume all players have complete information about payoffs; that is, all the players know the payoffs associated with each outcome.
+This assumption could be relaxed if at least one player were difficult for the other players to model (e.g., Vladimir Putin), but in many situations the players will know each other well enough that their information about payoffs will be complete.
+This is likely to be the case when the same leaders have been in place for a while so that they have interacted repeatedly.
+We assume the players know each other well for three reasons: (1) it is applicable in many situations; (2) it is simpler; and (3) it serves to highlight the issues on which we are focused.
+Thus, the only incomplete information in our strategic situation comes from the simultaneous move between the home government and secessionists.
+
 \subsection{Solution Concept}
 This game has a dynamic component, as the government and the secessionists move after seeing the patron's choice.
-Further, we assume that the players have symmetric information about payoffs; that is, all the players know the payoffs associated with each outcome.
-We will also assume that the players know each other well enough to be able to correctly predict each other's behavior; this is likely to be the case when the same leaders have been in place for a while so that they have interacted repeatedly.
-As this is a high-stakes `game,' all the players have strong incentives to seek out information and think carefully about what the other players will do before choosing the strategy that will maximize their expected payoffs, and this is common knowledge amongst the players.
-Finally, no commitment devices are available to enforce players' promises, so the players cannot make commitments that are incredible.
+Further, in the previous section we argue that the players have complete information about payoffs.
+Given our assumption that the players know each other well, it is also reasonable to assume that the players will be able to correctly predict each other's behavior.
+This is particularly true because this is a high-stakes `game,' so all the players have strong incentives to seek out information and think carefully about what the other players will do before choosing the strategy that will maximize their expected payoffs.
+As this is common knowledge amongst the players and no commitment devices are available to enforce players' promises, the players cannot make either threats or promises that are incredible.
 \\
 
-These assumptions (the game is dynamic, information in symmetric, and the actors have common knowledge of sequential rationality and can correctly predict each other's behavior) imply that Subgame Perfect Nash Equilibrium (SPNE) is the appropriate solution concept for this strategic situation (Watson p.
+These assumptions (the game has an important dynamic component, information about payoffs in complete, and the actors have common knowledge of sequential rationality and can correctly predict each other's behavior) imply that Subgame Perfect Nash Equilibrium (SPNE) is the appropriate solution concept for this strategic situation (Watson p.
 45 and p.
@@
 \subsection{Players and Strategy Spaces}
-Same as in Section 1.
+Same as in Section 2.
 
-\vskip.3in
-\subsection{Timing and Information}
-Same as in Section 1.
-
-\vskip.5in
 \subsection{Payoffs}
@@
 
-\vskip.5in
+\subsection{Timing and Information}
+Same as in Section 2.
+
+\subsection{Solution Concept}
+Same as in Section 2.
+
+\vskip.2in
 \subsection{Analysis}
@@
 
-\vskip.3in
-\subsection{Timing and Information}
-The patron makes its investment.
-This investment is seen by all other players.
-\textbf{Then the international community makes its investment.
-This investment is seen by all other players.} Then the home government and secessions move simultaneously.
-
-\vskip.5in
 \subsection{Payoffs}
@@
 \end{itemize}
+
+\subsection{Timing and Information}
+The patron makes its investment.
+This investment is seen by all other players.
+\textbf{Then the international community makes its investment.
+This investment is seen by all other players.} Then the home government and secessions move simultaneously.
+
+\subsection{Solution Concept}
+The addition of the international community does not change any of our assumptions related to the solution concept as long as the international community knows the other three players well and vice versa.
+Thus we continue to use Subgame Perfect Nash Equilibrium as our solution concept.
 
